### 2020-08-07 benchmarking/benchmarkingVM/ESS_round1_quant.tex
update

% Created by tikzDevice version 0.12.3 on 2020-08-07 14:28:34
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,289.08) rectangle (289.08,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,548.96) {\bfseries Comparison of cons (\textit{quant}) and nocons (\textit{quant})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,301.03) {ESS/chain [nocons (\textit{quant})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,433.62) {ESS/chain [cons (\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{gray}{0.91}

\path[draw=drawColor,line width= 0.1pt,line join=round,line cap=round,fill=fillColor] (  0.00,370.43) --
	(207.73,578.16) --
	(284.32,578.16) --
	(284.32,492.05) --
	(  0.00,207.73) --
	cycle;
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round,fill=fillColor] (  0.00,343.31) --
	(234.85,578.16) --
	(284.32,578.16) --
	(284.32,519.17) --
	(  0.00,234.85) --
	cycle;
\definecolor{fillColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (  0.00,316.20) --
	(261.96,578.16) --
	(284.32,578.16) --
	(284.32,546.28) --
	(  0.00,261.96) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (138.68,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,431.66) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (111.56,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,404.54) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.79,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92,377.43) {$4\times$};

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,528.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (154.78,458.84) --
	(163.04,458.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.35,457.22) --
	(159.35,460.34);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (159.35,455.61) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (224.94,510.20) --
	(228.38,510.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.73,506.80) --
	(226.73,513.10);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (226.73,506.17) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (197.40,510.73) --
	(197.94,510.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (197.67,509.26) --
	(197.67,512.10);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.67,506.70) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (150.10,455.56) --
	(157.51,455.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (154.15,454.17) --
	(154.15,456.86);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (154.15,451.53) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.84,449.55) --
	(170.33,449.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (166.15,447.40) --
	(166.15,451.49);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (166.15,446.64) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.37,520.33) --
	(233.01,520.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.69,519.95) --
	(232.69,520.70);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (232.69,517.43) {$\mu$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (224.96,506.23) --
	(230.80,506.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.10,504.49) --
	(228.10,507.82);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (228.10,503.00) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.75,524.40) --
	(212.92,524.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (210.95,522.92) --
	(210.95,525.77);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.95,520.36) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.29,515.90) --
	(230.63,515.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.46,515.51) --
	(230.46,516.29);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (230.46,511.87) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.29,493.62) --
	(212.37,493.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (210.89,492.26) --
	(210.89,494.89);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.89,489.59) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (207.98,501.27) --
	(226.00,501.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (219.00,498.36) --
	(219.00,503.79);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (219.00,498.36) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (132.74,453.24) --
	(141.29,453.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.48,446.20) --
	(137.48,458.41);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (137.48,450.02) {$p$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (219.83,509.99) --
	(224.29,509.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (222.19,509.30) --
	(222.19,510.65);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (222.19,505.95) {$L$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.03,507.85) --
	(141.88,507.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.05,507.28) --
	(140.05,508.40);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (140.05,503.81) {$r$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.17,452.60) --
	(141.38,452.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.70,447.13) --
	(137.70,456.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (137.70,448.56) {$h$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.75,425.78) --
	(152.95,425.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.15,422.66) --
	(144.15,428.46);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (144.15,422.87) {$\sigma$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.00,524.46) --
	(235.41,524.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.21,524.31) --
	(235.21,524.60);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (235.21,521.55) {$\mu$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.43,494.98) --
	(214.98,494.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (213.29,491.43) --
	(213.29,497.98);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (213.29,491.75) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (197.67,519.49) --
	(218.04,519.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (210.40,515.97) --
	(210.40,522.47);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.40,515.46) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.24,515.43) --
	(230.27,515.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.76,514.84) --
	(229.76,516.00);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (229.76,511.39) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.64,476.76) --
	(195.83,476.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (192.56,472.41) --
	(192.56,480.32);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.56,472.73) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.95,497.88) --
	(217.97,497.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.62,493.68) --
	(215.62,501.33);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (215.62,494.97) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.90,525.40) --
	(237.76,525.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.92,523.47) --
	(235.92,527.15);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (235.92,522.49) {$\mu$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (180.84,487.70) --
	(186.74,487.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.01,486.39) --
	(184.01,488.93);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.01,484.48) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.22,512.65) --
	(230.84,512.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.04,510.97) --
	(230.04,514.20);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (230.04,508.62) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.88,516.90) --
	(222.04,516.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.47,516.68) --
	(221.47,517.11);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (221.47,512.86) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (174.74,484.64) --
	(178.27,484.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.58,483.10) --
	(176.58,486.08);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (176.58,480.61) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (180.90,483.44) --
	(187.90,483.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.71,482.22) --
	(184.71,484.58);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.71,480.53) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.75,518.95) --
	(230.85,518.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.31,518.70) --
	(230.31,519.19);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (230.31,516.04) {$\mu$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (166.13,467.76) --
	(176.11,467.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.75,465.03) --
	(171.75,470.15);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (171.75,464.53) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.94,516.52) --
	(233.57,516.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.77,513.95) --
	(232.77,518.79);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (232.77,512.49) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.38,515.49) --
	(211.58,515.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.48,515.32) --
	(211.48,515.67);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.48,511.46) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.31,486.46) --
	(183.35,486.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.56,484.65) --
	(177.56,488.12);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (177.56,482.43) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.76,473.27) --
	(183.08,473.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (178.59,470.39) --
	(178.59,475.77);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.59,470.36) {$\sigma$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- (229.95,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 94.84,338.88) -- ( 94.84,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.88,338.88) -- (139.88,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.92,338.88) -- (184.92,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.95,338.88) -- (229.95,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 94.84,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (139.88,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (184.92,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (229.95,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,519.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 44.82,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,383.92) -- ( 44.82,383.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,428.96) -- ( 44.82,428.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,474.00) -- ( 44.82,474.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,519.03) -- ( 44.82,519.03);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,381.06) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,426.10) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,471.14) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,516.18) {10000};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,431.01) rectangle (189.67,426.03);
\definecolor{fillColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,421.05) rectangle (189.67,416.07);
\definecolor{fillColor}{RGB}{0,255,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,411.09) rectangle (189.67,406.11);
\definecolor{fillColor}{RGB}{255,255,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,401.13) rectangle (189.67,396.15);
\definecolor{fillColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,391.17) rectangle (189.67,386.19);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,381.21) rectangle (189.67,376.23);
\definecolor{fillColor}{RGB}{160,32,240}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,371.25) rectangle (189.67,366.27);
\definecolor{fillColor}{RGB}{0,100,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,361.29) rectangle (189.67,356.31);
\definecolor{fillColor}{RGB}{0,255,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,351.33) rectangle (189.67,346.35);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,425.66) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,415.70) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,405.74) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,395.78) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,385.82) {Moyle};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,375.86) {Ran};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,365.90) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,355.94) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,345.98) {simulated};
\end{scope}
\begin{scope}
\path[clip] (289.08,289.08) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,548.96) {\bfseries Comparison of adapt (\textit{quant}) and nocons (\textit{quant})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,301.03) {ESS/chain [nocons (\textit{quant})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,433.62) {ESS/chain [adapt (\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] (338.88,338.88) rectangle (528.36,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{gray}{0.91}

\path[draw=drawColor,line width= 0.1pt,line join=round,line cap=round,fill=fillColor] (157.53,239.42) --
	(496.27,578.16) --
	(573.70,578.16) --
	(573.70,491.81) --
	(157.53, 75.64) --
	cycle;
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round,fill=fillColor] (157.53,212.12) --
	(523.57,578.16) --
	(573.70,578.16) --
	(573.70,519.11) --
	(157.53,102.93) --
	cycle;
\definecolor{fillColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (157.53,184.82) --
	(550.86,578.16) --
	(573.70,578.16) --
	(573.70,546.40) --
	(157.53,130.23) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (428.30,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,432.20) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (401.01,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,404.91) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (372.05,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00,377.61) {$4\times$};

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,528.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.56,460.15) --
	(452.88,460.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (449.16,451.26) --
	(449.16,466.25);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.16,456.92) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.19,511.43) --
	(518.65,511.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (516.99,509.17) --
	(516.99,513.46);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (516.99,507.40) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.47,511.23) --
	(488.01,511.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.74,509.90) --
	(487.74,512.47);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (487.74,507.20) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (439.85,456.76) --
	(447.30,456.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.93,442.37) --
	(443.93,464.99);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (443.93,452.73) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (450.66,445.02) --
	(460.21,445.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.01,433.72) --
	(456.01,452.15);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (456.01,442.11) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.67,521.39) --
	(523.32,521.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.99,521.26) --
	(522.99,521.51);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (522.99,518.48) {$\mu$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.20,518.74) --
	(521.09,518.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (518.37,517.15) --
	(518.37,520.21);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (518.37,515.51) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.89,524.94) --
	(503.09,524.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.10,524.79) --
	(501.10,525.10);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (501.10,520.91) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (520.58,523.16) --
	(520.92,523.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (520.75,522.01) --
	(520.75,524.25);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (520.75,519.13) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.44,502.28) --
	(502.53,502.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.05,497.87) --
	(501.05,505.88);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (501.05,498.25) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.12,509.95) --
	(516.25,509.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (509.20,506.14) --
	(509.20,513.14);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (509.20,507.05) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.38,436.05) --
	(430.98,436.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.15,420.84) --
	(427.15,444.52);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (427.15,432.82) {$p$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.04,510.60) --
	(514.54,510.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (512.42,509.80) --
	(512.42,511.38);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (512.42,506.57) {$L$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.70,507.74) --
	(431.58,507.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.73,503.94) --
	(429.73,510.93);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (429.73,503.71) {$r$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.81,435.78) --
	(431.07,435.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.37,416.30) --
	(427.37,445.38);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (427.37,431.75) {$h$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.35,405.59) --
	(442.71,405.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.86,396.96) --
	(433.86,411.56);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (433.86,402.68) {$\sigma$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (525.32,525.14) --
	(525.72,525.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (525.52,524.81) --
	(525.52,525.46);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (525.52,522.23) {$\mu$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.59,504.01) --
	(505.16,504.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (503.46,501.14) --
	(503.46,506.51);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (503.46,500.78) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.73,516.99) --
	(508.25,516.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.55,514.10) --
	(500.55,519.51);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.55,512.95) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.52,521.82) --
	(520.55,521.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (520.04,521.15) --
	(520.04,522.46);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (520.04,517.78) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.65,484.41) --
	(485.88,484.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.59,480.26) --
	(482.59,487.83);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (482.59,480.37) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (503.11,505.33) --
	(508.17,505.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.81,502.96) --
	(505.81,507.44);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (505.81,502.42) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (524.20,526.64) --
	(528.09,526.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.24,525.57) --
	(526.24,527.65);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (526.24,523.73) {$\mu$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.79,487.52) --
	(476.73,487.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (473.99,482.71) --
	(473.99,491.39);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (473.99,484.30) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.49,514.36) --
	(521.12,514.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (520.32,512.66) --
	(520.32,515.93);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (520.32,510.33) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.10,518.21) --
	(512.27,518.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.70,516.35) --
	(511.70,519.91);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (511.70,514.18) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.66,484.99) --
	(468.20,484.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (466.51,478.51) --
	(466.51,489.86);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (466.51,480.96) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.85,484.00) --
	(477.90,484.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.69,481.54) --
	(474.69,486.20);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.69,481.10) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (520.03,520.25) --
	(521.13,520.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (520.59,519.78) --
	(520.59,520.70);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (520.59,517.34) {$\mu$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (455.99,472.35) --
	(466.03,472.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.64,469.91) --
	(461.64,474.53);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (461.64,469.13) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.24,517.24) --
	(523.87,517.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (523.07,515.10) --
	(523.07,519.18);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (523.07,513.21) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.54,517.24) --
	(501.74,517.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.64,517.01) --
	(501.64,517.47);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (501.64,513.21) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.19,486.19) --
	(473.32,486.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.49,483.53) --
	(467.49,488.53);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (467.49,482.16) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.66,476.31) --
	(473.05,476.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (468.53,474.09) --
	(468.53,478.31);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (468.53,473.41) {$\sigma$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (520.23,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.22,338.88) -- (384.22,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.56,338.88) -- (429.56,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.90,338.88) -- (474.90,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (520.23,338.88) -- (520.23,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (384.22,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (429.56,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (474.90,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (520.23,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,520.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (333.90,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,384.22) -- (333.90,384.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,429.56) -- (333.90,429.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,474.90) -- (333.90,474.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,520.23) -- (333.90,520.23);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,381.36) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,426.70) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,472.04) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,517.38) {10000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,259.88) {\bfseries Comparison of adapt (\textit{quant}) and cons (\textit{quant})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54, 11.95) {ESS/chain [cons (\textit{quant})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,144.54) {ESS/chain [adapt (\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80, 49.80) rectangle (239.28,239.28);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{gray}{0.91}

\path[draw=drawColor,line width= 0.1pt,line join=round,line cap=round,fill=fillColor] (  0.00, 81.35) --
	(284.32,365.67) --
	(284.32,202.97) --
	( 81.35,  0.00) --
	(  0.00,  0.00) --
	cycle;
\definecolor{fillColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round,fill=fillColor] (  0.00, 54.23) --
	(284.32,338.55) --
	(284.32,230.09) --
	( 54.23,  0.00) --
	(  0.00,  0.00) --
	cycle;
\definecolor{fillColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (  0.00, 27.12) --
	(284.32,311.43) --
	(284.32,257.20) --
	( 27.12,  0.00) --
	(  0.00,  0.00) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (138.68, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,142.58) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (111.56, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,115.46) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.79, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92, 88.35) {$4\times$};

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28,239.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (168.14,170.26) --
	(171.26,170.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.76,161.43) --
	(169.76,176.32);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.76,167.04) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.72,221.21) --
	(224.02,221.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.12,218.96) --
	(221.12,223.22);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (221.12,217.17) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.18,221.01) --
	(223.02,221.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.65,219.69) --
	(221.65,222.24);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (221.65,216.97) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (165.09,166.90) --
	(167.78,166.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (166.48,152.60) --
	(166.48,175.07);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (166.48,162.87) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.32,155.24) --
	(162.41,155.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.47,144.01) --
	(160.47,162.32);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.47,152.33) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.87,231.10) --
	(231.62,231.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.25,230.97) --
	(231.25,231.22);
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (231.25,228.19) {$\mu$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.41,228.47) --
	(218.74,228.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.15,226.88) --
	(217.15,229.93);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (217.15,225.24) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.84,234.63) --
	(236.69,234.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.32,234.47) --
	(235.32,234.79);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (235.32,230.60) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.43,232.86) --
	(227.21,232.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.82,231.71) --
	(226.82,233.94);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (226.82,228.83) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.18,212.12) --
	(205.81,212.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.54,207.74) --
	(204.54,215.70);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.54,208.09) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.28,219.74) --
	(214.71,219.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.19,215.95) --
	(212.19,222.91);
\definecolor{drawColor}{RGB}{255,165,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.19,216.83) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.12,146.32) --
	(169.33,146.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (164.16,131.22) --
	(164.16,154.74);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (164.16,143.10) {$p$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,220.39) --
	(221.57,220.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.91,219.58) --
	(220.91,221.16);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (220.91,216.35) {$L$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.20,217.55) --
	(219.32,217.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.77,213.77) --
	(218.77,220.71);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.77,213.51) {$r$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.05,146.06) --
	(167.79,146.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.52,126.71) --
	(163.52,155.60);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (163.52,142.03) {$h$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.58,116.07) --
	(139.38,116.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (136.70,107.50) --
	(136.70,122.00);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (136.70,113.16) {$\sigma$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.23,234.82) --
	(235.52,234.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.38,234.50) --
	(235.38,235.14);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (235.38,231.92) {$\mu$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (202.35,213.84) --
	(208.90,213.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (205.90,210.98) --
	(205.90,216.32);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.90,210.61) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.89,226.73) --
	(233.39,226.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.41,223.86) --
	(230.41,229.23);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (230.41,222.70) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.76,231.53) --
	(226.92,231.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.35,230.87) --
	(226.35,232.16);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (226.35,227.49) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (183.33,194.36) --
	(191.24,194.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.68,190.24) --
	(187.68,197.76);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.68,190.33) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.60,215.14) --
	(212.25,215.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.80,212.80) --
	(208.80,217.24);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.80,212.24) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.39,236.31) --
	(238.07,236.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.32,235.26) --
	(236.32,237.32);
\definecolor{drawColor}{RGB}{160,32,240}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (236.32,233.41) {$\mu$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (197.31,197.46) --
	(199.85,197.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.62,192.67) --
	(198.62,201.30);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (198.62,194.23) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.89,224.12) --
	(225.12,224.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.57,222.43) --
	(223.57,225.68);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (223.57,220.09) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.60,227.94) --
	(228.03,227.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.82,226.10) --
	(227.82,229.63);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.82,223.91) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.02,194.94) --
	(197.00,194.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.56,188.51) --
	(195.56,199.78);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.56,190.91) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (193.14,193.96) --
	(195.50,193.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.36,191.51) --
	(194.36,196.14);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.36,191.06) {$\sigma$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.62,229.97) --
	(230.11,229.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.87,229.50) --
	(229.87,230.42);
\definecolor{drawColor}{RGB}{0,100,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (229.87,227.06) {$\mu$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (175.95,182.39) --
	(181.07,182.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (178.68,179.96) --
	(178.68,184.55);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.68,179.16) {$p$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (224.87,226.98) --
	(229.71,226.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.44,224.85) --
	(227.44,228.90);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.44,222.95) {$L$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.24,226.98) --
	(226.59,226.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.41,226.76) --
	(226.41,227.21);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (226.41,222.95) {$r$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.57,196.14) --
	(199.04,196.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (197.38,193.50) --
	(197.38,198.46);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.38,192.10) {$h$};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.31,186.32) --
	(186.69,186.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.19,184.11) --
	(184.19,188.31);
\definecolor{drawColor}{RGB}{0,255,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.19,183.42) {$\sigma$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- (229.95, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 94.84, 49.80) -- ( 94.84, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.88, 49.80) -- (139.88, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.92, 49.80) -- (184.92, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.95, 49.80) -- (229.95, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 94.84, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (139.88, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (184.92, 31.87) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (229.95, 31.87) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80,229.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 44.82, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 94.84) -- ( 44.82, 94.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,139.88) -- ( 44.82,139.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,184.92) -- ( 44.82,184.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,229.95) -- ( 44.82,229.95);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 91.98) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,137.02) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,182.06) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,227.10) {10000};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.66] at (433.62,258.45) {\bfseries Speed comparison};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.66] at (433.62, 11.95) {Setting};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.66] at (307.01,144.54) {Speed ($10^6$ states/hr)};
\end{scope}
\begin{scope}
\path[clip] (338.88, 49.80) rectangle (528.36,239.28);
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.88, 78.53) -- (428.64, 77.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.60, 77.16) -- (516.36, 77.22);
\definecolor{fillColor}{RGB}{255,0,0}

\path[fill=fillColor] (345.90, 81.52) --
	(348.41, 77.16) --
	(343.38, 77.16) --
	cycle;

\path[fill=fillColor] (433.62, 77.16) circle (  1.87);

\path[fill=fillColor] (521.34, 77.23) circle (  1.87);
\definecolor{drawColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.86,117.42) -- (428.65,111.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.60,111.30) -- (516.36,110.92);
\definecolor{fillColor}{RGB}{255,165,0}

\path[fill=fillColor] (345.90,120.70) --
	(348.41,116.34) --
	(343.38,116.34) --
	cycle;

\path[fill=fillColor] (433.62,111.32) circle (  1.87);

\path[fill=fillColor] (521.34,110.89) circle (  1.87);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.88, 68.40) -- (428.64, 68.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.60, 68.43) -- (516.36, 68.52);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (345.90, 68.39) circle (  1.87);

\path[fill=fillColor] (433.62, 68.42) circle (  1.87);

\path[fill=fillColor] (521.34, 71.43) --
	(523.86, 67.07) --
	(518.83, 67.07) --
	cycle;
\definecolor{drawColor}{RGB}{160,32,240}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.88, 99.16) -- (428.64, 98.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.60, 98.01) -- (516.36, 96.50);
\definecolor{fillColor}{RGB}{160,32,240}

\path[fill=fillColor] (345.90,102.12) --
	(348.41, 97.77) --
	(343.38, 97.77) --
	cycle;

\path[fill=fillColor] (433.62, 98.11) circle (  1.87);

\path[fill=fillColor] (521.34, 96.40) circle (  1.87);
\definecolor{drawColor}{RGB}{0,100,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.88,103.24) -- (428.64,101.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.60,101.14) -- (516.36,100.31);
\definecolor{fillColor}{RGB}{0,100,0}

\path[fill=fillColor] (345.90,106.27) --
	(348.41,101.92) --
	(343.38,101.92) --
	cycle;

\path[fill=fillColor] (433.62,101.20) circle (  1.87);

\path[fill=fillColor] (521.34,100.26) circle (  1.87);
\definecolor{drawColor}{RGB}{0,255,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.88,109.78) -- (428.64,111.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.60,111.46) -- (516.36,111.39);
\definecolor{fillColor}{RGB}{0,255,255}

\path[fill=fillColor] (345.90,109.68) circle (  1.87);

\path[fill=fillColor] (433.62,114.37) --
	(436.14,110.02) --
	(431.10,110.02) --
	cycle;

\path[fill=fillColor] (521.34,111.38) circle (  1.87);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.90, 49.80) -- (521.34, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.90, 49.80) -- (345.90, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.62, 49.80) -- (433.62, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (521.34, 49.80) -- (521.34, 44.82);

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (348.76, 39.84) {nocons (\textit{quant})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (436.48, 39.84) {cons (\textit{quant})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (524.20, 39.84) {adapt (\textit{quant})};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 56.82) -- (338.88,216.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 56.82) -- (333.90, 56.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 96.66) -- (333.90, 96.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,136.50) -- (333.90,136.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,176.33) -- (333.90,176.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,216.17) -- (333.90,216.17);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (326.93, 56.82) {0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (326.93, 96.66) {10};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (326.93,136.50) {20};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (326.93,176.33) {30};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (326.93,216.17) {40};
\end{scope}
\end{tikzpicture}

\end{document}