### 2020-08-24 benchmarking/benchmarkingVM/ESS_round1_catquant.tex
results update

% Created by tikzDevice version 0.12.3.1 on 2020-08-24 19:04:32
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,289.08) rectangle (289.08,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,548.96) {\bfseries cons (\textit{quant}) vs. nocons (\textit{quant})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,301.03) {ESS/chain [nocons (\textit{quant})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,433.62) {ESS/chain [cons (\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,368.65) --
	(209.51,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,492.84) --
	(  0.00,209.51);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (  0.00,342.13) --
	(236.03,578.16);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,519.37) --
	(  0.00,236.03);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (  0.00,315.60) --
	(262.56,578.16);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,545.89) --
	(  0.00,262.56);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92,376.84) {$4\times$};

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.55,466.16) --
	(158.29,466.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (153.31,462.75) --
	(153.31,469.06);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (153.31,462.93) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.24,515.97) --
	(190.72,515.97);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.95,514.61) --
	(186.95,517.24);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.95,511.94) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (141.27,507.51) --
	(146.34,507.51);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (143.97,507.30) --
	(143.97,507.72);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (143.97,503.48) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (141.32,461.89) --
	(150.41,461.89);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.40,454.67) --
	(146.40,467.11);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (146.40,457.85) {$h$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (107.73,439.48) --
	(121.71,439.48);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (115.97,429.54) --
	(115.97,445.99);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (115.97,436.58) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.56,513.60) --
	(225.34,513.60);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.96,513.22) --
	(224.96,513.97);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (224.96,510.69) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (212.80,499.16) --
	(214.49,499.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (213.66,497.99) --
	(213.66,500.27);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (213.66,495.94) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.38,483.91) --
	(193.63,483.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.01,482.28) --
	(193.01,485.40);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.01,479.87) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (219.89,514.97) --
	(220.35,514.97);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (220.12,514.79) --
	(220.12,515.14);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (220.12,510.93) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (203.60,494.30) --
	(207.53,494.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (205.67,493.23) --
	(205.67,495.31);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.67,490.27) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (214.64,497.73) --
	(216.55,497.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (215.62,496.54) --
	(215.62,498.85);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (215.62,494.83) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.49,514.56) --
	(225.65,514.56);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.08,514.02) --
	(225.08,515.09);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.08,511.66) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (170.28,470.40) --
	(175.14,470.40);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.86,468.37) --
	(172.86,472.24);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (172.86,467.18) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (201.34,496.33) --
	(204.09,496.33);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (202.76,494.98) --
	(202.76,497.58);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.76,492.29) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (223.55,516.52) --
	(224.53,516.52);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.05,516.26) --
	(224.05,516.77);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (224.05,512.48) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (159.45,457.17) --
	(165.46,457.17);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (162.69,454.81) --
	(162.69,459.28);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (162.69,453.14) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (180.00,474.69) --
	(190.72,474.69);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.11,471.18) --
	(186.11,477.64);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.11,471.78) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.77,516.29) --
	(228.21,516.29);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.99,515.94) --
	(227.99,516.64);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.99,513.39) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.06,489.34) --
	(190.32,489.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.76,488.63) --
	(188.76,490.02);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.76,486.11) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.99,489.04) --
	(194.37,489.04);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.31,486.77) --
	(192.31,491.07);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.31,485.01) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (218.80,514.30) --
	(219.31,514.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (219.06,514.16) --
	(219.06,514.43);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (219.06,510.26) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (175.50,485.31) --
	(182.06,485.31);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.06,483.39) --
	(179.06,487.05);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.06,481.27) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (178.00,481.50) --
	(186.35,481.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.63,479.81) --
	(182.63,483.06);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.63,478.60) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.74,516.74) --
	(228.17,516.74);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.96,516.39) --
	(227.96,517.08);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.96,513.84) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (162.10,466.73) --
	(168.36,466.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.49,463.57) --
	(165.49,469.44);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.49,463.50) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (198.81,509.25) --
	(205.39,509.25);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (202.39,507.51) --
	(202.39,510.84);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.39,505.21) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (197.30,509.61) --
	(198.11,509.61);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (197.71,509.14) --
	(197.71,510.08);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.71,505.58) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (153.88,467.30) --
	(166.55,467.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (161.25,464.25) --
	(161.25,469.93);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (161.25,463.27) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (168.45,456.76) --
	(175.22,456.76);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.13,453.03) --
	(172.13,459.89);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (172.13,453.86) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.42,514.50) --
	(226.46,514.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.95,514.10) --
	(225.95,514.89);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.95,511.59) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (223.12,507.57) --
	(225.42,507.57);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.30,506.14) --
	(224.30,508.89);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (224.30,504.34) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (206.07,496.58) --
	(208.33,496.58);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (207.23,495.41) --
	(207.23,497.68);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.23,492.54) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (235.17,517.38) --
	(235.44,517.38);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (235.30,517.27) --
	(235.30,517.48);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (235.30,513.34) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (197.08,487.47) --
	(198.88,487.47);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (198.00,485.85) --
	(198.00,488.96);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (198.00,483.44) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (207.94,497.31) --
	(210.40,497.31);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (209.21,494.40) --
	(209.21,499.82);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (209.21,494.40) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.56,516.29) --
	(227.64,516.29);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.11,515.86) --
	(227.11,516.70);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.11,513.38) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (180.20,477.66) --
	(185.95,477.66);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.29,474.79) --
	(183.29,480.15);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.29,474.43) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.14,506.92) --
	(226.44,506.92);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.33,505.18) --
	(225.33,508.51);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.33,502.88) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (211.48,515.51) --
	(213.95,515.51);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (212.75,515.17) --
	(212.75,515.84);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.75,511.47) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.69,490.11) --
	(189.00,490.11);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.70,488.43) --
	(185.70,491.66);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.70,486.08) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (184.19,479.36) --
	(189.91,479.36);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.26,477.87) --
	(187.26,480.75);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.26,476.46) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.82,517.74) --
	(230.68,517.74);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (229.77,516.59) --
	(229.77,518.83);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (229.77,514.84) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (215.58,501.69) --
	(218.44,501.69);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (217.06,500.72) --
	(217.06,502.62);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (217.06,498.46) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.05,492.78) --
	(195.08,492.78);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.63,489.01) --
	(193.63,495.94);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.63,488.75) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.29,514.73) --
	(228.68,514.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.00,514.55) --
	(228.00,514.90);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (228.00,510.70) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.17,489.32) --
	(202.86,489.32);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (199.40,483.15) --
	(199.40,493.99);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.40,485.29) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.54,499.38) --
	(205.81,499.38);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.64,496.95) --
	(200.64,501.54);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.64,496.47) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (238.32,528.60) --
	(239.74,528.60);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (239.05,527.61) --
	(239.05,529.55);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (239.05,525.70) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.31,507.29) --
	(227.66,507.29);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.52,506.61) --
	(226.52,507.94);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (226.52,504.06) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (215.46,502.68) --
	(216.79,502.68);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (216.13,500.67) --
	(216.13,504.49);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (216.13,498.64) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (230.05,515.15) --
	(230.28,515.15);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (230.16,514.85) --
	(230.16,515.45);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (230.16,511.12) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (212.27,497.33) --
	(214.57,497.33);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (213.45,496.08) --
	(213.45,498.51);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (213.45,493.30) {$h$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (218.24,500.49) --
	(221.78,500.49);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (220.09,499.55) --
	(220.09,501.39);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (220.09,497.59) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (239.91,530.16) --
	(241.03,530.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (240.48,529.47) --
	(240.48,530.82);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (240.48,527.25) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- (226.02,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85,338.88) -- ( 93.85,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91,338.88) -- (137.91,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96,338.88) -- (181.96,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02,338.88) -- (226.02,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 44.82,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,382.93) -- ( 44.82,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,426.99) -- ( 44.82,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,471.04) -- ( 44.82,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,515.10) -- ( 44.82,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,431.01) rectangle (189.67,426.03);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,421.05) rectangle (189.67,416.07);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,411.09) rectangle (189.67,406.11);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,401.13) rectangle (189.67,396.15);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,391.17) rectangle (189.67,386.19);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,381.21) rectangle (189.67,376.23);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,371.25) rectangle (189.67,366.27);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,361.29) rectangle (189.67,356.31);
\definecolor{fillColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,351.33) rectangle (189.67,346.35);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,425.66) {Ran};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,415.70) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,405.74) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,395.78) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,385.82) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,375.86) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,365.90) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,355.94) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,345.98) {Moyle};
\end{scope}
\begin{scope}
\path[clip] (289.08,289.08) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,548.96) {\bfseries adapt (\textit{quant}) vs. nocons (\textit{quant})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,301.03) {ESS/chain [nocons (\textit{quant})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,433.62) {ESS/chain [adapt (\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] (338.88,338.88) rectangle (528.36,528.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.66,242.23) --
	(498.59,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,492.84) --
	(162.66, 83.09) --
	(162.66,242.23);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (162.66,215.71) --
	(525.11,578.16);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,519.37) --
	(162.66,109.61) --
	(162.66,215.71);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (162.66,189.19) --
	(551.64,578.16);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,545.89) --
	(162.66,136.14) --
	(162.66,189.19);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (425.98,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (399.46,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (371.28,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00,376.84) {$4\times$};

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (435.63,486.24) --
	(447.37,486.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (442.39,483.71) --
	(442.39,488.48);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (442.39,483.01) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (471.32,515.98) --
	(479.80,515.98);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (476.03,515.21) --
	(476.03,516.73);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.03,511.95) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (430.35,510.08) --
	(435.42,510.08);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (433.05,508.58) --
	(433.05,511.48);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (433.05,506.05) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (430.40,481.95) --
	(439.49,481.95);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (435.48,475.06) --
	(435.48,487.01);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (435.48,477.92) {$h$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (396.81,494.23) --
	(410.79,494.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (405.05,489.73) --
	(405.05,497.87);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (405.05,491.33) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (513.64,514.61) --
	(514.42,514.61);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.04,513.26) --
	(514.04,515.87);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (514.04,511.71) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (501.88,502.88) --
	(503.57,502.88);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (502.74,501.44) --
	(502.74,504.21);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (502.74,499.65) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.46,484.45) --
	(482.71,484.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (482.09,483.64) --
	(482.09,485.23);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (482.09,480.42) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (508.97,515.82) --
	(509.43,515.82);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (509.20,515.68) --
	(509.20,515.96);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (509.20,511.79) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (492.68,496.95) --
	(496.61,496.95);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (494.75,495.13) --
	(494.75,498.62);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (494.75,492.92) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (503.72,512.60) --
	(505.63,512.60);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (504.70,511.28) --
	(504.70,513.84);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (504.70,509.70) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (513.57,514.28) --
	(514.73,514.28);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.16,513.85) --
	(514.16,514.70);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (514.16,511.38) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (459.36,476.21) --
	(464.22,476.21);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (461.94,471.55) --
	(461.94,479.95);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (461.94,472.98) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (490.42,496.44) --
	(493.17,496.44);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (491.84,495.62) --
	(491.84,497.22);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (491.84,492.41) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (512.63,518.61) --
	(513.61,518.61);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (513.13,518.39) --
	(513.13,518.83);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (513.13,514.58) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (448.53,462.00) --
	(454.54,462.00);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (451.77,456.27) --
	(451.77,466.41);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (451.77,457.97) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (469.08,485.00) --
	(479.80,485.00);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (475.19,480.84) --
	(475.19,488.42);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (475.19,482.10) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (516.85,516.49) --
	(517.29,516.49);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (517.07,516.20) --
	(517.07,516.78);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (517.07,513.59) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (476.14,489.90) --
	(479.40,489.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (477.84,488.38) --
	(477.84,491.32);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (477.84,486.68) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (479.07,490.25) --
	(483.45,490.25);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.39,488.30) --
	(481.39,492.02);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (481.39,486.22) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (507.88,515.74) --
	(508.39,515.74);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (508.14,515.62) --
	(508.14,515.85);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (508.14,511.70) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (464.58,489.96) --
	(471.14,489.96);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (468.14,488.02) --
	(468.14,491.72);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (468.14,485.93) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (467.08,493.91) --
	(475.43,493.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (471.71,491.52) --
	(471.71,496.04);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (471.71,491.01) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (516.82,517.00) --
	(517.25,517.00);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (517.04,516.92) --
	(517.04,517.07);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (517.04,514.09) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (451.18,471.62) --
	(457.44,471.62);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (454.57,470.11) --
	(454.57,473.01);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (454.57,468.39) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (487.89,510.00) --
	(494.47,510.00);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (491.47,507.87) --
	(491.47,511.92);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (491.47,505.97) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (486.38,508.84) --
	(487.19,508.84);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (486.79,508.29) --
	(486.79,509.38);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (486.79,504.81) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (442.96,470.99) --
	(455.63,470.99);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (450.33,469.17) --
	(450.33,472.65);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (450.33,466.96) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (457.53,483.25) --
	(464.30,483.25);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (461.21,480.13) --
	(461.21,485.92);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (461.21,480.34) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.50,514.97) --
	(515.54,514.97);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (515.03,514.65) --
	(515.03,515.28);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (515.03,512.07) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (512.20,516.70) --
	(514.50,516.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (513.38,515.91) --
	(513.38,517.46);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (513.38,513.48) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (495.15,498.61) --
	(497.41,498.61);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (496.31,495.49) --
	(496.31,501.29);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (496.31,494.57) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (524.25,525.98) --
	(524.52,525.98);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (524.38,525.81) --
	(524.38,526.15);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (524.38,521.95) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (486.16,488.18) --
	(487.96,488.18);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (487.08,486.37) --
	(487.08,489.84);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (487.08,484.15) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (497.02,502.47) --
	(499.48,502.47);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (498.29,501.02) --
	(498.29,503.81);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (498.29,499.56) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (515.64,516.37) --
	(516.72,516.37);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (516.19,516.06) --
	(516.19,516.68);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (516.19,513.47) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (469.28,480.49) --
	(475.03,480.49);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (472.37,476.54) --
	(472.37,483.75);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (472.37,477.26) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (513.22,512.09) --
	(515.52,512.09);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.41,510.16) --
	(514.41,513.84);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (514.41,508.05) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (500.56,516.42) --
	(503.03,516.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (501.83,516.22) --
	(501.83,516.60);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (501.83,512.38) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (470.77,488.99) --
	(478.08,488.99);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (474.78,483.50) --
	(474.78,493.25);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.78,484.95) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (473.27,494.00) --
	(478.99,494.00);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (476.34,490.82) --
	(476.34,496.72);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.34,491.09) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (517.90,518.35) --
	(519.76,518.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (518.85,517.20) --
	(518.85,519.44);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (518.85,515.45) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (504.66,505.50) --
	(507.52,505.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (506.14,502.97) --
	(506.14,507.73);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (506.14,502.27) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.13,487.87) --
	(484.16,487.87);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (482.71,484.86) --
	(482.71,490.48);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (482.71,483.84) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (516.37,519.55) --
	(517.76,519.55);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (517.08,519.31) --
	(517.08,519.79);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (517.08,515.52) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (484.25,490.54) --
	(491.94,490.54);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (488.48,486.19) --
	(488.48,494.09);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (488.48,486.51) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (482.62,504.83) --
	(494.89,504.83);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (489.72,500.57) --
	(489.72,508.31);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (489.72,501.92) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (527.40,528.64) --
	(528.82,528.64);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (528.13,528.52) --
	(528.13,528.76);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (528.13,525.74) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.39,514.92) --
	(516.74,514.92);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (515.60,507.67) --
	(515.60,520.16);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (515.60,511.69) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (504.54,501.52) --
	(505.87,501.52);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (505.21,494.65) --
	(505.21,506.56);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (505.21,497.48) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (519.13,518.27) --
	(519.36,518.27);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (519.24,510.35) --
	(519.24,523.85);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (519.24,514.23) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (501.35,500.17) --
	(503.65,500.17);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (502.53,491.51) --
	(502.53,506.10);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (502.53,496.13) {$h$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (507.32,509.87) --
	(510.86,509.87);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (509.17,501.95) --
	(509.17,515.45);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (509.17,506.96) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (528.99,527.10) --
	(530.11,527.10);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (529.56,518.90) --
	(529.56,532.82);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (529.56,524.20) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (515.10,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.93,338.88) -- (382.93,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.99,338.88) -- (426.99,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.04,338.88) -- (471.04,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.10,338.88) -- (515.10,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (382.93,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (426.99,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (471.04,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (515.10,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (333.90,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,382.93) -- (333.90,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,426.99) -- (333.90,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,471.04) -- (333.90,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,515.10) -- (333.90,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,259.88) {\bfseries adapt (\textit{quant}) vs. cons (\textit{quant})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54, 11.95) {ESS/chain [cons (\textit{quant})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,144.54) {ESS/chain [adapt (\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80, 49.80) rectangle (239.28,239.28);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 79.57) --
	(283.33,362.91) --
	(283.33,203.76) --
	( 79.57,  0.00);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (  0.00, 53.05) --
	(283.33,336.38) --
	(283.33,230.29) --
	( 53.05,  0.00);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (  0.00, 26.52) --
	(283.33,309.86) --
	(283.33,256.81) --
	( 26.52,  0.00);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,140.80) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,114.28) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92, 87.76) {$4\times$};

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28,239.28);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (173.67,197.16) --
	(179.98,197.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (177.08,194.63) --
	(177.08,199.40);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (177.08,193.93) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.53,226.90) --
	(228.16,226.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.89,226.13) --
	(226.89,227.65);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (226.89,222.87) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (218.22,221.00) --
	(218.64,221.00);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (218.43,219.50) --
	(218.43,222.40);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.43,216.97) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.59,192.87) --
	(178.03,192.87);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.81,185.98) --
	(172.81,197.93);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (172.81,188.84) {$h$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (140.46,205.15) --
	(156.91,205.15);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (150.40,200.65) --
	(150.40,208.79);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (150.40,202.25) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.14,225.53) --
	(224.89,225.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.52,224.18) --
	(224.52,226.79);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (224.52,222.63) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.91,213.80) --
	(211.19,213.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (210.08,212.36) --
	(210.08,215.13);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.08,210.57) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.20,195.37) --
	(196.32,195.37);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.83,194.56) --
	(194.83,196.15);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.83,191.34) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.71,226.74) --
	(226.06,226.74);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.89,226.60) --
	(225.89,226.88);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.89,222.71) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (204.15,207.87) --
	(206.23,207.87);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (205.22,206.05) --
	(205.22,209.54);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.22,203.84) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (207.46,223.52) --
	(209.77,223.52);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.65,222.20) --
	(208.65,224.76);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.65,220.62) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.94,225.20) --
	(226.01,225.20);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.48,224.77) --
	(225.48,225.62);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.48,222.30) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.29,187.13) --
	(183.16,187.13);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.32,182.47) --
	(181.32,190.87);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.32,183.90) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (205.90,207.36) --
	(208.50,207.36);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (207.25,206.54) --
	(207.25,208.14);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.25,203.33) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.18,229.53) --
	(227.69,229.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.44,229.31) --
	(227.44,229.75);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.44,225.50) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.73,172.92) --
	(170.20,172.92);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (168.09,167.19) --
	(168.09,177.33);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (168.09,168.89) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.10,195.92) --
	(188.56,195.92);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.61,191.76) --
	(185.61,199.34);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.61,193.02) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.86,227.41) --
	(227.56,227.41);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.21,227.12) --
	(227.21,227.70);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.21,224.51) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (199.55,200.82) --
	(200.94,200.82);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.26,199.30) --
	(200.26,202.24);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.26,197.60) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (197.69,201.17) --
	(201.99,201.17);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (199.96,199.22) --
	(199.96,202.94);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.96,197.14) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.08,226.66) --
	(225.35,226.66);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.22,226.54) --
	(225.22,226.77);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.22,222.62) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.31,200.88) --
	(197.97,200.88);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (196.23,198.94) --
	(196.23,202.64);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (196.23,196.85) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.73,204.83) --
	(193.98,204.83);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.42,202.44) --
	(192.42,206.96);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.42,201.93) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.31,227.92) --
	(228.00,227.92);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.66,227.84) --
	(227.66,227.99);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.66,225.01) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (174.49,182.54) --
	(180.36,182.54);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (177.65,181.03) --
	(177.65,183.93);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (177.65,179.31) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (218.43,220.92) --
	(221.76,220.92);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (220.17,218.79) --
	(220.17,222.84);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (220.17,216.89) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (220.06,219.76) --
	(221.00,219.76);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (220.53,219.21) --
	(220.53,220.30);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (220.53,215.73) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (175.17,181.91) --
	(180.85,181.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (178.22,180.09) --
	(178.22,183.57);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.22,177.88) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (163.95,194.17) --
	(170.81,194.17);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (167.68,191.05) --
	(167.68,196.84);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (167.68,191.26) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.02,225.89) --
	(225.81,225.89);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.42,225.57) --
	(225.42,226.20);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.42,222.99) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (217.06,227.62) --
	(219.81,227.62);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (218.49,226.83) --
	(218.49,228.38);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.49,224.40) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (206.33,209.53) --
	(208.60,209.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (207.50,206.41) --
	(207.50,212.21);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.50,205.49) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.19,236.90) --
	(228.40,236.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.30,236.73) --
	(228.30,237.07);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (228.30,232.87) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (196.77,199.10) --
	(199.88,199.10);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (198.39,197.29) --
	(198.39,200.76);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (198.39,195.07) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (205.32,213.39) --
	(210.74,213.39);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.23,211.94) --
	(208.23,214.73);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.23,210.48) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.78,227.29) --
	(227.62,227.29);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.21,226.98) --
	(227.21,227.60);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.21,224.39) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.71,191.41) --
	(191.07,191.41);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.58,187.46) --
	(188.58,194.67);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.58,188.18) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (216.10,223.01) --
	(219.43,223.01);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (217.84,221.08) --
	(217.84,224.76);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (217.84,218.97) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.09,227.34) --
	(226.76,227.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.43,227.14) --
	(226.43,227.52);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (226.43,223.30) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (199.35,199.91) --
	(202.58,199.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (201.03,194.42) --
	(201.03,204.17);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (201.03,195.87) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.79,204.92) --
	(191.67,204.92);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.28,201.74) --
	(190.28,207.64);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (190.28,202.01) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.51,229.27) --
	(229.75,229.27);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.66,228.12) --
	(228.66,230.36);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (228.66,226.37) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (211.64,216.42) --
	(213.54,216.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (212.61,213.89) --
	(212.61,218.65);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.61,213.19) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (199.93,198.79) --
	(206.86,198.79);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (203.70,195.78) --
	(203.70,201.40);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.70,194.76) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.47,230.47) --
	(225.82,230.47);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.65,230.23) --
	(225.65,230.71);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.65,226.44) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.07,201.46) --
	(204.91,201.46);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.24,197.11) --
	(200.24,205.01);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.24,197.43) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (207.87,215.75) --
	(212.46,215.75);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (210.30,211.49) --
	(210.30,219.23);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.30,212.84) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (238.53,239.56) --
	(240.47,239.56);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (239.52,239.44) --
	(239.52,239.68);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (239.52,236.66) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (217.53,225.84) --
	(218.86,225.84);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (218.21,218.59) --
	(218.21,231.08);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.21,222.61) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (211.59,212.44) --
	(215.41,212.44);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (213.60,205.57) --
	(213.60,217.48);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (213.60,208.40) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.77,229.19) --
	(226.37,229.19);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.07,221.27) --
	(226.07,234.77);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (226.07,225.15) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (207.00,211.09) --
	(209.43,211.09);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.25,202.43) --
	(208.25,217.02);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.25,207.05) {$h$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (210.47,220.79) --
	(212.31,220.79);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (211.41,212.87) --
	(211.41,226.37);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.41,217.88) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (240.39,238.02) --
	(241.74,238.02);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (241.08,229.82) --
	(241.08,243.74);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (241.08,235.12) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- (226.02, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85, 49.80) -- ( 93.85, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91, 49.80) -- (137.91, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96, 49.80) -- (181.96, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02, 49.80) -- (226.02, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96, 31.87) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02, 31.87) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80,226.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 44.82, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 93.85) -- ( 44.82, 93.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,137.91) -- ( 44.82,137.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,181.96) -- ( 44.82,181.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,226.02) -- ( 44.82,226.02);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 91.00) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,135.05) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,179.11) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,223.16) {10000};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,259.88) {\bfseries adapt (\textit{cat}) vs. \textit{cat}};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62, 11.95) {ESS/chain [\textit{cat}]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,144.54) {ESS/chain [adapt (\textit{cat})]};
\end{scope}
\begin{scope}
\path[clip] (338.88, 49.80) rectangle (528.36,239.28);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.51,  0.00) --
	(572.41,362.91) --
	(572.41,203.76) --
	(368.65,  0.00);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (236.03,  0.00) --
	(572.41,336.38) --
	(572.41,230.29) --
	(342.13,  0.00);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (262.56,  0.00) --
	(572.41,309.86) --
	(572.41,256.81) --
	(315.60,  0.00);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (425.98, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,140.80) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (399.46, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,114.28) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (371.28, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00, 87.76) {$4\times$};

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (338.88, 49.80) -- (528.36,239.28);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (402.48,133.81) --
	(427.20,133.81);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (418.58,122.95) --
	(418.58,140.70);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (418.58,130.59) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (463.96,183.86) --
	(498.40,183.86);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (488.07,166.42) --
	(488.07,192.82);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (488.07,179.82) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (412.32,131.41) --
	(425.95,131.41);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (420.32,125.62) --
	(420.32,135.84);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (420.32,127.37) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (412.59,133.07) --
	(432.70,133.07);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (425.17,121.05) --
	(425.17,140.40);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (425.17,129.04) {$h$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (386.96,106.84) --
	(412.97,106.84);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (404.08, 99.96) --
	(404.08,111.88);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (404.08,103.93) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.29,225.53) --
	(514.97,225.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.63,225.34) --
	(514.63,225.72);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (514.63,222.63) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (497.79,213.90) --
	(503.46,213.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (500.83,212.35) --
	(500.83,215.34);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.83,210.68) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.90,193.69) --
	(482.27,193.69);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (480.66,192.49) --
	(480.66,194.82);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (480.66,189.66) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (507.93,220.22) --
	(508.43,220.22);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (508.18,220.02) --
	(508.18,220.41);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (508.18,216.18) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (491.57,205.97) --
	(495.12,205.97);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (493.43,203.33) --
	(493.43,208.29);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.43,201.94) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (503.22,214.26) --
	(504.66,214.26);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (503.95,213.02) --
	(503.95,215.41);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (503.95,211.35) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (515.97,224.96) --
	(517.47,224.96);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (516.73,224.23) --
	(516.73,225.66);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (516.73,222.06) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.25,189.86) --
	(484.33,189.86);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (482.85,187.61) --
	(482.85,191.86);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (482.85,186.63) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (485.11,203.25) --
	(489.02,203.25);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (487.16,201.42) --
	(487.16,204.91);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (487.16,199.21) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (511.92,224.51) --
	(513.09,224.51);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (512.51,223.97) --
	(512.51,225.03);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (512.51,220.48) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (456.53,174.40) --
	(465.54,174.40);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (461.56,168.49) --
	(461.56,178.92);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (461.56,170.37) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (471.24,182.70) --
	(477.85,182.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (474.83,179.96) --
	(474.83,185.10);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.83,179.80) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (519.16,227.69) --
	(519.69,227.69);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (519.43,227.51) --
	(519.43,227.87);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (519.43,224.79) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (465.86,186.69) --
	(466.93,186.69);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (466.40,185.13) --
	(466.40,188.14);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (466.40,183.47) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (476.81,194.33) --
	(481.05,194.33);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (479.05,192.40) --
	(479.05,196.08);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (479.05,190.29) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (503.73,219.06) --
	(506.09,219.06);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (504.95,218.53) --
	(504.95,219.58);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (504.95,215.03) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (457.61,179.76) --
	(463.97,179.76);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (461.06,175.67) --
	(461.06,183.13);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (461.06,175.73) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (461.22,186.48) --
	(469.53,186.48);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (465.82,164.01) --
	(465.82,196.53);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (465.82,183.58) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (518.70,228.09) --
	(519.30,228.09);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (519.00,227.61) --
	(519.00,228.56);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (519.00,225.19) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (440.05,149.07) --
	(446.60,149.07);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (443.61,140.42) --
	(443.61,155.00);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (443.61,145.84) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (482.66,197.96) --
	(497.10,197.96);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (491.21,193.52) --
	(491.21,201.57);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (491.21,193.93) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (482.31,198.48) --
	(483.02,198.48);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (482.67,197.65) --
	(482.67,199.28);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (482.67,194.45) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (435.88,141.67) --
	(444.13,141.67);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (440.45,137.70) --
	(440.45,144.96);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (440.45,137.63) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (430.12,152.12) --
	(446.95,152.12);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (440.33,138.07) --
	(440.33,160.13);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (440.33,149.21) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (515.31,226.85) --
	(516.96,226.85);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (516.15,226.06) --
	(516.15,227.60);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (516.15,223.94) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (511.01,226.91) --
	(512.84,226.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (511.95,226.11) --
	(511.95,227.68);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (511.95,223.69) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (491.35,207.58) --
	(497.18,207.58);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (494.48,206.11) --
	(494.48,208.95);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (494.48,203.55) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (523.54,236.94) --
	(523.81,236.94);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (523.67,236.77) --
	(523.67,237.11);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (523.67,232.91) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (482.77,199.05) --
	(485.89,199.05);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (484.39,196.91) --
	(484.39,200.98);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (484.39,195.02) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (497.32,213.65) --
	(500.98,213.65);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (499.24,211.50) --
	(499.24,215.58);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (499.24,210.75) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (517.89,227.34) --
	(518.87,227.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (518.38,227.05) --
	(518.38,227.63);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (518.38,224.44) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (455.83,171.67) --
	(462.39,171.67);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (459.39,168.44) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (511.24,226.95) --
	(515.97,226.95);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (513.75,222.92) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (493.31,212.27) --
	(493.57,212.27);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.44,208.24) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (461.13,178.96) --
	(463.91,178.96);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (462.57,174.93) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (468.52,174.11) --
	(474.54,174.11);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (471.77,171.20) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (516.74,230.25) --
	(519.35,230.25);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (518.09,227.35) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (487.89,210.95) --
	(497.51,210.95);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (493.30,206.38) --
	(493.30,214.63);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.30,207.72) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (464.41,188.45) --
	(478.88,188.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (472.98,185.46) --
	(472.98,191.04);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (472.98,184.42) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (510.18,227.14) --
	(512.82,227.14);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (511.55,226.45) --
	(511.55,227.79);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (511.55,223.10) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (472.41,191.50) --
	(476.82,191.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (474.74,184.11) --
	(474.74,196.81);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.74,187.46) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (474.64,199.45) --
	(482.47,199.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.95,190.32) --
	(478.95,205.61);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (478.95,196.55) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (528.55,239.61) --
	(529.80,239.61);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (529.19,236.91) --
	(529.19,241.97);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (529.19,236.70) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (507.92,228.14) --
	(510.30,228.14);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (509.15,227.35) --
	(509.15,228.89);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (509.15,224.91) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (498.70,215.49) --
	(502.10,215.49);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (500.48,214.12) --
	(500.48,216.77);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.48,211.46) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.09,231.37) --
	(515.01,231.37);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.56,231.18) --
	(514.56,231.57);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (514.56,227.34) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (495.91,214.59) --
	(499.36,214.59);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (497.71,213.26) --
	(497.71,215.83);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (497.71,210.56) {$h$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (500.08,220.23) --
	(503.50,220.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (501.87,219.79) --
	(501.87,220.66);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (501.87,217.32) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (530.30,241.17) --
	(532.04,241.17);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (531.19,240.33) --
	(531.19,241.98);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (531.19,238.27) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (515.10, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.93, 49.80) -- (382.93, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.99, 49.80) -- (426.99, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.04, 49.80) -- (471.04, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.10, 49.80) -- (515.10, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (382.93, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (426.99, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (471.04, 31.87) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (515.10, 31.87) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88,226.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (333.90, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 93.85) -- (333.90, 93.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,137.91) -- (333.90,137.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,181.96) -- (333.90,181.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,226.02) -- (333.90,226.02);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 91.00) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,135.05) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,179.11) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,223.16) {10000};
\end{scope}
\end{tikzpicture}

\end{document}