### 2020-08-24 benchmarking/benchmarkingVM/ESS_round1_real.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-24 19:04:05
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,289.08) rectangle (289.08,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,548.96) {\bfseries cons (\textit{real}) vs. nocons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,301.03) {ESS/chain [nocons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,433.62) {ESS/chain [cons (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,368.65) --
	(209.51,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,492.84) --
	(  0.00,209.51);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (  0.00,342.13) --
	(236.03,578.16);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,519.37) --
	(  0.00,236.03);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (  0.00,315.60) --
	(262.56,578.16);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,545.89) --
	(  0.00,262.56);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92,376.84) {$4\times$};

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (107.51,439.94) --
	(127.68,439.94);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (120.14,425.21) --
	(120.14,448.17);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (120.14,436.72) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (196.31,513.97) --
	(196.31,515.47);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (196.31,510.70) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (130.03,502.83) --
	(138.08,502.83);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (134.47,497.15) --
	(134.47,507.20);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (134.47,498.80) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (120.13,454.10) --
	(131.49,454.10);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (126.64,438.20) --
	(126.64,462.66);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (126.64,450.07) {$h$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (107.78,424.74) --
	(134.23,424.74);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (125.25,412.85) --
	(125.25,432.02);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (125.25,421.84) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.32,514.62) --
	(225.43,514.62);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.88,514.04) --
	(224.88,515.18);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (224.88,511.71) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (124.52,503.87) --
	(156.63,503.87);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.64,501.94) --
	(146.64,505.62);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (146.64,500.64) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (153.28,483.23) --
	(178.74,483.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (169.96,480.38) --
	(169.96,485.72);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.96,479.20) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (209.53,512.98) --
	(219.08,512.98);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (214.89,512.67) --
	(214.89,513.28);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (214.89,508.94) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (178.79,492.00) --
	(201.34,492.00);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.21,490.01) --
	(193.21,493.80);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.21,487.96) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (138.03,494.31) --
	(180.52,494.31);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (169.23,492.07) --
	(169.23,496.31);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.23,491.41) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.55,516.60) --
	(227.26,516.60);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.43,516.10) --
	(226.43,517.09);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (226.43,513.70) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (198.40,474.45) --
	(198.40,481.06);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (198.40,474.81) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (145.24,495.45) --
	(203.65,495.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.27,494.96) --
	(191.27,495.92);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.27,491.42) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (209.83,514.09) --
	(229.21,514.09);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (221.87,513.75) --
	(221.87,514.43);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (221.87,510.05) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.36,451.14) --
	(176.36,460.33);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (176.36,452.25) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (198.91,466.15) --
	(198.91,474.11);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (198.91,467.64) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.60,518.50) --
	(229.08,518.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.84,518.10) --
	(228.84,518.89);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (228.84,515.59) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (129.72,483.83) --
	(169.90,483.83);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (158.85,482.60) --
	(158.85,484.98);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (158.85,480.60) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (153.47,490.36) --
	(195.70,490.36);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (184.44,489.61) --
	(184.44,491.09);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.44,486.33) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (202.50,511.41) --
	(216.09,511.41);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (210.48,511.05) --
	(210.48,511.76);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.48,507.38) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (149.11,476.52) --
	(176.28,476.52);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (167.16,475.30) --
	(167.16,477.68);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (167.16,472.49) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (136.04,473.71) --
	(175.13,473.71);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (164.20,470.57) --
	(164.20,476.41);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (164.20,470.81) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.57,519.04) --
	(229.27,519.04);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.92,518.76) --
	(228.92,519.31);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (228.92,516.13) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (157.99,465.14) --
	(165.18,465.14);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (161.92,462.06) --
	(161.92,467.80);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (161.92,461.92) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (180.71,508.79) --
	(203.68,508.79);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.45,506.03) --
	(195.45,511.19);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.45,504.75) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.01,507.17) --
	(187.33,507.17);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.68,506.03) --
	(186.68,508.25);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.68,503.14) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (143.93,460.26) --
	(151.84,460.26);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (148.29,455.77) --
	(148.29,463.90);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (148.29,456.23) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (158.24,444.03) --
	(170.29,444.03);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.20,442.71) --
	(165.20,445.27);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.20,441.13) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.21,516.34) --
	(226.84,516.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.53,515.49) --
	(226.53,517.16);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (226.53,513.44) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.07,486.04) --
	(195.07,492.68);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.07,486.42) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.22,495.86) --
	(188.22,500.05);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.22,494.04) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.90,512.15) --
	(238.51,512.15);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (232.89,511.77) --
	(232.89,512.53);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (232.89,508.12) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.48,482.21) --
	(219.73,482.21);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.90,479.90) --
	(208.90,484.27);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.90,478.18) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (212.90,487.36) --
	(212.90,491.02);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.90,486.37) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.61,517.72) --
	(228.72,517.72);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.17,517.16) --
	(228.17,518.25);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (228.17,514.81) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (134.31,501.18) --
	(134.31,502.87);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (134.31,498.82) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (140.07,506.91) --
	(140.07,511.00);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (140.07,505.03) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (163.84,511.28) --
	(197.54,511.28);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.31,511.18) --
	(187.31,511.39);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.31,507.25) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (173.23,479.47) --
	(178.50,479.47);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.04,477.70) --
	(176.04,481.09);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (176.04,475.44) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (145.91,475.98) --
	(145.91,477.70);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (145.91,473.95) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.98,518.28) --
	(229.70,518.28);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.93,517.72) --
	(227.93,518.83);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.93,515.38) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (106.47,493.53) --
	(119.17,493.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (113.85,490.64) --
	(113.85,496.03);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (113.85,490.30) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (121.32,481.57) --
	(142.29,481.57);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (134.54,464.93) --
	(134.54,490.33);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (134.54,477.53) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (197.73,508.06) --
	(213.81,508.06);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (207.41,507.86) --
	(207.41,508.26);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.41,504.03) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (147.10,471.65) --
	(168.39,471.65);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.56,469.20) --
	(160.56,473.83);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.56,467.62) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (109.28,482.67) --
	(125.12,482.67);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (118.79,477.71) --
	(118.79,486.60);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (118.79,479.76) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (236.93,528.42) --
	(238.96,528.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (237.97,526.53) --
	(237.97,530.13);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (237.97,525.51) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.40,493.98) --
	(227.40,500.05);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.40,494.03) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (175.33,489.37) --
	(175.33,499.62);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (175.33,491.14) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.06,509.80) --
	(237.03,509.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.26,509.43) --
	(225.26,510.17);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.26,505.77) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (174.04,492.43) --
	(200.67,492.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.66,487.19) --
	(191.66,496.54);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.66,488.40) {$h$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (207.78,491.01) --
	(207.78,496.89);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.78,491.27) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (241.40,531.56) --
	(243.18,531.56);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (242.31,530.87) --
	(242.31,532.22);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (242.31,528.65) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- (226.02,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85,338.88) -- ( 93.85,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91,338.88) -- (137.91,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96,338.88) -- (181.96,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02,338.88) -- (226.02,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 44.82,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,382.93) -- ( 44.82,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,426.99) -- ( 44.82,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,471.04) -- ( 44.82,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,515.10) -- ( 44.82,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,431.01) rectangle (189.67,426.03);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,421.05) rectangle (189.67,416.07);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,411.09) rectangle (189.67,406.11);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,401.13) rectangle (189.67,396.15);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,391.17) rectangle (189.67,386.19);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,381.21) rectangle (189.67,376.23);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,371.25) rectangle (189.67,366.27);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,361.29) rectangle (189.67,356.31);
\definecolor{fillColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,351.33) rectangle (189.67,346.35);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,425.66) {Ran};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,415.70) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,405.74) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,395.78) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,385.82) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,375.86) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,365.90) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,355.94) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,345.98) {Moyle};
\end{scope}
\begin{scope}
\path[clip] (289.08,289.08) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,548.96) {\bfseries adapt (\textit{real}) vs. nocons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,301.03) {ESS/chain [nocons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,433.62) {ESS/chain [adapt (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] (338.88,338.88) rectangle (528.36,528.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.66,242.23) --
	(498.59,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,492.84) --
	(162.66, 83.09) --
	(162.66,242.23);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (162.66,215.71) --
	(525.11,578.16);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,519.37) --
	(162.66,109.61) --
	(162.66,215.71);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (162.66,189.19) --
	(551.64,578.16);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,545.89) --
	(162.66,136.14) --
	(162.66,189.19);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (425.98,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (399.46,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (371.28,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00,376.84) {$4\times$};

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (396.59,491.83) --
	(416.76,491.83);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (409.22,484.91) --
	(409.22,496.90);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (409.22,488.60) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (485.39,513.46) --
	(485.39,517.31);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (485.39,511.45) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (419.11,508.15) --
	(427.16,508.15);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (423.55,507.02) --
	(423.55,509.22);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (423.55,504.12) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (409.21,469.50) --
	(420.57,469.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (415.72,459.30) --
	(415.72,476.12);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (415.72,465.47) {$h$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (396.86,494.88) --
	(423.31,494.88);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (414.33,490.31) --
	(414.33,498.58);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (414.33,491.98) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (513.40,514.73) --
	(514.51,514.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (513.96,514.21) --
	(513.96,515.24);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (513.96,511.83) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (413.60,504.78) --
	(445.71,504.78);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (435.72,504.07) --
	(435.72,505.47);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (435.72,501.56) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (442.36,483.32) --
	(467.82,483.32);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (459.04,482.10) --
	(459.04,484.46);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (459.04,479.28) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (498.61,513.50) --
	(508.16,513.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (503.97,513.33) --
	(503.97,513.67);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (503.97,509.47) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (467.87,494.70) --
	(490.42,494.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (482.29,493.28) --
	(482.29,496.02);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (482.29,490.67) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (427.11,506.61) --
	(469.60,506.61);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (458.31,505.81) --
	(458.31,507.37);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (458.31,503.70) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.63,514.56) --
	(516.34,514.56);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (515.51,514.31) --
	(515.51,514.81);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (515.51,511.66) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (487.48,477.75) --
	(487.48,481.67);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (487.48,476.59) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (434.32,492.07) --
	(492.73,492.07);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (480.35,485.78) --
	(480.35,496.80);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (480.35,488.04) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (498.91,518.14) --
	(518.29,518.14);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (510.95,517.82) --
	(510.95,518.45);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (510.95,514.11) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (465.44,449.07) --
	(465.44,453.85);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (465.44,447.57) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (487.99,472.24) --
	(487.99,478.57);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (487.99,472.76) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (517.68,518.59) --
	(518.16,518.59);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (517.92,518.46) --
	(517.92,518.71);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (517.92,515.68) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (418.80,491.01) --
	(458.98,491.01);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (447.93,487.64) --
	(447.93,493.87);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (447.93,487.78) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (442.55,488.00) --
	(484.78,488.00);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (473.52,486.78) --
	(473.52,489.16);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (473.52,483.97) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (491.58,512.98) --
	(505.17,512.98);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (499.56,512.68) --
	(499.56,513.27);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (499.56,508.95) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (438.19,480.43) --
	(465.36,480.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (456.24,476.92) --
	(456.24,483.40);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (456.24,476.40) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (425.12,485.33) --
	(464.21,485.33);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (453.28,483.22) --
	(453.28,487.22);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (453.28,482.42) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (517.65,518.72) --
	(518.35,518.72);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (518.00,518.22) --
	(518.00,519.21);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (518.00,515.82) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (447.07,496.45) --
	(454.26,496.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (451.00,492.63) --
	(451.00,499.63);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (451.00,493.22) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (469.79,506.28) --
	(492.76,506.28);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (484.53,502.48) --
	(484.53,509.46);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (484.53,502.25) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (475.09,508.23) --
	(476.41,508.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (475.76,508.03) --
	(475.76,508.43);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (475.76,504.20) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (433.01,463.38) --
	(440.92,463.38);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (437.37,458.93) --
	(437.37,466.99);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (437.37,459.35) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (447.32,485.70) --
	(459.37,485.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (454.28,478.90) --
	(454.28,490.71);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (454.28,482.80) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (515.29,516.73) --
	(515.92,516.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (515.61,516.31) --
	(515.61,517.14);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (515.61,513.83) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (484.15,483.94) --
	(484.15,500.04);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (484.15,490.41) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (477.30,495.56) --
	(477.30,497.59);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (477.30,492.57) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (513.98,524.67) --
	(527.59,524.67);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (521.97,524.50) --
	(521.97,524.84);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (521.97,520.63) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (470.56,488.81) --
	(508.81,488.81);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (497.98,487.21) --
	(497.98,490.29);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (497.98,484.78) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (501.98,495.18) --
	(501.98,499.74);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (501.98,494.69) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (516.69,516.36) --
	(517.80,516.36);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (517.25,515.62) --
	(517.25,517.07);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (517.25,513.45) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (423.39,502.71) --
	(423.39,504.41);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (423.39,500.35) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (429.15,508.11) --
	(429.15,511.22);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (429.15,505.70) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (452.92,509.24) --
	(486.62,509.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (476.39,508.98) --
	(476.39,509.50);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.39,505.21) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (462.31,479.99) --
	(467.58,479.99);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (465.12,478.93) --
	(465.12,481.00);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (465.12,475.96) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (434.99,478.69) --
	(434.99,483.82);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (434.99,478.52) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (515.06,517.17) --
	(518.78,517.17);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (517.01,516.17) --
	(517.01,518.13);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (517.01,514.27) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (395.55,504.25) --
	(408.25,504.25);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (402.93,499.69) --
	(402.93,507.93);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (402.93,501.02) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (410.40,482.55) --
	(431.37,482.55);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (423.62,478.83) --
	(423.62,485.66);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (423.62,478.52) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (486.81,515.35) --
	(502.89,515.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (496.49,514.89) --
	(496.49,515.80);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (496.49,511.32) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (436.18,476.20) --
	(457.47,476.20);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (449.64,470.52) --
	(449.64,480.58);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.64,472.17) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (398.36,502.37) --
	(414.20,502.37);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (407.87,497.47) --
	(407.87,506.27);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (407.87,499.46) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (526.01,528.12) --
	(528.04,528.12);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (527.05,526.74) --
	(527.05,529.42);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (527.05,525.22) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (516.48,497.06) --
	(516.48,503.10);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (516.48,497.09) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (464.41,493.38) --
	(464.41,499.78);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (464.41,492.81) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.14,518.98) --
	(526.11,518.98);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.34,518.81) --
	(514.34,519.15);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (514.34,514.95) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (463.12,499.46) --
	(489.75,499.46);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (480.74,497.44) --
	(480.74,501.29);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (480.74,495.42) {$h$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (496.86,502.88) --
	(496.86,505.02);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (496.86,501.07) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (530.48,531.24) --
	(532.26,531.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (531.39,530.62) --
	(531.39,531.84);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (531.39,528.33) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (515.10,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.93,338.88) -- (382.93,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.99,338.88) -- (426.99,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.04,338.88) -- (471.04,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.10,338.88) -- (515.10,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (382.93,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (426.99,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (471.04,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (515.10,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (333.90,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,382.93) -- (333.90,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,426.99) -- (333.90,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,471.04) -- (333.90,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,515.10) -- (333.90,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,259.88) {\bfseries adapt (\textit{real}) vs. cons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54, 11.95) {ESS/chain [cons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,144.54) {ESS/chain [adapt (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80, 49.80) rectangle (239.28,239.28);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 79.57) --
	(283.33,362.91) --
	(283.33,203.76) --
	( 79.57,  0.00);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (  0.00, 53.05) --
	(283.33,336.38) --
	(283.33,230.29) --
	( 53.05,  0.00);

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] (  0.00, 26.52) --
	(283.33,309.86) --
	(283.33,256.81) --
	( 26.52,  0.00);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,140.80) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,114.28) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92, 87.76) {$4\times$};

\path[draw=drawColor,line width= 1.0pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28,239.28);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (136.13,202.75) --
	(159.09,202.75);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (150.86,195.83) --
	(150.86,207.82);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (150.86,199.52) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.89,226.40) --
	(226.39,226.40);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.65,224.38) --
	(225.65,228.23);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.65,222.37) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.07,219.07) --
	(218.12,219.07);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (213.75,217.94) --
	(213.75,220.14);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (213.75,215.04) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (149.12,180.42) --
	(173.58,180.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.02,170.22) --
	(165.02,187.04);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.02,176.39) {$h$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (123.77,205.80) --
	(142.94,205.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (135.66,201.23) --
	(135.66,209.50);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (135.66,202.90) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.96,225.65) --
	(226.10,225.65);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.54,225.13) --
	(225.54,226.16);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.54,222.75) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (212.86,215.70) --
	(216.54,215.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (214.79,214.99) --
	(214.79,216.39);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (214.79,212.48) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.30,194.24) --
	(196.64,194.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.15,193.02) --
	(194.15,195.38);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.15,190.20) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (223.59,224.42) --
	(224.20,224.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (223.90,224.25) --
	(223.90,224.59);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (223.90,220.39) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.93,205.62) --
	(204.72,205.62);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (202.92,204.20) --
	(202.92,206.94);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.92,201.59) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (202.99,217.53) --
	(207.23,217.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (205.23,216.73) --
	(205.23,218.29);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.23,214.62) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.02,225.48) --
	(228.01,225.48);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.52,225.23) --
	(227.52,225.73);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.52,222.58) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.37,190.73) --
	(191.98,190.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.96,188.67) --
	(188.96,192.59);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.96,187.51) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (205.88,202.99) --
	(206.84,202.99);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (206.37,196.70) --
	(206.37,207.72);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (206.37,198.96) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.67,229.06) --
	(225.35,229.06);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.01,228.74) --
	(225.01,229.37);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.01,225.03) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (162.06,162.53) --
	(171.25,162.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (167.20,159.99) --
	(167.20,164.77);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (167.20,158.49) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (177.07,186.59) --
	(185.03,186.59);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.46,183.16) --
	(181.46,189.49);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.46,183.68) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (229.02,229.51) --
	(229.81,229.51);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (229.42,229.38) --
	(229.42,229.63);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (229.42,226.60) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.52,201.93) --
	(195.90,201.93);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.75,198.56) --
	(194.75,204.79);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.75,198.70) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.53,198.92) --
	(202.01,198.92);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (201.28,197.70) --
	(201.28,200.08);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (201.28,194.89) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (221.97,223.90) --
	(222.68,223.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (222.33,223.60) --
	(222.33,224.19);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (222.33,219.87) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.22,191.35) --
	(188.60,191.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.44,187.84) --
	(187.44,194.32);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.44,187.32) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.49,196.25) --
	(187.33,196.25);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (184.63,194.14) --
	(184.63,198.14);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.63,193.34) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (229.68,229.64) --
	(230.23,229.64);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (229.96,229.14) --
	(229.96,230.13);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (229.96,226.74) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.98,207.37) --
	(178.72,207.37);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.06,203.55) --
	(176.06,210.55);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (176.06,204.14) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (216.95,217.20) --
	(222.11,217.20);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (219.71,213.40) --
	(219.71,220.38);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (219.71,213.17) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (216.95,219.15) --
	(219.17,219.15);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (218.09,218.95) --
	(218.09,219.35);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.09,215.12) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (166.69,174.30) --
	(174.82,174.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (171.18,169.85) --
	(171.18,177.91);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (171.18,170.27) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (153.63,196.62) --
	(156.19,196.62);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (154.95,189.82) --
	(154.95,201.63);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (154.95,193.72) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (226.41,227.65) --
	(228.08,227.65);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (227.26,227.23) --
	(227.26,228.06);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.26,224.75) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (196.96,204.56) --
	(203.60,204.56);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.57,194.86) --
	(200.57,210.96);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.57,201.33) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (206.78,207.52) --
	(210.97,207.52);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.99,206.48) --
	(208.99,208.51);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.99,203.49) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (222.69,235.59) --
	(223.45,235.59);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (223.07,235.42) --
	(223.07,235.76);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (223.07,231.55) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.82,199.73) --
	(195.19,199.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.13,198.13) --
	(193.13,201.21);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.13,195.70) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (198.28,208.52) --
	(201.94,208.52);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.19,206.10) --
	(200.19,210.66);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.19,205.61) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.08,227.28) --
	(229.17,227.28);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.64,226.54) --
	(228.64,227.99);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (228.64,224.37) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (212.10,214.50) --
	(213.79,214.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (212.97,213.63) --
	(212.97,215.33);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.97,211.27) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (217.83,220.65) --
	(221.92,220.65);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (219.98,219.03) --
	(219.98,222.14);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (219.98,216.62) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (222.10,220.16) --
	(222.31,220.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (222.20,219.90) --
	(222.20,220.42);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (222.20,216.13) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.62,190.91) --
	(192.01,190.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.39,189.85) --
	(190.39,191.92);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (190.39,186.88) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.90,192.35) --
	(188.62,192.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.78,189.61) --
	(187.78,194.74);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.78,189.44) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (228.64,228.09) --
	(229.75,228.09);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (229.20,227.09) --
	(229.20,229.05);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (229.20,225.19) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (201.56,215.17) --
	(206.95,215.17);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (204.45,210.61) --
	(204.45,218.85);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.45,211.94) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (175.85,193.47) --
	(201.25,193.47);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.49,189.75) --
	(192.49,196.58);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.49,189.44) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (218.78,226.27) --
	(219.18,226.27);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (218.98,225.81) --
	(218.98,226.72);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.98,222.24) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (180.12,187.12) --
	(184.75,187.12);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.57,181.44) --
	(182.57,191.50);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.57,183.09) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.63,213.29) --
	(197.52,213.29);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.59,208.39) --
	(193.59,217.19);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.59,210.38) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (237.45,239.04) --
	(241.05,239.04);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (239.34,237.66) --
	(239.34,240.34);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (239.34,236.14) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (204.90,211.23) --
	(210.97,211.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.18,207.98) --
	(208.18,214.02);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.18,208.01) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.29,207.77) --
	(210.54,207.77);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (206.09,204.30) --
	(206.09,210.70);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (206.09,203.73) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (220.35,229.90) --
	(221.09,229.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (220.72,229.73) --
	(220.72,230.07);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (220.72,225.87) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (198.11,210.38) --
	(207.46,210.38);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (203.35,208.36) --
	(203.35,212.21);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.35,206.34) {$h$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (201.93,214.90) --
	(207.81,214.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (205.10,213.80) --
	(205.10,215.94);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.10,211.99) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (241.79,242.16) --
	(243.14,242.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (242.48,241.54) --
	(242.48,242.76);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (242.48,239.25) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- (226.02, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85, 49.80) -- ( 93.85, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91, 49.80) -- (137.91, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96, 49.80) -- (181.96, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02, 49.80) -- (226.02, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96, 31.87) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02, 31.87) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80,226.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 44.82, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 93.85) -- ( 44.82, 93.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,137.91) -- ( 44.82,137.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,181.96) -- ( 44.82,181.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,226.02) -- ( 44.82,226.02);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 91.00) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,135.05) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,179.11) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,223.16) {10000};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,259.88) {\bfseries \texttt{AdaptiveOperatorSampler} weights};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62, 11.95) {Sequence length $L$ (kb)};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,144.54) {Learned operator weight};
\end{scope}
\begin{scope}
\path[clip] (338.88, 49.80) rectangle (528.36,239.28);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (485.73,232.77) circle (  1.87);

\path[fill=fillColor] (485.73, 55.06) --
	(488.24, 50.70) --
	(483.21, 50.70) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (394.78,149.36) circle (  1.87);

\path[fill=fillColor] (394.78, 80.07) --
	(397.29, 75.72) --
	(392.26, 75.72) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (390.04,184.14) circle (  1.87);

\path[fill=fillColor] (390.04, 69.81) --
	(392.55, 65.46) --
	(387.52, 65.46) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (404.25,211.12) circle (  1.87);

\path[fill=fillColor] (404.25, 60.31) --
	(406.77, 55.95) --
	(401.74, 55.95) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (433.62,198.59) circle (  1.87);

\path[fill=fillColor] (433.62, 71.59) --
	(436.14, 67.23) --
	(431.10, 67.23) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (359.72,131.32) circle (  1.87);

\path[fill=fillColor] (359.72, 82.69) --
	(362.24, 78.34) --
	(357.21, 78.34) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (346.46,125.73) circle (  1.87);

\path[fill=fillColor] (346.46, 76.98) --
	(348.97, 72.63) --
	(343.94, 72.63) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,113.48) --
	(339.26,113.88) --
	(339.64,114.28) --
	(340.02,114.68) --
	(340.40,115.09) --
	(340.78,115.49) --
	(341.16,115.90) --
	(341.54,116.31) --
	(341.92,116.71) --
	(342.30,117.12) --
	(342.68,117.53) --
	(343.06,117.95) --
	(343.44,118.36) --
	(343.82,118.77) --
	(344.20,119.19) --
	(344.58,119.60) --
	(344.96,120.02) --
	(345.34,120.44) --
	(345.71,120.86) --
	(346.09,121.28) --
	(346.47,121.70) --
	(346.85,122.12) --
	(347.23,122.55) --
	(347.61,122.97) --
	(347.99,123.39) --
	(348.37,123.82) --
	(348.75,124.25) --
	(349.13,124.67) --
	(349.51,125.10) --
	(349.89,125.53) --
	(350.27,125.96) --
	(350.65,126.39) --
	(351.03,126.83) --
	(351.41,127.26) --
	(351.79,127.69) --
	(352.17,128.12) --
	(352.55,128.56) --
	(352.93,128.99) --
	(353.31,129.43) --
	(353.69,129.87) --
	(354.07,130.30) --
	(354.45,130.74) --
	(354.83,131.18) --
	(355.21,131.62) --
	(355.59,132.06) --
	(355.97,132.50) --
	(356.35,132.94) --
	(356.73,133.38) --
	(357.11,133.82) --
	(357.49,134.27) --
	(357.87,134.71) --
	(358.25,135.15) --
	(358.63,135.60) --
	(359.01,136.04) --
	(359.38,136.48) --
	(359.76,136.93) --
	(360.14,137.37) --
	(360.52,137.82) --
	(360.90,138.26) --
	(361.28,138.71) --
	(361.66,139.16) --
	(362.04,139.60) --
	(362.42,140.05) --
	(362.80,140.50) --
	(363.18,140.94) --
	(363.56,141.39) --
	(363.94,141.84) --
	(364.32,142.28) --
	(364.70,142.73) --
	(365.08,143.18) --
	(365.46,143.63) --
	(365.84,144.08) --
	(366.22,144.52) --
	(366.60,144.97) --
	(366.98,145.42) --
	(367.36,145.87) --
	(367.74,146.31) --
	(368.12,146.76) --
	(368.50,147.21) --
	(368.88,147.66) --
	(369.26,148.10) --
	(369.64,148.55) --
	(370.02,149.00) --
	(370.40,149.44) --
	(370.78,149.89) --
	(371.16,150.34) --
	(371.54,150.78) --
	(371.92,151.23) --
	(372.30,151.67) --
	(372.68,152.12) --
	(373.05,152.56) --
	(373.43,153.01) --
	(373.81,153.45) --
	(374.19,153.89) --
	(374.57,154.34) --
	(374.95,154.78) --
	(375.33,155.22) --
	(375.71,155.66) --
	(376.09,156.11) --
	(376.47,156.55) --
	(376.85,156.99) --
	(377.23,157.43) --
	(377.61,157.87) --
	(377.99,158.30) --
	(378.37,158.74) --
	(378.75,159.18) --
	(379.13,159.62) --
	(379.51,160.05) --
	(379.89,160.49) --
	(380.27,160.92) --
	(380.65,161.36) --
	(381.03,161.79) --
	(381.41,162.22) --
	(381.79,162.65) --
	(382.17,163.08) --
	(382.55,163.51) --
	(382.93,163.94) --
	(383.31,164.37) --
	(383.69,164.80) --
	(384.07,165.23) --
	(384.45,165.65) --
	(384.83,166.08) --
	(385.21,166.50) --
	(385.59,166.93) --
	(385.97,167.35) --
	(386.34,167.77) --
	(386.72,168.19) --
	(387.10,168.61) --
	(387.48,169.03) --
	(387.86,169.44) --
	(388.24,169.86) --
	(388.62,170.28) --
	(389.00,170.69) --
	(389.38,171.10) --
	(389.76,171.51) --
	(390.14,171.93) --
	(390.52,172.34) --
	(390.90,172.74) --
	(391.28,173.15) --
	(391.66,173.56) --
	(392.04,173.96) --
	(392.42,174.37) --
	(392.80,174.77) --
	(393.18,175.17) --
	(393.56,175.57) --
	(393.94,175.97) --
	(394.32,176.37) --
	(394.70,176.76) --
	(395.08,177.16) --
	(395.46,177.55) --
	(395.84,177.95) --
	(396.22,178.34) --
	(396.60,178.73) --
	(396.98,179.12) --
	(397.36,179.50) --
	(397.74,179.89) --
	(398.12,180.27) --
	(398.50,180.66) --
	(398.88,181.04) --
	(399.26,181.42) --
	(399.64,181.80) --
	(400.01,182.18) --
	(400.39,182.55) --
	(400.77,182.93) --
	(401.15,183.30) --
	(401.53,183.67) --
	(401.91,184.05) --
	(402.29,184.41) --
	(402.67,184.78) --
	(403.05,185.15) --
	(403.43,185.51) --
	(403.81,185.88) --
	(404.19,186.24) --
	(404.57,186.60) --
	(404.95,186.96) --
	(405.33,187.31) --
	(405.71,187.67) --
	(406.09,188.02) --
	(406.47,188.38) --
	(406.85,188.73) --
	(407.23,189.08) --
	(407.61,189.43) --
	(407.99,189.77) --
	(408.37,190.12) --
	(408.75,190.46) --
	(409.13,190.80) --
	(409.51,191.14) --
	(409.89,191.48) --
	(410.27,191.82) --
	(410.65,192.15) --
	(411.03,192.49) --
	(411.41,192.82) --
	(411.79,193.15) --
	(412.17,193.48) --
	(412.55,193.81) --
	(412.93,194.13) --
	(413.31,194.46) --
	(413.68,194.78) --
	(414.06,195.10) --
	(414.44,195.42) --
	(414.82,195.74) --
	(415.20,196.05) --
	(415.58,196.37) --
	(415.96,196.68) --
	(416.34,196.99) --
	(416.72,197.30) --
	(417.10,197.61) --
	(417.48,197.92) --
	(417.86,198.22) --
	(418.24,198.52) --
	(418.62,198.83) --
	(419.00,199.13) --
	(419.38,199.42) --
	(419.76,199.72) --
	(420.14,200.02) --
	(420.52,200.31) --
	(420.90,200.60) --
	(421.28,200.89) --
	(421.66,201.18) --
	(422.04,201.47) --
	(422.42,201.75) --
	(422.80,202.03) --
	(423.18,202.32) --
	(423.56,202.60) --
	(423.94,202.88) --
	(424.32,203.15) --
	(424.70,203.43) --
	(425.08,203.70) --
	(425.46,203.97) --
	(425.84,204.25) --
	(426.22,204.51) --
	(426.60,204.78) --
	(426.97,205.05) --
	(427.35,205.31) --
	(427.73,205.58) --
	(428.11,205.84) --
	(428.49,206.10) --
	(428.87,206.35) --
	(429.25,206.61) --
	(429.63,206.86) --
	(430.01,207.12) --
	(430.39,207.37) --
	(430.77,207.62) --
	(431.15,207.87) --
	(431.53,208.11) --
	(431.91,208.36) --
	(432.29,208.60) --
	(432.67,208.85) --
	(433.05,209.09) --
	(433.43,209.33) --
	(433.81,209.56) --
	(434.19,209.80) --
	(434.57,210.03) --
	(434.95,210.27) --
	(435.33,210.50) --
	(435.71,210.73) --
	(436.09,210.96) --
	(436.47,211.18) --
	(436.85,211.41) --
	(437.23,211.63) --
	(437.61,211.86) --
	(437.99,212.08) --
	(438.37,212.30) --
	(438.75,212.51) --
	(439.13,212.73) --
	(439.51,212.95) --
	(439.89,213.16) --
	(440.27,213.37) --
	(440.64,213.58) --
	(441.02,213.79) --
	(441.40,214.00) --
	(441.78,214.21) --
	(442.16,214.41) --
	(442.54,214.61) --
	(442.92,214.82) --
	(443.30,215.02) --
	(443.68,215.22) --
	(444.06,215.41) --
	(444.44,215.61) --
	(444.82,215.81) --
	(445.20,216.00) --
	(445.58,216.19) --
	(445.96,216.38) --
	(446.34,216.57) --
	(446.72,216.76) --
	(447.10,216.95) --
	(447.48,217.13) --
	(447.86,217.32) --
	(448.24,217.50) --
	(448.62,217.68) --
	(449.00,217.86) --
	(449.38,218.04) --
	(449.76,218.22) --
	(450.14,218.39) --
	(450.52,218.57) --
	(450.90,218.74) --
	(451.28,218.91) --
	(451.66,219.09) --
	(452.04,219.26) --
	(452.42,219.42) --
	(452.80,219.59) --
	(453.18,219.76) --
	(453.56,219.92) --
	(453.93,220.09) --
	(454.31,220.25) --
	(454.69,220.41) --
	(455.07,220.57) --
	(455.45,220.73) --
	(455.83,220.89) --
	(456.21,221.04) --
	(456.59,221.20) --
	(456.97,221.35) --
	(457.35,221.50) --
	(457.73,221.66) --
	(458.11,221.81) --
	(458.49,221.96) --
	(458.87,222.10) --
	(459.25,222.25) --
	(459.63,222.40) --
	(460.01,222.54) --
	(460.39,222.69) --
	(460.77,222.83) --
	(461.15,222.97) --
	(461.53,223.11) --
	(461.91,223.25) --
	(462.29,223.39) --
	(462.67,223.52) --
	(463.05,223.66) --
	(463.43,223.80) --
	(463.81,223.93) --
	(464.19,224.06) --
	(464.57,224.19) --
	(464.95,224.32) --
	(465.33,224.45) --
	(465.71,224.58) --
	(466.09,224.71) --
	(466.47,224.84) --
	(466.85,224.96) --
	(467.23,225.09) --
	(467.60,225.21) --
	(467.98,225.33) --
	(468.36,225.46) --
	(468.74,225.58) --
	(469.12,225.70) --
	(469.50,225.81) --
	(469.88,225.93) --
	(470.26,226.05) --
	(470.64,226.16) --
	(471.02,226.28) --
	(471.40,226.39) --
	(471.78,226.51) --
	(472.16,226.62) --
	(472.54,226.73) --
	(472.92,226.84) --
	(473.30,226.95) --
	(473.68,227.06) --
	(474.06,227.17) --
	(474.44,227.27) --
	(474.82,227.38) --
	(475.20,227.48) --
	(475.58,227.59) --
	(475.96,227.69) --
	(476.34,227.79) --
	(476.72,227.89) --
	(477.10,228.00) --
	(477.48,228.10) --
	(477.86,228.19) --
	(478.24,228.29) --
	(478.62,228.39) --
	(479.00,228.49) --
	(479.38,228.58) --
	(479.76,228.68) --
	(480.14,228.77) --
	(480.52,228.87) --
	(480.90,228.96) --
	(481.27,229.05) --
	(481.65,229.14) --
	(482.03,229.23) --
	(482.41,229.32) --
	(482.79,229.41) --
	(483.17,229.50) --
	(483.55,229.58) --
	(483.93,229.67) --
	(484.31,229.76) --
	(484.69,229.84) --
	(485.07,229.93) --
	(485.45,230.01) --
	(485.83,230.09) --
	(486.21,230.18) --
	(486.59,230.26) --
	(486.97,230.34) --
	(487.35,230.42) --
	(487.73,230.50) --
	(488.11,230.58) --
	(488.49,230.65) --
	(488.87,230.73) --
	(489.25,230.81) --
	(489.63,230.89) --
	(490.01,230.96) --
	(490.39,231.04) --
	(490.77,231.11) --
	(491.15,231.18) --
	(491.53,231.26) --
	(491.91,231.33) --
	(492.29,231.40) --
	(492.67,231.47) --
	(493.05,231.54) --
	(493.43,231.61) --
	(493.81,231.68) --
	(494.19,231.75) --
	(494.56,231.82) --
	(494.94,231.89) --
	(495.32,231.95) --
	(495.70,232.02) --
	(496.08,232.08) --
	(496.46,232.15) --
	(496.84,232.21) --
	(497.22,232.28) --
	(497.60,232.34) --
	(497.98,232.40) --
	(498.36,232.47) --
	(498.74,232.53) --
	(499.12,232.59) --
	(499.50,232.65) --
	(499.88,232.71) --
	(500.26,232.77) --
	(500.64,232.83) --
	(501.02,232.89) --
	(501.40,232.95) --
	(501.78,233.00) --
	(502.16,233.06) --
	(502.54,233.12) --
	(502.92,233.17) --
	(503.30,233.23) --
	(503.68,233.28) --
	(504.06,233.34) --
	(504.44,233.39) --
	(504.82,233.45) --
	(505.20,233.50) --
	(505.58,233.55) --
	(505.96,233.60) --
	(506.34,233.66) --
	(506.72,233.71) --
	(507.10,233.76) --
	(507.48,233.81) --
	(507.86,233.86) --
	(508.23,233.91) --
	(508.61,233.96) --
	(508.99,234.01) --
	(509.37,234.05) --
	(509.75,234.10) --
	(510.13,234.15) --
	(510.51,234.20) --
	(510.89,234.24) --
	(511.27,234.29) --
	(511.65,234.34) --
	(512.03,234.38) --
	(512.41,234.43) --
	(512.79,234.47) --
	(513.17,234.51) --
	(513.55,234.56) --
	(513.93,234.60) --
	(514.31,234.64) --
	(514.69,234.69) --
	(515.07,234.73) --
	(515.45,234.77) --
	(515.83,234.81) --
	(516.21,234.85) --
	(516.59,234.89) --
	(516.97,234.93) --
	(517.35,234.97) --
	(517.73,235.01) --
	(518.11,235.05) --
	(518.49,235.09) --
	(518.87,235.13) --
	(519.25,235.17) --
	(519.63,235.21) --
	(520.01,235.24) --
	(520.39,235.28) --
	(520.77,235.32) --
	(521.15,235.35) --
	(521.53,235.39) --
	(521.90,235.43) --
	(522.28,235.46) --
	(522.66,235.50) --
	(523.04,235.53) --
	(523.42,235.57) --
	(523.80,235.60) --
	(524.18,235.63) --
	(524.56,235.67) --
	(524.94,235.70) --
	(525.32,235.73) --
	(525.70,235.77) --
	(526.08,235.80) --
	(526.46,235.83) --
	(526.84,235.86) --
	(527.22,235.90) --
	(527.60,235.93) --
	(527.98,235.96) --
	(528.36,235.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 82.37) --
	(339.26, 82.25) --
	(339.64, 82.13) --
	(340.02, 82.00) --
	(340.40, 81.88) --
	(340.78, 81.75) --
	(341.16, 81.63) --
	(341.54, 81.51) --
	(341.92, 81.39) --
	(342.30, 81.27) --
	(342.68, 81.14) --
	(343.06, 81.02) --
	(343.44, 80.90) --
	(343.82, 80.78) --
	(344.20, 80.66) --
	(344.58, 80.54) --
	(344.96, 80.42) --
	(345.34, 80.31) --
	(345.71, 80.19) --
	(346.09, 80.07) --
	(346.47, 79.95) --
	(346.85, 79.83) --
	(347.23, 79.72) --
	(347.61, 79.60) --
	(347.99, 79.48) --
	(348.37, 79.37) --
	(348.75, 79.25) --
	(349.13, 79.14) --
	(349.51, 79.02) --
	(349.89, 78.91) --
	(350.27, 78.80) --
	(350.65, 78.68) --
	(351.03, 78.57) --
	(351.41, 78.46) --
	(351.79, 78.34) --
	(352.17, 78.23) --
	(352.55, 78.12) --
	(352.93, 78.01) --
	(353.31, 77.90) --
	(353.69, 77.79) --
	(354.07, 77.68) --
	(354.45, 77.57) --
	(354.83, 77.46) --
	(355.21, 77.35) --
	(355.59, 77.24) --
	(355.97, 77.13) --
	(356.35, 77.02) --
	(356.73, 76.91) --
	(357.11, 76.81) --
	(357.49, 76.70) --
	(357.87, 76.59) --
	(358.25, 76.49) --
	(358.63, 76.38) --
	(359.01, 76.28) --
	(359.38, 76.17) --
	(359.76, 76.07) --
	(360.14, 75.96) --
	(360.52, 75.86) --
	(360.90, 75.75) --
	(361.28, 75.65) --
	(361.66, 75.55) --
	(362.04, 75.44) --
	(362.42, 75.34) --
	(362.80, 75.24) --
	(363.18, 75.14) --
	(363.56, 75.04) --
	(363.94, 74.93) --
	(364.32, 74.83) --
	(364.70, 74.73) --
	(365.08, 74.63) --
	(365.46, 74.53) --
	(365.84, 74.43) --
	(366.22, 74.33) --
	(366.60, 74.24) --
	(366.98, 74.14) --
	(367.36, 74.04) --
	(367.74, 73.94) --
	(368.12, 73.84) --
	(368.50, 73.75) --
	(368.88, 73.65) --
	(369.26, 73.55) --
	(369.64, 73.46) --
	(370.02, 73.36) --
	(370.40, 73.27) --
	(370.78, 73.17) --
	(371.16, 73.08) --
	(371.54, 72.98) --
	(371.92, 72.89) --
	(372.30, 72.80) --
	(372.68, 72.70) --
	(373.05, 72.61) --
	(373.43, 72.52) --
	(373.81, 72.42) --
	(374.19, 72.33) --
	(374.57, 72.24) --
	(374.95, 72.15) --
	(375.33, 72.06) --
	(375.71, 71.97) --
	(376.09, 71.88) --
	(376.47, 71.79) --
	(376.85, 71.70) --
	(377.23, 71.61) --
	(377.61, 71.52) --
	(377.99, 71.43) --
	(378.37, 71.34) --
	(378.75, 71.25) --
	(379.13, 71.16) --
	(379.51, 71.08) --
	(379.89, 70.99) --
	(380.27, 70.90) --
	(380.65, 70.82) --
	(381.03, 70.73) --
	(381.41, 70.64) --
	(381.79, 70.56) --
	(382.17, 70.47) --
	(382.55, 70.39) --
	(382.93, 70.30) --
	(383.31, 70.22) --
	(383.69, 70.13) --
	(384.07, 70.05) --
	(384.45, 69.97) --
	(384.83, 69.88) --
	(385.21, 69.80) --
	(385.59, 69.72) --
	(385.97, 69.63) --
	(386.34, 69.55) --
	(386.72, 69.47) --
	(387.10, 69.39) --
	(387.48, 69.31) --
	(387.86, 69.23) --
	(388.24, 69.15) --
	(388.62, 69.07) --
	(389.00, 68.99) --
	(389.38, 68.91) --
	(389.76, 68.83) --
	(390.14, 68.75) --
	(390.52, 68.67) --
	(390.90, 68.59) --
	(391.28, 68.51) --
	(391.66, 68.43) --
	(392.04, 68.36) --
	(392.42, 68.28) --
	(392.80, 68.20) --
	(393.18, 68.13) --
	(393.56, 68.05) --
	(393.94, 67.97) --
	(394.32, 67.90) --
	(394.70, 67.82) --
	(395.08, 67.75) --
	(395.46, 67.67) --
	(395.84, 67.60) --
	(396.22, 67.52) --
	(396.60, 67.45) --
	(396.98, 67.37) --
	(397.36, 67.30) --
	(397.74, 67.23) --
	(398.12, 67.15) --
	(398.50, 67.08) --
	(398.88, 67.01) --
	(399.26, 66.94) --
	(399.64, 66.86) --
	(400.01, 66.79) --
	(400.39, 66.72) --
	(400.77, 66.65) --
	(401.15, 66.58) --
	(401.53, 66.51) --
	(401.91, 66.44) --
	(402.29, 66.37) --
	(402.67, 66.30) --
	(403.05, 66.23) --
	(403.43, 66.16) --
	(403.81, 66.09) --
	(404.19, 66.02) --
	(404.57, 65.95) --
	(404.95, 65.88) --
	(405.33, 65.81) --
	(405.71, 65.75) --
	(406.09, 65.68) --
	(406.47, 65.61) --
	(406.85, 65.55) --
	(407.23, 65.48) --
	(407.61, 65.41) --
	(407.99, 65.35) --
	(408.37, 65.28) --
	(408.75, 65.21) --
	(409.13, 65.15) --
	(409.51, 65.08) --
	(409.89, 65.02) --
	(410.27, 64.95) --
	(410.65, 64.89) --
	(411.03, 64.83) --
	(411.41, 64.76) --
	(411.79, 64.70) --
	(412.17, 64.63) --
	(412.55, 64.57) --
	(412.93, 64.51) --
	(413.31, 64.45) --
	(413.68, 64.38) --
	(414.06, 64.32) --
	(414.44, 64.26) --
	(414.82, 64.20) --
	(415.20, 64.14) --
	(415.58, 64.07) --
	(415.96, 64.01) --
	(416.34, 63.95) --
	(416.72, 63.89) --
	(417.10, 63.83) --
	(417.48, 63.77) --
	(417.86, 63.71) --
	(418.24, 63.65) --
	(418.62, 63.59) --
	(419.00, 63.53) --
	(419.38, 63.48) --
	(419.76, 63.42) --
	(420.14, 63.36) --
	(420.52, 63.30) --
	(420.90, 63.24) --
	(421.28, 63.18) --
	(421.66, 63.13) --
	(422.04, 63.07) --
	(422.42, 63.01) --
	(422.80, 62.96) --
	(423.18, 62.90) --
	(423.56, 62.84) --
	(423.94, 62.79) --
	(424.32, 62.73) --
	(424.70, 62.68) --
	(425.08, 62.62) --
	(425.46, 62.56) --
	(425.84, 62.51) --
	(426.22, 62.45) --
	(426.60, 62.40) --
	(426.97, 62.35) --
	(427.35, 62.29) --
	(427.73, 62.24) --
	(428.11, 62.18) --
	(428.49, 62.13) --
	(428.87, 62.08) --
	(429.25, 62.02) --
	(429.63, 61.97) --
	(430.01, 61.92) --
	(430.39, 61.87) --
	(430.77, 61.81) --
	(431.15, 61.76) --
	(431.53, 61.71) --
	(431.91, 61.66) --
	(432.29, 61.61) --
	(432.67, 61.56) --
	(433.05, 61.51) --
	(433.43, 61.45) --
	(433.81, 61.40) --
	(434.19, 61.35) --
	(434.57, 61.30) --
	(434.95, 61.25) --
	(435.33, 61.20) --
	(435.71, 61.15) --
	(436.09, 61.10) --
	(436.47, 61.06) --
	(436.85, 61.01) --
	(437.23, 60.96) --
	(437.61, 60.91) --
	(437.99, 60.86) --
	(438.37, 60.81) --
	(438.75, 60.77) --
	(439.13, 60.72) --
	(439.51, 60.67) --
	(439.89, 60.62) --
	(440.27, 60.58) --
	(440.64, 60.53) --
	(441.02, 60.48) --
	(441.40, 60.44) --
	(441.78, 60.39) --
	(442.16, 60.34) --
	(442.54, 60.30) --
	(442.92, 60.25) --
	(443.30, 60.21) --
	(443.68, 60.16) --
	(444.06, 60.11) --
	(444.44, 60.07) --
	(444.82, 60.02) --
	(445.20, 59.98) --
	(445.58, 59.94) --
	(445.96, 59.89) --
	(446.34, 59.85) --
	(446.72, 59.80) --
	(447.10, 59.76) --
	(447.48, 59.72) --
	(447.86, 59.67) --
	(448.24, 59.63) --
	(448.62, 59.59) --
	(449.00, 59.54) --
	(449.38, 59.50) --
	(449.76, 59.46) --
	(450.14, 59.42) --
	(450.52, 59.37) --
	(450.90, 59.33) --
	(451.28, 59.29) --
	(451.66, 59.25) --
	(452.04, 59.21) --
	(452.42, 59.16) --
	(452.80, 59.12) --
	(453.18, 59.08) --
	(453.56, 59.04) --
	(453.93, 59.00) --
	(454.31, 58.96) --
	(454.69, 58.92) --
	(455.07, 58.88) --
	(455.45, 58.84) --
	(455.83, 58.80) --
	(456.21, 58.76) --
	(456.59, 58.72) --
	(456.97, 58.68) --
	(457.35, 58.64) --
	(457.73, 58.60) --
	(458.11, 58.57) --
	(458.49, 58.53) --
	(458.87, 58.49) --
	(459.25, 58.45) --
	(459.63, 58.41) --
	(460.01, 58.37) --
	(460.39, 58.34) --
	(460.77, 58.30) --
	(461.15, 58.26) --
	(461.53, 58.22) --
	(461.91, 58.19) --
	(462.29, 58.15) --
	(462.67, 58.11) --
	(463.05, 58.08) --
	(463.43, 58.04) --
	(463.81, 58.00) --
	(464.19, 57.97) --
	(464.57, 57.93) --
	(464.95, 57.89) --
	(465.33, 57.86) --
	(465.71, 57.82) --
	(466.09, 57.79) --
	(466.47, 57.75) --
	(466.85, 57.72) --
	(467.23, 57.68) --
	(467.60, 57.65) --
	(467.98, 57.61) --
	(468.36, 57.58) --
	(468.74, 57.54) --
	(469.12, 57.51) --
	(469.50, 57.47) --
	(469.88, 57.44) --
	(470.26, 57.41) --
	(470.64, 57.37) --
	(471.02, 57.34) --
	(471.40, 57.31) --
	(471.78, 57.27) --
	(472.16, 57.24) --
	(472.54, 57.21) --
	(472.92, 57.17) --
	(473.30, 57.14) --
	(473.68, 57.11) --
	(474.06, 57.08) --
	(474.44, 57.04) --
	(474.82, 57.01) --
	(475.20, 56.98) --
	(475.58, 56.95) --
	(475.96, 56.92) --
	(476.34, 56.88) --
	(476.72, 56.85) --
	(477.10, 56.82) --
	(477.48, 56.79) --
	(477.86, 56.76) --
	(478.24, 56.73) --
	(478.62, 56.70) --
	(479.00, 56.67) --
	(479.38, 56.64) --
	(479.76, 56.61) --
	(480.14, 56.57) --
	(480.52, 56.54) --
	(480.90, 56.51) --
	(481.27, 56.48) --
	(481.65, 56.45) --
	(482.03, 56.43) --
	(482.41, 56.40) --
	(482.79, 56.37) --
	(483.17, 56.34) --
	(483.55, 56.31) --
	(483.93, 56.28) --
	(484.31, 56.25) --
	(484.69, 56.22) --
	(485.07, 56.19) --
	(485.45, 56.16) --
	(485.83, 56.14) --
	(486.21, 56.11) --
	(486.59, 56.08) --
	(486.97, 56.05) --
	(487.35, 56.02) --
	(487.73, 56.00) --
	(488.11, 55.97) --
	(488.49, 55.94) --
	(488.87, 55.91) --
	(489.25, 55.89) --
	(489.63, 55.86) --
	(490.01, 55.83) --
	(490.39, 55.80) --
	(490.77, 55.78) --
	(491.15, 55.75) --
	(491.53, 55.72) --
	(491.91, 55.70) --
	(492.29, 55.67) --
	(492.67, 55.64) --
	(493.05, 55.62) --
	(493.43, 55.59) --
	(493.81, 55.57) --
	(494.19, 55.54) --
	(494.56, 55.51) --
	(494.94, 55.49) --
	(495.32, 55.46) --
	(495.70, 55.44) --
	(496.08, 55.41) --
	(496.46, 55.39) --
	(496.84, 55.36) --
	(497.22, 55.34) --
	(497.60, 55.31) --
	(497.98, 55.29) --
	(498.36, 55.26) --
	(498.74, 55.24) --
	(499.12, 55.21) --
	(499.50, 55.19) --
	(499.88, 55.17) --
	(500.26, 55.14) --
	(500.64, 55.12) --
	(501.02, 55.09) --
	(501.40, 55.07) --
	(501.78, 55.05) --
	(502.16, 55.02) --
	(502.54, 55.00) --
	(502.92, 54.98) --
	(503.30, 54.95) --
	(503.68, 54.93) --
	(504.06, 54.91) --
	(504.44, 54.88) --
	(504.82, 54.86) --
	(505.20, 54.84) --
	(505.58, 54.82) --
	(505.96, 54.79) --
	(506.34, 54.77) --
	(506.72, 54.75) --
	(507.10, 54.73) --
	(507.48, 54.70) --
	(507.86, 54.68) --
	(508.23, 54.66) --
	(508.61, 54.64) --
	(508.99, 54.62) --
	(509.37, 54.59) --
	(509.75, 54.57) --
	(510.13, 54.55) --
	(510.51, 54.53) --
	(510.89, 54.51) --
	(511.27, 54.49) --
	(511.65, 54.47) --
	(512.03, 54.45) --
	(512.41, 54.42) --
	(512.79, 54.40) --
	(513.17, 54.38) --
	(513.55, 54.36) --
	(513.93, 54.34) --
	(514.31, 54.32) --
	(514.69, 54.30) --
	(515.07, 54.28) --
	(515.45, 54.26) --
	(515.83, 54.24) --
	(516.21, 54.22) --
	(516.59, 54.20) --
	(516.97, 54.18) --
	(517.35, 54.16) --
	(517.73, 54.14) --
	(518.11, 54.12) --
	(518.49, 54.10) --
	(518.87, 54.08) --
	(519.25, 54.06) --
	(519.63, 54.04) --
	(520.01, 54.02) --
	(520.39, 54.01) --
	(520.77, 53.99) --
	(521.15, 53.97) --
	(521.53, 53.95) --
	(521.90, 53.93) --
	(522.28, 53.91) --
	(522.66, 53.89) --
	(523.04, 53.87) --
	(523.42, 53.86) --
	(523.80, 53.84) --
	(524.18, 53.82) --
	(524.56, 53.80) --
	(524.94, 53.78) --
	(525.32, 53.76) --
	(525.70, 53.75) --
	(526.08, 53.73) --
	(526.46, 53.71) --
	(526.84, 53.69) --
	(527.22, 53.68) --
	(527.60, 53.66) --
	(527.98, 53.64) --
	(528.36, 53.62);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (528.36, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.25, 49.80) -- (386.25, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.62, 49.80) -- (433.62, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.99, 49.80) -- (480.99, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.36, 49.80) -- (528.36, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88, 31.87) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (386.25, 31.87) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (433.62, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (480.99, 31.87) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (528.36, 31.87) {20};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88,239.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (333.90, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 87.70) -- (333.90, 87.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,125.59) -- (333.90,125.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,163.49) -- (333.90,163.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,201.38) -- (333.90,201.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,239.28) -- (333.90,239.28);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 46.94) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 84.84) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,122.73) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,160.63) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,198.53) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,236.42) {1.0};
\end{scope}
\end{tikzpicture}

\end{document}