### 2020-08-26 benchmarking/benchmarkingVM/ESS_round1_real.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-25 15:32:33
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,289.08) rectangle (289.08,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,548.96) {\bfseries cons (\textit{real}) vs. nocons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,301.03) {ESS/hr [nocons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,433.62) {ESS/hr [cons (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] (  0.00,368.65) --
	(209.51,578.16);

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,492.84) --
	(  0.00,209.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,342.13) --
	(236.03,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,519.37) --
	(  0.00,236.03);

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round] (  0.00,315.60) --
	(262.56,578.16);

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,545.89) --
	(  0.00,262.56);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92,376.84) {$4\times$};

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,528.36);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (120.87,491.63) --
	(153.32,491.63);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (143.28,489.68) --
	(143.28,493.39);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (143.28,488.40) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (149.77,470.99) --
	(175.39,470.99);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (166.58,468.16) --
	(166.58,473.45);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (166.58,466.95) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (206.16,500.74) --
	(215.69,500.74);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (211.51,500.34) --
	(211.51,501.13);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.51,496.70) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (175.45,479.76) --
	(197.91,479.76);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.80,477.75) --
	(189.80,481.58);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (189.80,475.72) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (134.12,482.07) --
	(177.22,482.07);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.87,479.79) --
	(165.87,484.11);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.87,479.17) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (222.16,504.36) --
	(223.91,504.36);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (223.06,503.90) --
	(223.06,504.81);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (223.06,501.45) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (180.77,450.33) --
	(180.77,456.96);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (180.77,450.70) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (127.28,471.34) --
	(186.05,471.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (173.66,470.84) --
	(173.66,471.83);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (173.66,467.31) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.13,489.98) --
	(211.57,489.98);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (204.21,489.60) --
	(204.21,490.35);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.21,485.95) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (158.71,426.99) --
	(158.71,436.27);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (158.71,428.15) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.28,442.01) --
	(181.28,450.04);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.28,443.54) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (210.92,494.39) --
	(211.41,494.39);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (211.17,494.00) --
	(211.17,494.78);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.17,491.49) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (111.69,458.54) --
	(151.89,458.54);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (140.84,457.29) --
	(140.84,459.71);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (140.84,455.31) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (135.21,465.08) --
	(177.74,465.08);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (166.44,464.31) --
	(166.44,465.81);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (166.44,461.04) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (184.37,486.12) --
	(198.07,486.12);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.42,485.75) --
	(192.42,486.49);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.42,482.09) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (131.08,451.24) --
	(158.21,451.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (149.10,449.98) --
	(149.10,452.42);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (149.10,447.21) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (118.18,448.42) --
	(157.07,448.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.16,445.27) --
	(146.16,451.13);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (146.16,445.52) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (210.39,493.75) --
	(211.29,493.75);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (210.85,493.47) --
	(210.85,494.02);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.85,490.85) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (123.38,424.42) --
	(130.68,424.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (127.38,421.26) --
	(127.38,427.12);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (127.38,421.19) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.20,468.06) --
	(169.06,468.06);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.86,465.26) --
	(160.86,470.50);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.86,464.02) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (151.44,466.43) --
	(152.81,466.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (152.13,465.29) --
	(152.13,467.52);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (152.13,462.40) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (109.33,419.53) --
	(117.33,419.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (113.75,415.02) --
	(113.75,423.18);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (113.75,415.50) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (123.65,403.30) --
	(135.78,403.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (130.66,401.93) --
	(130.66,404.58);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (130.66,400.39) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.69,475.61) --
	(192.26,475.61);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.98,474.76) --
	(191.98,476.41);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.98,472.70) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.12,466.92) --
	(187.12,473.64);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.12,467.35) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (180.14,476.78) --
	(180.14,480.98);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (180.14,474.96) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (216.84,493.07) --
	(230.43,493.07);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.81,492.69) --
	(224.81,493.45);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (224.81,489.04) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (173.46,463.13) --
	(211.65,463.13);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.82,460.82) --
	(200.82,465.20);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.82,459.10) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (204.80,468.25) --
	(204.80,471.97);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.80,467.30) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (219.49,498.64) --
	(220.69,498.64);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (220.10,498.09) --
	(220.10,499.17);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (220.10,495.73) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (128.17,481.37) --
	(128.17,483.93);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (128.17,479.47) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (133.95,488.01) --
	(133.95,491.15);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (133.95,485.61) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (157.10,491.91) --
	(191.40,491.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.09,491.31) --
	(181.09,492.50);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.09,487.88) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (166.81,460.10) --
	(172.28,460.10);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (169.74,458.25) --
	(169.74,461.79);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.74,456.07) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (139.80,456.53) --
	(139.80,458.39);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (139.80,454.58) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (219.72,498.92) --
	(223.33,498.92);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (221.61,498.04) --
	(221.61,499.76);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (221.61,496.01) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 77.66,454.16) --
	( 90.48,454.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 85.13,451.25) --
	( 85.13,456.69);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 85.13,450.94) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 92.57,442.19) --
	(113.59,442.19);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (105.83,425.62) --
	(105.83,450.94);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (105.83,438.16) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (168.95,468.69) --
	(185.12,468.69);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (178.70,468.48) --
	(178.70,468.90);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.70,464.66) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (118.32,432.28) --
	(139.71,432.28);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (131.86,429.85) --
	(131.86,434.45);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (131.86,428.25) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 80.49,443.30) --
	( 96.43,443.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 90.07,438.32) --
	( 90.07,447.26);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 90.07,440.40) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.10,489.05) --
	(210.28,489.05);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (209.22,487.17) --
	(209.22,490.76);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (209.22,486.14) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (211.14,461.92) --
	(211.14,468.04);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.14,462.00) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (159.01,457.26) --
	(159.01,467.68);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (159.01,459.14) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.24,477.77) --
	(220.71,477.77);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.91,477.50) --
	(208.91,478.04);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.91,473.74) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (157.73,460.39) --
	(184.13,460.39);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (175.16,455.16) --
	(175.16,464.50);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (175.16,456.36) {$h$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.46,458.99) --
	(191.46,464.85);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.46,459.24) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (224.99,499.53) --
	(226.68,499.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (225.86,498.82) --
	(225.86,500.21);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.86,496.62) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- (226.02,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85,338.88) -- ( 93.85,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91,338.88) -- (137.91,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96,338.88) -- (181.96,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02,338.88) -- (226.02,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 44.82,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,382.93) -- ( 44.82,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,426.99) -- ( 44.82,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,471.04) -- ( 44.82,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,515.10) -- ( 44.82,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,431.01) rectangle (189.67,426.03);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,421.05) rectangle (189.67,416.07);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,411.09) rectangle (189.67,406.11);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,401.13) rectangle (189.67,396.15);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,391.17) rectangle (189.67,386.19);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,381.21) rectangle (189.67,376.23);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,371.25) rectangle (189.67,366.27);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,361.29) rectangle (189.67,356.31);
\definecolor{fillColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,351.33) rectangle (189.67,346.35);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,425.66) {Ran};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,415.70) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,405.74) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,395.78) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,385.82) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,375.86) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,365.90) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,355.94) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,345.98) {Moyle};
\end{scope}
\begin{scope}
\path[clip] (289.08,289.08) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,548.96) {\bfseries adapt (\textit{real}) vs. nocons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,301.03) {ESS/hr [nocons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,433.62) {ESS/hr [adapt (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] (338.88,338.88) rectangle (528.36,528.36);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] (162.66,242.23) --
	(498.59,578.16);

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,492.84) --
	(162.66, 83.09) --
	(162.66,242.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.66,215.71) --
	(525.11,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,519.37) --
	(162.66,109.61) --
	(162.66,215.71);

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round] (162.66,189.19) --
	(551.64,578.16);

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,545.89) --
	(162.66,136.14) --
	(162.66,189.19);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (425.98,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (399.46,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (371.28,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00,376.84) {$4\times$};

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,528.36);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (409.95,498.53) --
	(442.40,498.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (432.36,495.69) --
	(432.36,501.00);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (432.36,495.30) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (438.85,477.19) --
	(464.47,477.19);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (455.66,473.20) --
	(455.66,480.49);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (455.66,473.16) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (495.24,507.29) --
	(504.77,507.29);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (500.59,504.13) --
	(500.59,510.00);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.59,503.26) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (464.53,488.44) --
	(486.99,488.44);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.88,485.34) --
	(478.88,491.10);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (478.88,484.40) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (423.20,500.32) --
	(466.30,500.32);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (454.95,497.70) --
	(454.95,502.63);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (454.95,497.42) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (511.24,508.32) --
	(512.99,508.32);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (512.14,505.44) --
	(512.14,510.82);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (512.14,505.41) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (469.85,459.47) --
	(469.85,463.29);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (469.85,458.25) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (416.36,473.75) --
	(475.13,473.75);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (462.74,467.42) --
	(462.74,478.51);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (462.74,469.72) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.21,499.81) --
	(500.65,499.81);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (493.29,499.49) --
	(493.29,500.13);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.29,495.78) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (447.79,430.75) --
	(447.79,435.51);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (447.79,429.24) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (470.36,453.95) --
	(470.36,460.21);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (470.36,454.43) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (500.00,500.26) --
	(500.49,500.26);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (500.25,500.10) --
	(500.25,500.41);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.25,497.35) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (400.77,472.45) --
	(440.97,472.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (429.92,469.05) --
	(429.92,475.33);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (429.92,469.22) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (424.29,469.44) --
	(466.82,469.44);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (455.52,468.18) --
	(455.52,470.63);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (455.52,465.41) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (473.45,494.42) --
	(487.15,494.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.50,494.08) --
	(481.50,494.75);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (481.50,490.38) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (420.16,461.88) --
	(447.29,461.88);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (438.18,458.33) --
	(438.18,464.86);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (438.18,457.84) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (407.26,466.76) --
	(446.15,466.76);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (435.24,464.72) --
	(435.24,468.61);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (435.24,463.86) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (499.47,500.16) --
	(500.37,500.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (499.93,499.71) --
	(499.93,500.60);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (499.93,497.25) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (412.46,461.86) --
	(419.76,461.86);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (416.46,457.98) --
	(416.46,465.08);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (416.46,458.63) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (435.28,471.68) --
	(458.14,471.68);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (449.94,467.85) --
	(449.94,474.87);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.94,467.65) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (440.52,473.63) --
	(441.89,473.63);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (441.21,473.45) --
	(441.21,473.81);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (441.21,469.59) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (398.41,428.77) --
	(406.41,428.77);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (402.83,424.39) --
	(402.83,432.32);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (402.83,424.73) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (412.73,451.10) --
	(424.86,451.10);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (419.74,444.27) --
	(419.74,456.13);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (419.74,448.20) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (480.77,482.13) --
	(481.34,482.13);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.06,481.71) --
	(481.06,482.53);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (481.06,479.22) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (476.20,469.20) --
	(476.20,486.87);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.20,476.78) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (469.22,479.54) --
	(469.22,485.69);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (469.22,478.82) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (505.92,510.87) --
	(519.51,510.87);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (513.89,508.22) --
	(513.89,513.20);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (513.89,506.84) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (462.54,475.07) --
	(500.73,475.07);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (489.90,471.46) --
	(489.90,478.10);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (489.90,471.03) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (493.88,480.19) --
	(493.88,486.83);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.88,480.89) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (508.57,502.56) --
	(509.77,502.56);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (509.18,499.82) --
	(509.18,504.96);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (509.18,499.66) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (417.25,495.96) --
	(417.25,497.69);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (417.25,493.62) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (423.03,501.50) --
	(423.03,504.37);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (423.03,498.95) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (446.18,502.51) --
	(480.48,502.51);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (470.17,502.30) --
	(470.17,502.71);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (470.17,498.47) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (455.89,473.26) --
	(461.36,473.26);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (458.82,472.23) --
	(458.82,474.23);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (458.82,469.22) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (428.88,471.94) --
	(428.88,477.11);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (428.88,471.79) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (508.80,510.43) --
	(512.41,510.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (510.69,509.55) --
	(510.69,511.28);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (510.69,507.53) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (366.74,474.68) --
	(379.56,474.68);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (374.21,470.11) --
	(374.21,478.38);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (374.21,471.46) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (381.65,452.98) --
	(402.67,452.98);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (394.91,449.28) --
	(394.91,456.08);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (394.91,448.95) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (458.03,485.78) --
	(474.20,485.78);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (467.78,485.35) --
	(467.78,486.21);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (467.78,481.75) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (407.40,446.62) --
	(428.79,446.62);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (420.94,441.01) --
	(420.94,450.95);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (420.94,442.59) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (369.57,472.80) --
	(385.51,472.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (379.15,467.90) --
	(379.15,476.70);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (379.15,469.90) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (497.18,498.56) --
	(499.36,498.56);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (498.30,497.16) --
	(498.30,499.86);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (498.30,495.65) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (500.22,478.54) --
	(500.22,484.51);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.22,478.53) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (448.09,474.79) --
	(448.09,481.27);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (448.09,474.27) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (461.32,500.43) --
	(509.79,500.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (497.99,500.31) --
	(497.99,500.55);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (497.99,496.40) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (446.81,480.91) --
	(473.21,480.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (464.24,478.84) --
	(464.24,482.78);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (464.24,476.88) {$h$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (480.54,484.34) --
	(480.54,486.46);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (480.54,482.53) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.07,512.69) --
	(515.76,512.69);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (514.94,512.04) --
	(514.94,513.32);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (514.94,509.78) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (515.10,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.93,338.88) -- (382.93,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.99,338.88) -- (426.99,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.04,338.88) -- (471.04,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.10,338.88) -- (515.10,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (382.93,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (426.99,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (471.04,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (515.10,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (333.90,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,382.93) -- (333.90,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,426.99) -- (333.90,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,471.04) -- (333.90,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,515.10) -- (333.90,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,259.88) {\bfseries adapt (\textit{real}) vs. cons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54, 11.95) {ESS/hr [cons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,144.54) {ESS/hr [adapt (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80, 49.80) rectangle (239.28,239.28);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] (  0.00, 79.57) --
	(283.33,362.91) --
	(283.33,203.76) --
	( 79.57,  0.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 53.05) --
	(283.33,336.38) --
	(283.33,230.29) --
	( 53.05,  0.00);

\path[draw=drawColor,line width= 0.7pt,line join=round,line cap=round] (  0.00, 26.52) --
	(283.33,309.86) --
	(283.33,256.81) --
	( 26.52,  0.00);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,140.80) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,114.28) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92, 87.76) {$4\times$};

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28,239.28);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.60,209.45) --
	(204.31,209.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (202.55,206.61) --
	(202.55,211.92);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.55,206.22) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.08,188.11) --
	(184.37,188.11);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.91,184.12) --
	(181.91,191.41);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.91,184.08) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (211.26,218.21) --
	(212.05,218.21);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (211.66,215.05) --
	(211.66,220.92);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.66,214.18) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.67,199.36) --
	(192.50,199.36);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.68,196.26) --
	(190.68,202.02);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (190.68,195.32) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.71,211.24) --
	(195.03,211.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.99,208.62) --
	(192.99,213.55);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.99,208.34) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (214.82,219.24) --
	(215.73,219.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (215.28,216.36) --
	(215.28,221.74);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (215.28,216.33) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (161.25,172.40) --
	(167.88,172.40);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (164.85,170.39) --
	(164.85,174.21);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (164.85,169.17) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.76,184.67) --
	(182.75,184.67);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.26,178.34) --
	(182.26,189.43);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.26,180.64) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.52,210.73) --
	(201.27,210.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.90,210.41) --
	(200.90,211.05);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.90,206.70) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (137.91,144.20) --
	(147.19,144.20);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (143.11,141.67) --
	(143.11,146.43);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (143.11,140.16) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (152.93,168.26) --
	(160.96,168.26);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (157.36,164.87) --
	(157.36,171.13);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (157.36,165.35) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (204.92,211.18) --
	(205.70,211.18);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (205.31,211.02) --
	(205.31,211.33);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.31,208.27) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (168.21,183.37) --
	(170.63,183.37);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (169.46,179.97) --
	(169.46,186.25);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.46,180.14) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (175.23,180.36) --
	(176.73,180.36);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.00,179.10) --
	(176.00,181.55);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (176.00,176.33) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (196.67,205.34) --
	(197.41,205.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (197.04,205.00) --
	(197.04,205.67);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.04,201.30) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.90,172.80) --
	(163.34,172.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (162.16,169.25) --
	(162.16,175.78);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (162.16,168.76) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (156.19,177.68) --
	(162.05,177.68);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (159.34,175.64) --
	(159.34,179.53);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (159.34,174.78) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (204.39,211.08) --
	(204.94,211.08);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (204.67,210.63) --
	(204.67,211.52);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.67,208.17) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (132.18,172.78) --
	(138.04,172.78);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (135.34,168.90) --
	(135.34,176.00);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (135.34,169.55) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.18,182.60) --
	(181.42,182.60);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (178.98,178.77) --
	(178.98,185.79);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.98,178.57) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.21,184.55) --
	(178.44,184.55);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (177.35,184.37) --
	(177.35,184.73);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (177.35,180.51) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (125.94,139.69) --
	(134.10,139.69);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (130.45,135.31) --
	(130.45,143.24);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (130.45,135.65) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (112.85,162.02) --
	(115.50,162.02);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (114.22,155.19) --
	(114.22,167.05);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (114.22,159.12) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.68,193.05) --
	(187.33,193.05);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.53,192.63) --
	(186.53,193.45);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.53,190.14) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (177.84,190.92) --
	(184.56,190.92);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.49,180.12) --
	(181.49,197.79);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.49,187.70) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.70,193.78) --
	(191.90,193.78);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.91,190.46) --
	(189.91,196.61);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (189.91,189.74) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (203.61,221.79) --
	(204.37,221.79);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (203.99,219.14) --
	(203.99,224.12);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.99,217.76) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (171.74,185.99) --
	(176.12,185.99);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (174.05,182.38) --
	(174.05,189.02);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.05,181.95) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.17,194.72) --
	(182.89,194.72);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.12,191.11) --
	(181.12,197.75);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.12,191.81) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (209.01,213.48) --
	(210.09,213.48);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (209.56,210.74) --
	(209.56,215.88);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (209.56,210.58) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.29,207.77) --
	(194.85,207.77);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.61,206.88) --
	(193.61,208.61);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.61,204.54) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (198.93,213.91) --
	(202.07,213.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (200.57,212.42) --
	(200.57,215.29);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.57,209.87) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (202.23,213.43) --
	(203.42,213.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (202.83,213.22) --
	(202.83,213.63);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.83,209.39) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (169.17,184.18) --
	(172.71,184.18);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (171.02,183.15) --
	(171.02,185.15);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (171.02,180.14) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (167.45,185.62) --
	(169.31,185.62);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (168.40,182.86) --
	(168.40,188.03);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (168.40,182.71) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (208.96,221.35) --
	(210.68,221.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (209.84,220.47) --
	(209.84,222.20);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (209.84,218.45) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (162.17,185.60) --
	(167.61,185.60);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.08,181.03) --
	(165.08,189.30);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.08,182.38) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (136.54,163.90) --
	(161.86,163.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (153.11,160.20) --
	(153.11,167.00);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (153.11,159.87) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.40,196.70) --
	(179.82,196.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.61,196.27) --
	(179.61,197.13);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.61,192.67) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (140.77,157.54) --
	(145.37,157.54);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (143.20,151.93) --
	(143.20,161.87);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (143.20,153.51) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (149.24,183.72) --
	(158.18,183.72);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (154.22,178.82) --
	(154.22,187.62);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (154.22,180.82) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (198.09,209.48) --
	(201.68,209.48);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (199.97,208.08) --
	(199.97,210.78);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.97,206.57) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.84,192.68) --
	(178.96,192.68);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.15,189.46) --
	(176.15,195.43);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (176.15,189.45) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (168.18,189.23) --
	(178.60,189.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (174.09,185.71) --
	(174.09,192.19);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.09,185.19) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.42,211.35) --
	(188.96,211.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.69,211.23) --
	(188.69,211.47);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.69,207.32) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (166.08,191.83) --
	(175.42,191.83);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (171.31,189.76) --
	(171.31,193.70);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (171.31,187.80) {$h$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (169.91,196.35) --
	(175.77,196.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (173.06,195.26) --
	(173.06,197.38);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (173.06,193.45) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (209.74,223.61) --
	(211.13,223.61);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (210.45,222.96) --
	(210.45,224.24);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.45,220.70) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- (226.02, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85, 49.80) -- ( 93.85, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91, 49.80) -- (137.91, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96, 49.80) -- (181.96, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02, 49.80) -- (226.02, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96, 31.87) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02, 31.87) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80,226.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 44.82, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 93.85) -- ( 44.82, 93.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,137.91) -- ( 44.82,137.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,181.96) -- ( 44.82,181.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,226.02) -- ( 44.82,226.02);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 91.00) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,135.05) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,179.11) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,223.16) {10000};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,259.88) {\bfseries \texttt{AdaptiveOperatorSampler} weights};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62, 11.95) {Sequence length $L$ (kb)};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,144.54) {Learned operator weight};
\end{scope}
\begin{scope}
\path[clip] (338.88, 49.80) rectangle (528.36,239.28);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (485.73,232.77) circle (  1.87);

\path[fill=fillColor] (485.73, 55.06) --
	(488.24, 50.70) --
	(483.21, 50.70) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (356.88,165.38) circle (  1.87);

\path[fill=fillColor] (356.88, 80.00) --
	(359.40, 75.65) --
	(354.37, 75.65) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (394.78,149.36) circle (  1.87);

\path[fill=fillColor] (394.78, 80.07) --
	(397.29, 75.72) --
	(392.26, 75.72) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (390.04,184.14) circle (  1.87);

\path[fill=fillColor] (390.04, 69.81) --
	(392.55, 65.46) --
	(387.52, 65.46) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (404.25,211.12) circle (  1.87);

\path[fill=fillColor] (404.25, 60.31) --
	(406.77, 55.95) --
	(401.74, 55.95) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (359.72,111.61) circle (  1.87);

\path[fill=fillColor] (359.72, 94.33) --
	(362.24, 89.97) --
	(357.21, 89.97) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (433.62,198.59) circle (  1.87);

\path[fill=fillColor] (433.62, 71.59) --
	(436.14, 67.23) --
	(431.10, 67.23) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (359.72,131.32) circle (  1.87);

\path[fill=fillColor] (359.72, 82.69) --
	(362.24, 78.34) --
	(357.21, 78.34) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (346.46,125.73) circle (  1.87);

\path[fill=fillColor] (346.46, 76.98) --
	(348.97, 72.63) --
	(343.94, 72.63) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,115.46) --
	(339.26,115.86) --
	(339.64,116.25) --
	(340.02,116.65) --
	(340.40,117.05) --
	(340.78,117.45) --
	(341.16,117.85) --
	(341.54,118.25) --
	(341.92,118.65) --
	(342.30,119.06) --
	(342.68,119.46) --
	(343.06,119.87) --
	(343.44,120.27) --
	(343.82,120.68) --
	(344.20,121.09) --
	(344.58,121.50) --
	(344.96,121.91) --
	(345.34,122.32) --
	(345.71,122.73) --
	(346.09,123.14) --
	(346.47,123.56) --
	(346.85,123.97) --
	(347.23,124.39) --
	(347.61,124.80) --
	(347.99,125.22) --
	(348.37,125.64) --
	(348.75,126.06) --
	(349.13,126.48) --
	(349.51,126.90) --
	(349.89,127.32) --
	(350.27,127.74) --
	(350.65,128.16) --
	(351.03,128.58) --
	(351.41,129.01) --
	(351.79,129.43) --
	(352.17,129.85) --
	(352.55,130.28) --
	(352.93,130.71) --
	(353.31,131.13) --
	(353.69,131.56) --
	(354.07,131.99) --
	(354.45,132.42) --
	(354.83,132.84) --
	(355.21,133.27) --
	(355.59,133.70) --
	(355.97,134.13) --
	(356.35,134.56) --
	(356.73,134.99) --
	(357.11,135.43) --
	(357.49,135.86) --
	(357.87,136.29) --
	(358.25,136.72) --
	(358.63,137.15) --
	(359.01,137.59) --
	(359.38,138.02) --
	(359.76,138.45) --
	(360.14,138.89) --
	(360.52,139.32) --
	(360.90,139.76) --
	(361.28,140.19) --
	(361.66,140.63) --
	(362.04,141.06) --
	(362.42,141.50) --
	(362.80,141.93) --
	(363.18,142.37) --
	(363.56,142.80) --
	(363.94,143.24) --
	(364.32,143.67) --
	(364.70,144.11) --
	(365.08,144.54) --
	(365.46,144.98) --
	(365.84,145.41) --
	(366.22,145.85) --
	(366.60,146.28) --
	(366.98,146.72) --
	(367.36,147.16) --
	(367.74,147.59) --
	(368.12,148.03) --
	(368.50,148.46) --
	(368.88,148.89) --
	(369.26,149.33) --
	(369.64,149.76) --
	(370.02,150.20) --
	(370.40,150.63) --
	(370.78,151.07) --
	(371.16,151.50) --
	(371.54,151.93) --
	(371.92,152.36) --
	(372.30,152.80) --
	(372.68,153.23) --
	(373.05,153.66) --
	(373.43,154.09) --
	(373.81,154.52) --
	(374.19,154.95) --
	(374.57,155.38) --
	(374.95,155.81) --
	(375.33,156.24) --
	(375.71,156.67) --
	(376.09,157.10) --
	(376.47,157.53) --
	(376.85,157.95) --
	(377.23,158.38) --
	(377.61,158.81) --
	(377.99,159.23) --
	(378.37,159.66) --
	(378.75,160.08) --
	(379.13,160.50) --
	(379.51,160.93) --
	(379.89,161.35) --
	(380.27,161.77) --
	(380.65,162.19) --
	(381.03,162.61) --
	(381.41,163.03) --
	(381.79,163.45) --
	(382.17,163.87) --
	(382.55,164.28) --
	(382.93,164.70) --
	(383.31,165.11) --
	(383.69,165.53) --
	(384.07,165.94) --
	(384.45,166.36) --
	(384.83,166.77) --
	(385.21,167.18) --
	(385.59,167.59) --
	(385.97,168.00) --
	(386.34,168.41) --
	(386.72,168.81) --
	(387.10,169.22) --
	(387.48,169.63) --
	(387.86,170.03) --
	(388.24,170.43) --
	(388.62,170.84) --
	(389.00,171.24) --
	(389.38,171.64) --
	(389.76,172.04) --
	(390.14,172.44) --
	(390.52,172.83) --
	(390.90,173.23) --
	(391.28,173.62) --
	(391.66,174.02) --
	(392.04,174.41) --
	(392.42,174.80) --
	(392.80,175.19) --
	(393.18,175.58) --
	(393.56,175.97) --
	(393.94,176.36) --
	(394.32,176.74) --
	(394.70,177.13) --
	(395.08,177.51) --
	(395.46,177.89) --
	(395.84,178.27) --
	(396.22,178.65) --
	(396.60,179.03) --
	(396.98,179.41) --
	(397.36,179.79) --
	(397.74,180.16) --
	(398.12,180.53) --
	(398.50,180.91) --
	(398.88,181.28) --
	(399.26,181.65) --
	(399.64,182.01) --
	(400.01,182.38) --
	(400.39,182.75) --
	(400.77,183.11) --
	(401.15,183.47) --
	(401.53,183.83) --
	(401.91,184.19) --
	(402.29,184.55) --
	(402.67,184.91) --
	(403.05,185.26) --
	(403.43,185.62) --
	(403.81,185.97) --
	(404.19,186.32) --
	(404.57,186.67) --
	(404.95,187.02) --
	(405.33,187.37) --
	(405.71,187.71) --
	(406.09,188.06) --
	(406.47,188.40) --
	(406.85,188.74) --
	(407.23,189.08) --
	(407.61,189.42) --
	(407.99,189.76) --
	(408.37,190.09) --
	(408.75,190.43) --
	(409.13,190.76) --
	(409.51,191.09) --
	(409.89,191.42) --
	(410.27,191.75) --
	(410.65,192.08) --
	(411.03,192.40) --
	(411.41,192.73) --
	(411.79,193.05) --
	(412.17,193.37) --
	(412.55,193.69) --
	(412.93,194.00) --
	(413.31,194.32) --
	(413.68,194.64) --
	(414.06,194.95) --
	(414.44,195.26) --
	(414.82,195.57) --
	(415.20,195.88) --
	(415.58,196.18) --
	(415.96,196.49) --
	(416.34,196.79) --
	(416.72,197.10) --
	(417.10,197.40) --
	(417.48,197.70) --
	(417.86,197.99) --
	(418.24,198.29) --
	(418.62,198.58) --
	(419.00,198.88) --
	(419.38,199.17) --
	(419.76,199.46) --
	(420.14,199.75) --
	(420.52,200.03) --
	(420.90,200.32) --
	(421.28,200.60) --
	(421.66,200.89) --
	(422.04,201.17) --
	(422.42,201.45) --
	(422.80,201.72) --
	(423.18,202.00) --
	(423.56,202.27) --
	(423.94,202.55) --
	(424.32,202.82) --
	(424.70,203.09) --
	(425.08,203.36) --
	(425.46,203.62) --
	(425.84,203.89) --
	(426.22,204.15) --
	(426.60,204.41) --
	(426.97,204.68) --
	(427.35,204.93) --
	(427.73,205.19) --
	(428.11,205.45) --
	(428.49,205.70) --
	(428.87,205.96) --
	(429.25,206.21) --
	(429.63,206.46) --
	(430.01,206.71) --
	(430.39,206.95) --
	(430.77,207.20) --
	(431.15,207.44) --
	(431.53,207.69) --
	(431.91,207.93) --
	(432.29,208.17) --
	(432.67,208.41) --
	(433.05,208.64) --
	(433.43,208.88) --
	(433.81,209.11) --
	(434.19,209.35) --
	(434.57,209.58) --
	(434.95,209.81) --
	(435.33,210.03) --
	(435.71,210.26) --
	(436.09,210.49) --
	(436.47,210.71) --
	(436.85,210.93) --
	(437.23,211.15) --
	(437.61,211.37) --
	(437.99,211.59) --
	(438.37,211.81) --
	(438.75,212.02) --
	(439.13,212.24) --
	(439.51,212.45) --
	(439.89,212.66) --
	(440.27,212.87) --
	(440.64,213.08) --
	(441.02,213.28) --
	(441.40,213.49) --
	(441.78,213.69) --
	(442.16,213.90) --
	(442.54,214.10) --
	(442.92,214.30) --
	(443.30,214.50) --
	(443.68,214.69) --
	(444.06,214.89) --
	(444.44,215.09) --
	(444.82,215.28) --
	(445.20,215.47) --
	(445.58,215.66) --
	(445.96,215.85) --
	(446.34,216.04) --
	(446.72,216.23) --
	(447.10,216.41) --
	(447.48,216.60) --
	(447.86,216.78) --
	(448.24,216.96) --
	(448.62,217.14) --
	(449.00,217.32) --
	(449.38,217.50) --
	(449.76,217.67) --
	(450.14,217.85) --
	(450.52,218.02) --
	(450.90,218.20) --
	(451.28,218.37) --
	(451.66,218.54) --
	(452.04,218.71) --
	(452.42,218.88) --
	(452.80,219.04) --
	(453.18,219.21) --
	(453.56,219.37) --
	(453.93,219.54) --
	(454.31,219.70) --
	(454.69,219.86) --
	(455.07,220.02) --
	(455.45,220.18) --
	(455.83,220.33) --
	(456.21,220.49) --
	(456.59,220.65) --
	(456.97,220.80) --
	(457.35,220.95) --
	(457.73,221.10) --
	(458.11,221.25) --
	(458.49,221.40) --
	(458.87,221.55) --
	(459.25,221.70) --
	(459.63,221.85) --
	(460.01,221.99) --
	(460.39,222.13) --
	(460.77,222.28) --
	(461.15,222.42) --
	(461.53,222.56) --
	(461.91,222.70) --
	(462.29,222.84) --
	(462.67,222.98) --
	(463.05,223.11) --
	(463.43,223.25) --
	(463.81,223.38) --
	(464.19,223.52) --
	(464.57,223.65) --
	(464.95,223.78) --
	(465.33,223.91) --
	(465.71,224.04) --
	(466.09,224.17) --
	(466.47,224.29) --
	(466.85,224.42) --
	(467.23,224.55) --
	(467.60,224.67) --
	(467.98,224.79) --
	(468.36,224.92) --
	(468.74,225.04) --
	(469.12,225.16) --
	(469.50,225.28) --
	(469.88,225.40) --
	(470.26,225.51) --
	(470.64,225.63) --
	(471.02,225.75) --
	(471.40,225.86) --
	(471.78,225.98) --
	(472.16,226.09) --
	(472.54,226.20) --
	(472.92,226.31) --
	(473.30,226.42) --
	(473.68,226.53) --
	(474.06,226.64) --
	(474.44,226.75) --
	(474.82,226.86) --
	(475.20,226.96) --
	(475.58,227.07) --
	(475.96,227.17) --
	(476.34,227.28) --
	(476.72,227.38) --
	(477.10,227.48) --
	(477.48,227.59) --
	(477.86,227.69) --
	(478.24,227.79) --
	(478.62,227.88) --
	(479.00,227.98) --
	(479.38,228.08) --
	(479.76,228.18) --
	(480.14,228.27) --
	(480.52,228.37) --
	(480.90,228.46) --
	(481.27,228.55) --
	(481.65,228.65) --
	(482.03,228.74) --
	(482.41,228.83) --
	(482.79,228.92) --
	(483.17,229.01) --
	(483.55,229.10) --
	(483.93,229.19) --
	(484.31,229.27) --
	(484.69,229.36) --
	(485.07,229.45) --
	(485.45,229.53) --
	(485.83,229.62) --
	(486.21,229.70) --
	(486.59,229.78) --
	(486.97,229.87) --
	(487.35,229.95) --
	(487.73,230.03) --
	(488.11,230.11) --
	(488.49,230.19) --
	(488.87,230.27) --
	(489.25,230.35) --
	(489.63,230.43) --
	(490.01,230.50) --
	(490.39,230.58) --
	(490.77,230.66) --
	(491.15,230.73) --
	(491.53,230.81) --
	(491.91,230.88) --
	(492.29,230.95) --
	(492.67,231.03) --
	(493.05,231.10) --
	(493.43,231.17) --
	(493.81,231.24) --
	(494.19,231.31) --
	(494.56,231.38) --
	(494.94,231.45) --
	(495.32,231.52) --
	(495.70,231.59) --
	(496.08,231.66) --
	(496.46,231.72) --
	(496.84,231.79) --
	(497.22,231.86) --
	(497.60,231.92) --
	(497.98,231.99) --
	(498.36,232.05) --
	(498.74,232.11) --
	(499.12,232.18) --
	(499.50,232.24) --
	(499.88,232.30) --
	(500.26,232.36) --
	(500.64,232.42) --
	(501.02,232.48) --
	(501.40,232.54) --
	(501.78,232.60) --
	(502.16,232.66) --
	(502.54,232.72) --
	(502.92,232.78) --
	(503.30,232.84) --
	(503.68,232.89) --
	(504.06,232.95) --
	(504.44,233.01) --
	(504.82,233.06) --
	(505.20,233.12) --
	(505.58,233.17) --
	(505.96,233.23) --
	(506.34,233.28) --
	(506.72,233.33) --
	(507.10,233.38) --
	(507.48,233.44) --
	(507.86,233.49) --
	(508.23,233.54) --
	(508.61,233.59) --
	(508.99,233.64) --
	(509.37,233.69) --
	(509.75,233.74) --
	(510.13,233.79) --
	(510.51,233.84) --
	(510.89,233.89) --
	(511.27,233.94) --
	(511.65,233.98) --
	(512.03,234.03) --
	(512.41,234.08) --
	(512.79,234.12) --
	(513.17,234.17) --
	(513.55,234.21) --
	(513.93,234.26) --
	(514.31,234.30) --
	(514.69,234.35) --
	(515.07,234.39) --
	(515.45,234.44) --
	(515.83,234.48) --
	(516.21,234.52) --
	(516.59,234.57) --
	(516.97,234.61) --
	(517.35,234.65) --
	(517.73,234.69) --
	(518.11,234.73) --
	(518.49,234.77) --
	(518.87,234.81) --
	(519.25,234.85) --
	(519.63,234.89) --
	(520.01,234.93) --
	(520.39,234.97) --
	(520.77,235.01) --
	(521.15,235.05) --
	(521.53,235.08) --
	(521.90,235.12) --
	(522.28,235.16) --
	(522.66,235.20) --
	(523.04,235.23) --
	(523.42,235.27) --
	(523.80,235.30) --
	(524.18,235.34) --
	(524.56,235.38) --
	(524.94,235.41) --
	(525.32,235.45) --
	(525.70,235.48) --
	(526.08,235.51) --
	(526.46,235.55) --
	(526.84,235.58) --
	(527.22,235.61) --
	(527.60,235.65) --
	(527.98,235.68) --
	(528.36,235.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 86.82) --
	(339.26, 86.66) --
	(339.64, 86.51) --
	(340.02, 86.35) --
	(340.40, 86.20) --
	(340.78, 86.04) --
	(341.16, 85.89) --
	(341.54, 85.74) --
	(341.92, 85.58) --
	(342.30, 85.43) --
	(342.68, 85.28) --
	(343.06, 85.13) --
	(343.44, 84.98) --
	(343.82, 84.83) --
	(344.20, 84.68) --
	(344.58, 84.53) --
	(344.96, 84.38) --
	(345.34, 84.24) --
	(345.71, 84.09) --
	(346.09, 83.94) --
	(346.47, 83.80) --
	(346.85, 83.65) --
	(347.23, 83.50) --
	(347.61, 83.36) --
	(347.99, 83.21) --
	(348.37, 83.07) --
	(348.75, 82.93) --
	(349.13, 82.78) --
	(349.51, 82.64) --
	(349.89, 82.50) --
	(350.27, 82.36) --
	(350.65, 82.22) --
	(351.03, 82.08) --
	(351.41, 81.94) --
	(351.79, 81.80) --
	(352.17, 81.66) --
	(352.55, 81.52) --
	(352.93, 81.38) --
	(353.31, 81.24) --
	(353.69, 81.11) --
	(354.07, 80.97) --
	(354.45, 80.83) --
	(354.83, 80.70) --
	(355.21, 80.56) --
	(355.59, 80.43) --
	(355.97, 80.29) --
	(356.35, 80.16) --
	(356.73, 80.03) --
	(357.11, 79.89) --
	(357.49, 79.76) --
	(357.87, 79.63) --
	(358.25, 79.50) --
	(358.63, 79.37) --
	(359.01, 79.24) --
	(359.38, 79.11) --
	(359.76, 78.98) --
	(360.14, 78.85) --
	(360.52, 78.72) --
	(360.90, 78.59) --
	(361.28, 78.46) --
	(361.66, 78.34) --
	(362.04, 78.21) --
	(362.42, 78.08) --
	(362.80, 77.96) --
	(363.18, 77.83) --
	(363.56, 77.71) --
	(363.94, 77.58) --
	(364.32, 77.46) --
	(364.70, 77.33) --
	(365.08, 77.21) --
	(365.46, 77.09) --
	(365.84, 76.97) --
	(366.22, 76.85) --
	(366.60, 76.72) --
	(366.98, 76.60) --
	(367.36, 76.48) --
	(367.74, 76.36) --
	(368.12, 76.24) --
	(368.50, 76.12) --
	(368.88, 76.01) --
	(369.26, 75.89) --
	(369.64, 75.77) --
	(370.02, 75.65) --
	(370.40, 75.54) --
	(370.78, 75.42) --
	(371.16, 75.30) --
	(371.54, 75.19) --
	(371.92, 75.07) --
	(372.30, 74.96) --
	(372.68, 74.85) --
	(373.05, 74.73) --
	(373.43, 74.62) --
	(373.81, 74.51) --
	(374.19, 74.39) --
	(374.57, 74.28) --
	(374.95, 74.17) --
	(375.33, 74.06) --
	(375.71, 73.95) --
	(376.09, 73.84) --
	(376.47, 73.73) --
	(376.85, 73.62) --
	(377.23, 73.51) --
	(377.61, 73.40) --
	(377.99, 73.29) --
	(378.37, 73.19) --
	(378.75, 73.08) --
	(379.13, 72.97) --
	(379.51, 72.87) --
	(379.89, 72.76) --
	(380.27, 72.65) --
	(380.65, 72.55) --
	(381.03, 72.44) --
	(381.41, 72.34) --
	(381.79, 72.24) --
	(382.17, 72.13) --
	(382.55, 72.03) --
	(382.93, 71.93) --
	(383.31, 71.83) --
	(383.69, 71.72) --
	(384.07, 71.62) --
	(384.45, 71.52) --
	(384.83, 71.42) --
	(385.21, 71.32) --
	(385.59, 71.22) --
	(385.97, 71.12) --
	(386.34, 71.02) --
	(386.72, 70.92) --
	(387.10, 70.83) --
	(387.48, 70.73) --
	(387.86, 70.63) --
	(388.24, 70.53) --
	(388.62, 70.44) --
	(389.00, 70.34) --
	(389.38, 70.25) --
	(389.76, 70.15) --
	(390.14, 70.06) --
	(390.52, 69.96) --
	(390.90, 69.87) --
	(391.28, 69.77) --
	(391.66, 69.68) --
	(392.04, 69.59) --
	(392.42, 69.49) --
	(392.80, 69.40) --
	(393.18, 69.31) --
	(393.56, 69.22) --
	(393.94, 69.13) --
	(394.32, 69.04) --
	(394.70, 68.95) --
	(395.08, 68.86) --
	(395.46, 68.77) --
	(395.84, 68.68) --
	(396.22, 68.59) --
	(396.60, 68.50) --
	(396.98, 68.41) --
	(397.36, 68.32) --
	(397.74, 68.24) --
	(398.12, 68.15) --
	(398.50, 68.06) --
	(398.88, 67.98) --
	(399.26, 67.89) --
	(399.64, 67.80) --
	(400.01, 67.72) --
	(400.39, 67.63) --
	(400.77, 67.55) --
	(401.15, 67.47) --
	(401.53, 67.38) --
	(401.91, 67.30) --
	(402.29, 67.22) --
	(402.67, 67.13) --
	(403.05, 67.05) --
	(403.43, 66.97) --
	(403.81, 66.89) --
	(404.19, 66.81) --
	(404.57, 66.73) --
	(404.95, 66.64) --
	(405.33, 66.56) --
	(405.71, 66.48) --
	(406.09, 66.40) --
	(406.47, 66.33) --
	(406.85, 66.25) --
	(407.23, 66.17) --
	(407.61, 66.09) --
	(407.99, 66.01) --
	(408.37, 65.93) --
	(408.75, 65.86) --
	(409.13, 65.78) --
	(409.51, 65.70) --
	(409.89, 65.63) --
	(410.27, 65.55) --
	(410.65, 65.48) --
	(411.03, 65.40) --
	(411.41, 65.33) --
	(411.79, 65.25) --
	(412.17, 65.18) --
	(412.55, 65.10) --
	(412.93, 65.03) --
	(413.31, 64.96) --
	(413.68, 64.88) --
	(414.06, 64.81) --
	(414.44, 64.74) --
	(414.82, 64.67) --
	(415.20, 64.60) --
	(415.58, 64.52) --
	(415.96, 64.45) --
	(416.34, 64.38) --
	(416.72, 64.31) --
	(417.10, 64.24) --
	(417.48, 64.17) --
	(417.86, 64.10) --
	(418.24, 64.03) --
	(418.62, 63.97) --
	(419.00, 63.90) --
	(419.38, 63.83) --
	(419.76, 63.76) --
	(420.14, 63.69) --
	(420.52, 63.63) --
	(420.90, 63.56) --
	(421.28, 63.49) --
	(421.66, 63.43) --
	(422.04, 63.36) --
	(422.42, 63.29) --
	(422.80, 63.23) --
	(423.18, 63.16) --
	(423.56, 63.10) --
	(423.94, 63.03) --
	(424.32, 62.97) --
	(424.70, 62.91) --
	(425.08, 62.84) --
	(425.46, 62.78) --
	(425.84, 62.72) --
	(426.22, 62.65) --
	(426.60, 62.59) --
	(426.97, 62.53) --
	(427.35, 62.47) --
	(427.73, 62.40) --
	(428.11, 62.34) --
	(428.49, 62.28) --
	(428.87, 62.22) --
	(429.25, 62.16) --
	(429.63, 62.10) --
	(430.01, 62.04) --
	(430.39, 61.98) --
	(430.77, 61.92) --
	(431.15, 61.86) --
	(431.53, 61.80) --
	(431.91, 61.74) --
	(432.29, 61.68) --
	(432.67, 61.63) --
	(433.05, 61.57) --
	(433.43, 61.51) --
	(433.81, 61.45) --
	(434.19, 61.40) --
	(434.57, 61.34) --
	(434.95, 61.28) --
	(435.33, 61.23) --
	(435.71, 61.17) --
	(436.09, 61.11) --
	(436.47, 61.06) --
	(436.85, 61.00) --
	(437.23, 60.95) --
	(437.61, 60.89) --
	(437.99, 60.84) --
	(438.37, 60.78) --
	(438.75, 60.73) --
	(439.13, 60.68) --
	(439.51, 60.62) --
	(439.89, 60.57) --
	(440.27, 60.52) --
	(440.64, 60.46) --
	(441.02, 60.41) --
	(441.40, 60.36) --
	(441.78, 60.31) --
	(442.16, 60.25) --
	(442.54, 60.20) --
	(442.92, 60.15) --
	(443.30, 60.10) --
	(443.68, 60.05) --
	(444.06, 60.00) --
	(444.44, 59.95) --
	(444.82, 59.90) --
	(445.20, 59.85) --
	(445.58, 59.80) --
	(445.96, 59.75) --
	(446.34, 59.70) --
	(446.72, 59.65) --
	(447.10, 59.60) --
	(447.48, 59.55) --
	(447.86, 59.50) --
	(448.24, 59.46) --
	(448.62, 59.41) --
	(449.00, 59.36) --
	(449.38, 59.31) --
	(449.76, 59.27) --
	(450.14, 59.22) --
	(450.52, 59.17) --
	(450.90, 59.13) --
	(451.28, 59.08) --
	(451.66, 59.03) --
	(452.04, 58.99) --
	(452.42, 58.94) --
	(452.80, 58.90) --
	(453.18, 58.85) --
	(453.56, 58.81) --
	(453.93, 58.76) --
	(454.31, 58.72) --
	(454.69, 58.67) --
	(455.07, 58.63) --
	(455.45, 58.58) --
	(455.83, 58.54) --
	(456.21, 58.50) --
	(456.59, 58.45) --
	(456.97, 58.41) --
	(457.35, 58.37) --
	(457.73, 58.32) --
	(458.11, 58.28) --
	(458.49, 58.24) --
	(458.87, 58.20) --
	(459.25, 58.15) --
	(459.63, 58.11) --
	(460.01, 58.07) --
	(460.39, 58.03) --
	(460.77, 57.99) --
	(461.15, 57.95) --
	(461.53, 57.91) --
	(461.91, 57.87) --
	(462.29, 57.83) --
	(462.67, 57.79) --
	(463.05, 57.75) --
	(463.43, 57.71) --
	(463.81, 57.67) --
	(464.19, 57.63) --
	(464.57, 57.59) --
	(464.95, 57.55) --
	(465.33, 57.51) --
	(465.71, 57.47) --
	(466.09, 57.43) --
	(466.47, 57.40) --
	(466.85, 57.36) --
	(467.23, 57.32) --
	(467.60, 57.28) --
	(467.98, 57.24) --
	(468.36, 57.21) --
	(468.74, 57.17) --
	(469.12, 57.13) --
	(469.50, 57.10) --
	(469.88, 57.06) --
	(470.26, 57.02) --
	(470.64, 56.99) --
	(471.02, 56.95) --
	(471.40, 56.91) --
	(471.78, 56.88) --
	(472.16, 56.84) --
	(472.54, 56.81) --
	(472.92, 56.77) --
	(473.30, 56.74) --
	(473.68, 56.70) --
	(474.06, 56.67) --
	(474.44, 56.63) --
	(474.82, 56.60) --
	(475.20, 56.56) --
	(475.58, 56.53) --
	(475.96, 56.50) --
	(476.34, 56.46) --
	(476.72, 56.43) --
	(477.10, 56.39) --
	(477.48, 56.36) --
	(477.86, 56.33) --
	(478.24, 56.30) --
	(478.62, 56.26) --
	(479.00, 56.23) --
	(479.38, 56.20) --
	(479.76, 56.17) --
	(480.14, 56.13) --
	(480.52, 56.10) --
	(480.90, 56.07) --
	(481.27, 56.04) --
	(481.65, 56.01) --
	(482.03, 55.97) --
	(482.41, 55.94) --
	(482.79, 55.91) --
	(483.17, 55.88) --
	(483.55, 55.85) --
	(483.93, 55.82) --
	(484.31, 55.79) --
	(484.69, 55.76) --
	(485.07, 55.73) --
	(485.45, 55.70) --
	(485.83, 55.67) --
	(486.21, 55.64) --
	(486.59, 55.61) --
	(486.97, 55.58) --
	(487.35, 55.55) --
	(487.73, 55.52) --
	(488.11, 55.49) --
	(488.49, 55.46) --
	(488.87, 55.44) --
	(489.25, 55.41) --
	(489.63, 55.38) --
	(490.01, 55.35) --
	(490.39, 55.32) --
	(490.77, 55.29) --
	(491.15, 55.27) --
	(491.53, 55.24) --
	(491.91, 55.21) --
	(492.29, 55.18) --
	(492.67, 55.16) --
	(493.05, 55.13) --
	(493.43, 55.10) --
	(493.81, 55.07) --
	(494.19, 55.05) --
	(494.56, 55.02) --
	(494.94, 54.99) --
	(495.32, 54.97) --
	(495.70, 54.94) --
	(496.08, 54.92) --
	(496.46, 54.89) --
	(496.84, 54.86) --
	(497.22, 54.84) --
	(497.60, 54.81) --
	(497.98, 54.79) --
	(498.36, 54.76) --
	(498.74, 54.74) --
	(499.12, 54.71) --
	(499.50, 54.69) --
	(499.88, 54.66) --
	(500.26, 54.64) --
	(500.64, 54.61) --
	(501.02, 54.59) --
	(501.40, 54.56) --
	(501.78, 54.54) --
	(502.16, 54.51) --
	(502.54, 54.49) --
	(502.92, 54.47) --
	(503.30, 54.44) --
	(503.68, 54.42) --
	(504.06, 54.40) --
	(504.44, 54.37) --
	(504.82, 54.35) --
	(505.20, 54.33) --
	(505.58, 54.30) --
	(505.96, 54.28) --
	(506.34, 54.26) --
	(506.72, 54.23) --
	(507.10, 54.21) --
	(507.48, 54.19) --
	(507.86, 54.17) --
	(508.23, 54.14) --
	(508.61, 54.12) --
	(508.99, 54.10) --
	(509.37, 54.08) --
	(509.75, 54.06) --
	(510.13, 54.03) --
	(510.51, 54.01) --
	(510.89, 53.99) --
	(511.27, 53.97) --
	(511.65, 53.95) --
	(512.03, 53.93) --
	(512.41, 53.91) --
	(512.79, 53.89) --
	(513.17, 53.86) --
	(513.55, 53.84) --
	(513.93, 53.82) --
	(514.31, 53.80) --
	(514.69, 53.78) --
	(515.07, 53.76) --
	(515.45, 53.74) --
	(515.83, 53.72) --
	(516.21, 53.70) --
	(516.59, 53.68) --
	(516.97, 53.66) --
	(517.35, 53.64) --
	(517.73, 53.62) --
	(518.11, 53.60) --
	(518.49, 53.58) --
	(518.87, 53.56) --
	(519.25, 53.54) --
	(519.63, 53.53) --
	(520.01, 53.51) --
	(520.39, 53.49) --
	(520.77, 53.47) --
	(521.15, 53.45) --
	(521.53, 53.43) --
	(521.90, 53.41) --
	(522.28, 53.39) --
	(522.66, 53.38) --
	(523.04, 53.36) --
	(523.42, 53.34) --
	(523.80, 53.32) --
	(524.18, 53.30) --
	(524.56, 53.28) --
	(524.94, 53.27) --
	(525.32, 53.25) --
	(525.70, 53.23) --
	(526.08, 53.21) --
	(526.46, 53.20) --
	(526.84, 53.18) --
	(527.22, 53.16) --
	(527.60, 53.14) --
	(527.98, 53.13) --
	(528.36, 53.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (528.36, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.25, 49.80) -- (386.25, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.62, 49.80) -- (433.62, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.99, 49.80) -- (480.99, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.36, 49.80) -- (528.36, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88, 31.87) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (386.25, 31.87) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (433.62, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (480.99, 31.87) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (528.36, 31.87) {20};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88,239.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (333.90, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 87.70) -- (333.90, 87.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,125.59) -- (333.90,125.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,163.49) -- (333.90,163.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,201.38) -- (333.90,201.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,239.28) -- (333.90,239.28);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 46.94) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 84.84) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,122.73) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,160.63) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,198.53) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,236.42) {1.0};
\end{scope}
\end{tikzpicture}

\end{document}