### 2020-08-28 benchmarking/benchmarkingVM/ESS_round1_real.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-28 13:28:47
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,289.08) rectangle (289.08,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,548.96) {\bfseries cons (\textit{real}) vs. nocons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,301.03) {ESS/$10^6$ states [nocons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,433.62) {ESS/$10^6$ states [cons (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (  0.00,403.16) --
	(175.00,578.16);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (302.44,578.16) --
	(302.44,477.44) --
	(  0.00,175.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (  0.00,365.13) --
	(213.03,578.16);

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (302.44,578.16) --
	(302.44,515.47) --
	(  0.00,213.03);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (  0.00,327.11) --
	(251.05,578.16);

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (302.44,578.16) --
	(302.44,553.49) --
	(  0.00,251.05);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (171.41,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,464.39) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (133.38,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,426.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.70,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92,388.34) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 78.41,410.95) --
	( 91.01,410.95);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 85.43,396.47) --
	( 85.43,420.38);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 85.43,407.72) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.48,467.57) --
	(182.33,467.57);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.56,466.01) --
	(179.56,469.05);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.56,463.54) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 81.27,472.87) --
	( 87.83,472.87);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 84.75,471.19) --
	( 84.75,474.46);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 84.75,468.84) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 77.34,424.98) --
	( 96.67,424.98);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 88.68,419.14) --
	( 88.68,429.79);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 88.68,420.94) {$h$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 77.52,382.73) --
	(108.13,382.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 96.89,359.01) --
	( 96.89,395.26);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 96.89,379.83) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.26,478.14) --
	(189.63,478.14);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.95,477.37) --
	(188.95,478.88);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.95,475.23) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (141.10,467.30) --
	(170.92,467.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (159.88,465.91) --
	(159.88,468.61);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (159.88,464.07) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (177.80,474.02) --
	(190.74,474.02);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.02,472.32) --
	(185.02,475.62);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.02,469.99) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (175.73,476.99) --
	(183.11,476.99);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.67,476.82) --
	(179.67,477.16);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.67,472.96) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (139.34,449.32) --
	(177.09,449.32);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (164.26,447.98) --
	(164.26,450.60);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (164.26,445.29) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (117.89,454.32) --
	(191.13,454.32);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (173.96,452.70) --
	(173.96,455.85);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (173.96,451.42) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.24,481.37) --
	(192.56,481.37);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.42,480.81) --
	(191.42,481.92);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.42,478.46) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (151.68,424.45) --
	(151.68,434.52);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (151.68,426.72) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.78,476.44) --
	(192.59,476.44);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (184.44,475.09) --
	(184.44,477.73);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.44,472.41) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.84,478.90) --
	(189.05,478.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.62,478.65) --
	(183.62,479.14);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.62,474.86) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (148.64,389.28) --
	(148.64,395.86);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (148.64,388.74) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (159.92,416.68) --
	(159.92,425.68);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (159.92,418.64) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.20,483.88) --
	(195.50,483.88);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.86,483.48) --
	(194.86,484.29);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.86,480.98) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (144.50,437.03) --
	(144.50,441.13);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (144.50,435.93) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (166.76,476.35) --
	(192.56,476.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.59,475.11) --
	(182.59,477.54);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.59,472.32) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (157.54,475.45) --
	(177.27,475.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (169.14,475.33) --
	(169.14,475.58);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.14,471.42) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (142.49,430.50) --
	(142.49,432.37);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (142.49,427.42) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.23,425.41) --
	(160.23,429.48);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.23,424.62) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.33,484.64) --
	(195.61,484.64);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.98,484.32) --
	(194.98,484.95);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.98,481.73) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (101.54,432.34) --
	(123.04,432.34);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (114.35,429.11) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (171.64,471.96) --
	(192.30,471.96);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.87,467.92) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (132.95,472.27) --
	(142.09,472.27);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (137.90,468.23) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 85.73,415.23) --
	(106.54,415.23);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 98.06,411.19) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 98.15,400.30) --
	(118.93,400.30);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (110.47,397.40) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.20,481.36) --
	(193.74,481.36);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.56,478.45) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 32.48,444.50) --
	(192.17,444.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (173.24,441.31) --
	(173.24,447.36);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (173.24,441.28) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (132.20,463.69) --
	(175.44,463.69);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (161.58,461.46) --
	(161.58,465.75);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (161.58,459.66) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.89,474.60) --
	(201.90,474.60);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (196.54,474.36) --
	(196.54,474.83);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (196.54,470.56) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.45,434.24) --
	(175.79,434.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (164.87,432.34) --
	(164.87,436.02);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (164.87,430.21) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (138.81,444.30) --
	(205.67,444.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.96,442.70) --
	(188.96,445.81);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.96,441.39) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.98,483.57) --
	(193.61,483.57);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.81,482.98) --
	(192.81,484.15);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.81,480.67) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (139.00,463.79) --
	(201.79,463.79);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.42,461.47) --
	(185.42,465.93);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.42,460.57) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.22,472.84) --
	(191.06,472.84);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.99,472.16) --
	(186.99,473.50);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.99,468.81) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (158.42,475.45) --
	(161.91,475.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.22,475.26) --
	(160.22,475.63);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.22,471.41) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (127.39,437.43) --
	(187.18,437.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (171.11,435.37) --
	(171.11,439.35);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (171.11,433.40) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (131.09,432.16) --
	(204.47,432.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.28,430.29) --
	(187.28,433.90);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.28,429.25) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.52,484.51) --
	(195.71,484.51);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.16,483.14) --
	(194.16,485.82);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.16,481.61) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 87.75,444.22) --
	(129.17,444.22);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (115.63,442.73) --
	(115.63,445.63);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (115.63,440.99) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (119.18,453.67) --
	(160.39,453.67);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.89,450.11) --
	(146.89,456.83);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (146.89,449.64) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.34,455.68) --
	(169.74,455.68);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.47,455.40) --
	(160.47,455.96);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.47,451.65) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 79.88,412.46) --
	(117.32,412.46);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (104.54,406.29) --
	(104.54,417.50);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (104.54,408.43) {$h$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 90.78,434.42) --
	(132.68,434.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (119.05,432.86) --
	(119.05,435.90);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (119.05,431.52) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.48,481.72) --
	(193.39,481.72);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.50,481.16) --
	(191.50,482.26);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.50,478.81) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.95,338.88) -- ( 93.95,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.96,338.88) -- (112.96,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.11,338.88) -- (157.11,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.12,338.88) -- (176.12,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.27,338.88) -- (220.27,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.28,338.88) -- (239.28,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.95,320.95) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (112.96,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (157.11,320.95) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (176.12,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (220.27,320.95) {500};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,528.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 44.82,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,383.03) -- ( 44.82,383.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,402.04) -- ( 44.82,402.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,446.19) -- ( 44.82,446.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,465.20) -- ( 44.82,465.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,509.35) -- ( 44.82,509.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,528.36) -- ( 44.82,528.36);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,380.17) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,399.18) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,443.33) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,462.34) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,506.49) {500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,525.50) {1000};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,431.01) rectangle (189.67,426.03);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,421.05) rectangle (189.67,416.07);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,411.09) rectangle (189.67,406.11);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,401.13) rectangle (189.67,396.15);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,391.17) rectangle (189.67,386.19);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,381.21) rectangle (189.67,376.23);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,371.25) rectangle (189.67,366.27);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,361.29) rectangle (189.67,356.31);
\definecolor{fillColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,351.33) rectangle (189.67,346.35);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,425.66) {Ran};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,415.70) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,405.74) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,395.78) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,385.82) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,375.86) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,365.90) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,355.94) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,345.98) {Moyle};
\end{scope}
\begin{scope}
\path[clip] (289.08,289.08) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,548.96) {\bfseries adapt (\textit{real}) vs. nocons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,301.03) {ESS/$10^6$ states [nocons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,433.62) {ESS/$10^6$ states [adapt (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] (338.88,338.88) rectangle (528.36,528.36);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] ( 86.24,200.32) --
	(464.08,578.16);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (578.16,464.08) --
	(114.08,  0.00);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] ( 86.24,  0.00) --
	( 86.24,200.32);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] ( 86.24,162.29) --
	(502.11,578.16);

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (578.16,502.11) --
	( 86.24, 10.19) --
	( 86.24,162.29);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] ( 86.24,124.27) --
	(540.13,578.16);

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (578.16,540.13) --
	( 86.24, 48.21) --
	( 86.24,124.27);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (460.49,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,464.39) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (422.46,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,426.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (382.78,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00,388.34) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (367.49,467.04) --
	(380.09,467.04);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (374.51,464.05) --
	(374.51,469.74);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (374.51,463.82) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (465.56,448.03) --
	(471.41,448.03);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (468.64,444.87) --
	(468.64,450.86);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (468.64,444.00) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (370.35,473.24) --
	(376.91,473.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (373.83,472.46) --
	(373.83,473.99);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (373.83,469.20) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (366.42,435.02) --
	(385.75,435.02);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (377.76,431.48) --
	(377.76,438.16);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (377.76,430.99) {$h$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (366.60,468.13) --
	(397.21,468.13);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (385.97,465.86) --
	(385.97,470.24);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (385.97,465.23) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (477.34,477.42) --
	(478.71,477.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.03,476.86) --
	(478.03,477.97);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (478.03,474.52) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (430.18,471.54) --
	(460.00,471.54);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (448.96,470.69) --
	(448.96,472.36);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (448.96,468.31) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (466.88,474.80) --
	(479.82,474.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (474.10,472.77) --
	(474.10,476.70);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.10,470.77) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (464.81,479.57) --
	(472.19,479.57);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (468.75,479.29) --
	(468.75,479.85);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (468.75,475.54) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (428.42,454.70) --
	(466.17,454.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (453.34,453.16) --
	(453.34,456.15);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (453.34,450.66) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (406.97,473.34) --
	(480.21,473.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (463.04,472.63) --
	(463.04,474.04);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (463.04,470.44) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (479.32,481.45) --
	(481.64,481.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (480.50,480.91) --
	(480.50,481.98);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (480.50,478.54) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (440.76,432.68) --
	(440.76,439.12);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (440.76,432.86) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (461.86,479.26) --
	(481.67,479.26);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (473.52,478.07) --
	(473.52,480.41);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (473.52,475.23) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (465.92,482.87) --
	(478.13,482.87);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (472.70,482.55) --
	(472.70,483.18);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (472.70,478.84) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (437.72,397.21) --
	(437.72,405.84);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (437.72,397.83) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (449.00,429.23) --
	(449.00,436.09);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.00,429.97) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (483.28,484.21) --
	(484.58,484.21);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (483.94,483.81) --
	(483.94,484.61);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.94,481.31) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (433.58,448.44) --
	(433.58,452.06);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (433.58,447.08) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (455.84,473.88) --
	(481.64,473.88);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (471.67,472.84) --
	(471.67,474.88);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (471.67,469.85) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (446.62,475.99) --
	(466.35,475.99);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (458.22,475.70) --
	(458.22,476.27);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (458.22,471.95) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (431.57,433.40) --
	(431.57,438.03);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (431.57,431.78) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (449.31,443.26) --
	(449.31,447.88);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.31,442.77) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (483.41,484.53) --
	(484.69,484.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (484.06,484.08) --
	(484.06,484.98);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (484.06,481.63) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (390.62,463.50) --
	(412.12,463.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (403.43,460.57) --
	(403.43,466.15);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (403.43,460.27) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (460.72,461.08) --
	(481.38,461.08);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (472.95,460.33) --
	(472.95,461.81);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (472.95,457.05) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (422.03,470.75) --
	(431.17,470.75);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (426.98,470.01) --
	(426.98,471.47);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (426.98,466.72) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (374.81,416.60) --
	(395.62,416.60);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (387.14,412.44) --
	(387.14,420.21);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (387.14,412.57) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (387.23,446.61) --
	(408.01,446.61);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (399.55,444.30) --
	(399.55,448.73);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (399.55,443.70) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.28,481.27) --
	(482.82,481.27);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (480.64,480.85) --
	(480.64,481.68);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (480.64,478.36) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (321.56,455.89) --
	(481.25,455.89);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (462.32,452.11) --
	(462.32,459.20);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (462.32,452.66) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (421.28,478.12) --
	(464.52,478.12);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (450.66,477.19) --
	(450.66,479.03);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (450.66,474.09) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.97,495.81) --
	(490.98,495.81);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (485.62,495.66) --
	(485.62,495.97);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (485.62,491.78) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (435.53,446.43) --
	(464.87,446.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (453.95,444.35) --
	(453.95,448.35);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (453.95,442.39) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (427.89,460.23) --
	(494.75,460.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.04,459.37) --
	(478.04,461.07);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (478.04,457.33) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.06,483.43) --
	(482.69,483.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.89,483.01) --
	(481.89,483.84);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (481.89,480.52) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (428.08,510.61) --
	(490.87,510.61);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (474.50,506.59) --
	(474.50,514.11);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.50,507.38) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (471.30,510.72) --
	(480.14,510.72);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (476.07,507.87) --
	(476.07,513.30);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.07,506.68) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (416.47,483.27) --
	(476.26,483.27);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (460.19,480.40) --
	(460.19,485.87);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (460.19,479.24) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (420.17,487.76) --
	(493.55,487.76);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (476.36,484.08) --
	(476.36,491.00);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.36,484.85) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.60,529.08) --
	(484.79,529.08);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (483.24,525.48) --
	(483.24,532.27);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.24,526.18) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.03,338.88) -- (383.03,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (402.04,338.88) -- (402.04,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (446.19,338.88) -- (446.19,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.20,338.88) -- (465.20,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (509.35,338.88) -- (509.35,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.36,338.88) -- (528.36,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (383.03,320.95) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (402.04,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (446.19,320.95) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (465.20,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (509.35,320.95) {500};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,528.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (333.90,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,383.03) -- (333.90,383.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,402.04) -- (333.90,402.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,446.19) -- (333.90,446.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,465.20) -- (333.90,465.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,509.35) -- (333.90,509.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,528.36) -- (333.90,528.36);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,380.17) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,399.18) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,443.33) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,462.34) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,506.49) {500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,525.50) {1000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,259.88) {\bfseries adapt (\textit{real}) vs. cons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54, 11.95) {ESS/$10^6$ states [cons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,144.54) {ESS/$10^6$ states [adapt (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80, 49.80) rectangle (239.28,239.28);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (  0.00,114.08) --
	(302.44,416.52) --
	(302.44,188.36) --
	(114.08,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (  0.00, 76.05) --
	(302.44,378.49) --
	(302.44,226.39) --
	( 76.05,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (  0.00, 38.03) --
	(302.44,340.47) --
	(302.44,264.41) --
	( 38.03,  0.00);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (171.41, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,175.31) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (133.38, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,137.28) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.70, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92, 99.26) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28,239.28);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (107.39,177.96) --
	(131.30,177.96);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (121.87,174.97) --
	(121.87,180.66);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (121.87,174.74) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.93,158.95) --
	(179.97,158.95);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (178.49,155.79) --
	(178.49,161.78);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.49,154.92) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.11,184.16) --
	(185.38,184.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.79,183.38) --
	(183.79,184.91);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.79,180.12) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (130.06,145.94) --
	(140.71,145.94);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (135.90,142.40) --
	(135.90,149.08);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (135.90,141.91) {$h$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 69.93,179.05) --
	(106.18,179.05);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 93.65,176.78) --
	( 93.65,181.16);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 93.65,176.15) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.29,188.34) --
	(189.80,188.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.06,187.78) --
	(189.06,188.89);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (189.06,185.44) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.83,182.46) --
	(179.53,182.46);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (178.22,181.61) --
	(178.22,183.28);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.22,179.23) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.24,185.72) --
	(186.54,185.72);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (184.94,183.69) --
	(184.94,187.62);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.94,181.69) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.74,190.49) --
	(188.08,190.49);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.91,190.21) --
	(187.91,190.77);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.91,186.46) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (158.90,165.62) --
	(161.52,165.62);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.24,164.08) --
	(160.24,167.07);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.24,161.58) {$h$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (163.62,184.26) --
	(166.77,184.26);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.24,183.55) --
	(165.24,184.96);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.24,181.36) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.73,192.37) --
	(192.84,192.37);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.29,191.83) --
	(192.29,192.90);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.29,189.46) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (135.37,147.01) --
	(145.44,147.01);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (140.87,143.60) --
	(140.87,150.04);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (140.87,143.78) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.01,190.18) --
	(188.65,190.18);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.36,188.99) --
	(187.36,191.33);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.36,186.15) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.57,193.79) --
	(190.06,193.79);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.82,193.47) --
	(189.82,194.10);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (189.82,189.76) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (100.20,112.78) --
	(106.78,112.78);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (103.69,108.13) --
	(103.69,116.76);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (103.69,108.75) {$h$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (127.60,143.80) --
	(136.60,143.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (132.47,140.15) --
	(132.47,147.01);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (132.47,140.89) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.40,195.13) --
	(195.21,195.13);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.80,194.73) --
	(194.80,195.53);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.80,192.23) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (147.95,161.23) --
	(152.05,161.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (150.07,159.36) --
	(150.07,162.98);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (150.07,158.00) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.03,184.80) --
	(188.46,184.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.27,183.76) --
	(187.27,185.80);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.27,180.77) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.25,186.91) --
	(186.50,186.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.37,186.62) --
	(186.37,187.19);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.37,182.87) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (141.42,146.73) --
	(143.29,146.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (142.37,144.32) --
	(142.37,148.95);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (142.37,142.70) {$h$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (136.33,156.59) --
	(140.40,156.59);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (138.44,154.18) --
	(138.44,158.80);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (138.44,153.69) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.24,195.45) --
	(195.87,195.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.56,195.00) --
	(195.56,195.90);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.56,192.55) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (143.26,171.49) --
	(143.26,177.07);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (143.26,171.19) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.88,171.25) --
	(182.88,172.73);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.88,167.97) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.19,180.93) --
	(183.19,182.39);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.19,177.64) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (126.15,123.36) --
	(126.15,131.13);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (126.15,123.49) {$h$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (111.22,155.22) --
	(111.22,159.65);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (111.22,154.62) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.28,191.77) --
	(192.28,192.60);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.28,189.28) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (152.23,166.81) --
	(158.28,166.81);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (155.42,163.03) --
	(155.42,170.12);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (155.42,163.58) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.38,189.04) --
	(176.67,189.04);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (174.61,188.11) --
	(174.61,189.95);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.61,185.01) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.28,206.73) --
	(185.75,206.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.52,206.58) --
	(185.52,206.89);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.52,202.70) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (143.26,157.35) --
	(146.94,157.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (145.16,155.27) --
	(145.16,159.27);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (145.16,153.31) {$h$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (153.62,171.15) --
	(156.73,171.15);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (155.22,170.29) --
	(155.22,171.99);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (155.22,168.25) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.90,194.35) --
	(195.07,194.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.49,193.93) --
	(194.49,194.76);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.49,191.44) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.39,221.53) --
	(176.85,221.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (174.71,217.51) --
	(174.71,225.03);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.71,218.30) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.08,221.64) --
	(184.42,221.64);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.76,218.79) --
	(183.76,224.22);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.76,217.60) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.29,194.19) --
	(150.27,194.19);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (148.35,191.32) --
	(148.35,196.79);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (148.35,190.16) {$h$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (141.21,198.68) --
	(144.82,198.68);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (143.08,195.00) --
	(143.08,201.92);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (143.08,195.77) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.06,240.00) --
	(196.74,240.00);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.43,236.40) --
	(195.43,243.19);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.43,237.10) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.95, 49.80) -- ( 93.95, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.96, 49.80) -- (112.96, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.11, 49.80) -- (157.11, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.12, 49.80) -- (176.12, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.27, 49.80) -- (220.27, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.28, 49.80) -- (239.28, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.95, 31.87) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (112.96, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (157.11, 31.87) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (176.12, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (220.27, 31.87) {500};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80,239.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 44.82, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 93.95) -- ( 44.82, 93.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,112.96) -- ( 44.82,112.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,157.11) -- ( 44.82,157.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,176.12) -- ( 44.82,176.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,220.27) -- ( 44.82,220.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,239.28) -- ( 44.82,239.28);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 91.09) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,110.10) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,154.25) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,173.26) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,217.41) {500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,236.42) {1000};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,259.88) {\bfseries \texttt{AdaptiveOperatorSampler} weights};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62, 11.95) {Sequence length $L$ (kb)};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,144.54) {Learned operator weight};
\end{scope}
\begin{scope}
\path[clip] (338.88, 49.80) rectangle (528.36,239.28);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (485.73,235.47) circle (  1.87);

\path[fill=fillColor] (485.73, 54.16) --
	(488.24, 49.81) --
	(483.21, 49.81) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (356.88,164.91) circle (  1.87);

\path[fill=fillColor] (356.88, 79.47) --
	(359.40, 75.11) --
	(354.37, 75.11) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (394.78,165.18) circle (  1.87);

\path[fill=fillColor] (394.78, 79.59) --
	(397.29, 75.23) --
	(392.26, 75.23) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (390.04,192.04) circle (  1.87);

\path[fill=fillColor] (390.04, 71.74) --
	(392.55, 67.38) --
	(387.52, 67.38) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (404.25,216.71) circle (  1.87);

\path[fill=fillColor] (404.25, 59.65) --
	(406.77, 55.29) --
	(401.74, 55.29) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (359.72,107.30) circle (  1.87);

\path[fill=fillColor] (359.72, 91.05) --
	(362.24, 86.70) --
	(357.21, 86.70) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (433.62,204.25) circle (  1.87);

\path[fill=fillColor] (433.62, 68.39) --
	(436.14, 64.04) --
	(431.10, 64.04) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (338.88,114.47) --
	(339.26,114.92) --
	(339.64,115.37) --
	(340.02,115.83) --
	(340.40,116.28) --
	(340.78,116.74) --
	(341.16,117.20) --
	(341.54,117.66) --
	(341.92,118.12) --
	(342.30,118.58) --
	(342.68,119.05) --
	(343.06,119.51) --
	(343.44,119.98) --
	(343.82,120.45) --
	(344.20,120.92) --
	(344.58,121.39) --
	(344.96,121.86) --
	(345.34,122.34) --
	(345.71,122.81) --
	(346.09,123.29) --
	(346.47,123.76) --
	(346.85,124.24) --
	(347.23,124.72) --
	(347.61,125.20) --
	(347.99,125.68) --
	(348.37,126.16) --
	(348.75,126.65) --
	(349.13,127.13) --
	(349.51,127.62) --
	(349.89,128.10) --
	(350.27,128.59) --
	(350.65,129.08) --
	(351.03,129.56) --
	(351.41,130.05) --
	(351.79,130.54) --
	(352.17,131.03) --
	(352.55,131.53) --
	(352.93,132.02) --
	(353.31,132.51) --
	(353.69,133.01) --
	(354.07,133.50) --
	(354.45,133.99) --
	(354.83,134.49) --
	(355.21,134.99) --
	(355.59,135.48) --
	(355.97,135.98) --
	(356.35,136.48) --
	(356.73,136.98) --
	(357.11,137.47) --
	(357.49,137.97) --
	(357.87,138.47) --
	(358.25,138.97) --
	(358.63,139.47) --
	(359.01,139.97) --
	(359.38,140.47) --
	(359.76,140.97) --
	(360.14,141.47) --
	(360.52,141.98) --
	(360.90,142.48) --
	(361.28,142.98) --
	(361.66,143.48) --
	(362.04,143.98) --
	(362.42,144.48) --
	(362.80,144.98) --
	(363.18,145.49) --
	(363.56,145.99) --
	(363.94,146.49) --
	(364.32,146.99) --
	(364.70,147.49) --
	(365.08,147.99) --
	(365.46,148.49) --
	(365.84,148.99) --
	(366.22,149.49) --
	(366.60,149.99) --
	(366.98,150.49) --
	(367.36,150.99) --
	(367.74,151.49) --
	(368.12,151.99) --
	(368.50,152.49) --
	(368.88,152.99) --
	(369.26,153.48) --
	(369.64,153.98) --
	(370.02,154.48) --
	(370.40,154.97) --
	(370.78,155.47) --
	(371.16,155.96) --
	(371.54,156.45) --
	(371.92,156.95) --
	(372.30,157.44) --
	(372.68,157.93) --
	(373.05,158.42) --
	(373.43,158.91) --
	(373.81,159.40) --
	(374.19,159.89) --
	(374.57,160.38) --
	(374.95,160.87) --
	(375.33,161.35) --
	(375.71,161.84) --
	(376.09,162.32) --
	(376.47,162.81) --
	(376.85,163.29) --
	(377.23,163.77) --
	(377.61,164.25) --
	(377.99,164.73) --
	(378.37,165.21) --
	(378.75,165.68) --
	(379.13,166.16) --
	(379.51,166.64) --
	(379.89,167.11) --
	(380.27,167.58) --
	(380.65,168.05) --
	(381.03,168.52) --
	(381.41,168.99) --
	(381.79,169.46) --
	(382.17,169.92) --
	(382.55,170.39) --
	(382.93,170.85) --
	(383.31,171.32) --
	(383.69,171.78) --
	(384.07,172.24) --
	(384.45,172.69) --
	(384.83,173.15) --
	(385.21,173.60) --
	(385.59,174.06) --
	(385.97,174.51) --
	(386.34,174.96) --
	(386.72,175.41) --
	(387.10,175.86) --
	(387.48,176.30) --
	(387.86,176.75) --
	(388.24,177.19) --
	(388.62,177.63) --
	(389.00,178.07) --
	(389.38,178.51) --
	(389.76,178.94) --
	(390.14,179.38) --
	(390.52,179.81) --
	(390.90,180.24) --
	(391.28,180.67) --
	(391.66,181.10) --
	(392.04,181.53) --
	(392.42,181.95) --
	(392.80,182.37) --
	(393.18,182.79) --
	(393.56,183.21) --
	(393.94,183.63) --
	(394.32,184.04) --
	(394.70,184.46) --
	(395.08,184.87) --
	(395.46,185.28) --
	(395.84,185.69) --
	(396.22,186.09) --
	(396.60,186.50) --
	(396.98,186.90) --
	(397.36,187.30) --
	(397.74,187.70) --
	(398.12,188.09) --
	(398.50,188.49) --
	(398.88,188.88) --
	(399.26,189.27) --
	(399.64,189.66) --
	(400.01,190.05) --
	(400.39,190.43) --
	(400.77,190.82) --
	(401.15,191.20) --
	(401.53,191.57) --
	(401.91,191.95) --
	(402.29,192.33) --
	(402.67,192.70) --
	(403.05,193.07) --
	(403.43,193.44) --
	(403.81,193.81) --
	(404.19,194.17) --
	(404.57,194.53) --
	(404.95,194.89) --
	(405.33,195.25) --
	(405.71,195.61) --
	(406.09,195.97) --
	(406.47,196.32) --
	(406.85,196.67) --
	(407.23,197.02) --
	(407.61,197.36) --
	(407.99,197.71) --
	(408.37,198.05) --
	(408.75,198.39) --
	(409.13,198.73) --
	(409.51,199.07) --
	(409.89,199.40) --
	(410.27,199.73) --
	(410.65,200.06) --
	(411.03,200.39) --
	(411.41,200.72) --
	(411.79,201.04) --
	(412.17,201.36) --
	(412.55,201.68) --
	(412.93,202.00) --
	(413.31,202.32) --
	(413.68,202.63) --
	(414.06,202.94) --
	(414.44,203.25) --
	(414.82,203.56) --
	(415.20,203.87) --
	(415.58,204.17) --
	(415.96,204.47) --
	(416.34,204.77) --
	(416.72,205.07) --
	(417.10,205.37) --
	(417.48,205.66) --
	(417.86,205.95) --
	(418.24,206.24) --
	(418.62,206.53) --
	(419.00,206.82) --
	(419.38,207.10) --
	(419.76,207.38) --
	(420.14,207.66) --
	(420.52,207.94) --
	(420.90,208.21) --
	(421.28,208.49) --
	(421.66,208.76) --
	(422.04,209.03) --
	(422.42,209.30) --
	(422.80,209.57) --
	(423.18,209.83) --
	(423.56,210.09) --
	(423.94,210.35) --
	(424.32,210.61) --
	(424.70,210.87) --
	(425.08,211.12) --
	(425.46,211.37) --
	(425.84,211.63) --
	(426.22,211.87) --
	(426.60,212.12) --
	(426.97,212.37) --
	(427.35,212.61) --
	(427.73,212.85) --
	(428.11,213.09) --
	(428.49,213.33) --
	(428.87,213.57) --
	(429.25,213.80) --
	(429.63,214.03) --
	(430.01,214.26) --
	(430.39,214.49) --
	(430.77,214.72) --
	(431.15,214.95) --
	(431.53,215.17) --
	(431.91,215.39) --
	(432.29,215.61) --
	(432.67,215.83) --
	(433.05,216.05) --
	(433.43,216.26) --
	(433.81,216.47) --
	(434.19,216.69) --
	(434.57,216.90) --
	(434.95,217.10) --
	(435.33,217.31) --
	(435.71,217.52) --
	(436.09,217.72) --
	(436.47,217.92) --
	(436.85,218.12) --
	(437.23,218.32) --
	(437.61,218.51) --
	(437.99,218.71) --
	(438.37,218.90) --
	(438.75,219.09) --
	(439.13,219.28) --
	(439.51,219.47) --
	(439.89,219.66) --
	(440.27,219.85) --
	(440.64,220.03) --
	(441.02,220.21) --
	(441.40,220.39) --
	(441.78,220.57) --
	(442.16,220.75) --
	(442.54,220.93) --
	(442.92,221.10) --
	(443.30,221.27) --
	(443.68,221.45) --
	(444.06,221.62) --
	(444.44,221.78) --
	(444.82,221.95) --
	(445.20,222.12) --
	(445.58,222.28) --
	(445.96,222.45) --
	(446.34,222.61) --
	(446.72,222.77) --
	(447.10,222.93) --
	(447.48,223.08) --
	(447.86,223.24) --
	(448.24,223.40) --
	(448.62,223.55) --
	(449.00,223.70) --
	(449.38,223.85) --
	(449.76,224.00) --
	(450.14,224.15) --
	(450.52,224.30) --
	(450.90,224.44) --
	(451.28,224.58) --
	(451.66,224.73) --
	(452.04,224.87) --
	(452.42,225.01) --
	(452.80,225.15) --
	(453.18,225.29) --
	(453.56,225.42) --
	(453.93,225.56) --
	(454.31,225.69) --
	(454.69,225.83) --
	(455.07,225.96) --
	(455.45,226.09) --
	(455.83,226.22) --
	(456.21,226.35) --
	(456.59,226.47) --
	(456.97,226.60) --
	(457.35,226.72) --
	(457.73,226.85) --
	(458.11,226.97) --
	(458.49,227.09) --
	(458.87,227.21) --
	(459.25,227.33) --
	(459.63,227.45) --
	(460.01,227.56) --
	(460.39,227.68) --
	(460.77,227.79) --
	(461.15,227.91) --
	(461.53,228.02) --
	(461.91,228.13) --
	(462.29,228.24) --
	(462.67,228.35) --
	(463.05,228.46) --
	(463.43,228.57) --
	(463.81,228.68) --
	(464.19,228.78) --
	(464.57,228.89) --
	(464.95,228.99) --
	(465.33,229.09) --
	(465.71,229.19) --
	(466.09,229.29) --
	(466.47,229.39) --
	(466.85,229.49) --
	(467.23,229.59) --
	(467.60,229.69) --
	(467.98,229.78) --
	(468.36,229.88) --
	(468.74,229.97) --
	(469.12,230.07) --
	(469.50,230.16) --
	(469.88,230.25) --
	(470.26,230.34) --
	(470.64,230.43) --
	(471.02,230.52) --
	(471.40,230.61) --
	(471.78,230.69) --
	(472.16,230.78) --
	(472.54,230.87) --
	(472.92,230.95) --
	(473.30,231.03) --
	(473.68,231.12) --
	(474.06,231.20) --
	(474.44,231.28) --
	(474.82,231.36) --
	(475.20,231.44) --
	(475.58,231.52) --
	(475.96,231.60) --
	(476.34,231.68) --
	(476.72,231.75) --
	(477.10,231.83) --
	(477.48,231.91) --
	(477.86,231.98) --
	(478.24,232.05) --
	(478.62,232.13) --
	(479.00,232.20) --
	(479.38,232.27) --
	(479.76,232.34) --
	(480.14,232.41) --
	(480.52,232.48) --
	(480.90,232.55) --
	(481.27,232.62) --
	(481.65,232.69) --
	(482.03,232.75) --
	(482.41,232.82) --
	(482.79,232.89) --
	(483.17,232.95) --
	(483.55,233.02) --
	(483.93,233.08) --
	(484.31,233.14) --
	(484.69,233.21) --
	(485.07,233.27) --
	(485.45,233.33) --
	(485.83,233.39) --
	(486.21,233.45) --
	(486.59,233.51) --
	(486.97,233.57) --
	(487.35,233.63) --
	(487.73,233.68) --
	(488.11,233.74) --
	(488.49,233.80) --
	(488.87,233.85) --
	(489.25,233.91) --
	(489.63,233.96) --
	(490.01,234.02) --
	(490.39,234.07) --
	(490.77,234.13) --
	(491.15,234.18) --
	(491.53,234.23) --
	(491.91,234.28) --
	(492.29,234.33) --
	(492.67,234.39) --
	(493.05,234.44) --
	(493.43,234.49) --
	(493.81,234.53) --
	(494.19,234.58) --
	(494.56,234.63) --
	(494.94,234.68) --
	(495.32,234.73) --
	(495.70,234.77) --
	(496.08,234.82) --
	(496.46,234.87) --
	(496.84,234.91) --
	(497.22,234.96) --
	(497.60,235.00) --
	(497.98,235.04) --
	(498.36,235.09) --
	(498.74,235.13) --
	(499.12,235.17) --
	(499.50,235.22) --
	(499.88,235.26) --
	(500.26,235.30) --
	(500.64,235.34) --
	(501.02,235.38) --
	(501.40,235.42) --
	(501.78,235.46) --
	(502.16,235.50) --
	(502.54,235.54) --
	(502.92,235.58) --
	(503.30,235.62) --
	(503.68,235.65) --
	(504.06,235.69) --
	(504.44,235.73) --
	(504.82,235.77) --
	(505.20,235.80) --
	(505.58,235.84) --
	(505.96,235.87) --
	(506.34,235.91) --
	(506.72,235.94) --
	(507.10,235.98) --
	(507.48,236.01) --
	(507.86,236.05) --
	(508.23,236.08) --
	(508.61,236.11) --
	(508.99,236.15) --
	(509.37,236.18) --
	(509.75,236.21) --
	(510.13,236.24) --
	(510.51,236.27) --
	(510.89,236.30) --
	(511.27,236.34) --
	(511.65,236.37) --
	(512.03,236.40) --
	(512.41,236.43) --
	(512.79,236.46) --
	(513.17,236.49) --
	(513.55,236.51) --
	(513.93,236.54) --
	(514.31,236.57) --
	(514.69,236.60) --
	(515.07,236.63) --
	(515.45,236.65) --
	(515.83,236.68) --
	(516.21,236.71) --
	(516.59,236.74) --
	(516.97,236.76) --
	(517.35,236.79) --
	(517.73,236.81) --
	(518.11,236.84) --
	(518.49,236.87) --
	(518.87,236.89) --
	(519.25,236.92) --
	(519.63,236.94) --
	(520.01,236.96) --
	(520.39,236.99) --
	(520.77,237.01) --
	(521.15,237.04) --
	(521.53,237.06) --
	(521.90,237.08) --
	(522.28,237.11) --
	(522.66,237.13) --
	(523.04,237.15) --
	(523.42,237.17) --
	(523.80,237.19) --
	(524.18,237.22) --
	(524.56,237.24) --
	(524.94,237.26) --
	(525.32,237.28) --
	(525.70,237.30) --
	(526.08,237.32) --
	(526.46,237.34) --
	(526.84,237.36) --
	(527.22,237.38) --
	(527.60,237.40) --
	(527.98,237.42) --
	(528.36,237.44);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (338.88, 93.87) --
	(339.26, 93.64) --
	(339.64, 93.40) --
	(340.02, 93.17) --
	(340.40, 92.94) --
	(340.78, 92.71) --
	(341.16, 92.48) --
	(341.54, 92.26) --
	(341.92, 92.03) --
	(342.30, 91.80) --
	(342.68, 91.58) --
	(343.06, 91.35) --
	(343.44, 91.13) --
	(343.82, 90.91) --
	(344.20, 90.69) --
	(344.58, 90.46) --
	(344.96, 90.24) --
	(345.34, 90.02) --
	(345.71, 89.81) --
	(346.09, 89.59) --
	(346.47, 89.37) --
	(346.85, 89.16) --
	(347.23, 88.94) --
	(347.61, 88.73) --
	(347.99, 88.51) --
	(348.37, 88.30) --
	(348.75, 88.09) --
	(349.13, 87.88) --
	(349.51, 87.67) --
	(349.89, 87.46) --
	(350.27, 87.25) --
	(350.65, 87.05) --
	(351.03, 86.84) --
	(351.41, 86.64) --
	(351.79, 86.43) --
	(352.17, 86.23) --
	(352.55, 86.02) --
	(352.93, 85.82) --
	(353.31, 85.62) --
	(353.69, 85.42) --
	(354.07, 85.22) --
	(354.45, 85.02) --
	(354.83, 84.83) --
	(355.21, 84.63) --
	(355.59, 84.43) --
	(355.97, 84.24) --
	(356.35, 84.04) --
	(356.73, 83.85) --
	(357.11, 83.66) --
	(357.49, 83.47) --
	(357.87, 83.28) --
	(358.25, 83.09) --
	(358.63, 82.90) --
	(359.01, 82.71) --
	(359.38, 82.52) --
	(359.76, 82.33) --
	(360.14, 82.15) --
	(360.52, 81.96) --
	(360.90, 81.78) --
	(361.28, 81.60) --
	(361.66, 81.41) --
	(362.04, 81.23) --
	(362.42, 81.05) --
	(362.80, 80.87) --
	(363.18, 80.69) --
	(363.56, 80.52) --
	(363.94, 80.34) --
	(364.32, 80.16) --
	(364.70, 79.99) --
	(365.08, 79.81) --
	(365.46, 79.64) --
	(365.84, 79.46) --
	(366.22, 79.29) --
	(366.60, 79.12) --
	(366.98, 78.95) --
	(367.36, 78.78) --
	(367.74, 78.61) --
	(368.12, 78.44) --
	(368.50, 78.27) --
	(368.88, 78.11) --
	(369.26, 77.94) --
	(369.64, 77.78) --
	(370.02, 77.61) --
	(370.40, 77.45) --
	(370.78, 77.29) --
	(371.16, 77.13) --
	(371.54, 76.96) --
	(371.92, 76.80) --
	(372.30, 76.64) --
	(372.68, 76.49) --
	(373.05, 76.33) --
	(373.43, 76.17) --
	(373.81, 76.01) --
	(374.19, 75.86) --
	(374.57, 75.70) --
	(374.95, 75.55) --
	(375.33, 75.40) --
	(375.71, 75.24) --
	(376.09, 75.09) --
	(376.47, 74.94) --
	(376.85, 74.79) --
	(377.23, 74.64) --
	(377.61, 74.49) --
	(377.99, 74.34) --
	(378.37, 74.20) --
	(378.75, 74.05) --
	(379.13, 73.91) --
	(379.51, 73.76) --
	(379.89, 73.62) --
	(380.27, 73.47) --
	(380.65, 73.33) --
	(381.03, 73.19) --
	(381.41, 73.05) --
	(381.79, 72.91) --
	(382.17, 72.77) --
	(382.55, 72.63) --
	(382.93, 72.49) --
	(383.31, 72.35) --
	(383.69, 72.22) --
	(384.07, 72.08) --
	(384.45, 71.94) --
	(384.83, 71.81) --
	(385.21, 71.68) --
	(385.59, 71.54) --
	(385.97, 71.41) --
	(386.34, 71.28) --
	(386.72, 71.15) --
	(387.10, 71.02) --
	(387.48, 70.89) --
	(387.86, 70.76) --
	(388.24, 70.63) --
	(388.62, 70.50) --
	(389.00, 70.37) --
	(389.38, 70.25) --
	(389.76, 70.12) --
	(390.14, 70.00) --
	(390.52, 69.87) --
	(390.90, 69.75) --
	(391.28, 69.63) --
	(391.66, 69.50) --
	(392.04, 69.38) --
	(392.42, 69.26) --
	(392.80, 69.14) --
	(393.18, 69.02) --
	(393.56, 68.90) --
	(393.94, 68.78) --
	(394.32, 68.67) --
	(394.70, 68.55) --
	(395.08, 68.43) --
	(395.46, 68.32) --
	(395.84, 68.20) --
	(396.22, 68.09) --
	(396.60, 67.97) --
	(396.98, 67.86) --
	(397.36, 67.75) --
	(397.74, 67.64) --
	(398.12, 67.53) --
	(398.50, 67.42) --
	(398.88, 67.31) --
	(399.26, 67.20) --
	(399.64, 67.09) --
	(400.01, 66.98) --
	(400.39, 66.87) --
	(400.77, 66.76) --
	(401.15, 66.66) --
	(401.53, 66.55) --
	(401.91, 66.45) --
	(402.29, 66.34) --
	(402.67, 66.24) --
	(403.05, 66.14) --
	(403.43, 66.03) --
	(403.81, 65.93) --
	(404.19, 65.83) --
	(404.57, 65.73) --
	(404.95, 65.63) --
	(405.33, 65.53) --
	(405.71, 65.43) --
	(406.09, 65.33) --
	(406.47, 65.23) --
	(406.85, 65.13) --
	(407.23, 65.04) --
	(407.61, 64.94) --
	(407.99, 64.84) --
	(408.37, 64.75) --
	(408.75, 64.65) --
	(409.13, 64.56) --
	(409.51, 64.47) --
	(409.89, 64.37) --
	(410.27, 64.28) --
	(410.65, 64.19) --
	(411.03, 64.10) --
	(411.41, 64.01) --
	(411.79, 63.91) --
	(412.17, 63.82) --
	(412.55, 63.74) --
	(412.93, 63.65) --
	(413.31, 63.56) --
	(413.68, 63.47) --
	(414.06, 63.38) --
	(414.44, 63.30) --
	(414.82, 63.21) --
	(415.20, 63.12) --
	(415.58, 63.04) --
	(415.96, 62.95) --
	(416.34, 62.87) --
	(416.72, 62.79) --
	(417.10, 62.70) --
	(417.48, 62.62) --
	(417.86, 62.54) --
	(418.24, 62.46) --
	(418.62, 62.37) --
	(419.00, 62.29) --
	(419.38, 62.21) --
	(419.76, 62.13) --
	(420.14, 62.05) --
	(420.52, 61.98) --
	(420.90, 61.90) --
	(421.28, 61.82) --
	(421.66, 61.74) --
	(422.04, 61.66) --
	(422.42, 61.59) --
	(422.80, 61.51) --
	(423.18, 61.44) --
	(423.56, 61.36) --
	(423.94, 61.29) --
	(424.32, 61.21) --
	(424.70, 61.14) --
	(425.08, 61.06) --
	(425.46, 60.99) --
	(425.84, 60.92) --
	(426.22, 60.85) --
	(426.60, 60.78) --
	(426.97, 60.70) --
	(427.35, 60.63) --
	(427.73, 60.56) --
	(428.11, 60.49) --
	(428.49, 60.42) --
	(428.87, 60.35) --
	(429.25, 60.29) --
	(429.63, 60.22) --
	(430.01, 60.15) --
	(430.39, 60.08) --
	(430.77, 60.02) --
	(431.15, 59.95) --
	(431.53, 59.88) --
	(431.91, 59.82) --
	(432.29, 59.75) --
	(432.67, 59.69) --
	(433.05, 59.62) --
	(433.43, 59.56) --
	(433.81, 59.49) --
	(434.19, 59.43) --
	(434.57, 59.37) --
	(434.95, 59.31) --
	(435.33, 59.24) --
	(435.71, 59.18) --
	(436.09, 59.12) --
	(436.47, 59.06) --
	(436.85, 59.00) --
	(437.23, 58.94) --
	(437.61, 58.88) --
	(437.99, 58.82) --
	(438.37, 58.76) --
	(438.75, 58.70) --
	(439.13, 58.64) --
	(439.51, 58.59) --
	(439.89, 58.53) --
	(440.27, 58.47) --
	(440.64, 58.41) --
	(441.02, 58.36) --
	(441.40, 58.30) --
	(441.78, 58.24) --
	(442.16, 58.19) --
	(442.54, 58.13) --
	(442.92, 58.08) --
	(443.30, 58.02) --
	(443.68, 57.97) --
	(444.06, 57.92) --
	(444.44, 57.86) --
	(444.82, 57.81) --
	(445.20, 57.76) --
	(445.58, 57.70) --
	(445.96, 57.65) --
	(446.34, 57.60) --
	(446.72, 57.55) --
	(447.10, 57.50) --
	(447.48, 57.45) --
	(447.86, 57.40) --
	(448.24, 57.35) --
	(448.62, 57.30) --
	(449.00, 57.25) --
	(449.38, 57.20) --
	(449.76, 57.15) --
	(450.14, 57.10) --
	(450.52, 57.05) --
	(450.90, 57.00) --
	(451.28, 56.96) --
	(451.66, 56.91) --
	(452.04, 56.86) --
	(452.42, 56.81) --
	(452.80, 56.77) --
	(453.18, 56.72) --
	(453.56, 56.68) --
	(453.93, 56.63) --
	(454.31, 56.59) --
	(454.69, 56.54) --
	(455.07, 56.50) --
	(455.45, 56.45) --
	(455.83, 56.41) --
	(456.21, 56.36) --
	(456.59, 56.32) --
	(456.97, 56.28) --
	(457.35, 56.23) --
	(457.73, 56.19) --
	(458.11, 56.15) --
	(458.49, 56.11) --
	(458.87, 56.06) --
	(459.25, 56.02) --
	(459.63, 55.98) --
	(460.01, 55.94) --
	(460.39, 55.90) --
	(460.77, 55.86) --
	(461.15, 55.82) --
	(461.53, 55.78) --
	(461.91, 55.74) --
	(462.29, 55.70) --
	(462.67, 55.66) --
	(463.05, 55.62) --
	(463.43, 55.58) --
	(463.81, 55.54) --
	(464.19, 55.50) --
	(464.57, 55.47) --
	(464.95, 55.43) --
	(465.33, 55.39) --
	(465.71, 55.35) --
	(466.09, 55.32) --
	(466.47, 55.28) --
	(466.85, 55.24) --
	(467.23, 55.21) --
	(467.60, 55.17) --
	(467.98, 55.13) --
	(468.36, 55.10) --
	(468.74, 55.06) --
	(469.12, 55.03) --
	(469.50, 54.99) --
	(469.88, 54.96) --
	(470.26, 54.92) --
	(470.64, 54.89) --
	(471.02, 54.85) --
	(471.40, 54.82) --
	(471.78, 54.79) --
	(472.16, 54.75) --
	(472.54, 54.72) --
	(472.92, 54.69) --
	(473.30, 54.65) --
	(473.68, 54.62) --
	(474.06, 54.59) --
	(474.44, 54.56) --
	(474.82, 54.53) --
	(475.20, 54.49) --
	(475.58, 54.46) --
	(475.96, 54.43) --
	(476.34, 54.40) --
	(476.72, 54.37) --
	(477.10, 54.34) --
	(477.48, 54.31) --
	(477.86, 54.28) --
	(478.24, 54.25) --
	(478.62, 54.22) --
	(479.00, 54.19) --
	(479.38, 54.16) --
	(479.76, 54.13) --
	(480.14, 54.10) --
	(480.52, 54.07) --
	(480.90, 54.04) --
	(481.27, 54.01) --
	(481.65, 53.99) --
	(482.03, 53.96) --
	(482.41, 53.93) --
	(482.79, 53.90) --
	(483.17, 53.87) --
	(483.55, 53.85) --
	(483.93, 53.82) --
	(484.31, 53.79) --
	(484.69, 53.77) --
	(485.07, 53.74) --
	(485.45, 53.71) --
	(485.83, 53.69) --
	(486.21, 53.66) --
	(486.59, 53.63) --
	(486.97, 53.61) --
	(487.35, 53.58) --
	(487.73, 53.56) --
	(488.11, 53.53) --
	(488.49, 53.51) --
	(488.87, 53.48) --
	(489.25, 53.46) --
	(489.63, 53.43) --
	(490.01, 53.41) --
	(490.39, 53.38) --
	(490.77, 53.36) --
	(491.15, 53.33) --
	(491.53, 53.31) --
	(491.91, 53.29) --
	(492.29, 53.26) --
	(492.67, 53.24) --
	(493.05, 53.22) --
	(493.43, 53.19) --
	(493.81, 53.17) --
	(494.19, 53.15) --
	(494.56, 53.13) --
	(494.94, 53.10) --
	(495.32, 53.08) --
	(495.70, 53.06) --
	(496.08, 53.04) --
	(496.46, 53.01) --
	(496.84, 52.99) --
	(497.22, 52.97) --
	(497.60, 52.95) --
	(497.98, 52.93) --
	(498.36, 52.91) --
	(498.74, 52.89) --
	(499.12, 52.87) --
	(499.50, 52.84) --
	(499.88, 52.82) --
	(500.26, 52.80) --
	(500.64, 52.78) --
	(501.02, 52.76) --
	(501.40, 52.74) --
	(501.78, 52.72) --
	(502.16, 52.70) --
	(502.54, 52.68) --
	(502.92, 52.66) --
	(503.30, 52.64) --
	(503.68, 52.62) --
	(504.06, 52.61) --
	(504.44, 52.59) --
	(504.82, 52.57) --
	(505.20, 52.55) --
	(505.58, 52.53) --
	(505.96, 52.51) --
	(506.34, 52.49) --
	(506.72, 52.48) --
	(507.10, 52.46) --
	(507.48, 52.44) --
	(507.86, 52.42) --
	(508.23, 52.40) --
	(508.61, 52.39) --
	(508.99, 52.37) --
	(509.37, 52.35) --
	(509.75, 52.33) --
	(510.13, 52.32) --
	(510.51, 52.30) --
	(510.89, 52.28) --
	(511.27, 52.27) --
	(511.65, 52.25) --
	(512.03, 52.23) --
	(512.41, 52.22) --
	(512.79, 52.20) --
	(513.17, 52.18) --
	(513.55, 52.17) --
	(513.93, 52.15) --
	(514.31, 52.13) --
	(514.69, 52.12) --
	(515.07, 52.10) --
	(515.45, 52.09) --
	(515.83, 52.07) --
	(516.21, 52.06) --
	(516.59, 52.04) --
	(516.97, 52.03) --
	(517.35, 52.01) --
	(517.73, 52.00) --
	(518.11, 51.98) --
	(518.49, 51.97) --
	(518.87, 51.95) --
	(519.25, 51.94) --
	(519.63, 51.92) --
	(520.01, 51.91) --
	(520.39, 51.89) --
	(520.77, 51.88) --
	(521.15, 51.86) --
	(521.53, 51.85) --
	(521.90, 51.84) --
	(522.28, 51.82) --
	(522.66, 51.81) --
	(523.04, 51.80) --
	(523.42, 51.78) --
	(523.80, 51.77) --
	(524.18, 51.75) --
	(524.56, 51.74) --
	(524.94, 51.73) --
	(525.32, 51.72) --
	(525.70, 51.70) --
	(526.08, 51.69) --
	(526.46, 51.68) --
	(526.84, 51.66) --
	(527.22, 51.65) --
	(527.60, 51.64) --
	(527.98, 51.63) --
	(528.36, 51.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (528.36, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.25, 49.80) -- (386.25, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.62, 49.80) -- (433.62, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.99, 49.80) -- (480.99, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.36, 49.80) -- (528.36, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88, 31.87) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (386.25, 31.87) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (433.62, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (480.99, 31.87) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (528.36, 31.87) {20};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88,239.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (333.90, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 87.70) -- (333.90, 87.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,125.59) -- (333.90,125.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,163.49) -- (333.90,163.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,201.38) -- (333.90,201.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,239.28) -- (333.90,239.28);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 46.94) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 84.84) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,122.73) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,160.63) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,198.53) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,236.42) {1.0};
\end{scope}
\end{tikzpicture}

\end{document}