### 2020-08-29 benchmarking/benchmarkingVM/ESS_round1_real.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-29 17:33:51
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,289.08) rectangle (289.08,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,548.96) {\bfseries cons (\textit{real}) vs. nocons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,301.03) {ESS/$10^6$ states [nocons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,433.62) {ESS/$10^6$ states [cons (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (  0.00,403.16) --
	(175.00,578.16);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (302.44,578.16) --
	(302.44,477.44) --
	(  0.00,175.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (  0.00,365.13) --
	(213.03,578.16);

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (302.44,578.16) --
	(302.44,515.47) --
	(  0.00,213.03);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (  0.00,327.11) --
	(251.05,578.16);

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (302.44,578.16) --
	(302.44,553.49) --
	(  0.00,251.05);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (171.41,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,464.39) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (133.38,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,426.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.70,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92,388.34) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 83.02,404.70) --
	( 97.04,404.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 90.92,398.69) --
	( 90.92,409.62);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 90.92,401.47) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.01,467.31) --
	(183.45,467.31);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.98,465.30) --
	(179.98,469.19);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.98,463.28) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 80.73,472.59) --
	( 84.67,472.59);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 82.77,471.35) --
	( 82.77,473.77);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 82.77,468.56) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 69.02,424.21) --
	( 77.50,424.21);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 73.59,418.49) --
	( 73.59,428.93);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 73.59,420.17) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 91.37,374.35) --
	(105.23,374.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 99.16,369.99) --
	( 99.16,378.12);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 99.16,371.45) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.87,480.03) --
	(190.98,480.03);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.43,479.66) --
	(190.43,480.40);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (190.43,477.13) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (141.10,467.30) --
	(170.92,467.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (159.88,465.91) --
	(159.88,468.61);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (159.88,464.07) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (177.80,474.02) --
	(190.74,474.02);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.02,472.32) --
	(185.02,475.62);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.02,469.99) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (175.73,476.99) --
	(183.11,476.99);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.67,476.82) --
	(179.67,477.16);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.67,472.96) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (139.57,454.32) --
	(189.04,454.32);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (174.21,453.15) --
	(174.21,455.44);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.21,450.29) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (117.89,454.32) --
	(191.13,454.32);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (173.96,452.70) --
	(173.96,455.85);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (173.96,451.42) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.24,481.37) --
	(192.56,481.37);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.42,480.81) --
	(191.42,481.92);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.42,478.46) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (151.68,424.45) --
	(151.68,434.52);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (151.68,426.72) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.78,476.44) --
	(192.59,476.44);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (184.44,475.09) --
	(184.44,477.73);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.44,472.41) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.84,478.90) --
	(189.05,478.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.62,478.65) --
	(183.62,479.14);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.62,474.86) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (155.04,402.37) --
	(155.04,408.04);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (155.04,401.32) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (159.92,416.68) --
	(159.92,425.68);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (159.92,418.64) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.20,483.88) --
	(195.50,483.88);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.86,483.48) --
	(194.86,484.29);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.86,480.98) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (144.50,437.03) --
	(144.50,441.13);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (144.50,435.93) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (166.76,476.35) --
	(192.56,476.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.59,475.11) --
	(182.59,477.54);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.59,472.32) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (157.54,475.45) --
	(177.27,475.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (169.14,475.33) --
	(169.14,475.58);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.14,471.42) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (148.29,435.17) --
	(148.29,438.68);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (148.29,432.95) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.23,425.41) --
	(160.23,429.48);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.23,424.62) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.33,484.64) --
	(195.61,484.64);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.98,484.32) --
	(194.98,484.95);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.98,481.73) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (106.98,425.81) --
	(121.30,425.81);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (115.07,420.45) --
	(115.07,430.30);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (115.07,422.59) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (178.17,470.70) --
	(191.58,470.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.68,470.14) --
	(185.68,471.25);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.68,466.67) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (135.16,472.44) --
	(141.55,472.44);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (138.54,472.04) --
	(138.54,472.83);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (138.54,468.40) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 89.10,417.71) --
	(103.63,417.71);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 97.32,415.03) --
	( 97.32,420.16);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 97.32,413.68) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (104.73,394.81) --
	(121.56,394.81);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (114.41,389.48) --
	(114.41,399.26);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (114.41,391.90) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.32,481.31) --
	(193.49,481.31);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.95,480.90) --
	(191.95,481.72);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.95,478.41) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 32.48,444.50) --
	(192.17,444.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (173.24,441.31) --
	(173.24,447.36);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (173.24,441.28) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (132.20,463.69) --
	(175.44,463.69);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (161.58,461.46) --
	(161.58,465.75);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (161.58,459.66) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.89,474.60) --
	(201.90,474.60);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (196.54,474.36) --
	(196.54,474.83);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (196.54,470.56) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (180.57,454.48) --
	(205.36,454.48);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.68,453.18) --
	(195.68,455.71);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.68,450.44) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (138.81,444.30) --
	(205.67,444.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.96,442.70) --
	(188.96,445.81);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.96,441.39) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.98,483.57) --
	(193.61,483.57);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.81,482.98) --
	(192.81,484.15);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.81,480.67) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (139.00,463.79) --
	(201.79,463.79);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.42,461.47) --
	(185.42,465.93);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.42,460.57) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.22,472.84) --
	(191.06,472.84);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.99,472.16) --
	(186.99,473.50);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.99,468.81) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (158.42,475.45) --
	(161.91,475.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.22,475.26) --
	(160.22,475.63);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.22,471.41) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (120.33,428.75) --
	(203.84,428.75);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.10,426.86) --
	(186.10,430.51);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.10,424.71) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (131.09,432.16) --
	(204.47,432.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.28,430.29) --
	(187.28,433.90);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.28,429.25) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.52,484.51) --
	(195.71,484.51);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.16,483.14) --
	(194.16,485.82);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.16,481.61) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (133.67,442.76) --
	(133.67,445.15);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (133.67,440.75) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (114.67,452.24) --
	(158.12,452.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (144.22,448.41) --
	(144.22,455.61);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (144.22,448.21) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.34,455.71) --
	(169.74,455.71);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.47,455.45) --
	(160.47,455.96);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.47,451.67) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (108.92,431.10) --
	(154.18,431.10);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (139.99,427.06) --
	(139.99,434.62);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (139.99,427.07) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (160.33,433.08) --
	(160.33,435.62);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.33,431.48) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.30,481.70) --
	(193.01,481.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.22,481.20) --
	(191.22,482.19);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.22,478.80) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.00,440.60) --
	(176.00,444.07);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (176.00,439.16) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (100.89,439.35) --
	(139.08,439.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (126.16,436.47) --
	(126.16,441.96);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (126.16,435.32) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (144.61,456.39) --
	(200.58,456.39);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (184.92,456.18) --
	(184.92,456.60);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.92,452.36) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (179.08,440.70) --
	(208.08,440.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (197.25,439.24) --
	(197.25,442.09);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.25,436.67) {$l$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (134.06,438.18) --
	(205.41,438.18);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (188.36,436.36) --
	(188.36,439.88);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.36,435.27) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.91,484.86) --
	(195.41,484.86);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.19,484.41) --
	(194.19,485.31);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.19,481.96) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.95,338.88) -- ( 93.95,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.96,338.88) -- (112.96,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.11,338.88) -- (157.11,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.12,338.88) -- (176.12,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.27,338.88) -- (220.27,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.28,338.88) -- (239.28,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.95,320.95) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (112.96,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (157.11,320.95) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (176.12,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (220.27,320.95) {500};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,528.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 44.82,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,383.03) -- ( 44.82,383.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,402.04) -- ( 44.82,402.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,446.19) -- ( 44.82,446.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,465.20) -- ( 44.82,465.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,509.35) -- ( 44.82,509.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,528.36) -- ( 44.82,528.36);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,380.17) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,399.18) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,443.33) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,462.34) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,506.49) {500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,525.50) {1000};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,431.01) rectangle (189.67,426.03);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,421.05) rectangle (189.67,416.07);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,411.09) rectangle (189.67,406.11);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,401.13) rectangle (189.67,396.15);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,391.17) rectangle (189.67,386.19);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,381.21) rectangle (189.67,376.23);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,371.25) rectangle (189.67,366.27);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,361.29) rectangle (189.67,356.31);
\definecolor{fillColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,351.33) rectangle (189.67,346.35);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,425.66) {Ran};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,415.70) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,405.74) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,395.78) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,385.82) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,375.86) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,365.90) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,355.94) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,345.98) {Moyle};
\end{scope}
\begin{scope}
\path[clip] (289.08,289.08) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,548.96) {\bfseries adapt (\textit{real}) vs. nocons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,301.03) {ESS/$10^6$ states [nocons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,433.62) {ESS/$10^6$ states [adapt (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] (338.88,338.88) rectangle (528.36,528.36);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] ( 86.24,200.32) --
	(464.08,578.16);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (578.16,464.08) --
	(114.08,  0.00);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] ( 86.24,  0.00) --
	( 86.24,200.32);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] ( 86.24,162.29) --
	(502.11,578.16);

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (578.16,502.11) --
	( 86.24, 10.19) --
	( 86.24,162.29);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] ( 86.24,124.27) --
	(540.13,578.16);

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (578.16,540.13) --
	( 86.24, 48.21) --
	( 86.24,124.27);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (460.49,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,464.39) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (422.46,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,426.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (382.78,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00,388.34) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (372.10,469.76) --
	(386.12,469.76);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (380.00,467.35) --
	(380.00,471.98);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (380.00,466.53) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (465.09,450.78) --
	(472.53,450.78);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (469.06,448.97) --
	(469.06,452.47);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (469.06,446.75) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (369.81,473.72) --
	(373.75,473.72);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (371.85,473.22) --
	(371.85,474.21);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (371.85,469.68) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (358.10,438.88) --
	(366.58,438.88);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (362.67,436.62) --
	(362.67,440.96);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (362.67,434.84) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (380.45,467.10) --
	(394.31,467.10);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (388.24,465.29) --
	(388.24,468.80);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (388.24,464.20) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.95,479.42) --
	(480.06,479.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (479.51,478.97) --
	(479.51,479.86);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (479.51,476.51) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (430.18,471.62) --
	(460.00,471.62);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (448.96,470.85) --
	(448.96,472.37);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (448.96,468.40) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (466.88,474.53) --
	(479.82,474.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (474.10,472.61) --
	(474.10,476.32);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.10,470.49) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (464.81,479.56) --
	(472.19,479.56);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (468.75,479.31) --
	(468.75,479.81);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (468.75,475.53) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (428.65,458.11) --
	(478.12,458.11);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (463.29,457.04) --
	(463.29,459.15);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (463.29,454.08) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (406.97,473.51) --
	(480.21,473.51);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (463.04,472.79) --
	(463.04,474.21);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (463.04,470.60) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (479.32,481.54) --
	(481.64,481.54);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (480.50,481.02) --
	(480.50,482.04);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (480.50,478.63) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (440.76,432.68) --
	(440.76,439.12);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (440.76,432.86) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (461.86,479.26) --
	(481.67,479.26);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (473.52,478.07) --
	(473.52,480.41);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (473.52,475.23) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (465.92,482.87) --
	(478.13,482.87);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (472.70,482.55) --
	(472.70,483.18);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (472.70,478.84) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (444.12,411.64) --
	(444.12,421.72);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (444.12,413.11) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (449.00,429.23) --
	(449.00,436.09);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.00,429.97) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (483.28,484.21) --
	(484.58,484.21);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (483.94,483.81) --
	(483.94,484.61);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.94,481.31) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (433.58,448.44) --
	(433.58,452.06);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (433.58,447.08) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (455.84,473.88) --
	(481.64,473.88);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (471.67,472.84) --
	(471.67,474.88);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (471.67,469.85) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (446.62,475.99) --
	(466.35,475.99);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (458.22,475.70) --
	(458.22,476.27);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (458.22,471.95) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (437.37,439.79) --
	(437.37,442.98);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (437.37,437.39) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (449.31,443.26) --
	(449.31,447.88);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.31,442.77) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (483.41,484.53) --
	(484.69,484.53);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (484.06,484.08) --
	(484.06,484.98);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (484.06,481.63) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (396.06,463.48) --
	(410.38,463.48);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (404.15,461.30) --
	(404.15,465.50);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (404.15,460.25) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (467.25,461.16) --
	(480.66,461.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (474.76,460.75) --
	(474.76,461.56);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.76,457.13) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (424.24,470.95) --
	(430.63,470.95);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (427.62,470.38) --
	(427.62,471.51);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (427.62,466.92) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (378.18,422.68) --
	(392.71,422.68);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (386.40,420.09) --
	(386.40,425.05);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (386.40,418.65) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (393.81,445.42) --
	(410.64,445.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (403.49,442.80) --
	(403.49,447.81);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (403.49,442.52) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (479.40,481.50) --
	(482.57,481.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.03,481.12) --
	(481.03,481.86);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (481.03,478.59) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (321.56,455.89) --
	(481.25,455.89);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (462.32,452.11) --
	(462.32,459.20);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (462.32,452.66) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (421.28,478.12) --
	(464.52,478.12);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (450.66,477.19) --
	(450.66,479.03);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (450.66,474.09) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.97,495.81) --
	(490.98,495.81);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (485.62,495.66) --
	(485.62,495.97);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (485.62,491.78) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (469.65,476.49) --
	(494.44,476.49);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (484.76,475.16) --
	(484.76,477.77);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (484.76,472.46) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (427.89,460.23) --
	(494.75,460.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.04,459.37) --
	(478.04,461.07);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (478.04,457.33) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.06,483.43) --
	(482.69,483.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.89,483.01) --
	(481.89,483.84);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (481.89,480.52) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (428.08,467.23) --
	(490.87,467.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (474.50,465.85) --
	(474.50,468.55);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.50,464.01) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (471.30,464.98) --
	(480.14,464.98);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (476.07,463.73) --
	(476.07,466.18);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.07,460.95) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (447.50,472.08) --
	(450.99,472.08);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (449.30,471.77) --
	(449.30,472.39);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.30,468.05) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (409.41,426.58) --
	(492.92,426.58);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (475.18,423.54) --
	(475.18,429.32);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (475.18,422.55) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (420.17,444.15) --
	(493.55,444.15);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (476.36,442.54) --
	(476.36,445.67);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.36,441.24) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.60,485.16) --
	(484.79,485.16);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (483.24,484.79) --
	(483.24,485.52);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.24,482.25) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (422.75,450.99) --
	(422.75,454.38);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (422.75,449.51) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (403.75,455.09) --
	(447.20,455.09);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (433.30,448.17) --
	(433.30,460.61);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (433.30,451.05) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (435.42,464.34) --
	(458.82,464.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (449.55,463.86) --
	(449.55,464.81);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.55,460.30) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (398.00,441.30) --
	(443.26,441.30);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (429.07,439.87) --
	(429.07,442.66);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (429.07,437.27) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (449.41,451.50) --
	(449.41,454.19);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.41,449.97) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (478.38,482.14) --
	(482.09,482.14);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (480.30,481.65) --
	(480.30,482.62);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (480.30,479.23) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (465.08,446.82) --
	(465.08,451.65);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (465.08,446.11) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (389.97,443.85) --
	(428.16,443.85);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (415.24,441.95) --
	(415.24,445.64);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (415.24,439.82) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (433.69,468.32) --
	(489.66,468.32);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (474.00,468.11) --
	(474.00,468.53);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.00,464.29) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (468.16,461.78) --
	(497.16,461.78);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (486.33,461.02) --
	(486.33,462.52);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (486.33,457.75) {$l$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (423.14,452.09) --
	(494.49,452.09);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (477.44,449.26) --
	(477.44,454.65);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (477.44,449.18) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (481.99,484.04) --
	(484.49,484.04);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (483.27,483.19) --
	(483.27,484.87);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.27,481.14) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.03,338.88) -- (383.03,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (402.04,338.88) -- (402.04,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (446.19,338.88) -- (446.19,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.20,338.88) -- (465.20,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (509.35,338.88) -- (509.35,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.36,338.88) -- (528.36,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (383.03,320.95) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (402.04,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (446.19,320.95) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (465.20,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (509.35,320.95) {500};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,528.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (333.90,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,383.03) -- (333.90,383.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,402.04) -- (333.90,402.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,446.19) -- (333.90,446.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,465.20) -- (333.90,465.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,509.35) -- (333.90,509.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,528.36) -- (333.90,528.36);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,380.17) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,399.18) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,443.33) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,462.34) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,506.49) {500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,525.50) {1000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,259.88) {\bfseries adapt (\textit{real}) vs. cons (\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54, 11.95) {ESS/$10^6$ states [cons (\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,144.54) {ESS/$10^6$ states [adapt (\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80, 49.80) rectangle (239.28,239.28);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (  0.00,114.08) --
	(302.44,416.52) --
	(302.44,188.36) --
	(114.08,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (  0.00, 76.05) --
	(302.44,378.49) --
	(302.44,226.39) --
	( 76.05,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (  0.00, 38.03) --
	(302.44,340.47) --
	(302.44,264.41) --
	( 38.03,  0.00);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (171.41, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,175.31) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (133.38, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,137.28) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.70, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92, 99.26) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28,239.28);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (109.61,180.68) --
	(120.54,180.68);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (115.62,178.27) --
	(115.62,182.90);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (115.62,177.45) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.22,161.70) --
	(180.11,161.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (178.23,159.89) --
	(178.23,163.39);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.23,157.67) {$L$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.27,184.64) --
	(184.69,184.64);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.51,184.14) --
	(183.51,185.13);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.51,180.60) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (129.41,149.80) --
	(139.85,149.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (135.13,147.54) --
	(135.13,151.88);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (135.13,145.76) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 80.91,178.02) --
	( 89.04,178.02);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] ( 85.27,176.21) --
	( 85.27,179.72);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at ( 85.27,175.12) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.58,190.34) --
	(191.32,190.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (190.95,189.89) --
	(190.95,190.78);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (190.95,187.43) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (176.83,182.54) --
	(179.53,182.54);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (178.22,181.77) --
	(178.22,183.29);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.22,179.32) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.24,185.45) --
	(186.54,185.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (184.94,183.53) --
	(184.94,187.24);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.94,181.41) {$L$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.74,190.48) --
	(188.08,190.48);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.91,190.23) --
	(187.91,190.73);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.91,186.45) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (164.07,169.03) --
	(166.36,169.03);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.24,167.96) --
	(165.24,170.07);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.24,165.00) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (163.62,184.43) --
	(166.77,184.43);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.24,183.71) --
	(165.24,185.13);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.24,181.52) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.73,192.46) --
	(192.84,192.46);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.29,191.94) --
	(192.29,192.96);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.29,189.55) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (135.37,147.01) --
	(145.44,147.01);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (140.87,143.60) --
	(140.87,150.04);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (140.87,143.78) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.01,190.18) --
	(188.65,190.18);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.36,188.99) --
	(187.36,191.33);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.36,186.15) {$L$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.57,193.79) --
	(190.06,193.79);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (189.82,193.47) --
	(189.82,194.10);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (189.82,189.76) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (113.29,128.06) --
	(118.96,128.06);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (116.27,122.56) --
	(116.27,132.64);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (116.27,124.03) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (127.60,143.80) --
	(136.60,143.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (132.47,140.15) --
	(132.47,147.01);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (132.47,140.89) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.40,195.13) --
	(195.21,195.13);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.80,194.73) --
	(194.80,195.53);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.80,192.23) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (147.95,161.23) --
	(152.05,161.23);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (150.07,159.36) --
	(150.07,162.98);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (150.07,158.00) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.03,184.80) --
	(188.46,184.80);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (187.27,183.76) --
	(187.27,185.80);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.27,180.77) {$L$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.25,186.91) --
	(186.50,186.91);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.37,186.62) --
	(186.37,187.19);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.37,182.87) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (146.09,152.35) --
	(149.60,152.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (147.90,150.71) --
	(147.90,153.90);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (147.90,148.31) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (136.33,156.59) --
	(140.40,156.59);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (138.44,154.18) --
	(138.44,158.80);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (138.44,153.69) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.24,195.45) --
	(195.87,195.45);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.56,195.00) --
	(195.56,195.90);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.56,192.55) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (131.37,174.40) --
	(141.22,174.40);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (136.73,172.22) --
	(136.73,176.42);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (136.73,171.17) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.06,172.08) --
	(182.17,172.08);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (181.62,171.67) --
	(181.62,172.48);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.62,168.05) {$L$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (182.96,181.87) --
	(183.75,181.87);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.36,181.30) --
	(183.36,182.43);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.36,177.84) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (125.95,133.60) --
	(131.08,133.60);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (128.63,131.01) --
	(128.63,135.97);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (128.63,129.57) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (100.40,156.34) --
	(110.18,156.34);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (105.73,153.72) --
	(105.73,158.73);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (105.73,153.44) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (191.82,192.42) --
	(192.64,192.42);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.23,192.04) --
	(192.23,192.78);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.23,189.51) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (152.23,166.81) --
	(158.28,166.81);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (155.42,163.03) --
	(155.42,170.12);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (155.42,163.58) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.38,189.04) --
	(176.67,189.04);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (174.61,188.11) --
	(174.61,189.95);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.61,185.01) {$L$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.28,206.73) --
	(185.75,206.73);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (185.52,206.58) --
	(185.52,206.89);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.52,202.70) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (164.10,187.41) --
	(166.63,187.41);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (165.40,186.08) --
	(165.40,188.69);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.40,183.38) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (153.62,171.15) --
	(156.73,171.15);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (155.22,170.29) --
	(155.22,171.99);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (155.22,168.25) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (193.90,194.35) --
	(195.07,194.35);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.49,193.93) --
	(194.49,194.76);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.49,191.44) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (172.39,178.15) --
	(176.85,178.15);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (174.71,176.77) --
	(174.71,179.47);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.71,174.93) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.08,175.90) --
	(184.42,175.90);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (183.76,174.65) --
	(183.76,177.10);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.76,171.87) {$L$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.18,183.00) --
	(186.55,183.00);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (186.37,182.69) --
	(186.37,183.31);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.37,178.97) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (137.78,137.50) --
	(141.43,137.50);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (139.67,134.46) --
	(139.67,140.24);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (139.67,133.47) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (141.21,155.07) --
	(144.82,155.07);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (143.08,153.46) --
	(143.08,156.59);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (143.08,152.16) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (194.06,196.08) --
	(196.74,196.08);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.43,195.71) --
	(195.43,196.44);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.43,193.17) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (153.68,163.66) --
	(156.07,163.66);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (154.90,161.91) --
	(154.90,165.30);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (154.90,160.43) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (159.33,166.01) --
	(166.53,166.01);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (163.16,159.09) --
	(163.16,171.53);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (163.16,161.97) {$L$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (166.37,175.26) --
	(166.88,175.26);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (166.63,174.78) --
	(166.63,175.73);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (166.63,171.22) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (137.98,152.22) --
	(145.54,152.22);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (142.02,150.79) --
	(142.02,153.58);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (142.02,148.19) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (144.00,163.79) --
	(146.54,163.79);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (145.30,162.42) --
	(145.30,165.11);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (145.30,160.89) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.12,193.06) --
	(193.11,193.06);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (192.62,192.57) --
	(192.62,193.54);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.62,190.15) {$\kappa$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (151.52,160.26) --
	(154.99,160.26);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (153.31,157.74) --
	(153.31,162.57);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (153.31,157.03) {$p$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (147.39,154.77) --
	(152.88,154.77);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (150.27,152.87) --
	(150.27,156.56);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (150.27,150.74) {$L$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (167.10,179.24) --
	(167.52,179.24);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (167.31,179.03) --
	(167.31,179.45);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (167.31,175.21) {$r$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (150.16,172.70) --
	(153.01,172.70);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (151.62,171.94) --
	(151.62,173.44);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (151.62,168.67) {$l$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (147.28,163.01) --
	(150.80,163.01);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (149.10,160.18) --
	(149.10,165.57);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (149.10,160.10) {$\sigma$};
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.33,194.96) --
	(196.23,194.96);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (195.78,194.11) --
	(195.78,195.79);
\definecolor{drawColor}{RGB}{68,68,68}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.78,192.06) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.95, 49.80) -- ( 93.95, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.96, 49.80) -- (112.96, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.11, 49.80) -- (157.11, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.12, 49.80) -- (176.12, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.27, 49.80) -- (220.27, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.28, 49.80) -- (239.28, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.95, 31.87) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (112.96, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (157.11, 31.87) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (176.12, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (220.27, 31.87) {500};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80,239.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 44.82, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 93.95) -- ( 44.82, 93.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,112.96) -- ( 44.82,112.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,157.11) -- ( 44.82,157.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,176.12) -- ( 44.82,176.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,220.27) -- ( 44.82,220.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,239.28) -- ( 44.82,239.28);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 91.09) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,110.10) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,154.25) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,173.26) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,217.41) {500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,236.42) {1000};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,259.88) {\bfseries \texttt{AdaptiveOperatorSampler} weights};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62, 11.95) {Sequence length $L$ (kb)};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,144.54) {Learned operator weight};
\end{scope}
\begin{scope}
\path[clip] (338.88, 49.80) rectangle (528.36,239.28);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (485.73,235.38) circle (  1.87);

\path[fill=fillColor] (485.73, 54.21) --
	(488.24, 49.85) --
	(483.21, 49.85) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (356.88,164.91) circle (  1.87);

\path[fill=fillColor] (356.88, 79.47) --
	(359.40, 75.11) --
	(354.37, 75.11) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (394.78,165.18) circle (  1.87);

\path[fill=fillColor] (394.78, 79.59) --
	(397.29, 75.23) --
	(392.26, 75.23) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (390.04,192.04) circle (  1.87);

\path[fill=fillColor] (390.04, 71.74) --
	(392.55, 67.38) --
	(387.52, 67.38) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (404.25,216.69) circle (  1.87);

\path[fill=fillColor] (404.25, 59.59) --
	(406.77, 55.23) --
	(401.74, 55.23) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (359.72,107.30) circle (  1.87);

\path[fill=fillColor] (359.72, 91.05) --
	(362.24, 86.70) --
	(357.21, 86.70) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (433.62,203.93) circle (  1.87);

\path[fill=fillColor] (433.62, 68.64) --
	(436.14, 64.29) --
	(431.10, 64.29) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (359.72,138.32) circle (  1.87);

\path[fill=fillColor] (359.72, 88.29) --
	(362.24, 83.94) --
	(357.21, 83.94) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (346.46,129.20) circle (  1.87);

\path[fill=fillColor] (346.46, 90.01) --
	(348.97, 85.65) --
	(343.94, 85.65) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (338.88,115.69) --
	(339.26,116.14) --
	(339.64,116.59) --
	(340.02,117.04) --
	(340.40,117.49) --
	(340.78,117.94) --
	(341.16,118.39) --
	(341.54,118.85) --
	(341.92,119.30) --
	(342.30,119.76) --
	(342.68,120.22) --
	(343.06,120.68) --
	(343.44,121.14) --
	(343.82,121.60) --
	(344.20,122.07) --
	(344.58,122.53) --
	(344.96,123.00) --
	(345.34,123.46) --
	(345.71,123.93) --
	(346.09,124.40) --
	(346.47,124.87) --
	(346.85,125.34) --
	(347.23,125.81) --
	(347.61,126.28) --
	(347.99,126.76) --
	(348.37,127.23) --
	(348.75,127.71) --
	(349.13,128.18) --
	(349.51,128.66) --
	(349.89,129.14) --
	(350.27,129.62) --
	(350.65,130.10) --
	(351.03,130.58) --
	(351.41,131.06) --
	(351.79,131.54) --
	(352.17,132.02) --
	(352.55,132.51) --
	(352.93,132.99) --
	(353.31,133.47) --
	(353.69,133.96) --
	(354.07,134.45) --
	(354.45,134.93) --
	(354.83,135.42) --
	(355.21,135.91) --
	(355.59,136.39) --
	(355.97,136.88) --
	(356.35,137.37) --
	(356.73,137.86) --
	(357.11,138.35) --
	(357.49,138.84) --
	(357.87,139.33) --
	(358.25,139.82) --
	(358.63,140.31) --
	(359.01,140.80) --
	(359.38,141.29) --
	(359.76,141.78) --
	(360.14,142.27) --
	(360.52,142.76) --
	(360.90,143.25) --
	(361.28,143.74) --
	(361.66,144.24) --
	(362.04,144.73) --
	(362.42,145.22) --
	(362.80,145.71) --
	(363.18,146.20) --
	(363.56,146.69) --
	(363.94,147.18) --
	(364.32,147.68) --
	(364.70,148.17) --
	(365.08,148.66) --
	(365.46,149.15) --
	(365.84,149.64) --
	(366.22,150.13) --
	(366.60,150.62) --
	(366.98,151.11) --
	(367.36,151.59) --
	(367.74,152.08) --
	(368.12,152.57) --
	(368.50,153.06) --
	(368.88,153.55) --
	(369.26,154.03) --
	(369.64,154.52) --
	(370.02,155.01) --
	(370.40,155.49) --
	(370.78,155.97) --
	(371.16,156.46) --
	(371.54,156.94) --
	(371.92,157.42) --
	(372.30,157.91) --
	(372.68,158.39) --
	(373.05,158.87) --
	(373.43,159.35) --
	(373.81,159.83) --
	(374.19,160.31) --
	(374.57,160.78) --
	(374.95,161.26) --
	(375.33,161.74) --
	(375.71,162.21) --
	(376.09,162.68) --
	(376.47,163.16) --
	(376.85,163.63) --
	(377.23,164.10) --
	(377.61,164.57) --
	(377.99,165.04) --
	(378.37,165.51) --
	(378.75,165.97) --
	(379.13,166.44) --
	(379.51,166.90) --
	(379.89,167.37) --
	(380.27,167.83) --
	(380.65,168.29) --
	(381.03,168.75) --
	(381.41,169.21) --
	(381.79,169.67) --
	(382.17,170.12) --
	(382.55,170.58) --
	(382.93,171.03) --
	(383.31,171.48) --
	(383.69,171.94) --
	(384.07,172.39) --
	(384.45,172.83) --
	(384.83,173.28) --
	(385.21,173.73) --
	(385.59,174.17) --
	(385.97,174.61) --
	(386.34,175.05) --
	(386.72,175.49) --
	(387.10,175.93) --
	(387.48,176.37) --
	(387.86,176.80) --
	(388.24,177.24) --
	(388.62,177.67) --
	(389.00,178.10) --
	(389.38,178.53) --
	(389.76,178.96) --
	(390.14,179.38) --
	(390.52,179.81) --
	(390.90,180.23) --
	(391.28,180.65) --
	(391.66,181.07) --
	(392.04,181.49) --
	(392.42,181.90) --
	(392.80,182.32) --
	(393.18,182.73) --
	(393.56,183.14) --
	(393.94,183.55) --
	(394.32,183.96) --
	(394.70,184.36) --
	(395.08,184.77) --
	(395.46,185.17) --
	(395.84,185.57) --
	(396.22,185.97) --
	(396.60,186.36) --
	(396.98,186.76) --
	(397.36,187.15) --
	(397.74,187.54) --
	(398.12,187.93) --
	(398.50,188.32) --
	(398.88,188.70) --
	(399.26,189.09) --
	(399.64,189.47) --
	(400.01,189.85) --
	(400.39,190.23) --
	(400.77,190.60) --
	(401.15,190.98) --
	(401.53,191.35) --
	(401.91,191.72) --
	(402.29,192.09) --
	(402.67,192.46) --
	(403.05,192.82) --
	(403.43,193.18) --
	(403.81,193.55) --
	(404.19,193.90) --
	(404.57,194.26) --
	(404.95,194.62) --
	(405.33,194.97) --
	(405.71,195.32) --
	(406.09,195.67) --
	(406.47,196.02) --
	(406.85,196.36) --
	(407.23,196.71) --
	(407.61,197.05) --
	(407.99,197.39) --
	(408.37,197.72) --
	(408.75,198.06) --
	(409.13,198.39) --
	(409.51,198.73) --
	(409.89,199.06) --
	(410.27,199.38) --
	(410.65,199.71) --
	(411.03,200.03) --
	(411.41,200.35) --
	(411.79,200.67) --
	(412.17,200.99) --
	(412.55,201.31) --
	(412.93,201.62) --
	(413.31,201.93) --
	(413.68,202.24) --
	(414.06,202.55) --
	(414.44,202.86) --
	(414.82,203.16) --
	(415.20,203.47) --
	(415.58,203.77) --
	(415.96,204.06) --
	(416.34,204.36) --
	(416.72,204.66) --
	(417.10,204.95) --
	(417.48,205.24) --
	(417.86,205.53) --
	(418.24,205.81) --
	(418.62,206.10) --
	(419.00,206.38) --
	(419.38,206.66) --
	(419.76,206.94) --
	(420.14,207.22) --
	(420.52,207.49) --
	(420.90,207.77) --
	(421.28,208.04) --
	(421.66,208.31) --
	(422.04,208.58) --
	(422.42,208.84) --
	(422.80,209.11) --
	(423.18,209.37) --
	(423.56,209.63) --
	(423.94,209.89) --
	(424.32,210.15) --
	(424.70,210.40) --
	(425.08,210.65) --
	(425.46,210.90) --
	(425.84,211.15) --
	(426.22,211.40) --
	(426.60,211.65) --
	(426.97,211.89) --
	(427.35,212.13) --
	(427.73,212.37) --
	(428.11,212.61) --
	(428.49,212.85) --
	(428.87,213.08) --
	(429.25,213.32) --
	(429.63,213.55) --
	(430.01,213.78) --
	(430.39,214.01) --
	(430.77,214.23) --
	(431.15,214.46) --
	(431.53,214.68) --
	(431.91,214.90) --
	(432.29,215.12) --
	(432.67,215.34) --
	(433.05,215.56) --
	(433.43,215.77) --
	(433.81,215.98) --
	(434.19,216.19) --
	(434.57,216.40) --
	(434.95,216.61) --
	(435.33,216.82) --
	(435.71,217.02) --
	(436.09,217.23) --
	(436.47,217.43) --
	(436.85,217.63) --
	(437.23,217.82) --
	(437.61,218.02) --
	(437.99,218.22) --
	(438.37,218.41) --
	(438.75,218.60) --
	(439.13,218.79) --
	(439.51,218.98) --
	(439.89,219.17) --
	(440.27,219.35) --
	(440.64,219.54) --
	(441.02,219.72) --
	(441.40,219.90) --
	(441.78,220.08) --
	(442.16,220.26) --
	(442.54,220.44) --
	(442.92,220.61) --
	(443.30,220.79) --
	(443.68,220.96) --
	(444.06,221.13) --
	(444.44,221.30) --
	(444.82,221.47) --
	(445.20,221.63) --
	(445.58,221.80) --
	(445.96,221.96) --
	(446.34,222.13) --
	(446.72,222.29) --
	(447.10,222.45) --
	(447.48,222.61) --
	(447.86,222.76) --
	(448.24,222.92) --
	(448.62,223.07) --
	(449.00,223.23) --
	(449.38,223.38) --
	(449.76,223.53) --
	(450.14,223.68) --
	(450.52,223.83) --
	(450.90,223.97) --
	(451.28,224.12) --
	(451.66,224.26) --
	(452.04,224.40) --
	(452.42,224.55) --
	(452.80,224.69) --
	(453.18,224.83) --
	(453.56,224.96) --
	(453.93,225.10) --
	(454.31,225.24) --
	(454.69,225.37) --
	(455.07,225.50) --
	(455.45,225.63) --
	(455.83,225.77) --
	(456.21,225.90) --
	(456.59,226.02) --
	(456.97,226.15) --
	(457.35,226.28) --
	(457.73,226.40) --
	(458.11,226.53) --
	(458.49,226.65) --
	(458.87,226.77) --
	(459.25,226.89) --
	(459.63,227.01) --
	(460.01,227.13) --
	(460.39,227.25) --
	(460.77,227.36) --
	(461.15,227.48) --
	(461.53,227.59) --
	(461.91,227.71) --
	(462.29,227.82) --
	(462.67,227.93) --
	(463.05,228.04) --
	(463.43,228.15) --
	(463.81,228.26) --
	(464.19,228.36) --
	(464.57,228.47) --
	(464.95,228.58) --
	(465.33,228.68) --
	(465.71,228.78) --
	(466.09,228.89) --
	(466.47,228.99) --
	(466.85,229.09) --
	(467.23,229.19) --
	(467.60,229.29) --
	(467.98,229.38) --
	(468.36,229.48) --
	(468.74,229.58) --
	(469.12,229.67) --
	(469.50,229.77) --
	(469.88,229.86) --
	(470.26,229.95) --
	(470.64,230.04) --
	(471.02,230.13) --
	(471.40,230.22) --
	(471.78,230.31) --
	(472.16,230.40) --
	(472.54,230.49) --
	(472.92,230.58) --
	(473.30,230.66) --
	(473.68,230.75) --
	(474.06,230.83) --
	(474.44,230.91) --
	(474.82,231.00) --
	(475.20,231.08) --
	(475.58,231.16) --
	(475.96,231.24) --
	(476.34,231.32) --
	(476.72,231.40) --
	(477.10,231.48) --
	(477.48,231.55) --
	(477.86,231.63) --
	(478.24,231.70) --
	(478.62,231.78) --
	(479.00,231.85) --
	(479.38,231.93) --
	(479.76,232.00) --
	(480.14,232.07) --
	(480.52,232.14) --
	(480.90,232.22) --
	(481.27,232.29) --
	(481.65,232.36) --
	(482.03,232.42) --
	(482.41,232.49) --
	(482.79,232.56) --
	(483.17,232.63) --
	(483.55,232.69) --
	(483.93,232.76) --
	(484.31,232.82) --
	(484.69,232.89) --
	(485.07,232.95) --
	(485.45,233.02) --
	(485.83,233.08) --
	(486.21,233.14) --
	(486.59,233.20) --
	(486.97,233.26) --
	(487.35,233.32) --
	(487.73,233.38) --
	(488.11,233.44) --
	(488.49,233.50) --
	(488.87,233.56) --
	(489.25,233.61) --
	(489.63,233.67) --
	(490.01,233.73) --
	(490.39,233.78) --
	(490.77,233.84) --
	(491.15,233.89) --
	(491.53,233.95) --
	(491.91,234.00) --
	(492.29,234.05) --
	(492.67,234.11) --
	(493.05,234.16) --
	(493.43,234.21) --
	(493.81,234.26) --
	(494.19,234.31) --
	(494.56,234.36) --
	(494.94,234.41) --
	(495.32,234.46) --
	(495.70,234.51) --
	(496.08,234.56) --
	(496.46,234.60) --
	(496.84,234.65) --
	(497.22,234.70) --
	(497.60,234.74) --
	(497.98,234.79) --
	(498.36,234.83) --
	(498.74,234.88) --
	(499.12,234.92) --
	(499.50,234.97) --
	(499.88,235.01) --
	(500.26,235.05) --
	(500.64,235.10) --
	(501.02,235.14) --
	(501.40,235.18) --
	(501.78,235.22) --
	(502.16,235.26) --
	(502.54,235.30) --
	(502.92,235.34) --
	(503.30,235.38) --
	(503.68,235.42) --
	(504.06,235.46) --
	(504.44,235.50) --
	(504.82,235.54) --
	(505.20,235.58) --
	(505.58,235.61) --
	(505.96,235.65) --
	(506.34,235.69) --
	(506.72,235.72) --
	(507.10,235.76) --
	(507.48,235.80) --
	(507.86,235.83) --
	(508.23,235.87) --
	(508.61,235.90) --
	(508.99,235.93) --
	(509.37,235.97) --
	(509.75,236.00) --
	(510.13,236.04) --
	(510.51,236.07) --
	(510.89,236.10) --
	(511.27,236.13) --
	(511.65,236.17) --
	(512.03,236.20) --
	(512.41,236.23) --
	(512.79,236.26) --
	(513.17,236.29) --
	(513.55,236.32) --
	(513.93,236.35) --
	(514.31,236.38) --
	(514.69,236.41) --
	(515.07,236.44) --
	(515.45,236.47) --
	(515.83,236.50) --
	(516.21,236.52) --
	(516.59,236.55) --
	(516.97,236.58) --
	(517.35,236.61) --
	(517.73,236.64) --
	(518.11,236.66) --
	(518.49,236.69) --
	(518.87,236.72) --
	(519.25,236.74) --
	(519.63,236.77) --
	(520.01,236.79) --
	(520.39,236.82) --
	(520.77,236.84) --
	(521.15,236.87) --
	(521.53,236.89) --
	(521.90,236.92) --
	(522.28,236.94) --
	(522.66,236.96) --
	(523.04,236.99) --
	(523.42,237.01) --
	(523.80,237.03) --
	(524.18,237.06) --
	(524.56,237.08) --
	(524.94,237.10) --
	(525.32,237.13) --
	(525.70,237.15) --
	(526.08,237.17) --
	(526.46,237.19) --
	(526.84,237.21) --
	(527.22,237.23) --
	(527.60,237.25) --
	(527.98,237.27) --
	(528.36,237.29);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (338.88, 93.77) --
	(339.26, 93.54) --
	(339.64, 93.31) --
	(340.02, 93.08) --
	(340.40, 92.85) --
	(340.78, 92.63) --
	(341.16, 92.40) --
	(341.54, 92.18) --
	(341.92, 91.95) --
	(342.30, 91.73) --
	(342.68, 91.51) --
	(343.06, 91.29) --
	(343.44, 91.07) --
	(343.82, 90.85) --
	(344.20, 90.63) --
	(344.58, 90.41) --
	(344.96, 90.19) --
	(345.34, 89.98) --
	(345.71, 89.76) --
	(346.09, 89.55) --
	(346.47, 89.33) --
	(346.85, 89.12) --
	(347.23, 88.91) --
	(347.61, 88.70) --
	(347.99, 88.49) --
	(348.37, 88.28) --
	(348.75, 88.07) --
	(349.13, 87.86) --
	(349.51, 87.66) --
	(349.89, 87.45) --
	(350.27, 87.25) --
	(350.65, 87.04) --
	(351.03, 86.84) --
	(351.41, 86.64) --
	(351.79, 86.43) --
	(352.17, 86.23) --
	(352.55, 86.03) --
	(352.93, 85.83) --
	(353.31, 85.64) --
	(353.69, 85.44) --
	(354.07, 85.24) --
	(354.45, 85.05) --
	(354.83, 84.85) --
	(355.21, 84.66) --
	(355.59, 84.46) --
	(355.97, 84.27) --
	(356.35, 84.08) --
	(356.73, 83.89) --
	(357.11, 83.70) --
	(357.49, 83.51) --
	(357.87, 83.32) --
	(358.25, 83.13) --
	(358.63, 82.95) --
	(359.01, 82.76) --
	(359.38, 82.58) --
	(359.76, 82.39) --
	(360.14, 82.21) --
	(360.52, 82.03) --
	(360.90, 81.84) --
	(361.28, 81.66) --
	(361.66, 81.48) --
	(362.04, 81.30) --
	(362.42, 81.13) --
	(362.80, 80.95) --
	(363.18, 80.77) --
	(363.56, 80.60) --
	(363.94, 80.42) --
	(364.32, 80.25) --
	(364.70, 80.07) --
	(365.08, 79.90) --
	(365.46, 79.73) --
	(365.84, 79.56) --
	(366.22, 79.39) --
	(366.60, 79.22) --
	(366.98, 79.05) --
	(367.36, 78.88) --
	(367.74, 78.71) --
	(368.12, 78.55) --
	(368.50, 78.38) --
	(368.88, 78.21) --
	(369.26, 78.05) --
	(369.64, 77.89) --
	(370.02, 77.72) --
	(370.40, 77.56) --
	(370.78, 77.40) --
	(371.16, 77.24) --
	(371.54, 77.08) --
	(371.92, 76.92) --
	(372.30, 76.77) --
	(372.68, 76.61) --
	(373.05, 76.45) --
	(373.43, 76.30) --
	(373.81, 76.14) --
	(374.19, 75.99) --
	(374.57, 75.83) --
	(374.95, 75.68) --
	(375.33, 75.53) --
	(375.71, 75.38) --
	(376.09, 75.23) --
	(376.47, 75.08) --
	(376.85, 74.93) --
	(377.23, 74.78) --
	(377.61, 74.63) --
	(377.99, 74.49) --
	(378.37, 74.34) --
	(378.75, 74.20) --
	(379.13, 74.05) --
	(379.51, 73.91) --
	(379.89, 73.77) --
	(380.27, 73.62) --
	(380.65, 73.48) --
	(381.03, 73.34) --
	(381.41, 73.20) --
	(381.79, 73.06) --
	(382.17, 72.92) --
	(382.55, 72.79) --
	(382.93, 72.65) --
	(383.31, 72.51) --
	(383.69, 72.38) --
	(384.07, 72.24) --
	(384.45, 72.11) --
	(384.83, 71.97) --
	(385.21, 71.84) --
	(385.59, 71.71) --
	(385.97, 71.58) --
	(386.34, 71.45) --
	(386.72, 71.31) --
	(387.10, 71.19) --
	(387.48, 71.06) --
	(387.86, 70.93) --
	(388.24, 70.80) --
	(388.62, 70.67) --
	(389.00, 70.55) --
	(389.38, 70.42) --
	(389.76, 70.30) --
	(390.14, 70.17) --
	(390.52, 70.05) --
	(390.90, 69.93) --
	(391.28, 69.80) --
	(391.66, 69.68) --
	(392.04, 69.56) --
	(392.42, 69.44) --
	(392.80, 69.32) --
	(393.18, 69.20) --
	(393.56, 69.09) --
	(393.94, 68.97) --
	(394.32, 68.85) --
	(394.70, 68.73) --
	(395.08, 68.62) --
	(395.46, 68.50) --
	(395.84, 68.39) --
	(396.22, 68.28) --
	(396.60, 68.16) --
	(396.98, 68.05) --
	(397.36, 67.94) --
	(397.74, 67.83) --
	(398.12, 67.71) --
	(398.50, 67.60) --
	(398.88, 67.50) --
	(399.26, 67.39) --
	(399.64, 67.28) --
	(400.01, 67.17) --
	(400.39, 67.06) --
	(400.77, 66.96) --
	(401.15, 66.85) --
	(401.53, 66.75) --
	(401.91, 66.64) --
	(402.29, 66.54) --
	(402.67, 66.43) --
	(403.05, 66.33) --
	(403.43, 66.23) --
	(403.81, 66.13) --
	(404.19, 66.02) --
	(404.57, 65.92) --
	(404.95, 65.82) --
	(405.33, 65.72) --
	(405.71, 65.62) --
	(406.09, 65.53) --
	(406.47, 65.43) --
	(406.85, 65.33) --
	(407.23, 65.23) --
	(407.61, 65.14) --
	(407.99, 65.04) --
	(408.37, 64.95) --
	(408.75, 64.85) --
	(409.13, 64.76) --
	(409.51, 64.66) --
	(409.89, 64.57) --
	(410.27, 64.48) --
	(410.65, 64.39) --
	(411.03, 64.30) --
	(411.41, 64.20) --
	(411.79, 64.11) --
	(412.17, 64.02) --
	(412.55, 63.93) --
	(412.93, 63.85) --
	(413.31, 63.76) --
	(413.68, 63.67) --
	(414.06, 63.58) --
	(414.44, 63.50) --
	(414.82, 63.41) --
	(415.20, 63.32) --
	(415.58, 63.24) --
	(415.96, 63.15) --
	(416.34, 63.07) --
	(416.72, 62.99) --
	(417.10, 62.90) --
	(417.48, 62.82) --
	(417.86, 62.74) --
	(418.24, 62.66) --
	(418.62, 62.57) --
	(419.00, 62.49) --
	(419.38, 62.41) --
	(419.76, 62.33) --
	(420.14, 62.25) --
	(420.52, 62.17) --
	(420.90, 62.10) --
	(421.28, 62.02) --
	(421.66, 61.94) --
	(422.04, 61.86) --
	(422.42, 61.79) --
	(422.80, 61.71) --
	(423.18, 61.63) --
	(423.56, 61.56) --
	(423.94, 61.48) --
	(424.32, 61.41) --
	(424.70, 61.34) --
	(425.08, 61.26) --
	(425.46, 61.19) --
	(425.84, 61.12) --
	(426.22, 61.04) --
	(426.60, 60.97) --
	(426.97, 60.90) --
	(427.35, 60.83) --
	(427.73, 60.76) --
	(428.11, 60.69) --
	(428.49, 60.62) --
	(428.87, 60.55) --
	(429.25, 60.48) --
	(429.63, 60.41) --
	(430.01, 60.34) --
	(430.39, 60.28) --
	(430.77, 60.21) --
	(431.15, 60.14) --
	(431.53, 60.08) --
	(431.91, 60.01) --
	(432.29, 59.94) --
	(432.67, 59.88) --
	(433.05, 59.81) --
	(433.43, 59.75) --
	(433.81, 59.69) --
	(434.19, 59.62) --
	(434.57, 59.56) --
	(434.95, 59.50) --
	(435.33, 59.43) --
	(435.71, 59.37) --
	(436.09, 59.31) --
	(436.47, 59.25) --
	(436.85, 59.19) --
	(437.23, 59.13) --
	(437.61, 59.07) --
	(437.99, 59.01) --
	(438.37, 58.95) --
	(438.75, 58.89) --
	(439.13, 58.83) --
	(439.51, 58.77) --
	(439.89, 58.71) --
	(440.27, 58.66) --
	(440.64, 58.60) --
	(441.02, 58.54) --
	(441.40, 58.49) --
	(441.78, 58.43) --
	(442.16, 58.37) --
	(442.54, 58.32) --
	(442.92, 58.26) --
	(443.30, 58.21) --
	(443.68, 58.15) --
	(444.06, 58.10) --
	(444.44, 58.04) --
	(444.82, 57.99) --
	(445.20, 57.94) --
	(445.58, 57.89) --
	(445.96, 57.83) --
	(446.34, 57.78) --
	(446.72, 57.73) --
	(447.10, 57.68) --
	(447.48, 57.63) --
	(447.86, 57.57) --
	(448.24, 57.52) --
	(448.62, 57.47) --
	(449.00, 57.42) --
	(449.38, 57.37) --
	(449.76, 57.32) --
	(450.14, 57.28) --
	(450.52, 57.23) --
	(450.90, 57.18) --
	(451.28, 57.13) --
	(451.66, 57.08) --
	(452.04, 57.03) --
	(452.42, 56.99) --
	(452.80, 56.94) --
	(453.18, 56.89) --
	(453.56, 56.85) --
	(453.93, 56.80) --
	(454.31, 56.76) --
	(454.69, 56.71) --
	(455.07, 56.67) --
	(455.45, 56.62) --
	(455.83, 56.58) --
	(456.21, 56.53) --
	(456.59, 56.49) --
	(456.97, 56.44) --
	(457.35, 56.40) --
	(457.73, 56.36) --
	(458.11, 56.31) --
	(458.49, 56.27) --
	(458.87, 56.23) --
	(459.25, 56.19) --
	(459.63, 56.14) --
	(460.01, 56.10) --
	(460.39, 56.06) --
	(460.77, 56.02) --
	(461.15, 55.98) --
	(461.53, 55.94) --
	(461.91, 55.90) --
	(462.29, 55.86) --
	(462.67, 55.82) --
	(463.05, 55.78) --
	(463.43, 55.74) --
	(463.81, 55.70) --
	(464.19, 55.66) --
	(464.57, 55.62) --
	(464.95, 55.59) --
	(465.33, 55.55) --
	(465.71, 55.51) --
	(466.09, 55.47) --
	(466.47, 55.43) --
	(466.85, 55.40) --
	(467.23, 55.36) --
	(467.60, 55.32) --
	(467.98, 55.29) --
	(468.36, 55.25) --
	(468.74, 55.22) --
	(469.12, 55.18) --
	(469.50, 55.14) --
	(469.88, 55.11) --
	(470.26, 55.07) --
	(470.64, 55.04) --
	(471.02, 55.00) --
	(471.40, 54.97) --
	(471.78, 54.94) --
	(472.16, 54.90) --
	(472.54, 54.87) --
	(472.92, 54.83) --
	(473.30, 54.80) --
	(473.68, 54.77) --
	(474.06, 54.74) --
	(474.44, 54.70) --
	(474.82, 54.67) --
	(475.20, 54.64) --
	(475.58, 54.61) --
	(475.96, 54.57) --
	(476.34, 54.54) --
	(476.72, 54.51) --
	(477.10, 54.48) --
	(477.48, 54.45) --
	(477.86, 54.42) --
	(478.24, 54.39) --
	(478.62, 54.36) --
	(479.00, 54.33) --
	(479.38, 54.30) --
	(479.76, 54.27) --
	(480.14, 54.24) --
	(480.52, 54.21) --
	(480.90, 54.18) --
	(481.27, 54.15) --
	(481.65, 54.12) --
	(482.03, 54.09) --
	(482.41, 54.06) --
	(482.79, 54.04) --
	(483.17, 54.01) --
	(483.55, 53.98) --
	(483.93, 53.95) --
	(484.31, 53.92) --
	(484.69, 53.90) --
	(485.07, 53.87) --
	(485.45, 53.84) --
	(485.83, 53.82) --
	(486.21, 53.79) --
	(486.59, 53.76) --
	(486.97, 53.74) --
	(487.35, 53.71) --
	(487.73, 53.68) --
	(488.11, 53.66) --
	(488.49, 53.63) --
	(488.87, 53.61) --
	(489.25, 53.58) --
	(489.63, 53.56) --
	(490.01, 53.53) --
	(490.39, 53.51) --
	(490.77, 53.48) --
	(491.15, 53.46) --
	(491.53, 53.43) --
	(491.91, 53.41) --
	(492.29, 53.39) --
	(492.67, 53.36) --
	(493.05, 53.34) --
	(493.43, 53.31) --
	(493.81, 53.29) --
	(494.19, 53.27) --
	(494.56, 53.24) --
	(494.94, 53.22) --
	(495.32, 53.20) --
	(495.70, 53.18) --
	(496.08, 53.15) --
	(496.46, 53.13) --
	(496.84, 53.11) --
	(497.22, 53.09) --
	(497.60, 53.06) --
	(497.98, 53.04) --
	(498.36, 53.02) --
	(498.74, 53.00) --
	(499.12, 52.98) --
	(499.50, 52.96) --
	(499.88, 52.94) --
	(500.26, 52.92) --
	(500.64, 52.89) --
	(501.02, 52.87) --
	(501.40, 52.85) --
	(501.78, 52.83) --
	(502.16, 52.81) --
	(502.54, 52.79) --
	(502.92, 52.77) --
	(503.30, 52.75) --
	(503.68, 52.73) --
	(504.06, 52.71) --
	(504.44, 52.69) --
	(504.82, 52.67) --
	(505.20, 52.66) --
	(505.58, 52.64) --
	(505.96, 52.62) --
	(506.34, 52.60) --
	(506.72, 52.58) --
	(507.10, 52.56) --
	(507.48, 52.54) --
	(507.86, 52.52) --
	(508.23, 52.51) --
	(508.61, 52.49) --
	(508.99, 52.47) --
	(509.37, 52.45) --
	(509.75, 52.43) --
	(510.13, 52.42) --
	(510.51, 52.40) --
	(510.89, 52.38) --
	(511.27, 52.36) --
	(511.65, 52.35) --
	(512.03, 52.33) --
	(512.41, 52.31) --
	(512.79, 52.30) --
	(513.17, 52.28) --
	(513.55, 52.26) --
	(513.93, 52.25) --
	(514.31, 52.23) --
	(514.69, 52.21) --
	(515.07, 52.20) --
	(515.45, 52.18) --
	(515.83, 52.17) --
	(516.21, 52.15) --
	(516.59, 52.13) --
	(516.97, 52.12) --
	(517.35, 52.10) --
	(517.73, 52.09) --
	(518.11, 52.07) --
	(518.49, 52.06) --
	(518.87, 52.04) --
	(519.25, 52.03) --
	(519.63, 52.01) --
	(520.01, 52.00) --
	(520.39, 51.98) --
	(520.77, 51.97) --
	(521.15, 51.95) --
	(521.53, 51.94) --
	(521.90, 51.92) --
	(522.28, 51.91) --
	(522.66, 51.90) --
	(523.04, 51.88) --
	(523.42, 51.87) --
	(523.80, 51.85) --
	(524.18, 51.84) --
	(524.56, 51.83) --
	(524.94, 51.81) --
	(525.32, 51.80) --
	(525.70, 51.79) --
	(526.08, 51.77) --
	(526.46, 51.76) --
	(526.84, 51.75) --
	(527.22, 51.73) --
	(527.60, 51.72) --
	(527.98, 51.71) --
	(528.36, 51.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (528.36, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.25, 49.80) -- (386.25, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.62, 49.80) -- (433.62, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.99, 49.80) -- (480.99, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.36, 49.80) -- (528.36, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88, 31.87) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (386.25, 31.87) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (433.62, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (480.99, 31.87) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (528.36, 31.87) {20};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88,239.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (333.90, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 87.70) -- (333.90, 87.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,125.59) -- (333.90,125.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,163.49) -- (333.90,163.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,201.38) -- (333.90,201.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,239.28) -- (333.90,239.28);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 46.94) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 84.84) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,122.73) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,160.63) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,198.53) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,236.42) {1.0};
\end{scope}
\end{tikzpicture}

\end{document}