### 2020-08-31 benchmarking/benchmarkingVM/ESS_round1_catquant.tex
Results update

% Created by tikzDevice version 0.12.3.1 on 2020-08-31 17:26:11
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,289.08) rectangle (289.08,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,548.96) {\bfseries cons(\textit{quant}) vs. nocons(\textit{quant})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,301.03) {ESS/hr [nocons(\textit{quant})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,433.62) {ESS/hr [cons(\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (  0.00,368.65) --
	(209.51,578.16);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,492.84) --
	(  0.00,209.51);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (  0.00,342.13) --
	(236.03,578.16);

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,519.37) --
	(  0.00,236.03);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (  0.00,315.60) --
	(262.56,578.16);

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,545.89) --
	(  0.00,262.56);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92,376.84) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.85,487.92) --
	(203.52,487.92);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.19,487.52) --
	(203.19,488.31);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.19,484.69) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.88,498.22) --
	(214.75,498.22);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.32,497.75) --
	(214.32,498.68);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (214.32,495.32) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.51,502.46) --
	(208.63,502.46);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.57,502.40) --
	(208.57,502.51);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.57,498.42) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.87,483.08) --
	(199.54,483.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.20,482.74) --
	(199.20,483.41);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.20,480.17) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.58,485.24) --
	(201.25,485.24);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.92,484.90) --
	(200.92,485.57);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.92,481.20) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.85,484.76) --
	(203.53,484.76);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.19,484.34) --
	(203.19,485.17);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.19,481.86) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.60,502.14) --
	(213.84,502.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.72,501.98) --
	(213.72,502.29);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (213.72,499.23) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (147.47,446.68) --
	(148.89,446.68);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (148.19,445.79) --
	(148.19,447.53);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (148.19,443.45) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.71,488.48) --
	(204.13,488.48);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.92,488.23) --
	(203.92,488.73);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.92,485.58) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.49,492.12) --
	(200.68,492.12);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.58,492.09) --
	(200.58,492.16);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.58,488.09) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (141.94,439.96) --
	(143.33,439.96);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (142.65,439.14) --
	(142.65,440.75);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (142.65,437.06) {$\bar{r}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (144.03,442.75) --
	(145.40,442.75);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (144.73,441.92) --
	(144.73,443.56);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (144.73,438.72) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (163.36,453.61) --
	(165.24,453.61);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (164.32,452.72) --
	(164.32,454.46);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (164.32,450.71) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.54,491.89) --
	(203.75,491.89);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.64,491.82) --
	(203.64,491.97);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.64,488.99) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (159.69,460.73) --
	(161.01,460.73);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (160.36,460.14) --
	(160.36,461.31);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.36,457.51) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.82,486.84) --
	(203.46,486.84);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.14,486.48) --
	(203.14,487.19);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.14,483.93) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.57,488.60) --
	(194.89,488.60);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.73,488.53) --
	(194.73,488.67);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.73,484.56) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (155.28,458.37) --
	(156.76,458.37);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (156.03,457.88) --
	(156.03,458.86);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (156.03,455.47) {$\bar{r}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (158.01,460.12) --
	(159.37,460.12);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (158.71,459.64) --
	(158.71,460.59);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (158.71,456.09) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (162.76,457.73) --
	(164.34,457.73);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (163.57,457.26) --
	(163.57,458.19);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (163.57,454.83) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.97,491.19) --
	(203.31,491.19);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.14,491.10) --
	(203.14,491.27);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.14,488.28) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (127.56,426.45) --
	(130.46,426.45);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (129.06,425.65) --
	(129.06,427.21);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (129.06,423.22) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (185.11,469.37) --
	(185.94,469.37);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (185.53,468.22) --
	(185.53,470.46);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.53,466.47) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (158.31,469.19) --
	(158.61,469.19);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (158.46,469.06) --
	(158.46,469.32);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (158.46,465.16) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (126.07,427.08) --
	(128.95,427.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (127.56,426.01) --
	(127.56,428.09);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (127.56,424.17) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (126.96,427.07) --
	(129.90,427.07);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (128.49,426.25) --
	(128.49,427.85);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (128.49,423.04) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (129.24,415.67) --
	(133.28,415.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (131.36,414.96) --
	(131.36,416.36);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (131.36,412.77) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (186.71,474.31) --
	(187.04,474.31);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (186.88,474.21) --
	(186.88,474.40);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.88,471.40) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.99,484.20) --
	(203.57,484.20);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.28,483.88) --
	(203.28,484.52);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.28,480.98) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.88,489.56) --
	(209.24,489.56);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.06,489.28) --
	(209.06,489.83);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (209.06,486.65) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (218.03,498.02) --
	(218.09,498.02);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (218.06,497.99) --
	(218.06,498.06);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.06,493.99) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.86,475.48) --
	(195.71,475.48);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (195.29,475.12) --
	(195.29,475.84);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.29,472.58) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.72,482.40) --
	(201.41,482.40);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.07,482.06) --
	(201.07,482.74);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (201.07,478.37) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.35,477.75) --
	(194.47,477.75);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.92,477.24) --
	(193.92,478.25);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.92,474.85) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.99,497.02) --
	(210.17,497.02);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (210.08,496.91) --
	(210.08,497.12);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.08,494.11) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (166.57,459.35) --
	(168.17,459.35);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (167.39,458.78) --
	(167.39,459.90);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (167.39,456.12) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (215.97,496.38) --
	(216.67,496.38);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (216.32,496.09) --
	(216.32,496.67);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (216.32,493.48) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.46,498.08) --
	(198.01,498.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.74,497.92) --
	(197.74,498.23);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.74,494.04) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (165.11,458.03) --
	(166.67,458.03);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (165.90,457.48) --
	(165.90,458.56);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.90,455.13) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (166.96,459.93) --
	(168.48,459.93);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (167.73,459.33) --
	(167.73,460.51);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (167.73,455.90) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (172.54,463.17) --
	(174.47,463.17);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (173.53,462.51) --
	(173.53,463.80);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (173.53,460.26) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.25,501.66) --
	(214.98,501.66);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.62,501.48) --
	(214.62,501.84);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (214.62,498.76) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (177.43,461.34) --
	(178.36,461.34);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (177.90,460.84) --
	(177.90,461.83);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (177.90,458.11) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.18,473.35) --
	(189.52,473.35);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.86,472.81) --
	(188.86,473.87);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.86,470.44) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (191.84,476.83) --
	(191.96,476.83);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (191.90,476.76) --
	(191.90,476.90);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.90,472.80) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (178.74,458.52) --
	(179.34,458.52);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.04,458.06) --
	(179.04,458.96);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.04,455.61) {$\bar{r}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (176.19,460.22) --
	(177.09,460.22);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (176.64,459.76) --
	(176.64,460.68);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (176.64,456.19) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.97,462.52) --
	(180.75,462.52);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.36,462.09) --
	(180.36,462.94);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (180.36,459.62) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.15,490.39) --
	(202.43,490.39);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.29,490.22) --
	(202.29,490.56);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.29,487.49) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.72,476.32) --
	(199.21,476.32);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.97,476.00) --
	(198.97,476.64);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (198.97,473.10) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (186.89,471.92) --
	(187.71,471.92);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.31,471.34) --
	(187.31,472.48);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.31,469.01) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.12,485.97) --
	(203.22,485.97);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.17,485.90) --
	(203.17,486.03);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.17,481.93) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (192.95,469.97) --
	(193.53,469.97);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.24,469.65) --
	(193.24,470.29);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.24,467.07) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.73,476.08) --
	(199.11,476.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.92,475.74) --
	(198.92,476.42);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (198.92,472.05) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (191.83,472.12) --
	(192.70,472.12);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (192.27,471.75) --
	(192.27,472.48);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.27,469.21) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.50,501.05) --
	(214.82,501.05);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.66,500.92) --
	(214.66,501.17);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (214.66,498.14) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- (226.02,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85,338.88) -- ( 93.85,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91,338.88) -- (137.91,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96,338.88) -- (181.96,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02,338.88) -- (226.02,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 44.82,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,382.93) -- ( 44.82,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,426.99) -- ( 44.82,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,471.04) -- ( 44.82,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,515.10) -- ( 44.82,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,421.05) rectangle (189.67,416.07);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,411.09) rectangle (189.67,406.11);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,401.13) rectangle (189.67,396.15);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,391.17) rectangle (189.67,386.19);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,381.21) rectangle (189.67,376.23);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,371.25) rectangle (189.67,366.27);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,361.29) rectangle (189.67,356.31);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,351.33) rectangle (189.67,346.35);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,415.70) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,405.74) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,395.78) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,385.82) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,375.86) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,365.90) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,355.94) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,345.98) {Moyle};
\end{scope}
\begin{scope}
\path[clip] (289.08,289.08) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,548.96) {\bfseries adapt(\textit{quant}) vs. nocons(\textit{quant})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,301.03) {ESS/hr [nocons(\textit{quant})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,433.62) {ESS/hr [adapt(\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] (338.88,338.88) rectangle (528.36,528.36);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (162.66,242.23) --
	(498.59,578.16);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,492.84) --
	(162.66, 83.09) --
	(162.66,242.23);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (162.66,215.71) --
	(525.11,578.16);

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,519.37) --
	(162.66,109.61) --
	(162.66,215.71);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (162.66,189.19) --
	(551.64,578.16);

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,545.89) --
	(162.66,136.14) --
	(162.66,189.19);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (425.98,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (399.46,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (371.28,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00,376.84) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.93,489.87) --
	(492.60,489.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.27,489.57) --
	(492.27,490.16);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (492.27,486.64) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (502.96,499.95) --
	(503.83,499.95);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.40,499.66) --
	(503.40,500.24);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (503.40,497.05) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (497.59,501.75) --
	(497.71,501.75);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (497.65,501.69) --
	(497.65,501.80);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (497.65,497.71) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (487.95,485.79) --
	(488.62,485.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (488.28,485.54) --
	(488.28,486.03);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (488.28,482.88) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (489.66,487.50) --
	(490.33,487.50);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (490.00,487.21) --
	(490.00,487.80);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (490.00,483.47) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.93,499.04) --
	(492.61,499.04);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.27,498.77) --
	(492.27,499.32);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (492.27,496.14) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (502.68,500.55) --
	(502.92,500.55);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (502.80,500.44) --
	(502.80,500.66);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (502.80,497.65) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (436.55,451.79) --
	(437.97,451.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (437.27,451.27) --
	(437.27,452.29);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (437.27,448.56) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.79,489.99) --
	(493.21,489.99);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.00,489.81) --
	(493.00,490.18);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.00,487.09) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (489.57,492.80) --
	(489.76,492.80);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (489.66,492.74) --
	(489.66,492.85);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (489.66,488.76) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (431.02,445.43) --
	(432.41,445.43);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (431.73,444.89) --
	(431.73,445.95);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (431.73,442.52) {$\bar{r}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (433.11,448.01) --
	(434.48,448.01);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (433.81,447.49) --
	(433.81,448.51);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (433.81,443.97) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (452.44,462.20) --
	(454.32,462.20);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (453.40,461.65) --
	(453.40,462.73);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (453.40,459.29) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.62,490.73) --
	(492.83,490.73);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.72,490.65) --
	(492.72,490.82);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (492.72,487.83) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (448.77,464.66) --
	(450.09,464.66);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (449.44,464.12) --
	(449.44,465.18);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.44,461.43) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.90,489.34) --
	(492.54,489.34);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.22,489.09) --
	(492.22,489.58);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (492.22,486.43) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.65,489.35) --
	(483.97,489.35);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.81,489.15) --
	(483.81,489.55);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.81,485.32) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (444.36,461.69) --
	(445.84,461.69);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (445.11,461.16) --
	(445.11,462.22);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (445.11,458.79) {$\bar{r}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (447.09,463.55) --
	(448.45,463.55);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (447.79,463.00) --
	(447.79,464.08);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (447.79,459.51) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (451.84,470.31) --
	(453.42,470.31);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (452.65,469.79) --
	(452.65,470.82);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (452.65,467.41) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.05,490.35) --
	(492.39,490.35);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.22,490.11) --
	(492.22,490.60);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (492.22,487.45) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (416.64,427.41) --
	(419.54,427.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (418.14,426.32) --
	(418.14,428.45);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (418.14,424.19) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (474.19,462.41) --
	(475.02,462.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (474.61,461.74) --
	(474.61,463.06);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.61,459.50) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (447.39,467.41) --
	(447.69,467.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (447.54,467.24) --
	(447.54,467.58);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (447.54,463.38) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (415.15,426.94) --
	(418.03,426.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (416.64,425.89) --
	(416.64,427.93);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (416.64,424.03) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (416.04,427.14) --
	(418.98,427.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (417.57,426.01) --
	(417.57,428.20);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (417.57,423.11) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (418.32,441.14) --
	(422.36,441.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (420.44,439.81) --
	(420.44,442.38);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (420.44,438.23) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (475.79,473.14) --
	(476.12,473.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (475.96,472.98) --
	(475.96,473.30);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (475.96,470.24) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.07,491.90) --
	(492.65,491.90);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.36,491.57) --
	(492.36,492.23);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (492.36,488.68) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (497.96,494.94) --
	(498.32,494.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (498.14,494.63) --
	(498.14,495.24);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (498.14,492.04) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (507.11,505.18) --
	(507.17,505.18);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (507.14,505.15) --
	(507.14,505.21);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (507.14,501.15) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.94,485.14) --
	(484.79,485.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (484.37,484.76) --
	(484.37,485.50);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (484.37,482.23) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (489.80,490.54) --
	(490.49,490.54);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (490.15,490.21) --
	(490.15,490.85);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (490.15,486.50) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (482.43,483.52) --
	(483.55,483.52);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.00,483.11) --
	(483.00,483.93);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.00,480.62) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (499.07,495.41) --
	(499.25,495.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (499.16,495.32) --
	(499.16,495.49);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (499.16,492.50) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (455.65,459.08) --
	(457.25,459.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (456.47,458.35) --
	(456.47,459.79);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (456.47,455.85) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.05,495.76) --
	(505.75,495.76);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.40,495.49) --
	(505.40,496.03);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (505.40,492.85) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (486.54,496.08) --
	(487.09,496.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (486.82,495.98) --
	(486.82,496.19);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (486.82,492.05) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (454.19,457.54) --
	(455.75,457.54);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (454.98,456.85) --
	(454.98,458.21);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (454.98,454.64) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (456.04,459.53) --
	(457.56,459.53);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (456.81,458.82) --
	(456.81,460.21);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (456.81,455.50) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (461.62,474.65) --
	(463.55,474.65);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (462.61,473.94) --
	(462.61,475.33);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (462.61,471.74) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.33,498.52) --
	(504.06,498.52);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.70,498.30) --
	(503.70,498.73);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (503.70,495.61) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (466.51,466.02) --
	(467.44,466.02);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (466.98,465.37) --
	(466.98,466.64);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (466.98,462.79) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (477.26,476.12) --
	(478.60,476.12);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (477.94,475.44) --
	(477.94,476.77);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (477.94,473.22) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (480.92,480.75) --
	(481.04,480.75);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (480.98,480.67) --
	(480.98,480.84);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (480.98,476.72) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (467.82,466.41) --
	(468.42,466.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (468.12,466.01) --
	(468.12,466.81);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (468.12,463.51) {$\bar{r}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (465.27,464.95) --
	(466.17,464.95);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (465.72,464.33) --
	(465.72,465.56);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (465.72,460.92) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (469.05,472.46) --
	(469.83,472.46);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (469.44,471.96) --
	(469.44,472.95);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (469.44,469.56) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.23,489.31) --
	(491.51,489.31);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.37,489.19) --
	(491.37,489.43);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (491.37,486.41) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (487.80,486.27) --
	(488.29,486.27);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (488.05,486.05) --
	(488.05,486.50);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (488.05,483.05) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (475.97,473.16) --
	(476.79,473.16);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (476.39,472.60) --
	(476.39,473.70);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.39,470.25) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.20,490.92) --
	(492.30,490.92);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.25,490.86) --
	(492.25,490.98);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (492.25,486.89) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (482.03,480.77) --
	(482.61,480.77);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (482.32,480.54) --
	(482.32,481.00);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (482.32,477.86) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (487.81,486.02) --
	(488.19,486.02);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (488.00,485.79) --
	(488.00,486.26);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (488.00,481.99) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (480.91,481.81) --
	(481.78,481.81);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (481.35,481.42) --
	(481.35,482.20);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (481.35,478.91) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.58,499.85) --
	(503.90,499.85);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.74,499.72) --
	(503.74,499.97);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (503.74,496.94) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (515.10,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.93,338.88) -- (382.93,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.99,338.88) -- (426.99,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.04,338.88) -- (471.04,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.10,338.88) -- (515.10,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (382.93,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (426.99,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (471.04,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (515.10,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (333.90,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,382.93) -- (333.90,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,426.99) -- (333.90,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,471.04) -- (333.90,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,515.10) -- (333.90,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,259.88) {\bfseries adapt(\textit{quant}) vs. cons(\textit{quant})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54, 11.95) {ESS/hr [cons(\textit{quant})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,144.54) {ESS/hr [adapt(\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80, 49.80) rectangle (239.28,239.28);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (  0.00, 79.57) --
	(283.33,362.91) --
	(283.33,203.76) --
	( 79.57,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (  0.00, 53.05) --
	(283.33,336.38) --
	(283.33,230.29) --
	( 53.05,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (  0.00, 26.52) --
	(283.33,309.86) --
	(283.33,256.81) --
	( 26.52,  0.00);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,140.80) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,114.28) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92, 87.76) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28,239.28);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.44,200.79) --
	(199.23,200.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.84,200.49) --
	(198.84,201.08);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (198.84,197.56) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.67,210.87) --
	(209.60,210.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.14,210.58) --
	(209.14,211.16);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (209.14,207.97) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.32,212.67) --
	(213.43,212.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.38,212.61) --
	(213.38,212.72);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (213.38,208.63) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.66,196.71) --
	(194.33,196.71);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.00,196.46) --
	(194.00,196.95);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.00,193.80) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (195.82,198.42) --
	(196.49,198.42);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (196.16,198.13) --
	(196.16,198.72);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (196.16,194.39) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (195.26,209.96) --
	(196.09,209.96);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (195.68,209.69) --
	(195.68,210.24);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.68,207.06) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.90,211.47) --
	(213.21,211.47);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.06,211.36) --
	(213.06,211.58);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (213.06,208.57) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (156.71,162.71) --
	(158.45,162.71);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (157.60,162.19) --
	(157.60,163.21);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (157.60,159.48) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.15,200.91) --
	(199.65,200.91);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.40,200.73) --
	(199.40,201.10);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.40,198.01) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.01,203.72) --
	(203.08,203.72);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.04,203.66) --
	(203.04,203.77);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.04,199.68) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (150.06,156.35) --
	(151.67,156.35);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (150.88,155.81) --
	(150.88,156.87);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (150.88,153.44) {$\bar{r}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (152.84,158.93) --
	(154.48,158.93);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (153.67,158.41) --
	(153.67,159.43);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (153.67,154.89) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (163.64,173.12) --
	(165.38,173.12);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (164.53,172.57) --
	(164.53,173.65);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (164.53,170.21) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.74,201.65) --
	(202.89,201.65);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.81,201.57) --
	(202.81,201.74);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.81,198.75) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.06,175.58) --
	(172.23,175.58);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.65,175.04) --
	(171.65,176.10);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (171.65,172.35) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.40,200.26) --
	(198.11,200.26);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.76,200.01) --
	(197.76,200.50);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.76,197.35) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.45,200.27) --
	(199.59,200.27);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.52,200.07) --
	(199.52,200.47);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.52,196.24) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.80,172.61) --
	(169.78,172.61);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (169.29,172.08) --
	(169.29,173.14);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.29,169.71) {$\bar{r}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (170.56,174.47) --
	(171.51,174.47);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.04,173.92) --
	(171.04,175.00);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (171.04,170.43) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.18,181.23) --
	(169.11,181.23);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.65,180.71) --
	(168.65,181.74);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (168.65,178.33) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.02,201.27) --
	(202.19,201.27);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.11,201.03) --
	(202.11,201.52);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.11,198.37) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (136.57,138.33) --
	(138.13,138.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (137.37,137.24) --
	(137.37,139.37);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (137.37,135.11) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.14,173.33) --
	(181.38,173.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.29,172.66) --
	(180.29,173.98);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (180.29,170.42) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.98,178.33) --
	(180.24,178.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.11,178.16) --
	(180.11,178.50);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (180.11,174.30) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (136.93,137.86) --
	(139.01,137.86);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (138.00,136.81) --
	(138.00,138.85);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (138.00,134.95) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (137.17,138.06) --
	(138.77,138.06);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (137.99,136.93) --
	(137.99,139.12);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (137.99,134.03) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (125.88,152.06) --
	(127.28,152.06);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (126.59,150.73) --
	(126.59,153.30);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (126.59,149.15) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (185.13,184.06) --
	(185.32,184.06);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (185.23,183.90) --
	(185.23,184.22);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.23,181.16) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.80,202.82) --
	(195.44,202.82);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (195.12,202.49) --
	(195.12,203.15);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.12,199.60) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.20,205.86) --
	(200.75,205.86);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.48,205.55) --
	(200.48,206.16);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.48,202.96) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.91,216.10) --
	(208.98,216.10);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.94,216.07) --
	(208.94,216.13);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.94,212.07) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (186.04,196.06) --
	(186.76,196.06);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (186.40,195.68) --
	(186.40,196.42);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.40,193.15) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (192.98,201.46) --
	(193.66,201.46);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.32,201.13) --
	(193.32,201.77);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.32,197.42) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.16,194.44) --
	(189.17,194.44);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.67,194.03) --
	(188.67,194.85);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.67,191.54) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (207.83,206.33) --
	(208.04,206.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (207.94,206.24) --
	(207.94,206.41);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.94,203.42) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (169.70,170.00) --
	(170.82,170.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (170.27,169.27) --
	(170.27,170.71);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (170.27,166.77) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (207.01,206.68) --
	(207.59,206.68);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (207.30,206.41) --
	(207.30,206.95);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.30,203.77) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.84,207.00) --
	(209.15,207.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.00,206.90) --
	(209.00,207.11);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (209.00,202.97) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.40,168.46) --
	(169.48,168.46);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.95,167.77) --
	(168.95,169.13);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (168.95,165.56) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (170.25,170.45) --
	(171.43,170.45);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (170.85,169.74) --
	(170.85,171.13);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (170.85,166.42) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (173.43,185.57) --
	(174.72,185.57);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (174.09,184.86) --
	(174.09,186.25);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.09,182.66) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.40,209.44) --
	(212.76,209.44);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.58,209.22) --
	(212.58,209.65);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.58,206.53) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.76,176.94) --
	(172.75,176.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (172.26,176.29) --
	(172.26,177.56);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (172.26,173.71) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (183.73,187.04) --
	(184.79,187.04);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (184.27,186.36) --
	(184.27,187.69);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.27,184.14) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.68,191.67) --
	(187.82,191.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.75,191.59) --
	(187.75,191.76);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.75,187.64) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.98,177.33) --
	(169.88,177.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (169.44,176.93) --
	(169.44,177.73);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.44,174.43) {$\bar{r}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (170.68,175.87) --
	(171.60,175.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.14,175.25) --
	(171.14,176.48);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (171.14,171.84) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (173.01,183.38) --
	(173.86,183.38);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (173.44,182.88) --
	(173.44,183.87);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (173.44,180.48) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.14,200.23) --
	(201.48,200.23);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.31,200.11) --
	(201.31,200.35);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (201.31,197.33) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (186.92,197.19) --
	(187.56,197.19);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.24,196.97) --
	(187.24,197.42);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.24,193.97) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.26,184.08) --
	(183.40,184.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.84,183.52) --
	(182.84,184.62);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.84,181.17) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (196.82,201.84) --
	(196.95,201.84);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (196.89,201.78) --
	(196.89,201.90);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (196.89,197.81) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.57,191.69) --
	(181.21,191.69);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.89,191.46) --
	(180.89,191.92);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (180.89,188.78) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (186.66,196.94) --
	(187.34,196.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.00,196.71) --
	(187.00,197.18);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.00,192.91) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.67,192.73) --
	(183.40,192.73);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (183.04,192.34) --
	(183.04,193.12);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.04,189.83) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.84,210.77) --
	(212.09,210.77);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.97,210.64) --
	(211.97,210.89);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.97,207.86) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- (226.02, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85, 49.80) -- ( 93.85, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91, 49.80) -- (137.91, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96, 49.80) -- (181.96, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02, 49.80) -- (226.02, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96, 31.87) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02, 31.87) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80,226.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 44.82, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 93.85) -- ( 44.82, 93.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,137.91) -- ( 44.82,137.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,181.96) -- ( 44.82,181.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,226.02) -- ( 44.82,226.02);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 91.00) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,135.05) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,179.11) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,223.16) {10000};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,259.88) {\bfseries adapt(\textit{cat}) vs. \textit{cat}};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62, 11.95) {ESS/hr [\textit{cat}]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,144.54) {ESS/hr [adapt(\textit{cat})]};
\end{scope}
\begin{scope}
\path[clip] (338.88, 49.80) rectangle (528.36,239.28);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (209.51,  0.00) --
	(572.41,362.91) --
	(572.41,203.76) --
	(368.65,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (236.03,  0.00) --
	(572.41,336.38) --
	(572.41,230.29) --
	(342.13,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (262.56,  0.00) --
	(572.41,309.86) --
	(572.41,256.81) --
	(315.60,  0.00);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (425.98, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,140.80) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (399.46, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,114.28) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (371.28, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00, 87.76) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (338.88, 49.80) -- (528.36,239.28);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.65,206.23) --
	(493.27,206.23);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.96,205.92) --
	(492.96,206.54);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (492.96,203.00) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.38,216.73) --
	(506.11,216.73);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.75,216.33) --
	(505.75,217.11);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (505.75,213.82) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (499.10,211.70) --
	(499.21,211.70);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (499.16,211.62) --
	(499.16,211.77);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (499.16,207.66) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (487.73,202.32) --
	(488.29,202.32);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (488.01,201.95) --
	(488.01,202.68);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (488.01,199.41) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (489.82,204.03) --
	(490.42,204.03);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (490.12,203.67) --
	(490.12,204.38);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (490.12,200.00) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.57,206.16) --
	(494.25,206.16);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.92,205.75) --
	(493.92,206.55);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.92,203.25) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (506.62,217.05) --
	(506.80,217.05);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (506.71,216.87) --
	(506.71,217.23);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (506.71,214.15) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (467.32,174.83) --
	(468.23,174.83);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (467.78,173.90) --
	(467.78,175.72);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (467.78,171.60) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.74,206.87) --
	(496.14,206.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.94,206.59) --
	(495.94,207.15);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (495.94,203.96) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.49,204.57) --
	(491.65,204.57);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.57,204.49) --
	(491.57,204.66);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (491.57,200.54) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (460.42,166.26) --
	(461.37,166.26);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (460.90,165.17) --
	(460.90,167.29);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (460.90,163.35) {$\bar{r}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (462.36,168.44) --
	(463.31,168.44);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (462.84,167.36) --
	(462.84,169.47);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (462.84,164.41) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (460.95,171.11) --
	(462.32,171.11);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (461.64,170.32) --
	(461.64,171.88);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (461.64,168.21) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (496.66,206.85) --
	(496.88,206.85);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (496.77,206.77) --
	(496.77,206.93);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (496.77,203.95) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (451.25,162.13) --
	(452.20,162.13);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (451.73,161.26) --
	(451.73,162.96);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (451.73,158.90) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.61,205.15) --
	(494.19,205.15);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.90,204.84) --
	(493.90,205.45);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.90,202.25) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (484.20,197.89) --
	(484.46,197.89);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (484.33,197.75) --
	(484.33,198.03);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (484.33,193.86) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (447.20,157.10) --
	(448.18,157.10);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (447.70,156.31) --
	(447.70,157.86);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (447.70,154.20) {$\bar{r}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (449.46,160.18) --
	(450.41,160.18);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (449.94,159.33) --
	(449.94,160.99);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.94,156.14) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (452.80,166.37) --
	(453.76,166.37);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (453.28,165.56) --
	(453.28,167.15);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (453.28,163.47) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.58,206.10) --
	(495.88,206.10);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.73,205.93) --
	(495.73,206.27);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (495.73,203.20) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (413.75,122.86) --
	(416.46,122.86);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (415.15,121.64) --
	(415.15,124.00);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (415.15,119.63) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (473.31,185.63) --
	(474.12,185.63);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (473.72,185.34) --
	(473.72,185.91);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (473.72,182.72) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (447.11,162.51) --
	(447.38,162.51);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (447.24,162.37) --
	(447.24,162.64);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (447.24,158.47) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (412.33,121.32) --
	(415.08,121.32);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (413.76,120.02) --
	(413.76,122.54);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (413.76,118.42) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (413.42,122.40) --
	(416.04,122.40);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (414.77,121.22) --
	(414.77,123.50);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (414.77,118.36) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (414.94,122.32) --
	(417.75,122.32);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (416.40,120.43) --
	(416.40,124.03);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (416.40,119.41) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (478.09,188.67) --
	(478.40,188.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (478.25,188.53) --
	(478.25,188.81);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (478.25,185.77) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.83,207.98) --
	(493.61,207.98);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.23,207.64) --
	(493.23,208.32);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.23,204.76) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (499.67,211.45) --
	(500.38,211.45);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (500.03,211.16) --
	(500.03,211.73);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.03,208.54) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (509.13,221.32) --
	(509.22,221.32);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (509.18,221.28) --
	(509.18,221.36);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (509.18,217.29) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (484.34,200.66) --
	(485.28,200.66);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (484.82,200.31) --
	(484.82,201.00);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (484.82,197.76) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (490.86,205.99) --
	(491.69,205.99);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.28,205.63) --
	(491.28,206.35);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (491.28,201.96) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (484.57,197.45) --
	(485.81,197.45);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (485.20,197.03) --
	(485.20,197.86);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (485.20,194.55) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (502.49,211.68) --
	(502.76,211.68);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (502.63,211.55) --
	(502.63,211.80);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (502.63,208.78) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (454.59,160.79) --
	(456.16,160.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (455.39,159.54) --
	(455.39,161.97);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (455.39,157.57) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.58,216.74) --
	(504.30,216.74);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.95,216.44) --
	(503.95,217.03);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (503.95,213.83) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.64,197.69) --
	(484.08,197.69);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.86,197.48) --
	(483.86,197.89);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.86,193.65) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (453.00,159.54) --
	(454.63,159.54);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (453.83,158.30) --
	(453.83,160.71);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (453.83,156.64) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (455.02,161.26) --
	(456.59,161.26);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (455.83,160.01) --
	(455.83,162.43);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (455.83,157.23) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (466.45,167.17) --
	(467.37,167.17);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (466.92,165.78) --
	(466.92,168.47);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (466.92,164.27) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.18,215.30) --
	(505.54,215.30);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.36,215.03) --
	(505.36,215.57);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (505.36,212.39) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (465.34,179.33) --
	(466.37,179.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (465.86,178.82) --
	(465.86,179.82);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (465.86,176.10) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (479.70,192.12) --
	(481.20,192.12);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (480.47,191.50) --
	(480.47,192.71);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (480.47,189.21) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (480.95,194.85) --
	(481.08,194.85);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (481.02,194.77) --
	(481.02,194.93);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (481.02,190.81) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (463.17,179.09) --
	(464.27,179.09);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (463.72,178.61) --
	(463.72,179.57);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (463.72,176.19) {$\bar{r}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (464.14,178.17) --
	(465.19,178.17);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (464.67,177.63) --
	(464.67,178.69);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (464.67,174.13) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (469.69,182.13) --
	(470.41,182.13);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (470.05,181.69) --
	(470.05,182.56);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (470.05,179.23) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (494.25,204.55) --
	(494.48,204.55);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (494.37,204.37) --
	(494.37,204.73);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (494.37,201.64) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (482.88,199.90) --
	(483.40,199.90);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.14,199.59) --
	(483.14,200.20);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.14,196.67) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (475.17,187.60) --
	(476.30,187.60);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (475.74,187.13) --
	(475.74,188.05);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (475.74,184.69) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (490.02,204.20) --
	(490.14,204.20);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (490.08,204.12) --
	(490.08,204.27);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (490.08,200.16) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (476.61,194.33) --
	(477.26,194.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (476.94,194.04) --
	(476.94,194.61);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.94,191.42) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (482.77,199.86) --
	(483.29,199.86);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.03,199.55) --
	(483.03,200.15);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.03,195.82) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (479.06,192.69) --
	(479.81,192.69);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (479.44,192.32) --
	(479.44,193.05);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (479.44,189.78) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.86,214.16) --
	(504.20,214.16);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (504.03,213.97) --
	(504.03,214.36);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (504.03,211.26) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (515.10, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.93, 49.80) -- (382.93, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.99, 49.80) -- (426.99, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.04, 49.80) -- (471.04, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.10, 49.80) -- (515.10, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (382.93, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (426.99, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (471.04, 31.87) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (515.10, 31.87) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88,226.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (333.90, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 93.85) -- (333.90, 93.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,137.91) -- (333.90,137.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,181.96) -- (333.90,181.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,226.02) -- (333.90,226.02);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 91.00) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,135.05) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,179.11) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,223.16) {10000};
\end{scope}
\end{tikzpicture}

\end{document}