### 2020-08-31 benchmarking/benchmarkingVM/ESS_round1_real.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-31 17:25:39
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,289.08) rectangle (289.08,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,548.96) {\bfseries cons(\textit{real}) vs. nocons(\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,301.03) {ESS/hr [nocons(\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,433.62) {ESS/hr [cons(\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (  0.00,368.65) --
	(209.51,578.16);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,492.84) --
	(  0.00,209.51);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (  0.00,342.13) --
	(236.03,578.16);

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,519.37) --
	(  0.00,236.03);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (  0.00,315.60) --
	(262.56,578.16);

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,545.89) --
	(  0.00,262.56);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92,376.84) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.28,494.66) --
	(204.36,494.66);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.56,494.38) --
	(201.56,494.93);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (201.56,491.43) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (218.52,499.39) --
	(220.75,499.39);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (219.67,499.07) --
	(219.67,499.70);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (219.67,496.48) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (215.91,501.59) --
	(219.27,501.59);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (217.66,501.55) --
	(217.66,501.62);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (217.66,497.55) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.70,482.86) --
	(203.04,482.86);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.44,482.57) --
	(201.44,483.15);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (201.44,479.96) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.98,485.79) --
	(219.03,485.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (215.04,485.56) --
	(215.04,486.02);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (215.04,481.76) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.84,485.65) --
	(216.47,485.65);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.53,485.34) --
	(211.53,485.96);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.53,482.75) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (223.18,504.72) --
	(223.70,504.72);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (223.44,504.58) --
	(223.44,504.85);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (223.44,501.81) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.15,457.84) --
	(189.63,457.84);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.54,456.70) --
	(182.54,458.92);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.54,454.61) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.33,490.22) --
	(209.40,490.22);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (207.93,489.91) --
	(207.93,490.52);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.93,487.31) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.72,491.74) --
	(206.73,491.74);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.75,491.68) --
	(205.75,491.80);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.75,487.70) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (151.21,440.14) --
	(153.87,440.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (152.58,439.22) --
	(152.58,441.01);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (152.58,437.23) {$\bar{r}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (169.17,441.93) --
	(192.67,441.93);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (184.32,441.01) --
	(184.32,442.81);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.32,437.90) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (174.07,451.54) --
	(196.64,451.54);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.51,450.41) --
	(188.51,452.61);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.51,448.64) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.19,495.41) --
	(212.45,495.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.32,495.32) --
	(212.32,495.49);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.32,492.50) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (167.41,462.49) --
	(183.10,462.49);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (176.82,461.97) --
	(176.82,462.99);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (176.82,459.26) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.58,488.87) --
	(206.74,488.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.34,488.55) --
	(204.34,489.19);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.34,485.97) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.01,488.24) --
	(196.83,488.24);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (195.02,488.18) --
	(195.02,488.29);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.02,484.20) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (161.20,460.37) --
	(165.98,460.37);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (163.74,459.93) --
	(163.74,460.81);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (163.74,457.47) {$\bar{r}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (170.46,461.23) --
	(184.95,461.23);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.04,460.77) --
	(179.04,461.68);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.04,457.20) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (173.29,454.48) --
	(196.02,454.48);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.85,454.03) --
	(187.85,454.92);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.85,451.57) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.76,494.45) --
	(212.07,494.45);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.92,494.36) --
	(211.92,494.55);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.92,491.55) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (139.82,437.49) --
	(142.69,437.49);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (141.31,435.71) --
	(141.31,439.11);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (141.31,434.26) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.50,469.39) --
	(191.19,469.39);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (189.89,469.21) --
	(189.89,469.56);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (189.89,466.49) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (156.82,470.66) --
	(158.12,470.66);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (157.48,470.52) --
	(157.48,470.79);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (157.48,466.62) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (128.56,432.56) --
	(131.74,432.56);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (130.21,431.55) --
	(130.21,433.52);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (130.21,429.66) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (128.30,432.21) --
	(131.22,432.21);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (129.82,431.35) --
	(129.82,433.04);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (129.82,428.18) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (140.69,415.66) --
	(144.04,415.66);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (142.44,413.79) --
	(142.44,417.37);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (142.44,412.76) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.07,476.87) --
	(194.73,476.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.40,476.71) --
	(194.40,477.02);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.40,473.96) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.41,472.87) --
	(214.60,472.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.46,472.13) --
	(209.46,473.58);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (209.46,469.64) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.07,486.45) --
	(205.48,486.45);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.74,486.03) --
	(201.74,486.85);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (201.74,483.54) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (220.79,493.62) --
	(223.70,493.62);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (222.30,493.57) --
	(222.30,493.68);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (222.30,489.59) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.05,472.03) --
	(211.88,472.03);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (207.97,471.72) --
	(207.97,472.32);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.97,469.12) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (220.70,479.63) --
	(226.20,479.63);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (223.64,479.32) --
	(223.64,479.93);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (223.64,475.59) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.79,472.59) --
	(223.65,472.59);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (218.98,472.14) --
	(218.98,473.03);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.98,469.68) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (220.29,499.77) --
	(220.89,499.77);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (220.59,499.63) --
	(220.59,499.92);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (220.59,496.87) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.55,484.09) --
	(218.19,484.09);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.89,483.58) --
	(212.89,484.58);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.89,480.86) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.90,490.53) --
	(218.08,490.53);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (216.56,490.26) --
	(216.56,490.80);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (216.56,487.63) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.42,491.99) --
	(208.44,491.99);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.71,491.95) --
	(203.71,492.04);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (203.71,487.96) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (173.25,458.16) --
	(191.52,458.16);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (184.49,457.57) --
	(184.49,458.73);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.49,455.25) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.50,458.69) --
	(221.94,458.69);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (216.71,458.03) --
	(216.71,459.32);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (216.71,454.65) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.78,461.44) --
	(220.04,461.44);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.54,460.91) --
	(214.54,461.96);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (214.54,458.54) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (221.97,498.57) --
	(222.68,498.57);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (222.33,498.31) --
	(222.33,498.83);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (222.33,495.67) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (159.51,466.05) --
	(176.34,466.05);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (169.72,465.68) --
	(169.72,466.40);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.72,462.82) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.65,471.60) --
	(178.58,471.60);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (174.25,470.81) --
	(174.25,472.36);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.25,468.70) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.28,473.47) --
	(197.14,473.47);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.22,473.41) --
	(193.22,473.54);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.22,469.44) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (163.90,455.11) --
	(172.62,455.11);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.75,454.43) --
	(168.75,455.77);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (168.75,452.21) {$\bar{r}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.98,456.65) --
	(181.96,456.65);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (177.61,455.82) --
	(177.61,457.45);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (177.61,452.62) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (172.32,459.17) --
	(197.10,459.17);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.47,458.78) --
	(188.47,459.55);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.47,456.27) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (210.15,491.65) --
	(211.06,491.65);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (210.61,491.52) --
	(210.61,491.78);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.61,488.74) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.13,470.57) --
	(219.00,470.57);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.29,470.19) --
	(213.29,470.95);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (213.29,467.35) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (174.44,469.22) --
	(182.16,469.22);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (178.68,468.65) --
	(178.68,469.78);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.68,466.32) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.32,480.03) --
	(222.77,480.03);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (218.62,479.97) --
	(218.62,480.08);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.62,475.99) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (186.37,463.32) --
	(203.09,463.32);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (196.50,462.98) --
	(196.50,463.66);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (196.50,460.42) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (224.68,469.61) --
	(230.80,469.61);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (227.99,469.27) --
	(227.99,469.95);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.99,465.58) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.23,467.52) --
	(225.36,467.52);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (220.24,467.13) --
	(220.24,467.89);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (220.24,464.61) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (224.79,499.66) --
	(225.47,499.66);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (225.13,499.50) --
	(225.13,499.81);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (225.13,496.75) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- (226.02,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85,338.88) -- ( 93.85,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91,338.88) -- (137.91,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96,338.88) -- (181.96,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02,338.88) -- (226.02,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 44.82,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,382.93) -- ( 44.82,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,426.99) -- ( 44.82,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,471.04) -- ( 44.82,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,515.10) -- ( 44.82,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,421.05) rectangle (189.67,416.07);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,411.09) rectangle (189.67,406.11);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,401.13) rectangle (189.67,396.15);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,391.17) rectangle (189.67,386.19);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,381.21) rectangle (189.67,376.23);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,371.25) rectangle (189.67,366.27);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,361.29) rectangle (189.67,356.31);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,351.33) rectangle (189.67,346.35);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,415.70) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,405.74) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,395.78) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,385.82) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,375.86) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,365.90) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,355.94) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,345.98) {Moyle};
\end{scope}
\begin{scope}
\path[clip] (289.08,289.08) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,548.96) {\bfseries adapt(\textit{real}) vs. nocons(\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,301.03) {ESS/hr [nocons(\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,433.62) {ESS/hr [adapt(\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] (338.88,338.88) rectangle (528.36,528.36);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (162.66,242.23) --
	(498.59,578.16);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,492.84) --
	(162.66, 83.09) --
	(162.66,242.23);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (162.66,215.71) --
	(525.11,578.16);

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,519.37) --
	(162.66,109.61) --
	(162.66,215.71);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (162.66,189.19) --
	(551.64,578.16);

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,545.89) --
	(162.66,136.14) --
	(162.66,189.19);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (425.98,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (399.46,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (371.28,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00,376.84) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (487.36,505.04) --
	(493.44,505.04);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (490.64,504.79) --
	(490.64,505.29);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (490.64,501.82) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (507.60,507.39) --
	(509.83,507.39);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (508.75,506.95) --
	(508.75,507.82);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (508.75,504.49) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (504.99,511.19) --
	(508.35,511.19);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (506.74,511.13) --
	(506.74,511.25);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (506.74,507.16) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (488.78,493.79) --
	(492.12,493.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (490.52,493.47) --
	(490.52,494.11);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (490.52,490.89) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (499.06,496.31) --
	(508.11,496.31);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (504.12,495.98) --
	(504.12,496.64);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (504.12,492.28) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.92,506.51) --
	(505.55,506.51);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (500.61,506.28) --
	(500.61,506.73);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.61,503.60) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (512.26,512.59) --
	(512.78,512.59);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (512.52,512.47) --
	(512.52,512.71);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (512.52,509.69) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (460.23,468.92) --
	(478.71,468.92);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (471.62,468.08) --
	(471.62,469.72);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (471.62,465.69) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.41,497.49) --
	(498.48,497.49);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (497.01,497.22) --
	(497.01,497.77);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (497.01,494.59) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.80,500.25) --
	(495.81,500.25);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (494.83,500.18) --
	(494.83,500.32);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (494.83,496.21) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (440.29,453.35) --
	(442.95,453.35);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (441.66,452.41) --
	(441.66,454.24);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (441.66,450.44) {$\bar{r}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (458.25,455.34) --
	(481.75,455.34);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (473.40,454.39) --
	(473.40,456.25);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (473.40,451.31) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (463.15,466.37) --
	(485.72,466.37);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (477.59,465.53) --
	(477.59,467.17);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (477.59,463.46) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (501.27,501.14) --
	(501.53,501.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (501.40,501.05) --
	(501.40,501.23);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (501.40,498.24) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (456.49,477.08) --
	(472.18,477.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (465.90,476.61) --
	(465.90,477.54);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (465.90,473.86) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (490.66,493.82) --
	(495.82,493.82);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.42,493.57) --
	(493.42,494.06);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.42,490.91) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (482.09,495.00) --
	(485.91,495.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (484.10,494.93) --
	(484.10,495.07);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (484.10,490.97) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (450.28,470.07) --
	(455.06,470.07);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (452.82,469.69) --
	(452.82,470.46);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (452.82,467.17) {$\bar{r}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (459.54,471.13) --
	(474.03,471.13);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (468.12,470.73) --
	(468.12,471.51);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (468.12,467.09) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (462.37,474.04) --
	(485.10,474.04);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (476.93,473.55) --
	(476.93,474.51);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.93,471.13) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (500.84,500.80) --
	(501.15,500.80);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (501.00,500.70) --
	(501.00,500.90);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (501.00,497.89) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (428.90,470.69) --
	(431.77,470.69);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (430.39,470.18) --
	(430.39,471.19);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (430.39,467.47) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (477.58,468.90) --
	(480.27,468.90);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (478.97,468.61) --
	(478.97,469.18);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (478.97,465.99) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (445.90,475.57) --
	(447.20,475.57);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (446.56,475.45) --
	(446.56,475.69);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (446.56,471.54) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (417.64,442.67) --
	(420.82,442.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (419.29,442.05) --
	(419.29,443.26);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (419.29,439.76) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (417.38,442.55) --
	(420.30,442.55);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (418.90,441.89) --
	(418.90,443.18);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (418.90,438.51) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (429.77,458.76) --
	(433.12,458.76);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (431.52,457.98) --
	(431.52,459.51);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (431.52,455.86) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.15,482.91) --
	(483.81,482.91);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.48,482.83) --
	(483.48,482.99);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.48,480.01) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.49,487.61) --
	(503.68,487.61);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (498.54,486.33) --
	(498.54,488.81);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (498.54,484.39) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (486.15,504.08) --
	(494.56,504.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (490.82,503.75) --
	(490.82,504.41);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (490.82,501.17) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (509.87,516.22) --
	(512.78,516.22);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (511.38,516.18) --
	(511.38,516.26);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (511.38,512.18) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.13,497.67) --
	(500.96,497.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (497.05,497.33) --
	(497.05,498.01);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (497.05,494.77) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (509.78,502.94) --
	(515.28,502.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (512.72,502.68) --
	(512.72,503.19);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (512.72,498.90) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (501.87,491.24) --
	(512.73,491.24);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (508.06,490.94) --
	(508.06,491.55);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (508.06,488.34) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (509.37,507.70) --
	(509.97,507.70);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (509.67,507.59) --
	(509.67,507.81);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (509.67,504.80) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (494.63,499.71) --
	(507.27,499.71);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (501.97,499.35) --
	(501.97,500.06);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (501.97,496.48) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.98,498.95) --
	(507.16,498.95);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.64,498.66) --
	(505.64,499.24);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (505.64,496.05) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (486.50,503.67) --
	(497.52,503.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.79,503.61) --
	(492.79,503.74);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (492.79,499.64) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (462.33,471.10) --
	(480.60,471.10);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (473.57,470.46) --
	(473.57,471.72);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (473.57,468.20) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (498.58,471.53) --
	(511.02,471.53);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.79,470.90) --
	(505.79,472.14);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (505.79,467.50) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.86,483.58) --
	(509.12,483.58);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.62,483.12) --
	(503.62,484.03);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (503.62,480.67) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (511.05,512.59) --
	(511.76,512.59);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (511.41,512.42) --
	(511.41,512.76);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (511.41,509.69) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (448.59,481.02) --
	(465.42,481.02);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (458.80,480.59) --
	(458.80,481.44);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (458.80,477.79) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (457.73,482.93) --
	(467.66,482.93);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (463.33,481.27) --
	(463.33,484.45);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (463.33,480.02) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (477.36,489.01) --
	(486.22,489.01);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (482.30,488.92) --
	(482.30,489.11);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (482.30,484.98) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (452.98,473.57) --
	(461.70,473.57);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (457.83,473.24) --
	(457.83,473.89);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (457.83,470.67) {$\bar{r}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (461.06,472.97) --
	(471.04,472.97);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (466.69,472.65) --
	(466.69,473.28);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (466.69,468.93) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (461.40,480.98) --
	(486.18,480.98);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (477.55,480.59) --
	(477.55,481.36);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (477.55,478.08) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (499.23,501.33) --
	(500.14,501.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (499.69,501.22) --
	(499.69,501.45);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (499.69,498.43) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (494.21,487.86) --
	(508.08,487.86);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (502.37,487.42) --
	(502.37,488.29);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (502.37,484.64) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (463.52,484.00) --
	(471.24,484.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (467.76,483.35) --
	(467.76,484.63);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (467.76,481.10) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (502.40,500.91) --
	(511.85,500.91);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (507.70,500.84) --
	(507.70,500.97);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (507.70,496.87) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (475.45,490.16) --
	(492.17,490.16);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (485.58,489.94) --
	(485.58,490.38);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (485.58,487.26) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (513.76,495.79) --
	(519.88,495.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (517.07,495.53) --
	(517.07,496.04);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (517.07,491.75) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (502.31,490.00) --
	(514.44,490.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (509.32,489.48) --
	(509.32,490.49);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (509.32,487.09) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (513.87,512.17) --
	(514.55,512.17);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (514.21,511.97) --
	(514.21,512.37);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (514.21,509.26) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (515.10,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.93,338.88) -- (382.93,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.99,338.88) -- (426.99,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.04,338.88) -- (471.04,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.10,338.88) -- (515.10,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (382.93,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (426.99,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (471.04,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (515.10,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (333.90,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,382.93) -- (333.90,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,426.99) -- (333.90,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,471.04) -- (333.90,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,515.10) -- (333.90,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,259.88) {\bfseries adapt(\textit{real}) vs. cons(\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54, 11.95) {ESS/hr [cons(\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,144.54) {ESS/hr [adapt(\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80, 49.80) rectangle (239.28,239.28);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (  0.00, 79.57) --
	(283.33,362.91) --
	(283.33,203.76) --
	( 79.57,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (  0.00, 53.05) --
	(283.33,336.38) --
	(283.33,230.29) --
	( 53.05,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (  0.00, 26.52) --
	(283.33,309.86) --
	(283.33,256.81) --
	( 26.52,  0.00);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,140.80) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,114.28) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92, 87.76) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28,239.28);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.30,215.96) --
	(205.85,215.96);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.58,215.71) --
	(205.58,216.21);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.58,212.74) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.99,218.31) --
	(210.62,218.31);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (210.31,217.87) --
	(210.31,218.74);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.31,215.41) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.47,222.11) --
	(212.54,222.11);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.51,222.05) --
	(212.51,222.17);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.51,218.08) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.49,204.71) --
	(194.07,204.71);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.78,204.39) --
	(193.78,205.03);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.78,201.81) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (196.48,207.23) --
	(196.94,207.23);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (196.71,206.90) --
	(196.71,207.56);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (196.71,203.20) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (196.26,217.43) --
	(196.88,217.43);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (196.57,217.20) --
	(196.57,217.65);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (196.57,214.52) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (215.50,223.51) --
	(215.77,223.51);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (215.64,223.39) --
	(215.64,223.63);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (215.64,220.61) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (167.62,179.84) --
	(169.84,179.84);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.76,179.00) --
	(168.76,180.64);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (168.76,176.61) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.83,208.41) --
	(201.44,208.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.14,208.14) --
	(201.14,208.69);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (201.14,205.51) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.60,211.17) --
	(202.72,211.17);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.66,211.10) --
	(202.66,211.24);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.66,207.13) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (150.14,164.27) --
	(151.93,164.27);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (151.06,163.33) --
	(151.06,165.16);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (151.06,161.36) {$\bar{r}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (151.93,166.26) --
	(153.73,166.26);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (152.85,165.31) --
	(152.85,167.17);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (152.85,162.23) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (161.33,177.29) --
	(163.53,177.29);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (162.46,176.45) --
	(162.46,178.09);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (162.46,174.38) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.24,212.06) --
	(206.41,212.06);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.33,211.97) --
	(206.33,212.15);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (206.33,209.16) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (172.89,188.00) --
	(173.91,188.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (173.41,187.53) --
	(173.41,188.46);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (173.41,184.78) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.47,204.74) --
	(200.11,204.74);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.79,204.49) --
	(199.79,204.98);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.79,201.83) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.10,205.92) --
	(199.21,205.92);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.16,205.85) --
	(199.16,205.99);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.16,201.89) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (170.85,180.99) --
	(171.73,180.99);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.29,180.61) --
	(171.29,181.38);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (171.29,178.09) {$\bar{r}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.69,182.05) --
	(172.60,182.05);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (172.15,181.65) --
	(172.15,182.43);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (172.15,178.01) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (164.95,184.96) --
	(165.84,184.96);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (165.40,184.47) --
	(165.40,185.43);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (165.40,182.05) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.28,211.72) --
	(205.47,211.72);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.37,211.62) --
	(205.37,211.82);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.37,208.81) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (146.63,181.61) --
	(150.03,181.61);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (148.41,181.10) --
	(148.41,182.11);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (148.41,178.39) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.13,179.82) --
	(180.48,179.82);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.31,179.53) --
	(180.31,180.10);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (180.31,176.91) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.44,186.49) --
	(181.71,186.49);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.58,186.37) --
	(181.58,186.61);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.58,182.46) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (142.47,153.59) --
	(144.44,153.59);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (143.48,152.97) --
	(143.48,154.18);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (143.48,150.68) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (142.27,153.47) --
	(143.96,153.47);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (143.13,152.81) --
	(143.13,154.10);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (143.13,149.43) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (124.71,169.68) --
	(128.29,169.68);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (126.58,168.90) --
	(126.58,170.43);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (126.58,166.78) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.63,193.83) --
	(187.94,193.83);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.79,193.75) --
	(187.79,193.91);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.79,190.93) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (183.05,198.53) --
	(184.50,198.53);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (183.79,197.25) --
	(183.79,199.73);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.79,195.31) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (196.95,215.00) --
	(197.77,215.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.37,214.67) --
	(197.37,215.33);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.37,212.09) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.49,227.14) --
	(204.60,227.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.54,227.10) --
	(204.54,227.18);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.54,223.10) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.64,208.59) --
	(183.24,208.59);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.95,208.25) --
	(182.95,208.93);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.95,205.69) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (190.24,213.86) --
	(190.85,213.86);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (190.55,213.60) --
	(190.55,214.11);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (190.55,209.82) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (183.06,202.16) --
	(183.95,202.16);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (183.51,201.86) --
	(183.51,202.47);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.51,199.26) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (210.55,218.62) --
	(210.84,218.62);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (210.69,218.51) --
	(210.69,218.73);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.69,215.72) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.50,210.63) --
	(195.50,210.63);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (195.01,210.27) --
	(195.01,210.98);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (195.01,207.40) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.18,209.87) --
	(201.72,209.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.45,209.58) --
	(201.45,210.16);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (201.45,206.97) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.87,214.59) --
	(202.96,214.59);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.91,214.53) --
	(202.91,214.66);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.91,210.56) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.49,182.02) --
	(169.65,182.02);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (169.08,181.38) --
	(169.08,182.64);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.08,179.12) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.95,182.45) --
	(170.24,182.45);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (169.61,181.82) --
	(169.61,183.06);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.61,178.42) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.83,194.50) --
	(172.88,194.50);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (172.36,194.04) --
	(172.36,194.95);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (172.36,191.59) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.23,223.51) --
	(209.75,223.51);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.49,223.34) --
	(209.49,223.68);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (209.49,220.61) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (176.60,191.94) --
	(177.32,191.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (176.97,191.51) --
	(176.97,192.36);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (176.97,188.71) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.73,193.85) --
	(183.28,193.85);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.52,192.19) --
	(182.52,195.37);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.52,190.94) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (184.33,199.93) --
	(184.46,199.93);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (184.39,199.84) --
	(184.39,200.03);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.39,195.90) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (165.35,184.49) --
	(166.69,184.49);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (166.03,184.16) --
	(166.03,184.81);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (166.03,181.59) {$\bar{r}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (166.74,183.89) --
	(168.37,183.89);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (167.57,183.57) --
	(167.57,184.20);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (167.57,179.85) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (169.70,191.90) --
	(170.47,191.90);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (170.09,191.51) --
	(170.09,192.28);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (170.09,189.00) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.44,212.25) --
	(202.70,212.25);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.57,212.14) --
	(202.57,212.37);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.57,209.35) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.11,198.78) --
	(181.87,198.78);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.49,198.34) --
	(181.49,199.21);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.49,195.56) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.57,194.92) --
	(180.70,194.92);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.14,194.27) --
	(180.14,195.55);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (180.14,192.02) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (190.89,211.83) --
	(191.00,211.83);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (190.95,211.76) --
	(190.95,211.89);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (190.95,207.79) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (173.90,201.08) --
	(174.58,201.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (174.24,200.86) --
	(174.24,201.30);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.24,198.18) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.19,206.71) --
	(180.87,206.71);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.53,206.45) --
	(180.53,206.96);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (180.53,202.67) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (178.05,200.92) --
	(178.81,200.92);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (178.44,200.40) --
	(178.44,201.41);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.44,198.01) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (210.42,223.09) --
	(210.73,223.09);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (210.58,222.89) --
	(210.58,223.29);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.58,220.18) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- (226.02, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85, 49.80) -- ( 93.85, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91, 49.80) -- (137.91, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96, 49.80) -- (181.96, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02, 49.80) -- (226.02, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96, 31.87) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02, 31.87) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80,226.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 44.82, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 93.85) -- ( 44.82, 93.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,137.91) -- ( 44.82,137.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,181.96) -- ( 44.82,181.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,226.02) -- ( 44.82,226.02);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 91.00) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,135.05) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,179.11) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,223.16) {10000};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,259.88) {\bfseries \texttt{AdaptiveOperatorSampler} weights};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62, 11.95) {Sequence length $L$ (kb)};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,144.54) {Learned operator weight};
\end{scope}
\begin{scope}
\path[clip] (338.88, 49.80) rectangle (528.36,239.28);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (356.88,165.22) circle (  1.87);

\path[fill=fillColor] (356.88, 79.54) --
	(359.40, 75.18) --
	(354.37, 75.18) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (394.78,165.34) circle (  1.87);

\path[fill=fillColor] (394.78, 79.41) --
	(397.29, 75.05) --
	(392.26, 75.05) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (390.04,192.10) circle (  1.87);

\path[fill=fillColor] (390.04, 71.80) --
	(392.55, 67.44) --
	(387.52, 67.44) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (404.25,216.47) circle (  1.87);

\path[fill=fillColor] (404.25, 59.60) --
	(406.77, 55.24) --
	(401.74, 55.24) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (359.72,107.37) circle (  1.87);

\path[fill=fillColor] (359.72, 91.15) --
	(362.24, 86.80) --
	(357.21, 86.80) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (433.62,203.66) circle (  1.87);

\path[fill=fillColor] (433.62, 68.80) --
	(436.14, 64.44) --
	(431.10, 64.44) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (359.72,137.48) circle (  1.87);

\path[fill=fillColor] (359.72, 88.39) --
	(362.24, 84.03) --
	(357.21, 84.03) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (346.46,129.05) circle (  1.87);

\path[fill=fillColor] (346.46, 89.93) --
	(348.97, 85.57) --
	(343.94, 85.57) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (435.32,150.27) circle (  2.15);

\path[fill=fillColor] (435.32,142.15) --
	(438.21,137.14) --
	(432.42,137.14) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,146.98) {\texttt{ConstantDistance}};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (443.91,135.53) {\texttt{RandomWalk}};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (338.88,117.32) --
	(339.26,117.75) --
	(339.64,118.18) --
	(340.02,118.61) --
	(340.40,119.05) --
	(340.78,119.48) --
	(341.16,119.92) --
	(341.54,120.35) --
	(341.92,120.79) --
	(342.30,121.23) --
	(342.68,121.67) --
	(343.06,122.11) --
	(343.44,122.55) --
	(343.82,123.00) --
	(344.20,123.44) --
	(344.58,123.89) --
	(344.96,124.33) --
	(345.34,124.78) --
	(345.71,125.23) --
	(346.09,125.68) --
	(346.47,126.13) --
	(346.85,126.58) --
	(347.23,127.03) --
	(347.61,127.48) --
	(347.99,127.93) --
	(348.37,128.39) --
	(348.75,128.84) --
	(349.13,129.30) --
	(349.51,129.75) --
	(349.89,130.21) --
	(350.27,130.67) --
	(350.65,131.13) --
	(351.03,131.59) --
	(351.41,132.05) --
	(351.79,132.51) --
	(352.17,132.97) --
	(352.55,133.43) --
	(352.93,133.89) --
	(353.31,134.35) --
	(353.69,134.81) --
	(354.07,135.28) --
	(354.45,135.74) --
	(354.83,136.20) --
	(355.21,136.67) --
	(355.59,137.13) --
	(355.97,137.60) --
	(356.35,138.06) --
	(356.73,138.53) --
	(357.11,139.00) --
	(357.49,139.46) --
	(357.87,139.93) --
	(358.25,140.40) --
	(358.63,140.86) --
	(359.01,141.33) --
	(359.38,141.80) --
	(359.76,142.27) --
	(360.14,142.73) --
	(360.52,143.20) --
	(360.90,143.67) --
	(361.28,144.14) --
	(361.66,144.61) --
	(362.04,145.07) --
	(362.42,145.54) --
	(362.80,146.01) --
	(363.18,146.48) --
	(363.56,146.94) --
	(363.94,147.41) --
	(364.32,147.88) --
	(364.70,148.35) --
	(365.08,148.81) --
	(365.46,149.28) --
	(365.84,149.75) --
	(366.22,150.21) --
	(366.60,150.68) --
	(366.98,151.14) --
	(367.36,151.61) --
	(367.74,152.08) --
	(368.12,152.54) --
	(368.50,153.00) --
	(368.88,153.47) --
	(369.26,153.93) --
	(369.64,154.40) --
	(370.02,154.86) --
	(370.40,155.32) --
	(370.78,155.78) --
	(371.16,156.24) --
	(371.54,156.70) --
	(371.92,157.16) --
	(372.30,157.62) --
	(372.68,158.08) --
	(373.05,158.54) --
	(373.43,159.00) --
	(373.81,159.45) --
	(374.19,159.91) --
	(374.57,160.36) --
	(374.95,160.82) --
	(375.33,161.27) --
	(375.71,161.73) --
	(376.09,162.18) --
	(376.47,162.63) --
	(376.85,163.08) --
	(377.23,163.53) --
	(377.61,163.98) --
	(377.99,164.42) --
	(378.37,164.87) --
	(378.75,165.32) --
	(379.13,165.76) --
	(379.51,166.21) --
	(379.89,166.65) --
	(380.27,167.09) --
	(380.65,167.53) --
	(381.03,167.97) --
	(381.41,168.41) --
	(381.79,168.85) --
	(382.17,169.28) --
	(382.55,169.72) --
	(382.93,170.15) --
	(383.31,170.59) --
	(383.69,171.02) --
	(384.07,171.45) --
	(384.45,171.88) --
	(384.83,172.31) --
	(385.21,172.74) --
	(385.59,173.16) --
	(385.97,173.59) --
	(386.34,174.01) --
	(386.72,174.43) --
	(387.10,174.85) --
	(387.48,175.27) --
	(387.86,175.69) --
	(388.24,176.11) --
	(388.62,176.52) --
	(389.00,176.93) --
	(389.38,177.35) --
	(389.76,177.76) --
	(390.14,178.17) --
	(390.52,178.58) --
	(390.90,178.98) --
	(391.28,179.39) --
	(391.66,179.79) --
	(392.04,180.19) --
	(392.42,180.60) --
	(392.80,180.99) --
	(393.18,181.39) --
	(393.56,181.79) --
	(393.94,182.18) --
	(394.32,182.58) --
	(394.70,182.97) --
	(395.08,183.36) --
	(395.46,183.75) --
	(395.84,184.13) --
	(396.22,184.52) --
	(396.60,184.90) --
	(396.98,185.29) --
	(397.36,185.67) --
	(397.74,186.04) --
	(398.12,186.42) --
	(398.50,186.80) --
	(398.88,187.17) --
	(399.26,187.54) --
	(399.64,187.91) --
	(400.01,188.28) --
	(400.39,188.65) --
	(400.77,189.02) --
	(401.15,189.38) --
	(401.53,189.74) --
	(401.91,190.10) --
	(402.29,190.46) --
	(402.67,190.82) --
	(403.05,191.17) --
	(403.43,191.53) --
	(403.81,191.88) --
	(404.19,192.23) --
	(404.57,192.58) --
	(404.95,192.93) --
	(405.33,193.27) --
	(405.71,193.61) --
	(406.09,193.96) --
	(406.47,194.30) --
	(406.85,194.63) --
	(407.23,194.97) --
	(407.61,195.30) --
	(407.99,195.64) --
	(408.37,195.97) --
	(408.75,196.30) --
	(409.13,196.62) --
	(409.51,196.95) --
	(409.89,197.27) --
	(410.27,197.60) --
	(410.65,197.92) --
	(411.03,198.23) --
	(411.41,198.55) --
	(411.79,198.87) --
	(412.17,199.18) --
	(412.55,199.49) --
	(412.93,199.80) --
	(413.31,200.11) --
	(413.68,200.41) --
	(414.06,200.72) --
	(414.44,201.02) --
	(414.82,201.32) --
	(415.20,201.62) --
	(415.58,201.92) --
	(415.96,202.21) --
	(416.34,202.51) --
	(416.72,202.80) --
	(417.10,203.09) --
	(417.48,203.38) --
	(417.86,203.66) --
	(418.24,203.95) --
	(418.62,204.23) --
	(419.00,204.51) --
	(419.38,204.79) --
	(419.76,205.07) --
	(420.14,205.35) --
	(420.52,205.62) --
	(420.90,205.89) --
	(421.28,206.16) --
	(421.66,206.43) --
	(422.04,206.70) --
	(422.42,206.97) --
	(422.80,207.23) --
	(423.18,207.49) --
	(423.56,207.75) --
	(423.94,208.01) --
	(424.32,208.27) --
	(424.70,208.52) --
	(425.08,208.78) --
	(425.46,209.03) --
	(425.84,209.28) --
	(426.22,209.53) --
	(426.60,209.77) --
	(426.97,210.02) --
	(427.35,210.26) --
	(427.73,210.50) --
	(428.11,210.75) --
	(428.49,210.98) --
	(428.87,211.22) --
	(429.25,211.46) --
	(429.63,211.69) --
	(430.01,211.92) --
	(430.39,212.15) --
	(430.77,212.38) --
	(431.15,212.61) --
	(431.53,212.83) --
	(431.91,213.06) --
	(432.29,213.28) --
	(432.67,213.50) --
	(433.05,213.72) --
	(433.43,213.94) --
	(433.81,214.15) --
	(434.19,214.37) --
	(434.57,214.58) --
	(434.95,214.79) --
	(435.33,215.00) --
	(435.71,215.21) --
	(436.09,215.42) --
	(436.47,215.62) --
	(436.85,215.83) --
	(437.23,216.03) --
	(437.61,216.23) --
	(437.99,216.43) --
	(438.37,216.63) --
	(438.75,216.82) --
	(439.13,217.02) --
	(439.51,217.21) --
	(439.89,217.40) --
	(440.27,217.59) --
	(440.64,217.78) --
	(441.02,217.97) --
	(441.40,218.16) --
	(441.78,218.34) --
	(442.16,218.52) --
	(442.54,218.71) --
	(442.92,218.89) --
	(443.30,219.06) --
	(443.68,219.24) --
	(444.06,219.42) --
	(444.44,219.59) --
	(444.82,219.77) --
	(445.20,219.94) --
	(445.58,220.11) --
	(445.96,220.28) --
	(446.34,220.45) --
	(446.72,220.61) --
	(447.10,220.78) --
	(447.48,220.94) --
	(447.86,221.11) --
	(448.24,221.27) --
	(448.62,221.43) --
	(449.00,221.59) --
	(449.38,221.75) --
	(449.76,221.90) --
	(450.14,222.06) --
	(450.52,222.21) --
	(450.90,222.37) --
	(451.28,222.52) --
	(451.66,222.67) --
	(452.04,222.82) --
	(452.42,222.96) --
	(452.80,223.11) --
	(453.18,223.26) --
	(453.56,223.40) --
	(453.93,223.54) --
	(454.31,223.69) --
	(454.69,223.83) --
	(455.07,223.97) --
	(455.45,224.10) --
	(455.83,224.24) --
	(456.21,224.38) --
	(456.59,224.51) --
	(456.97,224.65) --
	(457.35,224.78) --
	(457.73,224.91) --
	(458.11,225.04) --
	(458.49,225.17) --
	(458.87,225.30) --
	(459.25,225.43) --
	(459.63,225.55) --
	(460.01,225.68) --
	(460.39,225.80) --
	(460.77,225.93) --
	(461.15,226.05) --
	(461.53,226.17) --
	(461.91,226.29) --
	(462.29,226.41) --
	(462.67,226.53) --
	(463.05,226.64) --
	(463.43,226.76) --
	(463.81,226.87) --
	(464.19,226.99) --
	(464.57,227.10) --
	(464.95,227.21) --
	(465.33,227.32) --
	(465.71,227.43) --
	(466.09,227.54) --
	(466.47,227.65) --
	(466.85,227.76) --
	(467.23,227.87) --
	(467.60,227.97) --
	(467.98,228.08) --
	(468.36,228.18) --
	(468.74,228.28) --
	(469.12,228.38) --
	(469.50,228.49) --
	(469.88,228.59) --
	(470.26,228.68) --
	(470.64,228.78) --
	(471.02,228.88) --
	(471.40,228.98) --
	(471.78,229.07) --
	(472.16,229.17) --
	(472.54,229.26) --
	(472.92,229.36) --
	(473.30,229.45) --
	(473.68,229.54) --
	(474.06,229.63) --
	(474.44,229.72) --
	(474.82,229.81) --
	(475.20,229.90) --
	(475.58,229.99) --
	(475.96,230.07) --
	(476.34,230.16) --
	(476.72,230.24) --
	(477.10,230.33) --
	(477.48,230.41) --
	(477.86,230.50) --
	(478.24,230.58) --
	(478.62,230.66) --
	(479.00,230.74) --
	(479.38,230.82) --
	(479.76,230.90) --
	(480.14,230.98) --
	(480.52,231.06) --
	(480.90,231.13) --
	(481.27,231.21) --
	(481.65,231.29) --
	(482.03,231.36) --
	(482.41,231.44) --
	(482.79,231.51) --
	(483.17,231.58) --
	(483.55,231.66) --
	(483.93,231.73) --
	(484.31,231.80) --
	(484.69,231.87) --
	(485.07,231.94) --
	(485.45,232.01) --
	(485.83,232.08) --
	(486.21,232.15) --
	(486.59,232.21) --
	(486.97,232.28) --
	(487.35,232.35) --
	(487.73,232.41) --
	(488.11,232.48) --
	(488.49,232.54) --
	(488.87,232.61) --
	(489.25,232.67) --
	(489.63,232.73) --
	(490.01,232.79) --
	(490.39,232.86) --
	(490.77,232.92) --
	(491.15,232.98) --
	(491.53,233.04) --
	(491.91,233.10) --
	(492.29,233.16) --
	(492.67,233.21) --
	(493.05,233.27) --
	(493.43,233.33) --
	(493.81,233.39) --
	(494.19,233.44) --
	(494.56,233.50) --
	(494.94,233.55) --
	(495.32,233.61) --
	(495.70,233.66) --
	(496.08,233.72) --
	(496.46,233.77) --
	(496.84,233.82) --
	(497.22,233.87) --
	(497.60,233.92) --
	(497.98,233.98) --
	(498.36,234.03) --
	(498.74,234.08) --
	(499.12,234.13) --
	(499.50,234.18) --
	(499.88,234.22) --
	(500.26,234.27) --
	(500.64,234.32) --
	(501.02,234.37) --
	(501.40,234.42) --
	(501.78,234.46) --
	(502.16,234.51) --
	(502.54,234.55) --
	(502.92,234.60) --
	(503.30,234.64) --
	(503.68,234.69) --
	(504.06,234.73) --
	(504.44,234.78) --
	(504.82,234.82) --
	(505.20,234.86) --
	(505.58,234.90) --
	(505.96,234.95) --
	(506.34,234.99) --
	(506.72,235.03) --
	(507.10,235.07) --
	(507.48,235.11) --
	(507.86,235.15) --
	(508.23,235.19) --
	(508.61,235.23) --
	(508.99,235.27) --
	(509.37,235.31) --
	(509.75,235.35) --
	(510.13,235.38) --
	(510.51,235.42) --
	(510.89,235.46) --
	(511.27,235.50) --
	(511.65,235.53) --
	(512.03,235.57) --
	(512.41,235.60) --
	(512.79,235.64) --
	(513.17,235.67) --
	(513.55,235.71) --
	(513.93,235.74) --
	(514.31,235.78) --
	(514.69,235.81) --
	(515.07,235.84) --
	(515.45,235.88) --
	(515.83,235.91) --
	(516.21,235.94) --
	(516.59,235.98) --
	(516.97,236.01) --
	(517.35,236.04) --
	(517.73,236.07) --
	(518.11,236.10) --
	(518.49,236.13) --
	(518.87,236.16) --
	(519.25,236.19) --
	(519.63,236.22) --
	(520.01,236.25) --
	(520.39,236.28) --
	(520.77,236.31) --
	(521.15,236.34) --
	(521.53,236.37) --
	(521.90,236.40) --
	(522.28,236.42) --
	(522.66,236.45) --
	(523.04,236.48) --
	(523.42,236.51) --
	(523.80,236.53) --
	(524.18,236.56) --
	(524.56,236.59) --
	(524.94,236.61) --
	(525.32,236.64) --
	(525.70,236.66) --
	(526.08,236.69) --
	(526.46,236.71) --
	(526.84,236.74) --
	(527.22,236.76) --
	(527.60,236.79) --
	(527.98,236.81) --
	(528.36,236.84);

\path[draw=drawColor,draw opacity=0.60,line width= 0.4pt,line join=round,line cap=round] (338.88, 91.23) --
	(339.26, 91.04) --
	(339.64, 90.86) --
	(340.02, 90.67) --
	(340.40, 90.49) --
	(340.78, 90.30) --
	(341.16, 90.12) --
	(341.54, 89.94) --
	(341.92, 89.76) --
	(342.30, 89.58) --
	(342.68, 89.40) --
	(343.06, 89.22) --
	(343.44, 89.04) --
	(343.82, 88.86) --
	(344.20, 88.68) --
	(344.58, 88.50) --
	(344.96, 88.33) --
	(345.34, 88.15) --
	(345.71, 87.97) --
	(346.09, 87.80) --
	(346.47, 87.62) --
	(346.85, 87.45) --
	(347.23, 87.28) --
	(347.61, 87.10) --
	(347.99, 86.93) --
	(348.37, 86.76) --
	(348.75, 86.59) --
	(349.13, 86.42) --
	(349.51, 86.25) --
	(349.89, 86.08) --
	(350.27, 85.91) --
	(350.65, 85.75) --
	(351.03, 85.58) --
	(351.41, 85.41) --
	(351.79, 85.25) --
	(352.17, 85.08) --
	(352.55, 84.92) --
	(352.93, 84.75) --
	(353.31, 84.59) --
	(353.69, 84.43) --
	(354.07, 84.26) --
	(354.45, 84.10) --
	(354.83, 83.94) --
	(355.21, 83.78) --
	(355.59, 83.62) --
	(355.97, 83.46) --
	(356.35, 83.30) --
	(356.73, 83.14) --
	(357.11, 82.99) --
	(357.49, 82.83) --
	(357.87, 82.67) --
	(358.25, 82.52) --
	(358.63, 82.36) --
	(359.01, 82.21) --
	(359.38, 82.05) --
	(359.76, 81.90) --
	(360.14, 81.74) --
	(360.52, 81.59) --
	(360.90, 81.44) --
	(361.28, 81.29) --
	(361.66, 81.14) --
	(362.04, 80.99) --
	(362.42, 80.84) --
	(362.80, 80.69) --
	(363.18, 80.54) --
	(363.56, 80.39) --
	(363.94, 80.25) --
	(364.32, 80.10) --
	(364.70, 79.95) --
	(365.08, 79.81) --
	(365.46, 79.66) --
	(365.84, 79.52) --
	(366.22, 79.38) --
	(366.60, 79.23) --
	(366.98, 79.09) --
	(367.36, 78.95) --
	(367.74, 78.81) --
	(368.12, 78.67) --
	(368.50, 78.52) --
	(368.88, 78.38) --
	(369.26, 78.25) --
	(369.64, 78.11) --
	(370.02, 77.97) --
	(370.40, 77.83) --
	(370.78, 77.69) --
	(371.16, 77.56) --
	(371.54, 77.42) --
	(371.92, 77.29) --
	(372.30, 77.15) --
	(372.68, 77.02) --
	(373.05, 76.88) --
	(373.43, 76.75) --
	(373.81, 76.62) --
	(374.19, 76.49) --
	(374.57, 76.35) --
	(374.95, 76.22) --
	(375.33, 76.09) --
	(375.71, 75.96) --
	(376.09, 75.83) --
	(376.47, 75.70) --
	(376.85, 75.58) --
	(377.23, 75.45) --
	(377.61, 75.32) --
	(377.99, 75.19) --
	(378.37, 75.07) --
	(378.75, 74.94) --
	(379.13, 74.82) --
	(379.51, 74.69) --
	(379.89, 74.57) --
	(380.27, 74.45) --
	(380.65, 74.32) --
	(381.03, 74.20) --
	(381.41, 74.08) --
	(381.79, 73.96) --
	(382.17, 73.84) --
	(382.55, 73.71) --
	(382.93, 73.60) --
	(383.31, 73.48) --
	(383.69, 73.36) --
	(384.07, 73.24) --
	(384.45, 73.12) --
	(384.83, 73.00) --
	(385.21, 72.89) --
	(385.59, 72.77) --
	(385.97, 72.65) --
	(386.34, 72.54) --
	(386.72, 72.42) --
	(387.10, 72.31) --
	(387.48, 72.20) --
	(387.86, 72.08) --
	(388.24, 71.97) --
	(388.62, 71.86) --
	(389.00, 71.75) --
	(389.38, 71.63) --
	(389.76, 71.52) --
	(390.14, 71.41) --
	(390.52, 71.30) --
	(390.90, 71.19) --
	(391.28, 71.09) --
	(391.66, 70.98) --
	(392.04, 70.87) --
	(392.42, 70.76) --
	(392.80, 70.65) --
	(393.18, 70.55) --
	(393.56, 70.44) --
	(393.94, 70.34) --
	(394.32, 70.23) --
	(394.70, 70.13) --
	(395.08, 70.02) --
	(395.46, 69.92) --
	(395.84, 69.82) --
	(396.22, 69.71) --
	(396.60, 69.61) --
	(396.98, 69.51) --
	(397.36, 69.41) --
	(397.74, 69.31) --
	(398.12, 69.21) --
	(398.50, 69.11) --
	(398.88, 69.01) --
	(399.26, 68.91) --
	(399.64, 68.81) --
	(400.01, 68.71) --
	(400.39, 68.61) --
	(400.77, 68.52) --
	(401.15, 68.42) --
	(401.53, 68.32) --
	(401.91, 68.23) --
	(402.29, 68.13) --
	(402.67, 68.04) --
	(403.05, 67.94) --
	(403.43, 67.85) --
	(403.81, 67.76) --
	(404.19, 67.66) --
	(404.57, 67.57) --
	(404.95, 67.48) --
	(405.33, 67.39) --
	(405.71, 67.29) --
	(406.09, 67.20) --
	(406.47, 67.11) --
	(406.85, 67.02) --
	(407.23, 66.93) --
	(407.61, 66.84) --
	(407.99, 66.75) --
	(408.37, 66.66) --
	(408.75, 66.58) --
	(409.13, 66.49) --
	(409.51, 66.40) --
	(409.89, 66.31) --
	(410.27, 66.23) --
	(410.65, 66.14) --
	(411.03, 66.06) --
	(411.41, 65.97) --
	(411.79, 65.89) --
	(412.17, 65.80) --
	(412.55, 65.72) --
	(412.93, 65.63) --
	(413.31, 65.55) --
	(413.68, 65.47) --
	(414.06, 65.39) --
	(414.44, 65.30) --
	(414.82, 65.22) --
	(415.20, 65.14) --
	(415.58, 65.06) --
	(415.96, 64.98) --
	(416.34, 64.90) --
	(416.72, 64.82) --
	(417.10, 64.74) --
	(417.48, 64.66) --
	(417.86, 64.58) --
	(418.24, 64.50) --
	(418.62, 64.43) --
	(419.00, 64.35) --
	(419.38, 64.27) --
	(419.76, 64.20) --
	(420.14, 64.12) --
	(420.52, 64.04) --
	(420.90, 63.97) --
	(421.28, 63.89) --
	(421.66, 63.82) --
	(422.04, 63.74) --
	(422.42, 63.67) --
	(422.80, 63.60) --
	(423.18, 63.52) --
	(423.56, 63.45) --
	(423.94, 63.38) --
	(424.32, 63.30) --
	(424.70, 63.23) --
	(425.08, 63.16) --
	(425.46, 63.09) --
	(425.84, 63.02) --
	(426.22, 62.95) --
	(426.60, 62.88) --
	(426.97, 62.81) --
	(427.35, 62.74) --
	(427.73, 62.67) --
	(428.11, 62.60) --
	(428.49, 62.53) --
	(428.87, 62.46) --
	(429.25, 62.40) --
	(429.63, 62.33) --
	(430.01, 62.26) --
	(430.39, 62.20) --
	(430.77, 62.13) --
	(431.15, 62.06) --
	(431.53, 62.00) --
	(431.91, 61.93) --
	(432.29, 61.87) --
	(432.67, 61.80) --
	(433.05, 61.74) --
	(433.43, 61.67) --
	(433.81, 61.61) --
	(434.19, 61.55) --
	(434.57, 61.48) --
	(434.95, 61.42) --
	(435.33, 61.36) --
	(435.71, 61.29) --
	(436.09, 61.23) --
	(436.47, 61.17) --
	(436.85, 61.11) --
	(437.23, 61.05) --
	(437.61, 60.99) --
	(437.99, 60.93) --
	(438.37, 60.87) --
	(438.75, 60.81) --
	(439.13, 60.75) --
	(439.51, 60.69) --
	(439.89, 60.63) --
	(440.27, 60.57) --
	(440.64, 60.51) --
	(441.02, 60.46) --
	(441.40, 60.40) --
	(441.78, 60.34) --
	(442.16, 60.28) --
	(442.54, 60.23) --
	(442.92, 60.17) --
	(443.30, 60.11) --
	(443.68, 60.06) --
	(444.06, 60.00) --
	(444.44, 59.95) --
	(444.82, 59.89) --
	(445.20, 59.84) --
	(445.58, 59.78) --
	(445.96, 59.73) --
	(446.34, 59.67) --
	(446.72, 59.62) --
	(447.10, 59.57) --
	(447.48, 59.51) --
	(447.86, 59.46) --
	(448.24, 59.41) --
	(448.62, 59.36) --
	(449.00, 59.30) --
	(449.38, 59.25) --
	(449.76, 59.20) --
	(450.14, 59.15) --
	(450.52, 59.10) --
	(450.90, 59.05) --
	(451.28, 59.00) --
	(451.66, 58.95) --
	(452.04, 58.90) --
	(452.42, 58.85) --
	(452.80, 58.80) --
	(453.18, 58.75) --
	(453.56, 58.70) --
	(453.93, 58.65) --
	(454.31, 58.60) --
	(454.69, 58.56) --
	(455.07, 58.51) --
	(455.45, 58.46) --
	(455.83, 58.41) --
	(456.21, 58.37) --
	(456.59, 58.32) --
	(456.97, 58.27) --
	(457.35, 58.23) --
	(457.73, 58.18) --
	(458.11, 58.13) --
	(458.49, 58.09) --
	(458.87, 58.04) --
	(459.25, 58.00) --
	(459.63, 57.95) --
	(460.01, 57.91) --
	(460.39, 57.86) --
	(460.77, 57.82) --
	(461.15, 57.77) --
	(461.53, 57.73) --
	(461.91, 57.69) --
	(462.29, 57.64) --
	(462.67, 57.60) --
	(463.05, 57.56) --
	(463.43, 57.52) --
	(463.81, 57.47) --
	(464.19, 57.43) --
	(464.57, 57.39) --
	(464.95, 57.35) --
	(465.33, 57.31) --
	(465.71, 57.26) --
	(466.09, 57.22) --
	(466.47, 57.18) --
	(466.85, 57.14) --
	(467.23, 57.10) --
	(467.60, 57.06) --
	(467.98, 57.02) --
	(468.36, 56.98) --
	(468.74, 56.94) --
	(469.12, 56.90) --
	(469.50, 56.86) --
	(469.88, 56.82) --
	(470.26, 56.78) --
	(470.64, 56.75) --
	(471.02, 56.71) --
	(471.40, 56.67) --
	(471.78, 56.63) --
	(472.16, 56.59) --
	(472.54, 56.56) --
	(472.92, 56.52) --
	(473.30, 56.48) --
	(473.68, 56.44) --
	(474.06, 56.41) --
	(474.44, 56.37) --
	(474.82, 56.34) --
	(475.20, 56.30) --
	(475.58, 56.26) --
	(475.96, 56.23) --
	(476.34, 56.19) --
	(476.72, 56.16) --
	(477.10, 56.12) --
	(477.48, 56.09) --
	(477.86, 56.05) --
	(478.24, 56.02) --
	(478.62, 55.98) --
	(479.00, 55.95) --
	(479.38, 55.91) --
	(479.76, 55.88) --
	(480.14, 55.85) --
	(480.52, 55.81) --
	(480.90, 55.78) --
	(481.27, 55.75) --
	(481.65, 55.71) --
	(482.03, 55.68) --
	(482.41, 55.65) --
	(482.79, 55.61) --
	(483.17, 55.58) --
	(483.55, 55.55) --
	(483.93, 55.52) --
	(484.31, 55.49) --
	(484.69, 55.45) --
	(485.07, 55.42) --
	(485.45, 55.39) --
	(485.83, 55.36) --
	(486.21, 55.33) --
	(486.59, 55.30) --
	(486.97, 55.27) --
	(487.35, 55.24) --
	(487.73, 55.21) --
	(488.11, 55.18) --
	(488.49, 55.15) --
	(488.87, 55.12) --
	(489.25, 55.09) --
	(489.63, 55.06) --
	(490.01, 55.03) --
	(490.39, 55.00) --
	(490.77, 54.97) --
	(491.15, 54.94) --
	(491.53, 54.91) --
	(491.91, 54.89) --
	(492.29, 54.86) --
	(492.67, 54.83) --
	(493.05, 54.80) --
	(493.43, 54.77) --
	(493.81, 54.75) --
	(494.19, 54.72) --
	(494.56, 54.69) --
	(494.94, 54.66) --
	(495.32, 54.64) --
	(495.70, 54.61) --
	(496.08, 54.58) --
	(496.46, 54.55) --
	(496.84, 54.53) --
	(497.22, 54.50) --
	(497.60, 54.48) --
	(497.98, 54.45) --
	(498.36, 54.42) --
	(498.74, 54.40) --
	(499.12, 54.37) --
	(499.50, 54.35) --
	(499.88, 54.32) --
	(500.26, 54.30) --
	(500.64, 54.27) --
	(501.02, 54.25) --
	(501.40, 54.22) --
	(501.78, 54.20) --
	(502.16, 54.17) --
	(502.54, 54.15) --
	(502.92, 54.12) --
	(503.30, 54.10) --
	(503.68, 54.07) --
	(504.06, 54.05) --
	(504.44, 54.03) --
	(504.82, 54.00) --
	(505.20, 53.98) --
	(505.58, 53.96) --
	(505.96, 53.93) --
	(506.34, 53.91) --
	(506.72, 53.89) --
	(507.10, 53.86) --
	(507.48, 53.84) --
	(507.86, 53.82) --
	(508.23, 53.79) --
	(508.61, 53.77) --
	(508.99, 53.75) --
	(509.37, 53.73) --
	(509.75, 53.71) --
	(510.13, 53.68) --
	(510.51, 53.66) --
	(510.89, 53.64) --
	(511.27, 53.62) --
	(511.65, 53.60) --
	(512.03, 53.58) --
	(512.41, 53.55) --
	(512.79, 53.53) --
	(513.17, 53.51) --
	(513.55, 53.49) --
	(513.93, 53.47) --
	(514.31, 53.45) --
	(514.69, 53.43) --
	(515.07, 53.41) --
	(515.45, 53.39) --
	(515.83, 53.37) --
	(516.21, 53.35) --
	(516.59, 53.33) --
	(516.97, 53.31) --
	(517.35, 53.29) --
	(517.73, 53.27) --
	(518.11, 53.25) --
	(518.49, 53.23) --
	(518.87, 53.21) --
	(519.25, 53.19) --
	(519.63, 53.17) --
	(520.01, 53.15) --
	(520.39, 53.14) --
	(520.77, 53.12) --
	(521.15, 53.10) --
	(521.53, 53.08) --
	(521.90, 53.06) --
	(522.28, 53.04) --
	(522.66, 53.02) --
	(523.04, 53.01) --
	(523.42, 52.99) --
	(523.80, 52.97) --
	(524.18, 52.95) --
	(524.56, 52.93) --
	(524.94, 52.92) --
	(525.32, 52.90) --
	(525.70, 52.88) --
	(526.08, 52.86) --
	(526.46, 52.85) --
	(526.84, 52.83) --
	(527.22, 52.81) --
	(527.60, 52.80) --
	(527.98, 52.78) --
	(528.36, 52.76);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (528.36, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.25, 49.80) -- (386.25, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.62, 49.80) -- (433.62, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.99, 49.80) -- (480.99, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.36, 49.80) -- (528.36, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88, 31.87) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (386.25, 31.87) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (433.62, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (480.99, 31.87) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (528.36, 31.87) {20};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88,239.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (333.90, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 87.70) -- (333.90, 87.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,125.59) -- (333.90,125.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,163.49) -- (333.90,163.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,201.38) -- (333.90,201.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,239.28) -- (333.90,239.28);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 46.94) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 84.84) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,122.73) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,160.63) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,198.53) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,236.42) {1.0};
\end{scope}
\end{tikzpicture}

\end{document}