### 2020-08-31 benchmarking/benchmarkingVM/ESS_round2.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-31 17:26:30
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,289.08) rectangle (289.08,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,548.96) {\bfseries adapt(\textit{real}) vs. adapt(\textit{cat})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,301.03) {ESS/hr [adapt(\textit{cat})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,433.62) {ESS/hr [adapt(\textit{real})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (  0.00,368.65) --
	(209.51,578.16);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,492.84) --
	(  0.00,209.51);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (  0.00,342.13) --
	(236.03,578.16);

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,519.37) --
	(  0.00,236.03);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (  0.00,315.60) --
	(262.56,578.16);

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (283.33,578.16) --
	(283.33,545.89) --
	(  0.00,262.56);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92,376.84) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80,338.88) -- (239.28,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.92,505.04) --
	(206.54,505.04);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.23,504.79) --
	(206.23,505.29);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (206.23,501.82) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (216.33,507.39) --
	(217.11,507.39);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (216.73,506.95) --
	(216.73,507.82);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (216.73,504.49) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.62,511.19) --
	(211.77,511.19);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.70,511.13) --
	(211.70,511.25);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.70,507.16) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.95,493.79) --
	(202.68,493.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.32,493.47) --
	(202.32,494.11);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.32,490.89) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (203.67,496.31) --
	(204.38,496.31);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.03,495.98) --
	(204.03,496.64);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.03,492.28) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.75,506.51) --
	(206.55,506.51);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.16,506.28) --
	(206.16,506.73);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (206.16,503.60) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (216.87,512.59) --
	(217.23,512.59);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (217.05,512.47) --
	(217.05,512.71);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (217.05,509.69) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (173.90,468.92) --
	(175.72,468.92);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (174.83,468.08) --
	(174.83,469.72);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (174.83,465.69) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.59,497.49) --
	(207.15,497.49);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.87,497.22) --
	(206.87,497.77);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (206.87,494.59) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.49,500.25) --
	(204.66,500.25);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.57,500.18) --
	(204.57,500.32);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.57,496.21) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (165.17,453.35) --
	(167.29,453.35);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (166.26,452.41) --
	(166.26,454.24);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (166.26,450.44) {$\bar{r}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (167.36,455.34) --
	(169.47,455.34);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.44,454.39) --
	(168.44,456.25);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (168.44,451.31) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (170.32,466.37) --
	(171.88,466.37);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (171.11,465.53) --
	(171.11,467.17);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (171.11,463.46) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.77,501.14) --
	(206.93,501.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.85,501.05) --
	(206.85,501.23);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (206.85,498.24) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (161.26,477.08) --
	(162.96,477.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (162.13,476.61) --
	(162.13,477.54);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (162.13,473.86) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.84,493.82) --
	(205.45,493.82);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.15,493.57) --
	(205.15,494.06);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.15,490.91) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.75,495.00) --
	(198.03,495.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.89,494.93) --
	(197.89,495.07);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.89,490.97) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (156.31,470.07) --
	(157.86,470.07);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (157.10,469.69) --
	(157.10,470.46);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (157.10,467.17) {$\bar{r}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (159.33,471.13) --
	(160.99,471.13);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (160.18,470.73) --
	(160.18,471.51);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.18,467.09) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (165.56,474.04) --
	(167.15,474.04);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (166.37,473.55) --
	(166.37,474.51);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (166.37,471.13) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.93,500.80) --
	(206.27,500.80);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.10,500.70) --
	(206.10,500.90);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (206.10,497.89) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (121.64,470.69) --
	(124.00,470.69);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (122.86,470.18) --
	(122.86,471.19);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (122.86,467.47) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (185.34,468.90) --
	(185.91,468.90);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (185.63,468.61) --
	(185.63,469.18);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (185.63,465.99) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (162.37,475.57) --
	(162.64,475.57);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (162.51,475.45) --
	(162.51,475.69);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (162.51,471.54) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (120.02,442.67) --
	(122.54,442.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (121.32,442.05) --
	(121.32,443.26);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (121.32,439.76) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (121.22,442.55) --
	(123.50,442.55);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (122.40,441.89) --
	(122.40,443.18);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (122.40,438.51) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (120.43,458.76) --
	(124.03,458.76);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (122.32,457.98) --
	(122.32,459.51);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (122.32,455.86) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.53,482.91) --
	(188.81,482.91);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.67,482.83) --
	(188.67,482.99);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.67,480.01) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (207.64,487.61) --
	(208.32,487.61);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (207.98,486.33) --
	(207.98,488.81);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.98,484.39) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.16,504.08) --
	(211.73,504.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.45,503.75) --
	(211.45,504.41);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.45,501.17) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (221.28,516.22) --
	(221.36,516.22);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (221.32,516.18) --
	(221.32,516.26);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (221.32,512.18) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.31,497.67) --
	(201.00,497.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.66,497.33) --
	(200.66,498.01);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.66,494.77) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.63,502.94) --
	(206.35,502.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.99,502.68) --
	(205.99,503.19);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.99,498.90) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.03,491.24) --
	(197.86,491.24);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.45,490.94) --
	(197.45,491.55);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.45,488.34) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.55,507.70) --
	(211.80,507.70);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.68,507.59) --
	(211.68,507.81);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.68,504.80) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (159.54,499.71) --
	(161.97,499.71);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (160.79,499.35) --
	(160.79,500.06);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (160.79,496.48) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (216.44,498.95) --
	(217.03,498.95);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (216.74,498.66) --
	(216.74,499.24);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (216.74,496.05) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.48,503.67) --
	(197.89,503.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.69,503.61) --
	(197.69,503.74);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (197.69,499.64) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (158.30,471.10) --
	(160.71,471.10);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (159.54,470.46) --
	(159.54,471.72);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (159.54,468.20) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (160.01,471.53) --
	(162.43,471.53);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (161.26,470.90) --
	(161.26,472.14);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (161.26,467.50) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (165.78,483.58) --
	(168.47,483.58);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (167.17,483.12) --
	(167.17,484.03);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (167.17,480.67) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (215.03,512.59) --
	(215.57,512.59);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (215.30,512.42) --
	(215.30,512.76);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (215.30,509.69) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (178.82,481.02) --
	(179.82,481.02);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.33,480.59) --
	(179.33,481.44);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.33,477.79) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (191.50,482.93) --
	(192.71,482.93);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (192.12,481.27) --
	(192.12,484.45);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.12,480.02) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.77,489.01) --
	(194.93,489.01);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.85,488.92) --
	(194.85,489.11);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.85,484.98) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (178.61,473.57) --
	(179.57,473.57);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.09,473.24) --
	(179.09,473.89);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.09,470.67) {$\bar{r}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (177.63,472.97) --
	(178.69,472.97);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (178.17,472.65) --
	(178.17,473.28);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (178.17,468.93) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.69,480.98) --
	(182.56,480.98);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.13,480.59) --
	(182.13,481.36);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.13,478.08) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.37,501.33) --
	(204.73,501.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.55,501.22) --
	(204.55,501.45);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.55,498.43) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.59,487.86) --
	(200.20,487.86);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.90,487.42) --
	(199.90,488.29);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.90,484.64) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.13,484.00) --
	(188.05,484.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.60,483.35) --
	(187.60,484.63);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (187.60,481.10) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.12,500.91) --
	(204.27,500.91);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.20,500.84) --
	(204.20,500.97);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.20,496.87) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.04,490.16) --
	(194.61,490.16);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.33,489.94) --
	(194.33,490.38);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.33,487.26) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.55,495.79) --
	(200.15,495.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.86,495.53) --
	(199.86,496.04);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.86,491.75) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (192.32,490.00) --
	(193.05,490.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (192.69,489.48) --
	(192.69,490.49);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (192.69,487.09) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.97,512.17) --
	(214.36,512.17);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.16,511.97) --
	(214.16,512.37);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (214.16,509.26) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- (226.02,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85,338.88) -- ( 93.85,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91,338.88) -- (137.91,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96,338.88) -- (181.96,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02,338.88) -- (226.02,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 49.80,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,338.88) -- ( 44.82,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,382.93) -- ( 44.82,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,426.99) -- ( 44.82,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,471.04) -- ( 44.82,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,515.10) -- ( 44.82,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] ( 49.80,338.88) rectangle (239.28,528.36);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,421.05) rectangle (189.67,416.07);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,411.09) rectangle (189.67,406.11);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,401.13) rectangle (189.67,396.15);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,391.17) rectangle (189.67,386.19);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,381.21) rectangle (189.67,376.23);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,371.25) rectangle (189.67,366.27);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,361.29) rectangle (189.67,356.31);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.70,351.33) rectangle (189.67,346.35);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,415.70) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,405.74) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,395.78) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,385.82) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,375.86) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,365.90) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,355.94) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (197.14,345.98) {Moyle};
\end{scope}
\begin{scope}
\path[clip] (289.08,289.08) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,548.96) {\bfseries adapt(\textit{quant}) vs. adapt(\textit{cat})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,301.03) {ESS/hr [adapt(\textit{cat})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,433.62) {ESS/hr [adapt(\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] (338.88,338.88) rectangle (528.36,528.36);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (162.66,242.23) --
	(498.59,578.16);

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,492.84) --
	(162.66, 83.09) --
	(162.66,242.23);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (162.66,215.71) --
	(525.11,578.16);

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,519.37) --
	(162.66,109.61) --
	(162.66,215.71);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (162.66,189.19) --
	(551.64,578.16);

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (572.41,578.16) --
	(572.41,545.89) --
	(162.66,136.14) --
	(162.66,189.19);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (425.98,342.31) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,429.88) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (399.46,342.31) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.83,403.36) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (371.28,342.31) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (341.00,376.84) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] (338.88,338.88) -- (528.36,528.36);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.00,489.87) --
	(495.62,489.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.31,489.57) --
	(495.31,490.16);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (495.31,486.64) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.41,499.95) --
	(506.19,499.95);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.81,499.66) --
	(505.81,500.24);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (505.81,497.05) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (500.70,501.75) --
	(500.85,501.75);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (500.78,501.69) --
	(500.78,501.80);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.78,497.71) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.03,485.79) --
	(491.76,485.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (491.40,485.54) --
	(491.40,486.03);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (491.40,482.88) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (492.75,487.50) --
	(493.46,487.50);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.11,487.21) --
	(493.11,487.80);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.11,483.47) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (494.83,499.04) --
	(495.63,499.04);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.24,498.77) --
	(495.24,499.32);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (495.24,496.14) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.95,500.55) --
	(506.31,500.55);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (506.13,500.44) --
	(506.13,500.66);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (506.13,497.65) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (462.98,451.79) --
	(464.80,451.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (463.91,451.27) --
	(463.91,452.29);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (463.91,448.56) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.67,489.99) --
	(496.23,489.99);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.95,489.81) --
	(495.95,490.18);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (495.95,487.09) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.57,492.80) --
	(493.74,492.80);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.65,492.74) --
	(493.65,492.85);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.65,488.76) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (454.25,445.43) --
	(456.37,445.43);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (455.34,444.89) --
	(455.34,445.95);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (455.34,442.52) {$\bar{r}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (456.44,448.01) --
	(458.55,448.01);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (457.52,447.49) --
	(457.52,448.51);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (457.52,443.97) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (459.40,462.20) --
	(460.96,462.20);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (460.19,461.65) --
	(460.19,462.73);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (460.19,459.29) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.85,490.73) --
	(496.01,490.73);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.93,490.65) --
	(495.93,490.82);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (495.93,487.83) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (450.34,464.66) --
	(452.04,464.66);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (451.21,464.12) --
	(451.21,465.18);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (451.21,461.43) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.92,489.34) --
	(494.53,489.34);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (494.23,489.09) --
	(494.23,489.58);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (494.23,486.43) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (486.83,489.35) --
	(487.11,489.35);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (486.97,489.15) --
	(486.97,489.55);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (486.97,485.32) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (445.39,461.69) --
	(446.94,461.69);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (446.18,461.16) --
	(446.18,462.22);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (446.18,458.79) {$\bar{r}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (448.41,463.55) --
	(450.07,463.55);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (449.26,463.00) --
	(449.26,464.08);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.26,459.51) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (454.64,470.31) --
	(456.23,470.31);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (455.45,469.79) --
	(455.45,470.82);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (455.45,467.41) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.01,490.35) --
	(495.35,490.35);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.18,490.11) --
	(495.18,490.60);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (495.18,487.45) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (410.72,427.41) --
	(413.08,427.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (411.94,426.32) --
	(411.94,428.45);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (411.94,424.19) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (474.42,462.41) --
	(474.99,462.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (474.71,461.74) --
	(474.71,463.06);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (474.71,459.50) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (451.45,467.41) --
	(451.72,467.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (451.59,467.24) --
	(451.59,467.58);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (451.59,463.38) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (409.10,426.94) --
	(411.62,426.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (410.40,425.89) --
	(410.40,427.93);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (410.40,424.03) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (410.30,427.14) --
	(412.58,427.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (411.48,426.01) --
	(411.48,428.20);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (411.48,423.11) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (409.51,441.14) --
	(413.11,441.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (411.40,439.81) --
	(411.40,442.38);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (411.40,438.23) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (477.61,473.14) --
	(477.89,473.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (477.75,472.98) --
	(477.75,473.30);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (477.75,470.24) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (496.72,491.90) --
	(497.40,491.90);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (497.06,491.57) --
	(497.06,492.23);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (497.06,488.68) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (500.24,494.94) --
	(500.81,494.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (500.53,494.63) --
	(500.53,495.24);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.53,492.04) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (510.36,505.18) --
	(510.44,505.18);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (510.40,505.15) --
	(510.40,505.21);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (510.40,501.15) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (489.39,485.14) --
	(490.08,485.14);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (489.74,484.76) --
	(489.74,485.50);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (489.74,482.23) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (494.71,490.54) --
	(495.43,490.54);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (495.07,490.21) --
	(495.07,490.85);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (495.07,486.50) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (486.11,483.52) --
	(486.94,483.52);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (486.53,483.11) --
	(486.53,483.93);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (486.53,480.62) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (500.63,495.41) --
	(500.88,495.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (500.76,495.32) --
	(500.76,495.49);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (500.76,492.50) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (448.62,459.08) --
	(451.05,459.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (449.87,458.35) --
	(449.87,459.79);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (449.87,455.85) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.52,495.76) --
	(506.11,495.76);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (505.82,495.49) --
	(505.82,496.03);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (505.82,492.85) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (486.56,496.08) --
	(486.97,496.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (486.77,495.98) --
	(486.77,496.19);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (486.77,492.05) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (447.38,457.54) --
	(449.79,457.54);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (448.62,456.85) --
	(448.62,458.21);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (448.62,454.64) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (449.09,459.53) --
	(451.51,459.53);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (450.34,458.82) --
	(450.34,460.21);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (450.34,455.50) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (454.86,474.65) --
	(457.55,474.65);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (456.25,473.94) --
	(456.25,475.33);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (456.25,471.74) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (504.11,498.52) --
	(504.65,498.52);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (504.38,498.30) --
	(504.38,498.73);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (504.38,495.61) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (467.90,466.02) --
	(468.90,466.02);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (468.41,465.37) --
	(468.41,466.64);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (468.41,462.79) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (480.58,476.12) --
	(481.79,476.12);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (481.20,475.44) --
	(481.20,476.77);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (481.20,473.22) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.85,480.75) --
	(484.01,480.75);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.93,480.67) --
	(483.93,480.84);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.93,476.72) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (467.69,466.41) --
	(468.65,466.41);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (468.17,466.01) --
	(468.17,466.81);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (468.17,463.51) {$\bar{r}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (466.71,464.95) --
	(467.77,464.95);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (467.25,464.33) --
	(467.25,465.56);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (467.25,460.92) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (470.77,472.46) --
	(471.64,472.46);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (471.21,471.96) --
	(471.21,472.95);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (471.21,469.56) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.45,489.31) --
	(493.81,489.31);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.63,489.19) --
	(493.63,489.43);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.63,486.41) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (488.67,486.27) --
	(489.28,486.27);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (488.98,486.05) --
	(488.98,486.50);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (488.98,483.05) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (476.21,473.16) --
	(477.13,473.16);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (476.68,472.60) --
	(476.68,473.70);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (476.68,470.25) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.20,490.92) --
	(493.35,490.92);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (493.28,490.86) --
	(493.28,490.98);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (493.28,486.89) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.12,480.77) --
	(483.69,480.77);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (483.41,480.54) --
	(483.41,481.00);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (483.41,477.86) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (488.63,486.02) --
	(489.23,486.02);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (488.94,485.79) --
	(488.94,486.26);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (488.94,481.99) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (481.40,481.81) --
	(482.13,481.81);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (481.77,481.42) --
	(481.77,482.20);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (481.77,478.91) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.05,499.85) --
	(503.44,499.85);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (503.24,499.72) --
	(503.24,499.97);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (503.24,496.94) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (515.10,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.93,338.88) -- (382.93,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.99,338.88) -- (426.99,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.04,338.88) -- (471.04,333.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.10,338.88) -- (515.10,333.90);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88,320.95) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (382.93,320.95) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (426.99,320.95) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (471.04,320.95) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (515.10,320.95) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (338.88,515.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,338.88) -- (333.90,338.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,382.93) -- (333.90,382.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,426.99) -- (333.90,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,471.04) -- (333.90,471.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,515.10) -- (333.90,515.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,336.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,380.08) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,424.13) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,468.19) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,512.24) {10000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54,259.88) {\bfseries adapt(\textit{quant}) vs. adapt(\textit{real})};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (144.54, 11.95) {ESS/hr [adapt(\textit{real})]};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at ( 17.93,144.54) {ESS/hr [adapt(\textit{quant})]};
\end{scope}
\begin{scope}
\path[clip] ( 49.80, 49.80) rectangle (239.28,239.28);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.47,line width= 0.4pt,line join=round,line cap=round] (  0.00, 79.57) --
	(283.33,362.91) --
	(283.33,203.76) --
	( 79.57,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.67,line width= 0.6pt,line join=round,line cap=round] (  0.00, 53.05) --
	(283.33,336.38) --
	(283.33,230.29) --
	( 53.05,  0.00);
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.87,line width= 0.8pt,line join=round,line cap=round] (  0.00, 26.52) --
	(283.33,309.86) --
	(283.33,256.81) --
	( 26.52,  0.00);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (136.90, 53.23) {$\frac{1}{64} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,140.80) {$64\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (110.38, 53.23) {$\frac{1}{16} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 52.75,114.28) {$16\times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 82.20, 53.23) {$\frac{1}{4} \times$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 51.92, 87.76) {$4\times$};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,line width= 1.2pt,line join=round,line cap=round] ( 49.80, 49.80) -- (239.28,239.28);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (215.71,200.79) --
	(216.21,200.79);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (215.96,200.49) --
	(215.96,201.08);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (215.96,197.56) {$p$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (217.87,210.87) --
	(218.74,210.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (218.31,210.58) --
	(218.31,211.16);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.31,207.97) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (222.05,212.67) --
	(222.17,212.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (222.11,212.61) --
	(222.11,212.72);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (222.11,208.63) {$r$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.39,196.71) --
	(205.03,196.71);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.71,196.46) --
	(204.71,196.95);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.71,193.80) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.90,198.42) --
	(207.56,198.42);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (207.23,198.13) --
	(207.23,198.72);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (207.23,194.39) {$l$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (217.20,209.96) --
	(217.65,209.96);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (217.43,209.69) --
	(217.43,210.24);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (217.43,207.06) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (223.39,211.47) --
	(223.63,211.47);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (223.51,211.36) --
	(223.51,211.58);
\definecolor{drawColor}{RGB}{255,0,109}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (223.51,208.57) {$\kappa$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.00,162.71) --
	(180.64,162.71);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.84,162.19) --
	(179.84,163.21);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.84,159.48) {$p$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.14,200.91) --
	(208.69,200.91);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.41,200.73) --
	(208.41,201.10);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.41,198.01) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.10,203.72) --
	(211.24,203.72);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.17,203.66) --
	(211.17,203.77);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.17,199.68) {$r$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (163.33,156.35) --
	(165.16,156.35);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (164.27,155.81) --
	(164.27,156.87);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (164.27,153.44) {$\bar{r}$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (165.31,158.93) --
	(167.17,158.93);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (166.26,158.41) --
	(166.26,159.43);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (166.26,154.89) {$l$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (176.45,173.12) --
	(178.09,173.12);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (177.29,172.57) --
	(177.29,173.65);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (177.29,170.21) {$\sigma$};
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.97,201.65) --
	(212.15,201.65);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.06,201.57) --
	(212.06,201.74);
\definecolor{drawColor}{RGB}{253,104,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.06,198.75) {$\kappa$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (187.53,175.58) --
	(188.46,175.58);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (188.00,175.04) --
	(188.00,176.10);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (188.00,172.35) {$p$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.49,200.26) --
	(204.98,200.26);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (204.74,200.01) --
	(204.74,200.50);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (204.74,197.35) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.85,200.27) --
	(205.99,200.27);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (205.92,200.07) --
	(205.92,200.47);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (205.92,196.24) {$r$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.61,172.61) --
	(181.38,172.61);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (180.99,172.08) --
	(180.99,173.14);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (180.99,169.71) {$\bar{r}$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.65,174.47) --
	(182.43,174.47);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.05,173.92) --
	(182.05,175.00);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.05,170.43) {$l$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (184.47,181.23) --
	(185.43,181.23);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (184.96,180.71) --
	(184.96,181.74);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.96,178.33) {$\sigma$};
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.62,201.27) --
	(211.82,201.27);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.72,201.03) --
	(211.72,201.52);
\definecolor{drawColor}{RGB}{143,161,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.72,198.37) {$\kappa$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.10,138.33) --
	(182.11,138.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.61,137.24) --
	(181.61,139.37);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (181.61,135.11) {$p$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.53,173.33) --
	(180.10,173.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (179.82,172.66) --
	(179.82,173.98);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (179.82,170.42) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (186.37,178.33) --
	(186.61,178.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (186.49,178.16) --
	(186.49,178.50);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (186.49,174.30) {$r$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (152.97,137.86) --
	(154.18,137.86);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (153.59,136.81) --
	(153.59,138.85);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (153.59,134.95) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (152.81,138.06) --
	(154.10,138.06);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (153.47,136.93) --
	(153.47,139.12);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (153.47,134.03) {$l$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (168.90,152.06) --
	(170.43,152.06);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (169.68,150.73) --
	(169.68,153.30);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (169.68,149.15) {$\sigma$};
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.75,184.06) --
	(193.91,184.06);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.83,183.90) --
	(193.83,184.22);
\definecolor{drawColor}{RGB}{0,195,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.83,181.16) {$\kappa$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (197.25,202.82) --
	(199.73,202.82);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.53,202.49) --
	(198.53,203.15);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (198.53,199.60) {$p$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.67,205.86) --
	(215.33,205.86);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (215.00,205.55) --
	(215.00,206.16);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (215.00,202.96) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (227.10,216.10) --
	(227.18,216.10);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (227.14,216.07) --
	(227.14,216.13);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (227.14,212.07) {$r$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.25,196.06) --
	(208.93,196.06);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (208.59,195.68) --
	(208.59,196.42);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (208.59,193.15) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.60,201.46) --
	(214.11,201.46);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (213.86,201.13) --
	(213.86,201.77);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (213.86,197.42) {$l$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.86,194.44) --
	(202.47,194.44);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (202.16,194.03) --
	(202.16,194.85);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (202.16,191.54) {$\sigma$};
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (218.51,206.33) --
	(218.73,206.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (218.62,206.24) --
	(218.62,206.41);
\definecolor{drawColor}{RGB}{0,219,171}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (218.62,203.42) {$\kappa$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (210.27,170.00) --
	(210.98,170.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (210.63,169.27) --
	(210.63,170.71);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (210.63,166.77) {$p$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.58,206.68) --
	(210.16,206.68);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (209.87,206.41) --
	(209.87,206.95);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (209.87,203.77) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.53,207.00) --
	(214.66,207.00);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (214.59,206.90) --
	(214.59,207.11);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (214.59,202.97) {$r$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.38,168.46) --
	(182.64,168.46);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.02,167.77) --
	(182.02,169.13);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.02,165.56) {$\bar{r}$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (181.82,170.45) --
	(183.06,170.45);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (182.45,169.74) --
	(182.45,171.13);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (182.45,166.42) {$l$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.04,185.57) --
	(194.95,185.57);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.50,184.86) --
	(194.50,186.25);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.50,182.66) {$\sigma$};
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (223.34,209.44) --
	(223.68,209.44);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (223.51,209.22) --
	(223.51,209.65);
\definecolor{drawColor}{RGB}{0,216,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (223.51,206.53) {$\kappa$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (191.51,176.94) --
	(192.36,176.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (191.94,176.29) --
	(191.94,177.56);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.94,173.71) {$p$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (192.19,187.04) --
	(195.37,187.04);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (193.85,186.36) --
	(193.85,187.69);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (193.85,184.14) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.84,191.67) --
	(200.03,191.67);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (199.93,191.59) --
	(199.93,191.76);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (199.93,187.64) {$r$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (184.16,177.33) --
	(184.81,177.33);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (184.49,176.93) --
	(184.49,177.73);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (184.49,174.43) {$\bar{r}$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (183.57,175.87) --
	(184.20,175.87);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (183.89,175.25) --
	(183.89,176.48);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (183.89,171.84) {$l$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (191.51,183.38) --
	(192.28,183.38);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (191.90,182.88) --
	(191.90,183.87);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (191.90,180.48) {$\sigma$};
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.14,200.23) --
	(212.37,200.23);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (212.25,200.11) --
	(212.25,200.35);
\definecolor{drawColor}{RGB}{151,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (212.25,197.33) {$\kappa$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.34,197.19) --
	(199.21,197.19);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (198.78,196.97) --
	(198.78,197.42);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (198.78,193.97) {$p$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.27,184.08) --
	(195.55,184.08);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (194.92,183.52) --
	(194.92,184.62);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (194.92,181.17) {$\mathcal{L}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.76,201.84) --
	(211.89,201.84);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (211.83,201.78) --
	(211.83,201.90);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (211.83,197.81) {$r$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.86,191.69) --
	(201.30,191.69);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (201.08,191.46) --
	(201.08,191.92);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (201.08,188.78) {$\bar{r}$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.45,196.94) --
	(206.96,196.94);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (206.71,196.71) --
	(206.71,197.18);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (206.71,192.91) {$l$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.40,192.73) --
	(201.41,192.73);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (200.92,192.34) --
	(200.92,193.12);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (200.92,189.83) {$\sigma$};
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (222.89,210.77) --
	(223.29,210.77);

\path[draw=drawColor,draw opacity=0.33,line width= 0.4pt,line join=round,line cap=round] (223.09,210.64) --
	(223.09,210.89);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.16] at (223.09,207.86) {$\kappa$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- (226.02, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.85, 49.80) -- ( 93.85, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.91, 49.80) -- (137.91, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.96, 49.80) -- (181.96, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.02, 49.80) -- (226.02, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 49.80, 31.87) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 93.85, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (137.91, 31.87) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (181.96, 31.87) {1000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (226.02, 31.87) {10000};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 49.80,226.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 49.80) -- ( 44.82, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80, 93.85) -- ( 44.82, 93.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,137.91) -- ( 44.82,137.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,181.96) -- ( 44.82,181.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.80,226.02) -- ( 44.82,226.02);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 46.94) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84, 91.00) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,135.05) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,179.11) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at ( 39.84,223.16) {10000};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62,259.88) {\bfseries Leaf rate mixing vs sequence length};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (433.62, 11.95) {Sequence length $L$ (kb)};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.24] at (307.01,144.54) {ESS/hr of tip rates $r$};
\end{scope}
\begin{scope}
\path[clip] (338.88, 49.80) rectangle (528.36,239.28);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (356.88,127.23) --
	(356.88, 94.67);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (358.78,108.88) {$1.7\times $};
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (356.88,127.23) circle (  1.87);

\path[fill=fillColor] (356.88, 97.58) --
	(359.40, 93.22) --
	(354.37, 93.22) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (394.78, 93.40) --
	(394.78, 80.86);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (396.67, 85.06) {$1.4\times $};
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (394.78, 93.40) circle (  1.87);

\path[fill=fillColor] (394.78, 83.77) --
	(397.29, 79.41) --
	(392.26, 79.41) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (390.04, 82.94) --
	(390.04, 71.70);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (391.93, 75.24) {$1.5\times $};
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (390.04, 82.94) circle (  1.87);

\path[fill=fillColor] (390.04, 74.60) --
	(392.55, 70.24) --
	(387.52, 70.24) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (404.25, 61.80) --
	(404.25, 53.19);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (406.15, 55.42) {$3.5\times $};
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (404.25, 61.80) circle (  1.87);

\path[fill=fillColor] (404.25, 56.09) --
	(406.77, 51.74) --
	(401.74, 51.74) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (359.72,150.32) --
	(359.72,123.79);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (361.62,134.98) {$1.4\times $};
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (359.72,150.32) circle (  1.87);

\path[fill=fillColor] (359.72,126.70) --
	(362.24,122.34) --
	(357.21,122.34) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (433.62,101.96) --
	(433.62, 71.34);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (435.51, 84.58) {$2.4\times $};
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (433.62,101.96) circle (  1.87);

\path[fill=fillColor] (433.62, 74.24) --
	(436.14, 69.88) --
	(431.10, 69.88) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (359.72, 74.13) --
	(359.72, 68.32);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (361.62, 69.15) {$1.3\times $};
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (359.72, 74.13) circle (  1.87);

\path[fill=fillColor] (359.72, 71.22) --
	(362.24, 66.86) --
	(357.21, 66.86) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (346.46, 95.05) --
	(346.46, 80.15);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.83] at (348.35, 85.52) {$1.5\times $};
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (346.46, 95.05) circle (  1.87);

\path[fill=fillColor] (346.46, 83.05) --
	(348.97, 78.70) --
	(343.94, 78.70) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (465.27,227.83) circle (  2.15);

\path[fill=fillColor] (465.27,219.71) --
	(468.17,214.70) --
	(462.38,214.70) --
	cycle;

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (473.87,224.54) {adapt (\textit{real})};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.95] at (473.87,213.09) {adapt (\textit{cat})};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (528.36, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.25, 49.80) -- (386.25, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.62, 49.80) -- (433.62, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.99, 49.80) -- (480.99, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.36, 49.80) -- (528.36, 44.82);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (338.88, 31.87) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (386.25, 31.87) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (433.62, 31.87) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (480.99, 31.87) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.83] at (528.36, 31.87) {20};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (338.88,239.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 49.80) -- (333.90, 49.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88, 97.17) -- (333.90, 97.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,144.54) -- (333.90,144.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,191.91) -- (333.90,191.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.88,239.28) -- (333.90,239.28);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 46.94) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92, 94.31) {5000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,141.68) {10000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,189.05) {15000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.83] at (328.92,236.42) {20000};
\end{scope}
\end{tikzpicture}

\end{document}