### 2020-08-31 benchmarking/benchmarkingVM/ESS_round3.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-31 17:26:53
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,357.33);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (301.08,335.19) {\bfseries Comparison of Bactrian+adapt(\textit{real}) and adapt(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (  9.60,190.67) {Relative ESS/hr};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 48.00,173.19) --
	(474.88,173.19);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.10,246.09) --
	( 54.10,233.67);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] ( 54.10,239.88) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.20,249.96) --
	( 60.20,218.26);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] ( 60.20,234.11) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.29,226.82) --
	( 66.29,207.10);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] ( 66.29,216.96) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.39,210.32) --
	( 72.39,191.80);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] ( 72.39,201.06) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.49,205.70) --
	( 78.49,138.28);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] ( 78.49,171.99) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.59,250.06) --
	( 84.59,239.46);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] ( 84.59,244.76) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.69,223.96) --
	( 90.69,209.76);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] ( 90.69,216.86) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.79,229.79) --
	( 96.79,207.30);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] ( 96.79,218.55) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] ( 54.10,218.87) --
	(102.88,218.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 78.49,284.14) {18\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (108.98,  0.00) --
	(108.98,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (115.08,231.18) --
	(115.08,216.59);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (115.08,223.89) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.18,210.48) --
	(121.18,196.00);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (121.18,203.24) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.28,217.85) --
	(127.28,208.48);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (127.28,213.17) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.38,206.09) --
	(133.38,189.08);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (133.38,197.58) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.47,184.21) --
	(139.47,163.53);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (139.47,173.87) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (145.57,274.87) --
	(145.57,264.59);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (145.57,269.73) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.67,345.05) --
	(151.67,220.65);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (151.67,282.85) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.77,222.89) --
	(157.77,199.51);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (157.77,211.20) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (115.08,224.11) --
	(163.87,224.11);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (139.47,284.14) {20\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (169.97,  0.00) --
	(169.97,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.06,237.79) --
	(176.06,235.79);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (176.06,236.79) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.16,213.87) --
	(182.16,210.98);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (182.16,212.42) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.26,213.40) --
	(188.26,210.38);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (188.26,211.89) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.36,192.72) --
	(194.36,185.58);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (194.36,189.15) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.46,187.16) --
	(200.46,185.57);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (200.46,186.36) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.56,240.68) --
	(206.56,238.19);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (206.56,239.44) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.65,223.05) --
	(212.65,219.07);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (212.65,221.06) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.75,215.97) --
	(218.75,213.85);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (218.75,214.91) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (176.06,214.59) --
	(224.85,214.59);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (200.46,284.14) {16\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (230.95,  0.00) --
	(230.95,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.05,250.44) --
	(237.05,241.85);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (237.05,246.14) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.15,239.13) --
	(243.15,208.60);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (243.15,223.86) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.24,213.14) --
	(249.24,198.87);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (249.24,206.01) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.34,219.07) --
	(255.34,189.90);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (255.34,204.48) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,175.34) --
	(261.44,160.05);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (261.44,167.69) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (267.54,231.48) --
	(267.54,205.22);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (267.54,218.35) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.64,253.96) --
	(273.64,239.75);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (273.64,246.86) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.74,236.72) --
	(279.74,225.81);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (279.74,231.26) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (237.05,219.11) --
	(285.83,219.11);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (261.44,284.14) {18\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (291.93,  0.00) --
	(291.93,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.03,240.03) --
	(298.03,230.01);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (298.03,235.02) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.13,238.59) --
	(304.13,208.20);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (304.13,223.39) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.23,210.81) --
	(310.23,195.70);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (310.23,203.25) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.33,228.88) --
	(316.33,202.54);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (316.33,215.71) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.42,175.84) --
	(322.42,163.67);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (322.42,169.76) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.52,230.74) --
	(328.52,204.43);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (328.52,217.58) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.62,232.28) --
	(334.62,208.40);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (334.62,220.34) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.72,233.20) --
	(340.72,221.37);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (340.72,227.29) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (298.03,214.66) --
	(346.82,214.66);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (322.42,284.14) {16\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (352.92,  0.00) --
	(352.92,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.01,237.99) --
	(359.01,223.51);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (359.01,230.75) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.11,244.31) --
	(365.11,211.56);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (365.11,227.93) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (371.21,216.84) --
	(371.21,196.62);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (371.21,206.73) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.31,225.85) --
	(377.31,198.06);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (377.31,211.95) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.41,197.62) --
	(383.41,172.56);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (383.41,185.09) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.51,237.58) --
	(389.51,211.29);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (389.51,224.44) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.60,213.92) --
	(395.60,200.71);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (395.60,207.32) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.70,230.89) --
	(401.70,207.97);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (401.70,219.43) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (359.01,214.55) --
	(407.80,214.55);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (383.41,284.14) {16\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (413.90,  0.00) --
	(413.90,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.00,182.77) --
	(420.00,177.72);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (420.00,180.25) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.10,177.03) --
	(426.10,173.95);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (426.10,175.49) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (432.19,175.37) --
	(432.19,171.54);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (432.19,173.45) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.29,175.65) --
	(438.29,172.26);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (438.29,173.95) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.39,182.05) --
	(444.39,177.75);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (444.39,179.90) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (450.49,176.59) --
	(450.49,169.19);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (450.49,172.89) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.59,177.97) --
	(456.59,173.15);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (456.59,175.56) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.69,183.83) --
	(462.69,176.41);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (462.69,180.12) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (420.00,176.47) --
	(468.78,176.47);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (444.39,284.14) {1.2\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (474.88,  0.00) --
	(474.88,357.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- (554.16, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.49, 60.00) -- ( 78.49, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.47, 60.00) -- (139.47, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.46, 60.00) -- (200.46, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44, 60.00) -- (261.44, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.42, 60.00) -- (322.42, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.41, 60.00) -- (383.41, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.39, 60.00) -- (444.39, 54.00);

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at ( 80.56, 48.00) {Prior};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (141.54, 48.00) {Likelihood};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (202.52, 48.00) {Leaf rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (263.51, 48.00) {Mean branch rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (324.49, 48.00) {Tree length};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (385.47, 48.00) {Clock SD $\sigma$};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (446.46, 48.00) {HKY model $\kappa$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 48.00,286.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 42.00, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 92.88) -- ( 42.00, 92.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,135.92) -- ( 42.00,135.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,173.19) -- ( 42.00,173.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,210.47) -- ( 42.00,210.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,253.51) -- ( 42.00,253.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,286.39) -- ( 42.00,286.39);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 56.56) {$\frac{2}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 89.44) {$\frac{3}{4}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,132.47) {$\frac{7}{8}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,169.75) {$1\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,207.03) {$\frac{8}{7}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,250.06) {$\frac{4}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,282.94) {$\frac{3}{2}\times$};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,235.67) rectangle (494.39,229.67);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,223.67) rectangle (494.39,217.67);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,211.67) rectangle (494.39,205.67);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,199.67) rectangle (494.39,193.67);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,187.67) rectangle (494.39,181.67);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,175.67) rectangle (494.39,169.67);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,163.67) rectangle (494.39,157.67);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,151.67) rectangle (494.39,145.67);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,229.22) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,217.22) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,205.22) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,193.22) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,181.22) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,169.22) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,157.22) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,145.22) {Moyle};
\end{scope}
\end{tikzpicture}

\end{document}