### 2020-08-31 benchmarking/benchmarkingVM/ESS_round4b.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-31 17:27:18
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,357.33);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (301.08,335.19) {\bfseries Comparison of Bactrian+adapt(\textit{real}) and NER+Bactrian+adapt(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (  9.60,190.67) {Relative ESS/hr};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 48.00,173.19) --
	(474.88,173.19);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.10,186.05) --
	( 54.10,174.62);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] ( 54.10,180.33) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.20,164.25) --
	( 60.20,141.62);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] ( 60.20,152.94) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.29,167.28) --
	( 66.29,150.64);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] ( 66.29,158.96) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.39,166.76) --
	( 72.39,144.26);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] ( 72.39,155.51) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.49,162.02) --
	( 78.49, 98.30);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] ( 78.49,130.16) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.59,202.23) --
	( 84.59,189.46);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] ( 84.59,195.84) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.69,203.28) --
	( 90.69,181.04);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] ( 90.69,192.16) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.79,240.71) --
	( 96.79,222.93);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] ( 96.79,231.82) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] ( 54.10,176.33) --
	(102.88,176.33);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 78.49,284.14) {1.1\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (108.98,  0.00) --
	(108.98,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (115.08,192.84) --
	(115.08,179.45);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (115.08,186.14) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.18,182.97) --
	(121.18,171.24);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (121.18,177.11) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.28,184.85) --
	(127.28,171.12);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (127.28,177.99) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.38,190.36) --
	(133.38,166.22);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (133.38,178.29) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.47,169.54) --
	(139.47,147.07);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (139.47,158.31) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (145.57,195.26) --
	(145.57,186.84);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (145.57,191.05) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.67,257.34) --
	(151.67,213.88);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (151.67,235.61) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.77,242.16) --
	(157.77,219.28);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (157.77,230.72) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (115.08,193.08) --
	(163.87,193.08);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (139.47,284.14) {7.4\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (169.97,  0.00) --
	(169.97,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.06,183.40) --
	(176.06,181.91);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (176.06,182.66) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.16,178.32) --
	(182.16,176.29);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (182.16,177.30) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.26,173.36) --
	(188.26,171.68);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (188.26,172.52) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.36,167.30) --
	(194.36,164.77);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (194.36,166.03) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.46,175.58) --
	(200.46,173.46);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (200.46,174.52) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.56,188.62) --
	(206.56,187.01);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (206.56,187.82) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.65,191.38) --
	(212.65,187.66);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (212.65,189.52) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.75,221.94) --
	(218.75,219.83);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (218.75,220.89) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (176.06,184.37) --
	(224.85,184.37);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (200.46,284.14) {4.1\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (230.95,  0.00) --
	(230.95,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.05,181.65) --
	(237.05,172.54);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (237.05,177.10) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.15,159.01) --
	(243.15,132.62);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (243.15,145.81) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.24,177.17) --
	(249.24,157.93);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (249.24,167.55) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.34,189.97) --
	(255.34,147.35);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (255.34,168.66) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,177.18) --
	(261.44,164.80);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (261.44,170.99) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (267.54,208.69) --
	(267.54,185.32);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (267.54,197.01) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.64,218.20) --
	(273.64,198.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (273.64,208.22) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.74,223.20) --
	(279.74,214.78);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (279.74,218.99) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (237.05,182.72) --
	(285.83,182.72);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (261.44,284.14) {3.5\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (291.93,  0.00) --
	(291.93,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.03,178.61) --
	(298.03,167.68);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (298.03,173.15) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.13,162.68) --
	(304.13,139.42);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (304.13,151.05) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.23,176.48) --
	(310.23,158.35);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (310.23,167.41) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.33,195.13) --
	(316.33,158.87);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (316.33,177.00) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.42,176.14) --
	(322.42,165.55);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (322.42,170.84) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.52,208.50) --
	(328.52,185.79);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (328.52,197.14) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.62,224.16) --
	(334.62,192.88);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (334.62,208.52) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.72,226.54) --
	(340.72,218.12);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (340.72,222.33) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (298.03,184.32) --
	(346.82,184.32);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (322.42,284.14) {4.1\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (352.92,  0.00) --
	(352.92,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.01,182.24) --
	(359.01,167.41);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (359.01,174.82) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.11,157.24) --
	(365.11,131.60);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (365.11,144.42) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (371.21,169.99) --
	(371.21,153.88);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (371.21,161.94) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.31,168.23) --
	(377.31,146.43);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (377.31,157.33) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.41,174.11) --
	(383.41,150.84);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (383.41,162.48) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.51,193.90) --
	(389.51,166.47);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (389.51,180.18) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.60,203.16) --
	(395.60,181.99);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (395.60,192.57) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.70,239.78) --
	(401.70,221.53);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (401.70,230.66) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (359.01,176.71) --
	(407.80,176.71);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (383.41,284.14) {1.3\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (413.90,  0.00) --
	(413.90,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.00,189.40) --
	(420.00,183.99);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (420.00,186.70) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.10,179.73) --
	(426.10,176.02);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (426.10,177.87) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (432.19,178.20) --
	(432.19,174.37);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (432.19,176.28) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.29,173.45) --
	(438.29,169.46);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (438.29,171.45) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.39,177.55) --
	(444.39,172.82);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (444.39,175.18) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (450.49,198.72) --
	(450.49,192.07);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (450.49,195.40) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.59,191.62) --
	(456.59,186.36);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (456.59,188.99) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.69,229.29) --
	(462.69,223.00);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (462.69,226.15) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (420.00,187.76) --
	(468.78,187.76);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (444.39,284.14) {5.4\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (474.88,  0.00) --
	(474.88,357.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- (554.16, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.49, 60.00) -- ( 78.49, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.47, 60.00) -- (139.47, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.46, 60.00) -- (200.46, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44, 60.00) -- (261.44, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.42, 60.00) -- (322.42, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.41, 60.00) -- (383.41, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.39, 60.00) -- (444.39, 54.00);

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at ( 80.56, 48.00) {Prior};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (141.54, 48.00) {Likelihood};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (202.52, 48.00) {Leaf rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (263.51, 48.00) {Mean branch rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (324.49, 48.00) {Tree length};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (385.47, 48.00) {Clock SD $\sigma$};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (446.46, 48.00) {HKY model $\kappa$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 48.00,286.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 42.00, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 92.88) -- ( 42.00, 92.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,135.92) -- ( 42.00,135.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,173.19) -- ( 42.00,173.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,210.47) -- ( 42.00,210.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,253.51) -- ( 42.00,253.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,286.39) -- ( 42.00,286.39);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 56.56) {$\frac{2}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 89.44) {$\frac{3}{4}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,132.47) {$\frac{7}{8}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,169.75) {$1\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,207.03) {$\frac{8}{7}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,250.06) {$\frac{4}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,282.94) {$\frac{3}{2}\times$};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,235.67) rectangle (494.39,229.67);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,223.67) rectangle (494.39,217.67);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,211.67) rectangle (494.39,205.67);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,199.67) rectangle (494.39,193.67);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,187.67) rectangle (494.39,181.67);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,175.67) rectangle (494.39,169.67);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,163.67) rectangle (494.39,157.67);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,151.67) rectangle (494.39,145.67);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,229.22) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,217.22) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,205.22) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,193.22) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,181.22) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,169.22) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,157.22) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,145.22) {Moyle};
\end{scope}
\end{tikzpicture}

\end{document}