### 2020-08-31 benchmarking/benchmarkingVM/ESS_round5.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-31 17:27:26
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,357.33);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (301.08,335.19) {\bfseries Comparison of AVMVN+NER+Bactrian+adapt(\textit{real}) and NER+Bactrian+adapt(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (  9.60,190.67) {Relative ESS/hr};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 48.00,173.19) --
	(474.88,173.19);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.10,170.57) --
	( 54.10,160.89);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] ( 54.10,165.73) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.20,193.03) --
	( 60.20,169.27);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] ( 60.20,181.15) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.29,191.40) --
	( 66.29,174.61);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] ( 66.29,183.01) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.39,196.48) --
	( 72.39,170.97);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] ( 72.39,183.73) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.49,141.00) --
	( 78.49, 76.27);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] ( 78.49,108.63) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.59,183.38) --
	( 84.59,168.50);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] ( 84.59,175.94) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.69,170.18) --
	( 90.69,143.82);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] ( 90.69,157.00) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.79,196.84) --
	( 96.79,181.84);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] ( 96.79,189.34) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] ( 54.10,169.09) --
	(102.88,169.09);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 78.49,284.14) {-1.5\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (108.98,  0.00) --
	(108.98,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (115.08,194.11) --
	(115.08,179.13);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (115.08,186.62) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.18,185.55) --
	(121.18,175.29);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (121.18,180.42) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.28,214.73) --
	(127.28,201.99);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (127.28,208.36) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.38,249.42) --
	(133.38,231.42);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (133.38,240.42) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.47,159.92) --
	(139.47,140.75);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (139.47,150.33) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (145.57,214.01) --
	(145.57,205.08);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (145.57,209.54) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.67,200.10) --
	(151.67,107.60);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (151.67,153.85) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.77,178.44) --
	(157.77,154.84);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (157.77,166.64) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (115.08,188.55) --
	(163.87,188.55);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (139.47,284.14) {5.7\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (169.97,  0.00) --
	(169.97,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.06,172.66) --
	(176.06,170.84);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (176.06,171.75) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.16,169.79) --
	(182.16,167.75);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (182.16,168.77) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (188.26,189.75) --
	(188.26,187.44);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (188.26,188.59) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.36,203.19) --
	(194.36,199.69);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (194.36,201.44) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.46,171.28) --
	(200.46,169.79);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (200.46,170.53) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.56,184.57) --
	(206.56,182.27);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (206.56,183.42) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.65,163.32) --
	(212.65,160.43);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (212.65,161.87) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.75,155.78) --
	(218.75,154.17);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (218.75,154.97) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (176.06,175.53) --
	(224.85,175.53);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (200.46,284.14) {0.84\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (230.95,  0.00) --
	(230.95,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.05,218.83) --
	(237.05,206.46);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (237.05,212.65) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.15,242.97) --
	(243.15,220.02);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (243.15,231.49) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.24,208.35) --
	(249.24,192.32);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (249.24,200.33) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.34,176.56) --
	(255.34,142.47);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (255.34,159.52) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,187.62) --
	(261.44,174.61);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (261.44,181.12) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (267.54,206.77) --
	(267.54,181.77);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (267.54,194.27) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.64,229.57) --
	(273.64,209.79);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (273.64,219.68) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.74,216.31) --
	(279.74,206.19);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (279.74,211.25) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (237.05,202.10) --
	(285.83,202.10);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (261.44,284.14) {11\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (291.93,  0.00) --
	(291.93,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.03,209.96) --
	(298.03,198.22);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (298.03,204.09) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.13,238.47) --
	(304.13,217.31);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (304.13,227.89) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.23,204.06) --
	(310.23,189.08);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (310.23,196.57) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.33,171.91) --
	(316.33,144.86);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (316.33,158.38) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.42,183.17) --
	(322.42,172.59);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (322.42,177.88) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.52,206.07) --
	(328.52,181.21);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (328.52,193.64) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.62,210.76) --
	(334.62,182.47);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (334.62,196.61) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.72,208.61) --
	(340.72,197.60);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (340.72,203.11) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (298.03,195.41) --
	(346.82,195.41);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (322.42,284.14) {8.3\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (352.92,  0.00) --
	(352.92,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.01,174.36) --
	(359.01,162.95);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (359.01,168.66) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.11,189.08) --
	(365.11,165.88);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (365.11,177.48) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (371.21,190.22) --
	(371.21,174.54);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (371.21,182.38) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.31,180.24) --
	(377.31,143.99);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (377.31,162.11) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.41,146.90) --
	(383.41,126.33);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (383.41,136.61) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.51,194.96) --
	(389.51,169.70);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (389.51,182.33) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.60,175.08) --
	(395.60,145.39);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (395.60,160.24) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.70,198.72) --
	(401.70,181.55);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (401.70,190.14) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (359.01,170.44) --
	(407.80,170.44);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (383.41,284.14) {-0.98\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (413.90,  0.00) --
	(413.90,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.00,167.61) --
	(420.00,161.93);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (420.00,164.77) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.10,164.25) --
	(426.10,160.41);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (426.10,162.33) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (432.19,168.36) --
	(432.19,165.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (432.19,166.80) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.29,170.98) --
	(438.29,165.44);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (438.29,168.21) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.39,161.65) --
	(444.39,156.22);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (444.39,158.94) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (450.49,162.17) --
	(450.49,154.52);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (450.49,158.35) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.59,155.49) --
	(456.59,149.68);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (456.59,152.58) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.69,166.82) --
	(462.69,162.55);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (462.69,164.69) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (420.00,162.13) --
	(468.78,162.13);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (444.39,284.14) {-3.9\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (474.88,  0.00) --
	(474.88,357.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- (554.16, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.49, 60.00) -- ( 78.49, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.47, 60.00) -- (139.47, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.46, 60.00) -- (200.46, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44, 60.00) -- (261.44, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.42, 60.00) -- (322.42, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.41, 60.00) -- (383.41, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.39, 60.00) -- (444.39, 54.00);

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at ( 80.56, 48.00) {Prior};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (141.54, 48.00) {Likelihood};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (202.52, 48.00) {Leaf rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (263.51, 48.00) {Mean branch rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (324.49, 48.00) {Tree length};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (385.47, 48.00) {Clock SD $\sigma$};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (446.46, 48.00) {HKY model $\kappa$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 48.00,286.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 42.00, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 92.88) -- ( 42.00, 92.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,135.92) -- ( 42.00,135.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,173.19) -- ( 42.00,173.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,210.47) -- ( 42.00,210.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,253.51) -- ( 42.00,253.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,286.39) -- ( 42.00,286.39);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 56.56) {$\frac{2}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 89.44) {$\frac{3}{4}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,132.47) {$\frac{7}{8}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,169.75) {$1\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,207.03) {$\frac{8}{7}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,250.06) {$\frac{4}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,282.94) {$\frac{3}{2}\times$};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,235.67) rectangle (494.39,229.67);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,223.67) rectangle (494.39,217.67);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,211.67) rectangle (494.39,205.67);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,199.67) rectangle (494.39,193.67);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,187.67) rectangle (494.39,181.67);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,175.67) rectangle (494.39,169.67);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,163.67) rectangle (494.39,157.67);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,151.67) rectangle (494.39,145.67);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,229.22) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,217.22) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,205.22) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,193.22) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,181.22) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,169.22) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,157.22) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,145.22) {Moyle};
\end{scope}
\end{tikzpicture}

\end{document}