### 2020-08-31 benchmarking/benchmarkingVM/ESS_timesettings.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-31 17:27:56
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,357.33);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (301.08,328.59) {\bfseries Relative speed compared with \textit{cat}};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,184.67) {Relative samples/hr};
\end{scope}
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (552.96,308.13);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 49.20,184.67) --
	(458.76,184.67);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] ( 53.30,192.21) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] ( 57.39,192.33) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] ( 61.49,192.07) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] ( 65.58,189.41) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] ( 69.68,201.89) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] ( 73.77,192.33) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] ( 77.87,193.68) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] ( 81.96,194.75) circle (  2.25);

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] ( 53.30,193.62) --
	( 86.06,193.62);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at ( 69.68,297.24) {5.2\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] ( 90.16,  0.00) --
	( 90.16,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] ( 94.25,131.15) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] ( 98.35,142.51) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (102.44,137.12) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (106.54,133.51) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (110.63,115.30) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (114.73,104.29) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (118.83,121.85) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (122.92, 87.01) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] ( 94.25,122.43) --
	(127.02,122.43);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (110.63,297.24) {29\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (131.11,  0.00) --
	(131.11,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (135.21,210.25) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (139.30,199.78) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (143.40,200.44) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (147.49,195.49) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (151.59,212.96) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (155.69,234.21) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (159.78,210.45) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (163.88,228.46) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (135.21,211.98) --
	(167.97,211.98);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (151.59,297.24) {17\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (172.07,  0.00) --
	(172.07,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (176.16,137.47) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (180.26,147.34) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (184.36,140.19) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (188.45,140.07) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (192.55,140.34) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (196.64,103.65) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (200.74,122.50) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (204.83,113.14) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (176.16,131.17) --
	(208.93,131.17);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (192.55,297.24) {26\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (213.02,  0.00) --
	(213.02,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (217.12,221.94) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (221.22,222.13) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (225.31,218.65) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (229.41,207.34) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (233.50,217.59) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (237.60,198.68) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (241.69,213.67) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (245.79,189.78) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (217.12,211.56) --
	(249.88,211.56);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (233.50,297.24) {16\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (253.98,  0.00) --
	(253.98,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (258.08,229.80) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (262.17,227.84) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (266.27,228.47) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (270.36,216.45) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (274.46,237.37) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (278.55,221.39) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (282.65,219.64) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (286.75,212.76) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (258.08,224.38) --
	(290.84,224.38);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (274.46,297.24) {25\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (294.94,  0.00) --
	(294.94,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (299.03,243.36) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (303.13,239.88) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (307.22,240.25) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (311.32,228.38) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (315.41,250.89) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (319.51,247.98) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (323.61,232.84) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (327.70,224.69) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (299.03,238.74) --
	(331.80,238.74);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (315.41,297.24) {35\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (335.89,  0.00) --
	(335.89,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (339.99,133.85) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (344.08,146.75) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (348.18,141.24) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (352.28,140.02) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (356.37,135.64) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (360.47,103.87) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (364.56,119.49) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (368.66,109.12) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (339.99,129.36) --
	(372.75,129.36);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (356.37,297.24) {27\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (376.85,  0.00) --
	(376.85,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (380.94,141.02) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (385.04,149.13) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (389.14,142.97) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (393.23,139.61) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (397.33,137.57) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (401.42,114.39) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (405.52,129.13) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (409.61,140.48) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (380.94,137.06) --
	(413.71,137.06);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (397.33,297.24) {23\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (417.80,  0.00) --
	(417.80,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (421.90,147.24) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (426.00,156.87) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (430.09,147.29) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (434.19,143.53) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (438.28,144.92) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (442.38,124.70) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (446.47,142.61) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (450.57,147.18) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (421.90,144.49) --
	(454.67,144.49);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (438.28,297.24) {20\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (458.76,  0.00) --
	(458.76,357.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 61.20) -- (552.96, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 69.68, 61.20) -- ( 69.68, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.63, 61.20) -- (110.63, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.59, 61.20) -- (151.59, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (192.55, 61.20) -- (192.55, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.50, 61.20) -- (233.50, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.46, 61.20) -- (274.46, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.41, 61.20) -- (315.41, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.37, 61.20) -- (356.37, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (397.33, 61.20) -- (397.33, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.28, 61.20) -- (438.28, 55.20);

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at ( 71.74, 49.20) {adapt(\textit{cat})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (112.70, 49.20) {nocons(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (153.66, 49.20) {cons(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (194.61, 49.20) {adapt(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (235.57, 49.20) {nocons(\textit{quant})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (276.52, 49.20) {cons(\textit{quant})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (317.48, 49.20) {adapt(\textit{quant})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (358.44, 49.20) {Bactrian};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (399.39, 49.20) {NER};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (440.35, 49.20) {AVMVN};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 61.20) -- ( 49.20,308.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 61.20) -- ( 43.20, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,133.42) -- ( 43.20,133.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,184.67) -- ( 43.20,184.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,235.91) -- ( 43.20,235.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,308.13) -- ( 43.20,308.13);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20, 57.76) {$\frac{1}{2}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,129.98) {$\frac{3}{4}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,181.22) {$1\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,232.46) {$\frac{4}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,304.69) {$\frac{2}{1}\times$};
\end{scope}
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (552.96,308.13);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,229.67) rectangle (493.19,223.67);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,217.67) rectangle (493.19,211.67);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,205.67) rectangle (493.19,199.67);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,193.67) rectangle (493.19,187.67);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,181.67) rectangle (493.19,175.67);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,169.67) rectangle (493.19,163.67);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,157.67) rectangle (493.19,151.67);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,145.67) rectangle (493.19,139.67);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,223.22) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,211.22) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,199.22) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,187.22) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,175.22) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,163.22) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,151.22) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,139.22) {Moyle};
\end{scope}
\end{tikzpicture}

\end{document}