### 2020-09-01 RatesWCSS/WCSS_AVMVN.tex
complete draft

% Created by tikzDevice version 0.12.3.1 on 2020-09-01 10:43:33
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,385.44) rectangle (192.72,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,559.19) {\bfseries Tree height $h$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47,425.83) rectangle (176.09,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47,425.83) --
	(176.09,545.69);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.50,452.51) --
	( 65.50,455.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.05,431.38) --
	( 42.05,439.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.12,452.89) --
	( 66.12,455.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.24,447.06) --
	( 60.24,451.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.61,461.83) --
	( 78.61,465.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.79,442.39) --
	( 52.79,444.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.14,433.32) --
	( 46.14,441.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.48,430.31) --
	( 38.48,431.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.07,427.02) --
	( 34.07,427.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.14,434.49) --
	( 44.14,439.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.61,431.14) --
	( 39.61,432.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 47.43,436.82) --
	( 47.43,438.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.33,429.75) --
	( 37.33,430.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.22,483.96) --
	(128.22,523.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.70,437.65) --
	( 46.70,438.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.97,433.27) --
	( 42.97,435.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.46,442.55) --
	( 62.46,469.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.87,443.34) --
	( 54.87,449.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.97,429.18) --
	( 36.97,430.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.34,430.38) --
	( 38.34,431.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.56,441.34) --
	( 52.56,443.68);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.80,433.61) --
	( 51.80,438.31);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.44,447.20) --
	( 60.44,450.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.29,456.18) --
	( 77.29,473.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.08,443.45) --
	( 54.08,445.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.68,435.11) --
	( 44.68,436.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.95,431.17) --
	( 39.95,433.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.47,444.63) --
	( 56.47,447.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.90,462.65) --
	( 76.90,466.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.51,431.86) --
	( 41.51,433.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.22,438.12) --
	( 48.22,443.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.69,456.13) --
	( 74.69,462.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.27,427.08) --
	( 34.27,427.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 53.39,440.46) --
	( 53.39,444.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.21,432.77) --
	( 41.21,433.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.34,431.94) --
	( 40.34,432.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.55,433.22) --
	( 46.55,448.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.28,430.15) --
	( 38.28,430.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.51,454.21) --
	( 66.51,458.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.33,442.90) --
	( 54.33,444.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 69.08,455.12) --
	( 69.08,457.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.72,426.74) --
	( 33.72,427.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.02,441.49) --
	( 52.02,443.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 47.67,437.76) --
	( 47.67,439.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.25,460.87) --
	( 77.25,464.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.23,430.74) --
	( 39.23,432.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.69,433.38) --
	( 42.69,434.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.60,438.64) --
	( 48.60,440.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.71,437.34) --
	( 46.71,438.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.57,430.81) --
	( 42.57,435.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.32,431.71) --
	( 40.32,432.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.21,438.08) --
	( 48.21,440.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.64,431.69) --
	( 43.64,438.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.33,428.03) --
	( 35.33,429.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.32,428.82) --
	( 36.32,429.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.82,432.93) --
	( 41.82,434.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.77,468.35) --
	( 86.77,479.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 47.87,437.92) --
	( 47.87,445.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.72,429.85) --
	( 37.72,430.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.79,451.82) --
	( 65.79,454.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.12,430.15) --
	( 38.12,431.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.30,446.67) --
	( 59.30,448.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 47.93,437.19) --
	( 47.93,438.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.19,432.93) --
	( 44.19,435.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.98,477.76) --
	( 96.98,485.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.10,432.74) --
	( 42.10,461.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.15,431.78) --
	( 40.15,432.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.10,446.71) --
	( 59.10,449.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.93,441.54) --
	( 51.93,442.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.10,434.53) --
	( 44.10,436.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.92,433.02) --
	( 41.92,434.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.35,445.92) --
	( 60.35,451.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.27,427.66) --
	( 44.27,473.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.16,434.68) --
	( 44.16,435.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.86,457.86) --
	( 74.86,463.06);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.74,428.92) --
	( 35.74,430.63);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.93,431.40) --
	( 46.93,449.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.30,431.69) --
	( 41.30,433.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.15,428.25) --
	( 36.15,429.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.81,433.11) --
	( 44.81,439.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.80,436.40) --
	( 45.80,437.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.15,433.33) --
	( 42.15,434.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.82,443.24) --
	( 54.82,445.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.82,431.59) --
	( 41.82,436.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.92,431.04) --
	( 38.92,431.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.94,457.56) --
	( 70.94,465.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.35,435.12) --
	( 45.35,438.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.88,428.25) --
	( 36.88,431.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.49,443.61) --
	( 54.49,445.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.82,427.28) --
	( 36.82,430.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.84,432.89) --
	( 43.84,439.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.97,436.45) --
	( 46.97,440.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.46,431.48) --
	( 39.46,432.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.02,427.55) --
	( 35.02,428.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.33,433.76) --
	( 42.33,434.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.57,453.74) --
	( 68.57,458.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.58,458.93) --
	( 72.58,461.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.03,428.53) --
	( 36.03,429.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.61,429.73) --
	( 37.61,430.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 64.66,451.09) --
	( 64.66,488.18);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 65.50,453.86) circle (  1.49);

\path[fill=fillColor] ( 42.05,434.83) circle (  1.49);

\path[fill=fillColor] ( 66.12,454.14) circle (  1.49);

\path[fill=fillColor] ( 60.24,449.13) circle (  1.49);

\path[fill=fillColor] ( 78.61,463.72) circle (  1.49);

\path[fill=fillColor] ( 52.79,443.26) circle (  1.49);

\path[fill=fillColor] ( 46.14,436.79) circle (  1.49);

\path[fill=fillColor] ( 38.48,430.75) circle (  1.49);

\path[fill=fillColor] ( 34.07,427.19) circle (  1.49);

\path[fill=fillColor] ( 44.14,436.70) circle (  1.49);

\path[fill=fillColor] ( 39.61,431.66) circle (  1.49);

\path[fill=fillColor] ( 47.43,437.65) circle (  1.49);

\path[fill=fillColor] ( 37.33,430.25) circle (  1.49);

\path[fill=fillColor] (128.22,503.09) circle (  1.49);

\path[fill=fillColor] ( 46.70,438.26) circle (  1.49);

\path[fill=fillColor] ( 42.97,434.25) circle (  1.49);

\path[fill=fillColor] ( 62.46,454.66) circle (  1.49);

\path[fill=fillColor] ( 54.87,446.28) circle (  1.49);

\path[fill=fillColor] ( 36.97,429.75) circle (  1.49);

\path[fill=fillColor] ( 38.34,430.81) circle (  1.49);

\path[fill=fillColor] ( 52.56,442.53) circle (  1.49);

\path[fill=fillColor] ( 51.80,435.63) circle (  1.49);

\path[fill=fillColor] ( 60.44,448.71) circle (  1.49);

\path[fill=fillColor] ( 77.29,464.54) circle (  1.49);

\path[fill=fillColor] ( 54.08,444.37) circle (  1.49);

\path[fill=fillColor] ( 44.68,435.98) circle (  1.49);

\path[fill=fillColor] ( 39.95,432.24) circle (  1.49);

\path[fill=fillColor] ( 56.47,446.00) circle (  1.49);

\path[fill=fillColor] ( 76.90,464.60) circle (  1.49);

\path[fill=fillColor] ( 41.51,432.93) circle (  1.49);

\path[fill=fillColor] ( 48.22,440.69) circle (  1.49);

\path[fill=fillColor] ( 74.69,459.51) circle (  1.49);

\path[fill=fillColor] ( 34.27,427.27) circle (  1.49);

\path[fill=fillColor] ( 53.39,442.70) circle (  1.49);

\path[fill=fillColor] ( 41.21,433.19) circle (  1.49);

\path[fill=fillColor] ( 40.34,432.36) circle (  1.49);

\path[fill=fillColor] ( 46.55,439.53) circle (  1.49);

\path[fill=fillColor] ( 38.28,430.45) circle (  1.49);

\path[fill=fillColor] ( 66.51,456.26) circle (  1.49);

\path[fill=fillColor] ( 54.33,443.72) circle (  1.49);

\path[fill=fillColor] ( 69.08,456.42) circle (  1.49);

\path[fill=fillColor] ( 33.72,426.93) circle (  1.49);

\path[fill=fillColor] ( 52.02,442.58) circle (  1.49);

\path[fill=fillColor] ( 47.67,438.43) circle (  1.49);

\path[fill=fillColor] ( 77.25,462.87) circle (  1.49);

\path[fill=fillColor] ( 39.23,431.63) circle (  1.49);

\path[fill=fillColor] ( 42.69,434.04) circle (  1.49);

\path[fill=fillColor] ( 48.60,439.49) circle (  1.49);

\path[fill=fillColor] ( 46.71,437.94) circle (  1.49);

\path[fill=fillColor] ( 42.57,432.95) circle (  1.49);

\path[fill=fillColor] ( 40.32,432.10) circle (  1.49);

\path[fill=fillColor] ( 48.21,439.11) circle (  1.49);

\path[fill=fillColor] ( 43.64,434.87) circle (  1.49);

\path[fill=fillColor] ( 35.33,428.62) circle (  1.49);

\path[fill=fillColor] ( 36.32,429.10) circle (  1.49);

\path[fill=fillColor] ( 41.82,433.57) circle (  1.49);

\path[fill=fillColor] ( 86.77,473.75) circle (  1.49);

\path[fill=fillColor] ( 47.87,441.73) circle (  1.49);

\path[fill=fillColor] ( 37.72,430.25) circle (  1.49);

\path[fill=fillColor] ( 65.79,453.08) circle (  1.49);

\path[fill=fillColor] ( 38.12,430.65) circle (  1.49);

\path[fill=fillColor] ( 59.30,447.69) circle (  1.49);

\path[fill=fillColor] ( 47.93,438.07) circle (  1.49);

\path[fill=fillColor] ( 44.19,434.20) circle (  1.49);

\path[fill=fillColor] ( 96.98,481.57) circle (  1.49);

\path[fill=fillColor] ( 42.10,444.28) circle (  1.49);

\path[fill=fillColor] ( 40.15,432.13) circle (  1.49);

\path[fill=fillColor] ( 59.10,448.27) circle (  1.49);

\path[fill=fillColor] ( 51.93,442.20) circle (  1.49);

\path[fill=fillColor] ( 44.10,435.42) circle (  1.49);

\path[fill=fillColor] ( 41.92,433.67) circle (  1.49);

\path[fill=fillColor] ( 60.35,448.74) circle (  1.49);

\path[fill=fillColor] ( 44.27,443.40) circle (  1.49);

\path[fill=fillColor] ( 44.16,435.17) circle (  1.49);

\path[fill=fillColor] ( 74.86,460.42) circle (  1.49);

\path[fill=fillColor] ( 35.74,429.73) circle (  1.49);

\path[fill=fillColor] ( 46.93,438.90) circle (  1.49);

\path[fill=fillColor] ( 41.30,432.63) circle (  1.49);

\path[fill=fillColor] ( 36.15,428.79) circle (  1.49);

\path[fill=fillColor] ( 44.81,436.15) circle (  1.49);

\path[fill=fillColor] ( 45.80,436.96) circle (  1.49);

\path[fill=fillColor] ( 42.15,433.89) circle (  1.49);

\path[fill=fillColor] ( 54.82,444.12) circle (  1.49);

\path[fill=fillColor] ( 41.82,434.14) circle (  1.49);

\path[fill=fillColor] ( 38.92,431.40) circle (  1.49);

\path[fill=fillColor] ( 70.94,461.51) circle (  1.49);

\path[fill=fillColor] ( 45.35,436.63) circle (  1.49);

\path[fill=fillColor] ( 36.88,429.85) circle (  1.49);

\path[fill=fillColor] ( 54.49,444.39) circle (  1.49);

\path[fill=fillColor] ( 36.82,428.89) circle (  1.49);

\path[fill=fillColor] ( 43.84,435.53) circle (  1.49);

\path[fill=fillColor] ( 46.97,438.48) circle (  1.49);

\path[fill=fillColor] ( 39.46,432.02) circle (  1.49);

\path[fill=fillColor] ( 35.02,427.78) circle (  1.49);

\path[fill=fillColor] ( 42.33,434.30) circle (  1.49);

\path[fill=fillColor] ( 68.57,455.97) circle (  1.49);

\path[fill=fillColor] ( 72.58,460.32) circle (  1.49);

\path[fill=fillColor] ( 36.03,428.89) circle (  1.49);

\path[fill=fillColor] ( 37.61,430.05) circle (  1.49);

\path[fill=fillColor] ( 64.66,467.58) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,547.27) {Coverage: 98\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- (151.38,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 32.47,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.20,425.83) -- ( 62.20,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.93,425.83) -- ( 91.93,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.66,425.83) -- (121.66,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.38,425.83) -- (151.38,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47,411.58) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 62.20,411.58) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 91.93,411.58) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (121.66,411.58) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (151.38,411.58) {2.0};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 32.47,525.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 28.51,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,450.64) -- ( 28.51,450.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,475.45) -- ( 28.51,475.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,500.26) -- ( 28.51,500.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,525.07) -- ( 28.51,525.07);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,423.56) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,448.37) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,473.18) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,497.99) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,522.80) {2.0};
\end{scope}
\begin{scope}
\path[clip] (192.72,385.44) rectangle (385.44,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (297.00,559.19) {\bfseries Yule birth rate $\lambda$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (297.00,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (199.85,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (225.19,425.83) rectangle (368.81,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (225.19,425.83) --
	(368.81,545.69);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.99,428.72) --
	(232.99,431.90);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.04,432.04) --
	(240.04,441.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.81,427.89) --
	(228.81,430.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.44,428.91) --
	(231.44,432.38);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.85,427.16) --
	(228.85,428.63);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.73,430.21) --
	(232.73,435.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.11,432.04) --
	(238.11,439.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.85,435.06) --
	(240.85,445.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (288.03,462.02) --
	(288.03,504.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.45,431.73) --
	(237.45,438.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.16,434.47) --
	(241.16,444.22);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.10,431.55) --
	(240.10,438.10);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.37,440.20) --
	(245.37,456.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.76,426.73) --
	(226.76,427.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.39,429.79) --
	(231.39,434.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.67,434.51) --
	(241.67,444.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.17,427.70) --
	(230.17,431.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.31,429.11) --
	(232.31,432.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.41,441.94) --
	(264.41,460.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.04,434.90) --
	(247.04,445.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.93,430.13) --
	(234.93,435.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.37,430.83) --
	(234.37,436.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.76,428.43) --
	(229.76,431.36);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.27,427.87) --
	(231.27,430.36);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.99,429.86) --
	(233.99,434.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.44,430.24) --
	(234.44,435.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.82,435.46) --
	(239.82,446.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.16,429.71) --
	(231.16,434.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.50,427.32) --
	(228.50,428.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.13,435.95) --
	(242.13,447.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.68,428.96) --
	(230.68,432.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.86,428.45) --
	(229.86,431.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (309.02,465.14) --
	(309.02,512.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.18,429.17) --
	(232.18,432.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.93,433.15) --
	(238.93,441.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.28,433.05) --
	(242.28,441.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.91,428.98) --
	(231.91,436.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.63,436.42) --
	(246.63,448.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.43,428.44) --
	(230.43,431.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.62,429.37) --
	(231.62,433.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.70,427.84) --
	(229.70,430.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (289.02,475.95) --
	(289.02,537.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.36,430.61) --
	(233.36,436.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.50,431.04) --
	(235.50,436.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.96,427.32) --
	(227.96,428.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.25,439.53) --
	(253.25,455.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.30,433.52) --
	(240.30,442.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.47,430.53) --
	(235.47,435.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.84,431.18) --
	(235.84,437.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.15,434.44) --
	(240.15,445.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (244.67,434.85) --
	(244.67,445.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.58,430.02) --
	(233.58,434.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.06,431.76) --
	(237.06,439.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (258.46,444.77) --
	(258.46,468.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.31,444.97) --
	(262.31,466.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.53,432.75) --
	(237.53,440.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.94,427.60) --
	(228.94,429.54);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.30,429.38) --
	(236.30,433.83);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.67,437.92) --
	(243.67,451.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.85,427.98) --
	(228.85,430.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.82,438.99) --
	(247.82,454.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.79,428.14) --
	(230.79,430.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.52,430.02) --
	(232.52,434.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.71,432.36) --
	(234.71,439.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.74,426.93) --
	(227.74,428.15);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.13,427.89) --
	(246.13,438.03);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.95,433.06) --
	(237.95,441.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.89,428.40) --
	(229.89,431.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.75,428.87) --
	(231.75,432.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.49,431.67) --
	(239.49,438.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.38,431.37) --
	(234.38,437.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.71,428.62) --
	(231.71,431.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.28,426.42) --
	(241.28,448.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.45,432.03) --
	(240.45,439.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.76,427.30) --
	(227.76,428.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.65,442.89) --
	(255.65,463.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.61,428.68) --
	(238.61,438.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.31,433.41) --
	(239.31,442.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.83,447.95) --
	(257.83,475.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.62,431.50) --
	(235.62,439.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.37,431.16) --
	(237.37,437.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.47,432.40) --
	(241.47,439.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.24,428.93) --
	(232.24,432.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.72,433.33) --
	(243.72,443.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.72,435.31) --
	(241.72,445.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.86,427.40) --
	(227.86,429.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.59,430.94) --
	(235.59,436.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.44,438.37) --
	(253.44,460.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.18,428.22) --
	(229.18,430.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.07,439.11) --
	(255.07,476.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.28,430.29) --
	(234.28,435.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.10,429.99) --
	(235.10,434.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.06,436.73) --
	(249.06,449.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.87,452.20) --
	(261.87,483.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.19,432.42) --
	(240.19,439.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.99,428.72) --
	(228.99,431.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.48,427.43) --
	(228.48,429.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.45,448.04) --
	(276.45,474.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.17,439.06) --
	(243.17,453.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.40,427.25) --
	(229.40,429.76);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (232.99,430.22) circle (  1.49);

\path[fill=fillColor] (240.04,436.82) circle (  1.49);

\path[fill=fillColor] (228.81,429.01) circle (  1.49);

\path[fill=fillColor] (231.44,430.58) circle (  1.49);

\path[fill=fillColor] (228.85,427.88) circle (  1.49);

\path[fill=fillColor] (232.73,432.64) circle (  1.49);

\path[fill=fillColor] (238.11,435.73) circle (  1.49);

\path[fill=fillColor] (240.85,440.38) circle (  1.49);

\path[fill=fillColor] (288.03,481.88) circle (  1.49);

\path[fill=fillColor] (237.45,435.22) circle (  1.49);

\path[fill=fillColor] (241.16,439.33) circle (  1.49);

\path[fill=fillColor] (240.10,434.80) circle (  1.49);

\path[fill=fillColor] (245.37,447.87) circle (  1.49);

\path[fill=fillColor] (226.76,427.24) circle (  1.49);

\path[fill=fillColor] (231.39,431.98) circle (  1.49);

\path[fill=fillColor] (241.67,439.47) circle (  1.49);

\path[fill=fillColor] (230.17,429.38) circle (  1.49);

\path[fill=fillColor] (232.31,430.93) circle (  1.49);

\path[fill=fillColor] (264.41,450.77) circle (  1.49);

\path[fill=fillColor] (247.04,439.76) circle (  1.49);

\path[fill=fillColor] (234.93,432.53) circle (  1.49);

\path[fill=fillColor] (234.37,433.69) circle (  1.49);

\path[fill=fillColor] (229.76,429.87) circle (  1.49);

\path[fill=fillColor] (231.27,429.11) circle (  1.49);

\path[fill=fillColor] (233.99,431.98) circle (  1.49);

\path[fill=fillColor] (234.44,432.76) circle (  1.49);

\path[fill=fillColor] (239.82,440.82) circle (  1.49);

\path[fill=fillColor] (231.16,431.83) circle (  1.49);

\path[fill=fillColor] (228.50,428.11) circle (  1.49);

\path[fill=fillColor] (242.13,441.70) circle (  1.49);

\path[fill=fillColor] (230.68,430.80) circle (  1.49);

\path[fill=fillColor] (229.86,429.88) circle (  1.49);

\path[fill=fillColor] (309.02,488.18) circle (  1.49);

\path[fill=fillColor] (232.18,431.05) circle (  1.49);

\path[fill=fillColor] (238.93,437.20) circle (  1.49);

\path[fill=fillColor] (242.28,437.13) circle (  1.49);

\path[fill=fillColor] (231.91,432.46) circle (  1.49);

\path[fill=fillColor] (246.63,442.08) circle (  1.49);

\path[fill=fillColor] (230.43,429.86) circle (  1.49);

\path[fill=fillColor] (231.62,431.31) circle (  1.49);

\path[fill=fillColor] (229.70,428.95) circle (  1.49);

\path[fill=fillColor] (289.02,505.74) circle (  1.49);

\path[fill=fillColor] (233.36,433.33) circle (  1.49);

\path[fill=fillColor] (235.50,433.93) circle (  1.49);

\path[fill=fillColor] (227.96,428.09) circle (  1.49);

\path[fill=fillColor] (253.25,447.54) circle (  1.49);

\path[fill=fillColor] (240.30,437.74) circle (  1.49);

\path[fill=fillColor] (235.47,432.98) circle (  1.49);

\path[fill=fillColor] (235.84,434.14) circle (  1.49);

\path[fill=fillColor] (240.15,439.70) circle (  1.49);

\path[fill=fillColor] (244.67,439.98) circle (  1.49);

\path[fill=fillColor] (233.58,432.23) circle (  1.49);

\path[fill=fillColor] (237.06,435.42) circle (  1.49);

\path[fill=fillColor] (258.46,455.82) circle (  1.49);

\path[fill=fillColor] (262.31,455.63) circle (  1.49);

\path[fill=fillColor] (237.53,436.63) circle (  1.49);

\path[fill=fillColor] (228.94,428.54) circle (  1.49);

\path[fill=fillColor] (236.30,431.58) circle (  1.49);

\path[fill=fillColor] (243.67,444.78) circle (  1.49);

\path[fill=fillColor] (228.85,429.13) circle (  1.49);

\path[fill=fillColor] (247.82,446.42) circle (  1.49);

\path[fill=fillColor] (230.79,429.38) circle (  1.49);

\path[fill=fillColor] (232.52,432.36) circle (  1.49);

\path[fill=fillColor] (234.71,435.87) circle (  1.49);

\path[fill=fillColor] (227.74,427.53) circle (  1.49);

\path[fill=fillColor] (246.13,432.64) circle (  1.49);

\path[fill=fillColor] (237.95,437.10) circle (  1.49);

\path[fill=fillColor] (229.89,429.69) circle (  1.49);

\path[fill=fillColor] (231.75,430.57) circle (  1.49);

\path[fill=fillColor] (239.49,434.71) circle (  1.49);

\path[fill=fillColor] (234.38,434.52) circle (  1.49);

\path[fill=fillColor] (231.71,430.03) circle (  1.49);

\path[fill=fillColor] (241.28,435.18) circle (  1.49);

\path[fill=fillColor] (240.45,435.37) circle (  1.49);

\path[fill=fillColor] (227.76,428.12) circle (  1.49);

\path[fill=fillColor] (255.65,453.20) circle (  1.49);

\path[fill=fillColor] (238.61,433.36) circle (  1.49);

\path[fill=fillColor] (239.31,437.66) circle (  1.49);

\path[fill=fillColor] (257.83,461.88) circle (  1.49);

\path[fill=fillColor] (235.62,435.12) circle (  1.49);

\path[fill=fillColor] (237.37,433.89) circle (  1.49);

\path[fill=fillColor] (241.47,436.01) circle (  1.49);

\path[fill=fillColor] (232.24,430.65) circle (  1.49);

\path[fill=fillColor] (243.72,438.26) circle (  1.49);

\path[fill=fillColor] (241.72,440.36) circle (  1.49);

\path[fill=fillColor] (227.86,428.22) circle (  1.49);

\path[fill=fillColor] (235.59,433.81) circle (  1.49);

\path[fill=fillColor] (253.44,449.44) circle (  1.49);

\path[fill=fillColor] (229.18,429.48) circle (  1.49);

\path[fill=fillColor] (255.07,456.90) circle (  1.49);

\path[fill=fillColor] (234.28,433.00) circle (  1.49);

\path[fill=fillColor] (235.10,432.38) circle (  1.49);

\path[fill=fillColor] (249.06,443.05) circle (  1.49);

\path[fill=fillColor] (261.87,466.59) circle (  1.49);

\path[fill=fillColor] (240.19,435.95) circle (  1.49);

\path[fill=fillColor] (228.99,430.24) circle (  1.49);

\path[fill=fillColor] (228.48,428.31) circle (  1.49);

\path[fill=fillColor] (276.45,460.85) circle (  1.49);

\path[fill=fillColor] (243.17,446.15) circle (  1.49);

\path[fill=fillColor] (229.40,428.47) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (297.00,547.27) {Coverage: 94\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (343.12,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (225.19,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.50,425.83) -- (264.50,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (303.81,425.83) -- (303.81,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.12,425.83) -- (343.12,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (225.19,411.58) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (264.50,411.58) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (303.81,411.58) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (343.12,411.58) {150};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (225.19,524.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (221.23,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,458.64) -- (221.23,458.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,491.45) -- (221.23,491.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,524.25) -- (221.23,524.25);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,423.56) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,456.37) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,489.17) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,521.98) {150};
\end{scope}
\begin{scope}
\path[clip] (385.44,385.44) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (489.72,559.19) {\bfseries HKY model $\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (489.72,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (392.57,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (417.91,425.83) rectangle (561.53,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (417.91,425.83) --
	(561.53,545.69);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.04,426.68) --
	(419.04,426.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.85,460.96) --
	(465.85,471.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.30,432.39) --
	(426.30,433.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.79,426.53) --
	(418.79,426.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.27,438.16) --
	(433.27,439.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.66,427.92) --
	(420.66,428.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.71,429.43) --
	(422.71,429.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.94,430.23) --
	(423.94,431.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (440.52,440.96) --
	(440.52,449.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.39,426.91) --
	(419.39,427.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.71,428.96) --
	(421.71,429.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.33,427.81) --
	(420.33,428.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.29,431.01) --
	(424.29,432.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.51,428.65) --
	(421.51,428.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.35,426.93) --
	(419.35,427.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.38,427.73) --
	(420.38,428.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.03,434.90) --
	(429.03,435.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.42,427.92) --
	(420.42,428.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (509.55,485.38) --
	(509.55,528.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.76,427.36) --
	(419.76,427.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.82,427.29) --
	(419.82,427.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.79,428.87) --
	(421.79,429.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.40,429.34) --
	(422.40,429.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (430.89,435.82) --
	(430.89,436.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.03,429.89) --
	(423.03,430.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.23,426.07) --
	(418.23,426.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.57,428.63) --
	(421.57,429.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.94,432.09) --
	(425.94,432.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.90,426.64) --
	(418.90,426.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.04,428.19) --
	(421.04,428.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.70,428.15) --
	(420.70,428.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.16,425.99) --
	(418.16,426.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.86,429.91) --
	(423.86,431.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.66,429.47) --
	(422.66,429.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.30,429.17) --
	(422.30,429.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.51,428.61) --
	(421.51,429.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.10,426.75) --
	(419.10,426.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.50,432.38) --
	(426.50,433.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.78,431.40) --
	(424.78,431.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.53,427.90) --
	(420.53,428.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.26,431.60) --
	(425.26,432.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.95,432.40) --
	(426.95,435.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.96,432.66) --
	(426.96,433.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.68,426.38) --
	(418.68,426.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (450.68,450.55) --
	(450.68,453.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.97,433.34) --
	(427.97,435.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.16,434.21) --
	(429.16,435.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.98,427.39) --
	(419.98,427.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.49,427.87) --
	(420.49,428.16);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (451.53,455.32) --
	(451.53,466.32);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.59,427.13) --
	(419.59,427.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.32,434.23) --
	(428.32,435.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.15,426.78) --
	(419.15,426.94);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.51,430.63) --
	(425.51,431.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.23,427.66) --
	(420.23,428.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.54,432.04) --
	(425.54,433.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.40,429.26) --
	(422.40,429.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.03,427.44) --
	(420.03,427.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.66,427.18) --
	(419.66,427.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.21,426.05) --
	(418.21,426.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.69,426.38) --
	(418.69,426.55);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.84,430.82) --
	(423.84,431.30);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.39,455.44) --
	(453.39,461.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.53,430.09) --
	(423.53,430.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (447.79,449.83) --
	(447.79,452.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.79,426.48) --
	(418.79,426.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.94,427.37) --
	(419.94,427.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.66,426.38) --
	(418.66,426.48);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.16,428.25) --
	(421.16,428.52);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.78,428.01) --
	(420.78,428.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.65,429.46) --
	(422.65,429.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.52,428.69) --
	(421.52,428.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.13,431.67) --
	(425.13,433.11);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.89,432.50) --
	(425.89,433.56);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.82,426.53) --
	(418.82,426.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (434.49,437.00) --
	(434.49,440.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.24,428.42) --
	(421.24,428.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (441.10,442.75) --
	(441.10,446.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (448.30,446.18) --
	(448.30,456.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.39,426.19) --
	(418.39,426.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.44,431.76) --
	(425.44,432.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.79,456.64) --
	(459.79,465.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.49,428.75) --
	(421.49,429.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.46,432.74) --
	(426.46,433.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.88,433.60) --
	(427.88,435.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.75,427.29) --
	(419.75,427.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.68,432.34) --
	(426.68,433.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.77,427.98) --
	(420.77,428.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.59,441.04) --
	(436.59,442.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.55,428.40) --
	(421.55,429.15);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.68,440.67) --
	(435.68,443.09);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.14,427.58) --
	(420.14,427.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.45,433.13) --
	(427.45,434.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.23,430.75) --
	(424.23,432.34);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.62,428.98) --
	(421.62,429.44);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.16,429.93) --
	(423.16,430.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.92,429.00) --
	(421.92,429.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.95,430.98) --
	(424.95,432.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.34,426.18) --
	(418.34,426.31);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.10,429.42) --
	(422.10,429.80);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (419.04,426.74) circle (  1.49);

\path[fill=fillColor] (465.85,465.89) circle (  1.49);

\path[fill=fillColor] (426.30,432.71) circle (  1.49);

\path[fill=fillColor] (418.79,426.58) circle (  1.49);

\path[fill=fillColor] (433.27,438.84) circle (  1.49);

\path[fill=fillColor] (420.66,428.07) circle (  1.49);

\path[fill=fillColor] (422.71,429.70) circle (  1.49);

\path[fill=fillColor] (423.94,430.63) circle (  1.49);

\path[fill=fillColor] (440.52,445.29) circle (  1.49);

\path[fill=fillColor] (419.39,427.01) circle (  1.49);

\path[fill=fillColor] (421.71,429.24) circle (  1.49);

\path[fill=fillColor] (420.33,427.95) circle (  1.49);

\path[fill=fillColor] (424.29,431.60) circle (  1.49);

\path[fill=fillColor] (421.51,428.78) circle (  1.49);

\path[fill=fillColor] (419.35,427.01) circle (  1.49);

\path[fill=fillColor] (420.38,427.90) circle (  1.49);

\path[fill=fillColor] (429.03,435.41) circle (  1.49);

\path[fill=fillColor] (420.42,428.04) circle (  1.49);

\path[fill=fillColor] (509.55,505.74) circle (  1.49);

\path[fill=fillColor] (419.76,427.50) circle (  1.49);

\path[fill=fillColor] (419.82,427.39) circle (  1.49);

\path[fill=fillColor] (421.79,429.07) circle (  1.49);

\path[fill=fillColor] (422.40,429.52) circle (  1.49);

\path[fill=fillColor] (430.89,436.42) circle (  1.49);

\path[fill=fillColor] (423.03,430.13) circle (  1.49);

\path[fill=fillColor] (418.23,426.10) circle (  1.49);

\path[fill=fillColor] (421.57,428.89) circle (  1.49);

\path[fill=fillColor] (425.94,432.47) circle (  1.49);

\path[fill=fillColor] (418.90,426.70) circle (  1.49);

\path[fill=fillColor] (421.04,428.43) circle (  1.49);

\path[fill=fillColor] (420.70,428.28) circle (  1.49);

\path[fill=fillColor] (418.16,426.02) circle (  1.49);

\path[fill=fillColor] (423.86,430.71) circle (  1.49);

\path[fill=fillColor] (422.66,429.67) circle (  1.49);

\path[fill=fillColor] (422.30,429.44) circle (  1.49);

\path[fill=fillColor] (421.51,428.82) circle (  1.49);

\path[fill=fillColor] (419.10,426.83) circle (  1.49);

\path[fill=fillColor] (426.50,433.01) circle (  1.49);

\path[fill=fillColor] (424.78,431.70) circle (  1.49);

\path[fill=fillColor] (420.53,428.03) circle (  1.49);

\path[fill=fillColor] (425.26,431.87) circle (  1.49);

\path[fill=fillColor] (426.95,434.10) circle (  1.49);

\path[fill=fillColor] (426.96,433.13) circle (  1.49);

\path[fill=fillColor] (418.68,426.43) circle (  1.49);

\path[fill=fillColor] (450.68,452.01) circle (  1.49);

\path[fill=fillColor] (427.97,434.26) circle (  1.49);

\path[fill=fillColor] (429.16,434.96) circle (  1.49);

\path[fill=fillColor] (419.98,427.51) circle (  1.49);

\path[fill=fillColor] (420.49,428.02) circle (  1.49);

\path[fill=fillColor] (451.53,460.82) circle (  1.49);

\path[fill=fillColor] (419.59,427.25) circle (  1.49);

\path[fill=fillColor] (428.32,434.80) circle (  1.49);

\path[fill=fillColor] (419.15,426.86) circle (  1.49);

\path[fill=fillColor] (425.51,431.29) circle (  1.49);

\path[fill=fillColor] (420.23,427.90) circle (  1.49);

\path[fill=fillColor] (425.54,432.52) circle (  1.49);

\path[fill=fillColor] (422.40,429.42) circle (  1.49);

\path[fill=fillColor] (420.03,427.54) circle (  1.49);

\path[fill=fillColor] (419.66,427.32) circle (  1.49);

\path[fill=fillColor] (418.21,426.08) circle (  1.49);

\path[fill=fillColor] (418.69,426.47) circle (  1.49);

\path[fill=fillColor] (423.84,431.05) circle (  1.49);

\path[fill=fillColor] (453.39,458.41) circle (  1.49);

\path[fill=fillColor] (423.53,430.45) circle (  1.49);

\path[fill=fillColor] (447.79,451.13) circle (  1.49);

\path[fill=fillColor] (418.79,426.55) circle (  1.49);

\path[fill=fillColor] (419.94,427.50) circle (  1.49);

\path[fill=fillColor] (418.66,426.43) circle (  1.49);

\path[fill=fillColor] (421.16,428.39) circle (  1.49);

\path[fill=fillColor] (420.78,428.16) circle (  1.49);

\path[fill=fillColor] (422.65,429.73) circle (  1.49);

\path[fill=fillColor] (421.52,428.83) circle (  1.49);

\path[fill=fillColor] (425.13,432.38) circle (  1.49);

\path[fill=fillColor] (425.89,433.03) circle (  1.49);

\path[fill=fillColor] (418.82,426.57) circle (  1.49);

\path[fill=fillColor] (434.49,438.96) circle (  1.49);

\path[fill=fillColor] (421.24,428.61) circle (  1.49);

\path[fill=fillColor] (441.10,444.79) circle (  1.49);

\path[fill=fillColor] (448.30,451.20) circle (  1.49);

\path[fill=fillColor] (418.39,426.23) circle (  1.49);

\path[fill=fillColor] (425.44,432.18) circle (  1.49);

\path[fill=fillColor] (459.79,461.16) circle (  1.49);

\path[fill=fillColor] (421.49,428.91) circle (  1.49);

\path[fill=fillColor] (426.46,433.36) circle (  1.49);

\path[fill=fillColor] (427.88,434.38) circle (  1.49);

\path[fill=fillColor] (419.75,427.37) circle (  1.49);

\path[fill=fillColor] (426.68,432.81) circle (  1.49);

\path[fill=fillColor] (420.77,428.24) circle (  1.49);

\path[fill=fillColor] (436.59,441.94) circle (  1.49);

\path[fill=fillColor] (421.55,428.77) circle (  1.49);

\path[fill=fillColor] (435.68,441.82) circle (  1.49);

\path[fill=fillColor] (420.14,427.69) circle (  1.49);

\path[fill=fillColor] (427.45,433.95) circle (  1.49);

\path[fill=fillColor] (424.23,431.55) circle (  1.49);

\path[fill=fillColor] (421.62,429.21) circle (  1.49);

\path[fill=fillColor] (423.16,430.15) circle (  1.49);

\path[fill=fillColor] (421.92,429.15) circle (  1.49);

\path[fill=fillColor] (424.95,431.72) circle (  1.49);

\path[fill=fillColor] (418.34,426.24) circle (  1.49);

\path[fill=fillColor] (422.10,429.61) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (489.72,547.27) {Coverage: 92\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (548.61,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (417.91,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.58,425.83) -- (436.58,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (455.25,425.83) -- (455.25,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (473.93,425.83) -- (473.93,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (492.60,425.83) -- (492.60,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.27,425.83) -- (511.27,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (529.94,425.83) -- (529.94,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (548.61,425.83) -- (548.61,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (417.91,411.58) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (436.58,411.58) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (455.25,411.58) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (473.93,411.58) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (492.60,411.58) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (511.27,411.58) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (529.94,411.58) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (548.61,411.58) {70};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (417.91,534.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (413.95,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,441.41) -- (413.95,441.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,457.00) -- (413.95,457.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,472.58) -- (413.95,472.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,488.16) -- (413.95,488.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,503.74) -- (413.95,503.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,519.33) -- (413.95,519.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,534.91) -- (413.95,534.91);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,423.56) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,439.14) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,454.72) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,470.31) {30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,485.89) {40};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,501.47) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,517.05) {60};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,532.64) {70};
\end{scope}
\begin{scope}
\path[clip] (  0.00,192.72) rectangle (192.72,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,366.47) {\bfseries Nucleotide frequency $f$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47,233.11) rectangle (176.09,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47,233.11) --
	(176.09,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.52,254.87) --
	( 59.52,257.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 80.98,272.46) --
	( 80.98,276.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.46,265.70) --
	( 73.46,268.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (106.47,294.55) --
	(106.47,298.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.59,282.37) --
	( 93.59,284.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.32,276.04) --
	( 85.32,280.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.14,300.47) --
	(114.14,305.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.03,297.85) --
	(114.03,303.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.32,285.17) --
	( 99.32,291.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.70,275.49) --
	( 85.70,279.75);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (119.28,299.94) --
	(119.28,305.38);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.17,265.60) --
	( 74.17,269.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.64,273.68) --
	( 86.64,279.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.12,268.46) --
	( 76.12,270.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.26,271.34) --
	( 81.26,274.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.22,304.51) --
	(121.22,310.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.79,254.90) --
	( 60.79,257.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.73,278.22) --
	( 88.73,281.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (106.97,291.56) --
	(106.97,297.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.24,290.64) --
	(105.24,295.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.44,264.52) --
	( 72.44,267.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.86,279.95) --
	( 92.86,284.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.03,297.32) --
	(112.03,301.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.55,274.40) --
	( 84.55,277.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.64,279.36) --
	( 92.64,283.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.89,285.15) --
	( 96.89,289.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (108.82,294.57) --
	(108.82,300.15);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.57,260.01) --
	( 68.57,263.14);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.03,286.07) --
	( 98.03,288.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.02,291.50) --
	(104.02,297.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.85,280.06) --
	( 90.85,283.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.22,309.74) --
	(128.22,313.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.10,280.96) --
	( 95.10,287.47);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.15,254.13) --
	( 61.15,256.58);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.84,288.41) --
	(100.84,293.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.64,287.51) --
	(101.64,292.52);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.37,274.60) --
	( 81.37,278.39);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (108.47,295.10) --
	(108.47,300.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.02,269.10) --
	( 76.02,272.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.04,284.38) --
	( 95.04,288.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 82.97,273.68) --
	( 82.97,276.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.04,276.91) --
	( 89.04,283.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.13,287.34) --
	(100.13,291.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.15,276.01) --
	( 86.15,279.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.23,261.05) --
	( 66.23,263.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.83,274.47) --
	( 84.83,279.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.01,297.22) --
	(110.01,302.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.64,266.48) --
	( 72.64,269.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.63,288.63) --
	(103.63,293.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.17,265.66) --
	( 75.17,270.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.42,304.17) --
	(121.42,309.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.69,282.27) --
	( 95.69,286.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.56,266.91) --
	( 73.56,270.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.69,299.16) --
	(114.69,305.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.50,278.41) --
	( 92.50,284.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.37,287.82) --
	(101.37,292.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.95,277.19) --
	( 86.95,279.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.83,269.76) --
	( 77.83,273.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.72,276.88) --
	( 86.72,282.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (113.57,297.73) --
	(113.57,301.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.07,260.17) --
	( 66.07,264.53);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.56,286.72) --
	( 95.56,290.06);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.40,288.64) --
	(103.40,292.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.68,269.34) --
	( 78.68,273.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.18,283.62) --
	( 95.18,286.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.11,275.02) --
	( 85.11,279.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (107.03,292.61) --
	(107.03,297.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.85,264.58) --
	( 70.85,267.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.58,267.70) --
	( 74.58,270.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.56,283.44) --
	( 96.56,288.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.98,280.32) --
	( 89.98,284.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.98,292.03) --
	(104.98,295.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.19,266.15) --
	( 75.19,271.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.27,295.17) --
	(110.27,300.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.48,259.32) --
	( 65.48,261.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.91,263.41) --
	( 72.91,268.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.11,279.79) --
	( 90.11,284.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 94.00,280.95) --
	( 94.00,285.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.13,289.61) --
	(105.13,296.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.06,272.90) --
	( 84.06,276.85);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.09,286.33) --
	( 96.09,290.74);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.77,276.50) --
	( 89.77,281.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.41,298.37) --
	(114.41,302.47);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.74,286.31) --
	(103.74,291.71);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.57,296.29) --
	(112.57,301.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.47,271.54) --
	( 79.47,274.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.02,276.92) --
	( 88.02,280.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.94,265.50) --
	( 74.94,270.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.72,267.62) --
	( 76.72,270.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.57,269.12) --
	( 81.57,274.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 71.87,262.79) --
	( 71.87,266.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.18,296.09) --
	(110.18,300.46);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.41,267.45) --
	( 73.41,272.07);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.28,274.95) --
	( 85.28,280.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.24,276.17) --
	( 86.24,280.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.02,274.31) --
	( 83.02,277.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.89,266.12) --
	( 72.89,268.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.64,283.42) --
	( 93.64,289.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.83,289.09) --
	(103.83,294.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 64.27,258.93) --
	( 64.27,261.11);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 59.52,256.08) circle (  1.49);

\path[fill=fillColor] ( 80.98,274.42) circle (  1.49);

\path[fill=fillColor] ( 73.46,267.06) circle (  1.49);

\path[fill=fillColor] (106.47,296.53) circle (  1.49);

\path[fill=fillColor] ( 93.59,283.58) circle (  1.49);

\path[fill=fillColor] ( 85.32,278.22) circle (  1.49);

\path[fill=fillColor] (114.14,302.96) circle (  1.49);

\path[fill=fillColor] (114.03,300.67) circle (  1.49);

\path[fill=fillColor] ( 99.32,288.43) circle (  1.49);

\path[fill=fillColor] ( 85.70,277.63) circle (  1.49);

\path[fill=fillColor] (119.28,302.67) circle (  1.49);

\path[fill=fillColor] ( 74.17,267.38) circle (  1.49);

\path[fill=fillColor] ( 86.64,276.23) circle (  1.49);

\path[fill=fillColor] ( 76.12,269.49) circle (  1.49);

\path[fill=fillColor] ( 81.26,273.01) circle (  1.49);

\path[fill=fillColor] (121.22,307.35) circle (  1.49);

\path[fill=fillColor] ( 60.79,255.95) circle (  1.49);

\path[fill=fillColor] ( 88.73,279.90) circle (  1.49);

\path[fill=fillColor] (106.97,294.75) circle (  1.49);

\path[fill=fillColor] (105.24,293.36) circle (  1.49);

\path[fill=fillColor] ( 72.44,266.10) circle (  1.49);

\path[fill=fillColor] ( 92.86,282.00) circle (  1.49);

\path[fill=fillColor] (112.03,299.31) circle (  1.49);

\path[fill=fillColor] ( 84.55,275.86) circle (  1.49);

\path[fill=fillColor] ( 92.64,281.37) circle (  1.49);

\path[fill=fillColor] ( 96.89,287.26) circle (  1.49);

\path[fill=fillColor] (108.82,297.40) circle (  1.49);

\path[fill=fillColor] ( 68.57,261.61) circle (  1.49);

\path[fill=fillColor] ( 98.03,287.50) circle (  1.49);

\path[fill=fillColor] (104.02,294.29) circle (  1.49);

\path[fill=fillColor] ( 90.85,281.82) circle (  1.49);

\path[fill=fillColor] (128.22,311.66) circle (  1.49);

\path[fill=fillColor] ( 95.10,284.08) circle (  1.49);

\path[fill=fillColor] ( 61.15,255.36) circle (  1.49);

\path[fill=fillColor] (100.84,290.95) circle (  1.49);

\path[fill=fillColor] (101.64,290.02) circle (  1.49);

\path[fill=fillColor] ( 81.37,276.47) circle (  1.49);

\path[fill=fillColor] (108.47,297.83) circle (  1.49);

\path[fill=fillColor] ( 76.02,270.51) circle (  1.49);

\path[fill=fillColor] ( 95.04,286.24) circle (  1.49);

\path[fill=fillColor] ( 82.97,275.11) circle (  1.49);

\path[fill=fillColor] ( 89.04,280.23) circle (  1.49);

\path[fill=fillColor] (100.13,289.52) circle (  1.49);

\path[fill=fillColor] ( 86.15,277.94) circle (  1.49);

\path[fill=fillColor] ( 66.23,262.22) circle (  1.49);

\path[fill=fillColor] ( 84.83,277.21) circle (  1.49);

\path[fill=fillColor] (110.01,300.00) circle (  1.49);

\path[fill=fillColor] ( 72.64,268.11) circle (  1.49);

\path[fill=fillColor] (103.63,290.97) circle (  1.49);

\path[fill=fillColor] ( 75.17,267.96) circle (  1.49);

\path[fill=fillColor] (121.42,307.05) circle (  1.49);

\path[fill=fillColor] ( 95.69,284.22) circle (  1.49);

\path[fill=fillColor] ( 73.56,268.84) circle (  1.49);

\path[fill=fillColor] (114.69,302.29) circle (  1.49);

\path[fill=fillColor] ( 92.50,281.45) circle (  1.49);

\path[fill=fillColor] (101.37,290.10) circle (  1.49);

\path[fill=fillColor] ( 86.95,278.47) circle (  1.49);

\path[fill=fillColor] ( 77.83,271.40) circle (  1.49);

\path[fill=fillColor] ( 86.72,279.64) circle (  1.49);

\path[fill=fillColor] (113.57,299.49) circle (  1.49);

\path[fill=fillColor] ( 66.07,262.34) circle (  1.49);

\path[fill=fillColor] ( 95.56,288.42) circle (  1.49);

\path[fill=fillColor] (103.40,290.82) circle (  1.49);

\path[fill=fillColor] ( 78.68,271.55) circle (  1.49);

\path[fill=fillColor] ( 95.18,284.99) circle (  1.49);

\path[fill=fillColor] ( 85.11,277.09) circle (  1.49);

\path[fill=fillColor] (107.03,295.09) circle (  1.49);

\path[fill=fillColor] ( 70.85,265.93) circle (  1.49);

\path[fill=fillColor] ( 74.58,269.21) circle (  1.49);

\path[fill=fillColor] ( 96.56,285.72) circle (  1.49);

\path[fill=fillColor] ( 89.98,282.42) circle (  1.49);

\path[fill=fillColor] (104.98,293.79) circle (  1.49);

\path[fill=fillColor] ( 75.19,268.66) circle (  1.49);

\path[fill=fillColor] (110.27,297.62) circle (  1.49);

\path[fill=fillColor] ( 65.48,260.37) circle (  1.49);

\path[fill=fillColor] ( 72.91,266.15) circle (  1.49);

\path[fill=fillColor] ( 90.11,282.11) circle (  1.49);

\path[fill=fillColor] ( 94.00,283.18) circle (  1.49);

\path[fill=fillColor] (105.13,292.88) circle (  1.49);

\path[fill=fillColor] ( 84.06,274.85) circle (  1.49);

\path[fill=fillColor] ( 96.09,288.56) circle (  1.49);

\path[fill=fillColor] ( 89.77,278.88) circle (  1.49);

\path[fill=fillColor] (114.41,300.41) circle (  1.49);

\path[fill=fillColor] (103.74,288.94) circle (  1.49);

\path[fill=fillColor] (112.57,298.92) circle (  1.49);

\path[fill=fillColor] ( 79.47,272.79) circle (  1.49);

\path[fill=fillColor] ( 88.02,278.90) circle (  1.49);

\path[fill=fillColor] ( 74.94,267.89) circle (  1.49);

\path[fill=fillColor] ( 76.72,268.95) circle (  1.49);

\path[fill=fillColor] ( 81.57,271.81) circle (  1.49);

\path[fill=fillColor] ( 71.87,264.41) circle (  1.49);

\path[fill=fillColor] (110.18,298.16) circle (  1.49);

\path[fill=fillColor] ( 73.41,269.77) circle (  1.49);

\path[fill=fillColor] ( 85.28,277.79) circle (  1.49);

\path[fill=fillColor] ( 86.24,278.43) circle (  1.49);

\path[fill=fillColor] ( 83.02,276.03) circle (  1.49);

\path[fill=fillColor] ( 72.89,267.31) circle (  1.49);

\path[fill=fillColor] ( 93.64,286.52) circle (  1.49);

\path[fill=fillColor] (103.83,291.88) circle (  1.49);

\path[fill=fillColor] ( 64.27,260.00) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,354.55) {Coverage: 92\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- (174.42,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 32.47,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.13,233.11) -- ( 56.13,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.79,233.11) -- ( 79.79,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.45,233.11) -- (103.45,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.10,233.11) -- (127.10,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (150.76,233.11) -- (150.76,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (174.42,233.11) -- (174.42,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 56.13,218.86) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 79.79,218.86) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (103.45,218.86) {0.3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (127.10,218.86) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (150.76,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (174.42,218.86) {0.6};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 32.47,351.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 28.51,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,252.86) -- ( 28.51,252.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,272.60) -- ( 28.51,272.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,292.34) -- ( 28.51,292.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,312.09) -- ( 28.51,312.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,331.83) -- ( 28.51,331.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,351.58) -- ( 28.51,351.58);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,250.58) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,270.33) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,290.07) {0.3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,309.82) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,329.56) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,349.30) {0.6};
\end{scope}
\begin{scope}
\path[clip] (192.72,192.72) rectangle (385.44,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (297.00,366.47) {\bfseries Clock standard deviation $\sigma$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (297.00,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (199.85,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (225.19,233.11) rectangle (368.81,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (225.19,233.11) --
	(368.81,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.20,233.11) --
	(225.20,235.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.54,250.68) --
	(252.54,263.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.46,233.11) --
	(225.46,234.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.28,234.67) --
	(230.28,239.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.12,233.11) --
	(227.12,236.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.84,233.11) --
	(226.84,234.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.84,241.54) --
	(239.84,249.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.94,234.48) --
	(231.94,240.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.80,233.11) --
	(232.80,242.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.96,240.38) --
	(236.96,247.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.47,233.11) --
	(225.47,236.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.15,233.11) --
	(226.15,235.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.67,233.11) --
	(229.67,241.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.67,241.82) --
	(238.67,249.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.42,233.11) --
	(225.42,234.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.75,234.76) --
	(227.75,240.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.31,255.18) --
	(257.31,273.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.78,238.23) --
	(236.78,243.67);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.11,236.33) --
	(228.11,243.76);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.65,235.35) --
	(231.65,240.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.58,234.14) --
	(231.58,238.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.03,242.62) --
	(239.03,250.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.90,233.11) --
	(227.90,236.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.39,241.09) --
	(237.39,248.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.13,233.11) --
	(227.13,234.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.62,236.38) --
	(231.62,240.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.01,239.57) --
	(240.01,246.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.94,233.11) --
	(225.94,235.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.22,233.11) --
	(225.22,234.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.84,235.84) --
	(229.84,242.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.69,242.99) --
	(237.69,250.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.18,233.11) --
	(229.18,237.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.29,233.11) --
	(230.29,238.24);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.16,234.94) --
	(227.16,239.08);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.62,233.11) --
	(225.62,236.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.55,233.11) --
	(226.55,236.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.99,260.46) --
	(264.99,279.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.87,233.11) --
	(227.87,236.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.20,233.11) --
	(226.20,235.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.35,233.11) --
	(226.35,234.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.32,233.11) --
	(225.32,234.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.82,233.11) --
	(225.82,241.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.46,233.11) --
	(226.46,235.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) --
	(225.19,235.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.00,233.11) --
	(226.00,235.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.67,236.65) --
	(233.67,243.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.46,233.11) --
	(226.46,235.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.04,233.11) --
	(226.04,234.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.50,233.11) --
	(225.50,236.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (244.16,245.02) --
	(244.16,254.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.77,233.11) --
	(225.77,235.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.90,233.11) --
	(226.90,235.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.26,243.97) --
	(240.26,252.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.84,240.19) --
	(239.84,249.40);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.58,233.11) --
	(230.58,236.76);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.19,234.25) --
	(230.19,239.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.09,235.39) --
	(229.09,239.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.99,243.88) --
	(243.99,251.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.40,233.11) --
	(225.40,238.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.33,233.11) --
	(226.33,236.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.07,233.11) --
	(229.07,236.91);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.48,233.91) --
	(231.48,237.57);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.83,233.11) --
	(227.83,237.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.35,237.38) --
	(232.35,242.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.03,233.11) --
	(226.03,235.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.27,272.09) --
	(273.27,299.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.19,233.11) --
	(226.19,235.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.02,234.13) --
	(227.02,237.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.62,233.11) --
	(225.62,235.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.07,234.49) --
	(229.07,239.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.11,236.10) --
	(230.11,240.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.03,236.70) --
	(230.03,240.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.25,292.90) --
	(310.25,333.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.76,233.11) --
	(226.76,235.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.92,235.75) --
	(231.92,239.20);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.47,238.48) --
	(247.47,246.73);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.97,265.96) --
	(273.97,289.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.21,237.30) --
	(232.21,242.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.52,237.46) --
	(238.52,247.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.63,247.33) --
	(245.63,260.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.94,233.11) --
	(226.94,235.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.69,233.11) --
	(226.69,238.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) --
	(225.19,235.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.76,246.43) --
	(247.76,257.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.18,233.11) --
	(226.18,236.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.38,238.30) --
	(233.38,242.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.86,239.20) --
	(234.86,244.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.26,253.90) --
	(252.26,268.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.26,233.11) --
	(225.26,235.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.01,261.06) --
	(280.01,284.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.33,244.96) --
	(241.33,255.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.61,238.17) --
	(234.61,243.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.28,233.11) --
	(226.28,236.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.30,233.11) --
	(226.30,240.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.37,233.11) --
	(227.37,237.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.20,233.11) --
	(225.20,235.70);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.43,233.11) --
	(229.43,235.87);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.43,233.11) --
	(226.43,237.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.66,233.11) --
	(226.66,236.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (259.60,256.24) --
	(259.60,272.00);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (225.20,233.77) circle (  1.49);

\path[fill=fillColor] (252.54,256.86) circle (  1.49);

\path[fill=fillColor] (225.46,233.62) circle (  1.49);

\path[fill=fillColor] (230.28,236.84) circle (  1.49);

\path[fill=fillColor] (227.12,234.39) circle (  1.49);

\path[fill=fillColor] (226.84,233.65) circle (  1.49);

\path[fill=fillColor] (239.84,245.22) circle (  1.49);

\path[fill=fillColor] (231.94,237.39) circle (  1.49);

\path[fill=fillColor] (232.80,236.70) circle (  1.49);

\path[fill=fillColor] (236.96,243.86) circle (  1.49);

\path[fill=fillColor] (225.47,234.34) circle (  1.49);

\path[fill=fillColor] (226.15,233.85) circle (  1.49);

\path[fill=fillColor] (229.67,237.30) circle (  1.49);

\path[fill=fillColor] (238.67,245.53) circle (  1.49);

\path[fill=fillColor] (225.42,233.54) circle (  1.49);

\path[fill=fillColor] (227.75,237.73) circle (  1.49);

\path[fill=fillColor] (257.31,263.92) circle (  1.49);

\path[fill=fillColor] (236.78,240.86) circle (  1.49);

\path[fill=fillColor] (228.11,239.75) circle (  1.49);

\path[fill=fillColor] (231.65,237.99) circle (  1.49);

\path[fill=fillColor] (231.58,236.38) circle (  1.49);

\path[fill=fillColor] (239.03,246.25) circle (  1.49);

\path[fill=fillColor] (227.90,234.86) circle (  1.49);

\path[fill=fillColor] (237.39,244.68) circle (  1.49);

\path[fill=fillColor] (227.13,233.68) circle (  1.49);

\path[fill=fillColor] (231.62,238.48) circle (  1.49);

\path[fill=fillColor] (240.01,243.12) circle (  1.49);

\path[fill=fillColor] (225.94,234.01) circle (  1.49);

\path[fill=fillColor] (225.22,233.48) circle (  1.49);

\path[fill=fillColor] (229.84,239.19) circle (  1.49);

\path[fill=fillColor] (237.69,246.76) circle (  1.49);

\path[fill=fillColor] (229.18,235.86) circle (  1.49);

\path[fill=fillColor] (230.29,234.73) circle (  1.49);

\path[fill=fillColor] (227.16,236.96) circle (  1.49);

\path[fill=fillColor] (225.62,234.18) circle (  1.49);

\path[fill=fillColor] (226.55,234.26) circle (  1.49);

\path[fill=fillColor] (264.99,269.25) circle (  1.49);

\path[fill=fillColor] (227.87,234.25) circle (  1.49);

\path[fill=fillColor] (226.20,233.88) circle (  1.49);

\path[fill=fillColor] (226.35,233.67) circle (  1.49);

\path[fill=fillColor] (225.32,233.56) circle (  1.49);

\path[fill=fillColor] (225.82,235.93) circle (  1.49);

\path[fill=fillColor] (226.46,233.99) circle (  1.49);

\path[fill=fillColor] (225.19,233.78) circle (  1.49);

\path[fill=fillColor] (226.00,234.14) circle (  1.49);

\path[fill=fillColor] (233.67,239.87) circle (  1.49);

\path[fill=fillColor] (226.46,233.97) circle (  1.49);

\path[fill=fillColor] (226.04,233.68) circle (  1.49);

\path[fill=fillColor] (225.50,234.37) circle (  1.49);

\path[fill=fillColor] (244.16,249.71) circle (  1.49);

\path[fill=fillColor] (225.77,233.81) circle (  1.49);

\path[fill=fillColor] (226.90,234.06) circle (  1.49);

\path[fill=fillColor] (240.26,248.10) circle (  1.49);

\path[fill=fillColor] (239.84,244.62) circle (  1.49);

\path[fill=fillColor] (230.58,234.25) circle (  1.49);

\path[fill=fillColor] (230.19,236.90) circle (  1.49);

\path[fill=fillColor] (229.09,237.44) circle (  1.49);

\path[fill=fillColor] (243.99,247.69) circle (  1.49);

\path[fill=fillColor] (225.40,235.04) circle (  1.49);

\path[fill=fillColor] (226.33,234.49) circle (  1.49);

\path[fill=fillColor] (229.07,234.35) circle (  1.49);

\path[fill=fillColor] (231.48,235.69) circle (  1.49);

\path[fill=fillColor] (227.83,235.39) circle (  1.49);

\path[fill=fillColor] (232.35,239.75) circle (  1.49);

\path[fill=fillColor] (226.03,234.14) circle (  1.49);

\path[fill=fillColor] (273.27,285.26) circle (  1.49);

\path[fill=fillColor] (226.19,233.85) circle (  1.49);

\path[fill=fillColor] (227.02,235.81) circle (  1.49);

\path[fill=fillColor] (225.62,233.85) circle (  1.49);

\path[fill=fillColor] (229.07,237.09) circle (  1.49);

\path[fill=fillColor] (230.11,238.29) circle (  1.49);

\path[fill=fillColor] (230.03,238.60) circle (  1.49);

\path[fill=fillColor] (310.25,313.02) circle (  1.49);

\path[fill=fillColor] (226.76,233.91) circle (  1.49);

\path[fill=fillColor] (231.92,237.40) circle (  1.49);

\path[fill=fillColor] (247.47,242.53) circle (  1.49);

\path[fill=fillColor] (273.97,277.27) circle (  1.49);

\path[fill=fillColor] (232.21,239.84) circle (  1.49);

\path[fill=fillColor] (238.52,242.32) circle (  1.49);

\path[fill=fillColor] (245.63,253.43) circle (  1.49);

\path[fill=fillColor] (226.94,233.86) circle (  1.49);

\path[fill=fillColor] (226.69,235.88) circle (  1.49);

\path[fill=fillColor] (225.19,233.81) circle (  1.49);

\path[fill=fillColor] (247.76,251.83) circle (  1.49);

\path[fill=fillColor] (226.18,234.41) circle (  1.49);

\path[fill=fillColor] (233.38,240.50) circle (  1.49);

\path[fill=fillColor] (234.86,241.91) circle (  1.49);

\path[fill=fillColor] (252.26,261.14) circle (  1.49);

\path[fill=fillColor] (225.26,233.95) circle (  1.49);

\path[fill=fillColor] (280.01,271.74) circle (  1.49);

\path[fill=fillColor] (241.33,249.71) circle (  1.49);

\path[fill=fillColor] (234.61,240.53) circle (  1.49);

\path[fill=fillColor] (226.28,234.32) circle (  1.49);

\path[fill=fillColor] (226.30,235.76) circle (  1.49);

\path[fill=fillColor] (227.37,235.18) circle (  1.49);

\path[fill=fillColor] (225.20,234.02) circle (  1.49);

\path[fill=fillColor] (229.43,234.28) circle (  1.49);

\path[fill=fillColor] (226.43,234.43) circle (  1.49);

\path[fill=fillColor] (226.66,234.09) circle (  1.49);

\path[fill=fillColor] (259.60,263.79) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (297.00,354.55) {Coverage: 94\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (360.93,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (225.19,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (244.58,233.11) -- (244.58,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.97,233.11) -- (263.97,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.37,233.11) -- (283.37,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.76,233.11) -- (302.76,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.15,233.11) -- (322.15,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (341.54,233.11) -- (341.54,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (360.93,233.11) -- (360.93,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (225.19,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (244.58,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (263.97,218.86) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (283.37,218.86) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (302.76,218.86) {2.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (322.15,218.86) {2.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (341.54,218.86) {3.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (360.93,218.86) {3.5};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (225.19,346.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (221.23,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,249.30) -- (221.23,249.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,265.48) -- (221.23,265.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,281.66) -- (221.23,281.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,297.85) -- (221.23,297.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,314.03) -- (221.23,314.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,330.21) -- (221.23,330.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,346.40) -- (221.23,346.40);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,247.02) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,263.21) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,279.39) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,295.57) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,311.76) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,327.94) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,344.12) {3.5};
\end{scope}
\begin{scope}
\path[clip] (385.44,192.72) rectangle (578.16,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (489.72,366.47) {\bfseries Branch rate mean};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (489.72,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (392.57,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (417.91,233.11) rectangle (561.53,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (417.91,233.11) --
	(561.53,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.04,270.31) --
	(463.04,271.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.63,265.99) --
	(474.63,289.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.10,270.45) --
	(463.10,271.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.95,268.74) --
	(462.95,272.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.00,270.01) --
	(463.00,271.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.23,270.45) --
	(463.23,271.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.19,267.32) --
	(465.19,278.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.88,269.09) --
	(462.88,272.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.18,268.19) --
	(462.18,272.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.24,265.35) --
	(462.24,274.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.05,269.88) --
	(463.05,271.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.06,270.22) --
	(463.06,271.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.81,268.94) --
	(463.81,272.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.19,265.12) --
	(462.19,274.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.01,270.48) --
	(463.01,271.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.16,268.90) --
	(463.16,273.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.32,258.61) --
	(465.32,286.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.35,267.57) --
	(465.35,273.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.88,267.60) --
	(463.88,272.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.38,268.43) --
	(463.38,272.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.39,269.10) --
	(462.39,272.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.65,265.76) --
	(459.65,276.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.78,269.84) --
	(462.78,271.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (466.68,265.65) --
	(466.68,275.44);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.80,270.42) --
	(463.80,271.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.46,268.55) --
	(465.46,272.73);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.44,267.65) --
	(465.44,276.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.29,270.15) --
	(463.29,271.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.04,270.50) --
	(463.04,271.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.92,268.48) --
	(462.92,273.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.28,264.21) --
	(462.28,273.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.18,269.62) --
	(462.18,272.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.54,269.68) --
	(462.54,271.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.75,269.44) --
	(462.75,273.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.95,270.12) --
	(462.95,271.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.77,270.02) --
	(462.77,271.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (468.80,254.45) --
	(468.80,283.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.75,269.89) --
	(462.75,271.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.09,270.32) --
	(463.09,271.35);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.63,270.47) --
	(462.63,271.15);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.05,270.43) --
	(463.05,271.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.82,268.80) --
	(462.82,272.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.43,270.02) --
	(463.43,271.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.04,270.28) --
	(463.04,271.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.17,270.13) --
	(463.17,271.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.92,267.63) --
	(461.92,273.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.13,270.18) --
	(463.13,271.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.82,270.37) --
	(462.82,271.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.93,270.07) --
	(462.93,271.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.97,262.01) --
	(458.97,274.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.06,270.28) --
	(463.06,271.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.82,270.25) --
	(462.82,271.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.76,264.65) --
	(459.76,275.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.55,266.02) --
	(467.55,276.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.27,269.86) --
	(463.27,271.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.77,269.30) --
	(462.77,272.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.17,269.01) --
	(463.17,272.74);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (473.83,264.90) --
	(473.83,275.74);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.06,269.45) --
	(463.06,271.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.71,269.83) --
	(462.71,271.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.25,269.92) --
	(463.25,271.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.84,269.89) --
	(463.84,272.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.88,269.95) --
	(462.88,272.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.53,267.45) --
	(461.53,272.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.79,270.27) --
	(462.79,271.43);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (513.66,245.32) --
	(513.66,290.35);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.17,270.26) --
	(463.17,271.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.15,269.59) --
	(463.15,271.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.91,270.34) --
	(462.91,271.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.65,269.11) --
	(463.65,272.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.36,268.91) --
	(463.36,272.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.12,268.79) --
	(463.12,273.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.54,234.11) --
	(438.54,270.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.48,270.29) --
	(462.48,271.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.88,268.62) --
	(462.88,271.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.80,267.87) --
	(463.80,276.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (457.61,247.98) --
	(457.61,282.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.83,267.91) --
	(463.83,272.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.14,265.97) --
	(459.14,274.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.79,263.33) --
	(463.79,280.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.59,270.28) --
	(463.59,271.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.90,269.37) --
	(462.90,272.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.04,270.33) --
	(463.04,271.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (460.71,262.95) --
	(460.71,278.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.21,269.96) --
	(463.21,271.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.07,267.10) --
	(465.07,272.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.70,266.15) --
	(461.70,272.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.39,260.59) --
	(471.39,286.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.04,270.17) --
	(463.04,271.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.60,250.44) --
	(453.60,283.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.58,265.11) --
	(467.58,278.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.40,267.47) --
	(464.40,274.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.15,269.84) --
	(463.15,271.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.87,269.04) --
	(462.87,272.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.62,269.47) --
	(462.62,271.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.04,270.08) --
	(463.04,271.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.51,270.17) --
	(463.51,271.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.92,269.85) --
	(462.92,271.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.13,270.03) --
	(463.13,271.29);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.56,258.52) --
	(483.56,284.78);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (463.04,270.80) circle (  1.49);

\path[fill=fillColor] (474.63,278.14) circle (  1.49);

\path[fill=fillColor] (463.10,270.77) circle (  1.49);

\path[fill=fillColor] (462.95,270.53) circle (  1.49);

\path[fill=fillColor] (463.00,270.76) circle (  1.49);

\path[fill=fillColor] (463.23,270.79) circle (  1.49);

\path[fill=fillColor] (465.19,272.89) circle (  1.49);

\path[fill=fillColor] (462.88,270.78) circle (  1.49);

\path[fill=fillColor] (462.18,270.50) circle (  1.49);

\path[fill=fillColor] (462.24,269.80) circle (  1.49);

\path[fill=fillColor] (463.05,270.73) circle (  1.49);

\path[fill=fillColor] (463.06,270.76) circle (  1.49);

\path[fill=fillColor] (463.81,270.89) circle (  1.49);

\path[fill=fillColor] (462.19,269.92) circle (  1.49);

\path[fill=fillColor] (463.01,270.78) circle (  1.49);

\path[fill=fillColor] (463.16,270.95) circle (  1.49);

\path[fill=fillColor] (465.32,272.17) circle (  1.49);

\path[fill=fillColor] (465.35,270.68) circle (  1.49);

\path[fill=fillColor] (463.88,270.20) circle (  1.49);

\path[fill=fillColor] (463.38,270.47) circle (  1.49);

\path[fill=fillColor] (462.39,270.61) circle (  1.49);

\path[fill=fillColor] (459.65,271.15) circle (  1.49);

\path[fill=fillColor] (462.78,270.76) circle (  1.49);

\path[fill=fillColor] (466.68,270.41) circle (  1.49);

\path[fill=fillColor] (463.80,270.79) circle (  1.49);

\path[fill=fillColor] (465.46,270.63) circle (  1.49);

\path[fill=fillColor] (465.44,271.75) circle (  1.49);

\path[fill=fillColor] (463.29,270.79) circle (  1.49);

\path[fill=fillColor] (463.04,270.78) circle (  1.49);

\path[fill=fillColor] (462.92,270.87) circle (  1.49);

\path[fill=fillColor] (462.28,269.03) circle (  1.49);

\path[fill=fillColor] (462.18,271.00) circle (  1.49);

\path[fill=fillColor] (462.54,270.75) circle (  1.49);

\path[fill=fillColor] (462.75,271.15) circle (  1.49);

\path[fill=fillColor] (462.95,270.78) circle (  1.49);

\path[fill=fillColor] (462.77,270.73) circle (  1.49);

\path[fill=fillColor] (468.80,268.62) circle (  1.49);

\path[fill=fillColor] (462.75,270.72) circle (  1.49);

\path[fill=fillColor] (463.09,270.79) circle (  1.49);

\path[fill=fillColor] (462.63,270.79) circle (  1.49);

\path[fill=fillColor] (463.05,270.76) circle (  1.49);

\path[fill=fillColor] (462.82,270.70) circle (  1.49);

\path[fill=fillColor] (463.43,270.75) circle (  1.49);

\path[fill=fillColor] (463.04,270.75) circle (  1.49);

\path[fill=fillColor] (463.17,270.75) circle (  1.49);

\path[fill=fillColor] (461.92,270.63) circle (  1.49);

\path[fill=fillColor] (463.13,270.77) circle (  1.49);

\path[fill=fillColor] (462.82,270.77) circle (  1.49);

\path[fill=fillColor] (462.93,270.81) circle (  1.49);

\path[fill=fillColor] (458.97,268.35) circle (  1.49);

\path[fill=fillColor] (463.06,270.77) circle (  1.49);

\path[fill=fillColor] (462.82,270.81) circle (  1.49);

\path[fill=fillColor] (459.76,270.23) circle (  1.49);

\path[fill=fillColor] (467.55,271.17) circle (  1.49);

\path[fill=fillColor] (463.27,270.72) circle (  1.49);

\path[fill=fillColor] (462.77,270.95) circle (  1.49);

\path[fill=fillColor] (463.17,270.85) circle (  1.49);

\path[fill=fillColor] (473.83,270.44) circle (  1.49);

\path[fill=fillColor] (463.06,270.70) circle (  1.49);

\path[fill=fillColor] (462.71,270.66) circle (  1.49);

\path[fill=fillColor] (463.25,270.78) circle (  1.49);

\path[fill=fillColor] (463.84,270.95) circle (  1.49);

\path[fill=fillColor] (462.88,271.02) circle (  1.49);

\path[fill=fillColor] (461.53,270.19) circle (  1.49);

\path[fill=fillColor] (462.79,270.80) circle (  1.49);

\path[fill=fillColor] (513.66,267.09) circle (  1.49);

\path[fill=fillColor] (463.17,270.76) circle (  1.49);

\path[fill=fillColor] (463.15,270.76) circle (  1.49);

\path[fill=fillColor] (462.91,270.77) circle (  1.49);

\path[fill=fillColor] (463.65,270.84) circle (  1.49);

\path[fill=fillColor] (463.36,270.88) circle (  1.49);

\path[fill=fillColor] (463.12,270.98) circle (  1.49);

\path[fill=fillColor] (438.54,248.54) circle (  1.49);

\path[fill=fillColor] (462.48,270.78) circle (  1.49);

\path[fill=fillColor] (462.88,270.28) circle (  1.49);

\path[fill=fillColor] (463.80,271.81) circle (  1.49);

\path[fill=fillColor] (457.61,265.47) circle (  1.49);

\path[fill=fillColor] (463.83,270.47) circle (  1.49);

\path[fill=fillColor] (459.14,270.15) circle (  1.49);

\path[fill=fillColor] (463.79,271.75) circle (  1.49);

\path[fill=fillColor] (463.59,270.77) circle (  1.49);

\path[fill=fillColor] (462.90,270.72) circle (  1.49);

\path[fill=fillColor] (463.04,270.79) circle (  1.49);

\path[fill=fillColor] (460.71,270.63) circle (  1.49);

\path[fill=fillColor] (463.21,270.75) circle (  1.49);

\path[fill=fillColor] (465.07,269.90) circle (  1.49);

\path[fill=fillColor] (461.70,269.36) circle (  1.49);

\path[fill=fillColor] (471.39,273.40) circle (  1.49);

\path[fill=fillColor] (463.04,270.73) circle (  1.49);

\path[fill=fillColor] (453.60,266.85) circle (  1.49);

\path[fill=fillColor] (467.58,271.90) circle (  1.49);

\path[fill=fillColor] (464.40,270.82) circle (  1.49);

\path[fill=fillColor] (463.15,270.73) circle (  1.49);

\path[fill=fillColor] (462.87,270.68) circle (  1.49);

\path[fill=fillColor] (462.62,270.68) circle (  1.49);

\path[fill=fillColor] (463.04,270.76) circle (  1.49);

\path[fill=fillColor] (463.51,270.78) circle (  1.49);

\path[fill=fillColor] (462.92,270.77) circle (  1.49);

\path[fill=fillColor] (463.13,270.73) circle (  1.49);

\path[fill=fillColor] (483.56,272.10) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (489.72,354.55) {Coverage: 94\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (553.30,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (417.91,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (440.48,233.11) -- (440.48,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.04,233.11) -- (463.04,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.61,233.11) -- (485.61,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (508.17,233.11) -- (508.17,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (530.74,233.11) -- (530.74,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.30,233.11) -- (553.30,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (417.91,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (440.48,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (463.04,218.86) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (485.61,218.86) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (508.17,218.86) {2.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (530.74,218.86) {2.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (553.30,218.86) {3.0};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (417.91,346.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (413.95,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,251.94) -- (413.95,251.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,270.78) -- (413.95,270.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,289.61) -- (413.95,289.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,308.44) -- (413.95,308.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,327.27) -- (413.95,327.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,346.10) -- (413.95,346.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,249.67) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,268.50) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,287.33) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,306.17) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,325.00) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,343.83) {3.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (192.72,192.72);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,173.75) {\bfseries Branch rate variance};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28, 10.30) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,100.32) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47, 40.39) rectangle (176.09,160.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47, 40.39) --
	(176.09,160.25);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.19, 41.56) --
	( 37.19, 46.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.55, 40.39) --
	( 32.55, 40.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.08, 40.63) --
	( 33.08, 41.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.57, 40.39) --
	( 32.57, 40.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.61, 40.39) --
	( 32.61, 40.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.95, 40.57) --
	( 32.95, 41.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.53, 40.39) --
	( 32.53, 40.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.93, 40.65) --
	( 32.93, 41.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.68, 41.26) --
	( 36.68, 50.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.82, 40.48) --
	( 32.82, 40.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.41) --
	( 32.50, 40.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.59, 40.40) --
	( 32.59, 40.57);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.58, 40.39) --
	( 32.58, 40.48);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.14, 40.69) --
	( 33.14, 41.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.98, 40.60) --
	( 32.98, 41.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.60, 40.43) --
	( 32.60, 40.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.04, 40.53) --
	( 33.04, 41.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.53, 40.39) --
	( 32.53, 40.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.03, 40.71) --
	( 33.03, 41.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.51, 40.39) --
	( 32.51, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.54, 40.39) --
	( 32.54, 40.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.40) --
	( 32.48, 40.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.58, 41.31) --
	( 35.58, 52.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.64, 40.42) --
	( 32.64, 40.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.63, 40.84) --
	( 33.63, 42.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.04, 40.75) --
	( 33.04, 41.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.97, 40.54) --
	( 32.97, 41.26);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.54, 40.39) --
	( 32.54, 40.44);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.53, 40.39) --
	( 32.53, 40.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.52, 40.41) --
	( 32.52, 40.52);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.10, 40.78) --
	( 34.10, 41.55);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.44);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.56, 40.39) --
	( 32.56, 40.44);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.59, 40.45) --
	( 32.59, 40.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.22, 40.64) --
	(128.22, 78.11);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.51, 40.39) --
	( 32.51, 40.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.55, 40.42) --
	( 32.55, 40.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.53, 40.44) --
	( 32.53, 40.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.52, 40.39) --
	( 56.52,128.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.58, 40.42) --
	( 32.58, 40.50);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.74, 40.49) --
	( 33.74, 41.00);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.75, 41.16) --
	( 52.75, 87.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.64, 40.44) --
	( 32.64, 40.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.02, 40.44) --
	( 33.02, 41.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.45, 40.98) --
	( 33.45, 43.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.62, 40.97) --
	( 33.62, 42.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.74, 40.49) --
	( 32.74, 40.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.70, 40.52) --
	( 32.70, 40.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.82, 41.89) --
	( 46.82, 55.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.01, 41.33) --
	( 38.01, 70.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.51, 40.86) --
	( 33.51, 42.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.68, 40.49) --
	( 32.68, 40.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.42);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.42);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 47.89, 41.71) --
	( 47.89, 50.78);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 32.47, 40.39) circle (  1.49);

\path[fill=fillColor] ( 37.19, 43.69) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.55, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.39) circle (  1.49);

\path[fill=fillColor] ( 33.08, 40.91) circle (  1.49);

\path[fill=fillColor] ( 32.57, 40.46) circle (  1.49);

\path[fill=fillColor] ( 32.61, 40.46) circle (  1.49);

\path[fill=fillColor] ( 32.95, 40.80) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.53, 40.47) circle (  1.49);

\path[fill=fillColor] ( 32.93, 40.95) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.47) circle (  1.49);

\path[fill=fillColor] ( 36.68, 44.68) circle (  1.49);

\path[fill=fillColor] ( 32.82, 40.59) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.55) circle (  1.49);

\path[fill=fillColor] ( 32.59, 40.48) circle (  1.49);

\path[fill=fillColor] ( 32.58, 40.43) circle (  1.49);

\path[fill=fillColor] ( 33.14, 40.98) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.98, 40.86) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.60, 40.49) circle (  1.49);

\path[fill=fillColor] ( 33.04, 40.74) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.53, 40.52) circle (  1.49);

\path[fill=fillColor] ( 33.03, 41.02) circle (  1.49);

\path[fill=fillColor] ( 32.51, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.54, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 35.58, 45.96) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.64, 40.55) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 33.63, 41.42) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 33.04, 41.14) circle (  1.49);

\path[fill=fillColor] ( 32.97, 40.86) circle (  1.49);

\path[fill=fillColor] ( 32.54, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.53, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.52, 40.46) circle (  1.49);

\path[fill=fillColor] ( 34.10, 41.12) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.56, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.59, 40.54) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] (128.22, 53.78) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.51, 40.45) circle (  1.49);

\path[fill=fillColor] ( 32.55, 40.48) circle (  1.49);

\path[fill=fillColor] ( 32.53, 40.49) circle (  1.49);

\path[fill=fillColor] ( 56.52, 74.10) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.58, 40.45) circle (  1.49);

\path[fill=fillColor] ( 33.74, 40.72) circle (  1.49);

\path[fill=fillColor] ( 52.75, 56.50) circle (  1.49);

\path[fill=fillColor] ( 32.64, 40.55) circle (  1.49);

\path[fill=fillColor] ( 33.02, 40.70) circle (  1.49);

\path[fill=fillColor] ( 33.45, 41.85) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 33.62, 41.78) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.74, 40.58) circle (  1.49);

\path[fill=fillColor] ( 32.70, 40.66) circle (  1.49);

\path[fill=fillColor] ( 46.82, 47.45) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 38.01, 51.63) circle (  1.49);

\path[fill=fillColor] ( 33.51, 41.40) circle (  1.49);

\path[fill=fillColor] ( 32.68, 40.59) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 47.89, 45.38) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,161.83) {Coverage: 92\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- (159.11, 40.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 32.47, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 53.58, 40.39) -- ( 53.58, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.68, 40.39) -- ( 74.68, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.79, 40.39) -- ( 95.79, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (116.90, 40.39) -- (116.90, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.00, 40.39) -- (138.00, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.11, 40.39) -- (159.11, 36.43);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47, 26.14) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 53.58, 26.14) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 74.68, 26.14) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 95.79, 26.14) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (116.90, 26.14) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (138.00, 26.14) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (159.11, 26.14) {30};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 32.47,146.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 28.51, 40.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 58.01) -- ( 28.51, 58.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 75.62) -- ( 28.51, 75.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 93.23) -- ( 28.51, 93.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,110.85) -- ( 28.51,110.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,128.46) -- ( 28.51,128.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,146.08) -- ( 28.51,146.08);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 38.12) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 55.73) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 73.35) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 90.96) {15};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,108.58) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,126.19) {25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,143.80) {30};
\end{scope}
\end{tikzpicture}

\end{document}