### 2020-09-01 RatesWCSS/WCSS_cat_adaptive.tex

% Created by tikzDevice version 0.12.3.1 on 2020-09-01 10:42:19
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,385.44) rectangle (192.72,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,559.19) {\bfseries Tree height $h$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47,425.83) rectangle (176.09,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47,425.83) --
	(176.09,545.69);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.33,450.31) --
	( 62.33,452.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.13,431.81) --
	( 41.13,437.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.89,450.33) --
	( 62.89,452.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.58,445.50) --
	( 57.58,450.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.19,457.72) --
	( 74.19,462.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.84,440.45) --
	( 50.84,442.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.83,433.64) --
	( 44.83,439.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.91,430.04) --
	( 37.91,430.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.92,426.94) --
	( 33.92,427.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.02,433.88) --
	( 43.02,436.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.93,430.96) --
	( 38.93,432.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.00,436.10) --
	( 46.00,437.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.87,428.91) --
	( 36.87,429.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (119.03,484.29) --
	(119.03,529.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.34,436.05) --
	( 45.34,437.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.96,433.35) --
	( 41.96,434.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.58,439.80) --
	( 59.58,457.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.72,440.80) --
	( 52.72,447.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.54,428.92) --
	( 36.54,429.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.77,429.68) --
	( 37.77,430.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.63,439.54) --
	( 50.63,443.04);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.95,434.64) --
	( 49.95,440.19);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.75,445.13) --
	( 57.75,447.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.99,452.30) --
	( 72.99,464.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.01,440.84) --
	( 52.01,442.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.50,434.78) --
	( 43.50,436.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.23,430.05) --
	( 39.23,432.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.17,442.44) --
	( 54.17,445.51);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.64,459.66) --
	( 72.64,463.13);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.64,431.59) --
	( 40.64,433.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.71,436.56) --
	( 46.71,440.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.64,456.10) --
	( 70.64,462.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.10,427.08) --
	( 34.10,427.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.38,440.59) --
	( 51.38,443.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.37,432.08) --
	( 40.37,432.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.58,431.60) --
	( 39.58,432.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.20,431.96) --
	( 45.20,441.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.72,430.12) --
	( 37.72,431.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.25,450.86) --
	( 63.25,454.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.23,441.83) --
	( 52.23,443.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.57,451.95) --
	( 65.57,454.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.60,426.56) --
	( 33.60,426.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.15,438.74) --
	( 50.15,440.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.22,436.72) --
	( 46.22,437.90);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.95,459.87) --
	( 72.95,466.23);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.58,429.89) --
	( 38.58,430.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.71,432.91) --
	( 41.71,434.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 47.05,437.56) --
	( 47.05,439.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.34,435.60) --
	( 45.34,436.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.60,430.55) --
	( 41.60,434.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.57,431.62) --
	( 39.57,432.45);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.70,436.12) --
	( 46.70,437.61);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.57,431.85) --
	( 42.57,437.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.05,427.39) --
	( 35.05,428.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.95,428.59) --
	( 35.95,429.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.92,432.28) --
	( 40.92,433.34);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.56,460.81) --
	( 81.56,466.35);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.39,433.91) --
	( 46.39,438.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.22,429.64) --
	( 37.22,430.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.59,450.42) --
	( 62.59,453.57);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.58,428.98) --
	( 37.58,429.91);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.72,444.02) --
	( 56.72,447.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.45,437.13) --
	( 46.45,438.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.06,432.61) --
	( 43.06,434.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.79,471.66) --
	( 90.79,477.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.18,429.17) --
	( 41.18,434.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.41,431.38) --
	( 39.41,432.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.55,444.55) --
	( 56.55,446.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.06,440.09) --
	( 50.06,441.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.99,434.08) --
	( 42.99,434.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.01,432.74) --
	( 41.01,434.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.68,444.46) --
	( 57.68,447.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.14,429.51) --
	( 43.14,439.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.04,434.06) --
	( 43.04,434.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.80,453.73) --
	( 70.80,458.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.42,427.88) --
	( 35.42,429.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.54,432.36) --
	( 45.54,445.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.45,431.81) --
	( 40.45,433.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.80,428.09) --
	( 35.80,429.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.63,432.90) --
	( 43.63,440.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.52,435.21) --
	( 44.52,436.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.22,432.92) --
	( 41.22,433.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.67,442.21) --
	( 52.67,443.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.92,431.08) --
	( 40.92,434.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.30,430.49) --
	( 38.30,431.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 67.25,453.02) --
	( 67.25,459.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.11,434.00) --
	( 44.11,438.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.46,428.16) --
	( 36.46,431.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.38,442.33) --
	( 52.38,443.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.40,428.08) --
	( 36.40,435.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.74,431.95) --
	( 42.74,436.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.58,435.35) --
	( 45.58,438.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.79,430.81) --
	( 38.79,431.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.77,427.61) --
	( 34.77,428.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.38,432.62) --
	( 41.38,433.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.11,451.63) --
	( 65.11,455.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.73,454.66) --
	( 68.73,459.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.68,427.95) --
	( 35.68,428.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.11,429.41) --
	( 37.11,430.02);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.57,452.67) --
	( 61.57,509.88);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 62.33,451.56) circle (  1.49);

\path[fill=fillColor] ( 41.13,434.39) circle (  1.49);

\path[fill=fillColor] ( 62.89,451.49) circle (  1.49);

\path[fill=fillColor] ( 57.58,447.80) circle (  1.49);

\path[fill=fillColor] ( 74.19,459.87) circle (  1.49);

\path[fill=fillColor] ( 50.84,441.25) circle (  1.49);

\path[fill=fillColor] ( 44.83,436.30) circle (  1.49);

\path[fill=fillColor] ( 37.91,430.49) circle (  1.49);

\path[fill=fillColor] ( 33.92,427.17) circle (  1.49);

\path[fill=fillColor] ( 43.02,435.33) circle (  1.49);

\path[fill=fillColor] ( 38.93,431.52) circle (  1.49);

\path[fill=fillColor] ( 46.00,436.85) circle (  1.49);

\path[fill=fillColor] ( 36.87,429.38) circle (  1.49);

\path[fill=fillColor] (119.03,505.74) circle (  1.49);

\path[fill=fillColor] ( 45.34,436.56) circle (  1.49);

\path[fill=fillColor] ( 41.96,433.93) circle (  1.49);

\path[fill=fillColor] ( 59.58,447.71) circle (  1.49);

\path[fill=fillColor] ( 52.72,443.85) circle (  1.49);

\path[fill=fillColor] ( 36.54,429.25) circle (  1.49);

\path[fill=fillColor] ( 37.77,430.12) circle (  1.49);

\path[fill=fillColor] ( 50.63,441.23) circle (  1.49);

\path[fill=fillColor] ( 49.95,437.25) circle (  1.49);

\path[fill=fillColor] ( 57.75,446.19) circle (  1.49);

\path[fill=fillColor] ( 72.99,458.25) circle (  1.49);

\path[fill=fillColor] ( 52.01,441.87) circle (  1.49);

\path[fill=fillColor] ( 43.50,435.62) circle (  1.49);

\path[fill=fillColor] ( 39.23,431.36) circle (  1.49);

\path[fill=fillColor] ( 54.17,444.00) circle (  1.49);

\path[fill=fillColor] ( 72.64,461.34) circle (  1.49);

\path[fill=fillColor] ( 40.64,432.39) circle (  1.49);

\path[fill=fillColor] ( 46.71,438.45) circle (  1.49);

\path[fill=fillColor] ( 70.64,459.50) circle (  1.49);

\path[fill=fillColor] ( 34.10,427.26) circle (  1.49);

\path[fill=fillColor] ( 51.38,442.07) circle (  1.49);

\path[fill=fillColor] ( 40.37,432.46) circle (  1.49);

\path[fill=fillColor] ( 39.58,432.19) circle (  1.49);

\path[fill=fillColor] ( 45.20,436.05) circle (  1.49);

\path[fill=fillColor] ( 37.72,430.56) circle (  1.49);

\path[fill=fillColor] ( 63.25,452.61) circle (  1.49);

\path[fill=fillColor] ( 52.23,442.66) circle (  1.49);

\path[fill=fillColor] ( 65.57,453.41) circle (  1.49);

\path[fill=fillColor] ( 33.60,426.71) circle (  1.49);

\path[fill=fillColor] ( 50.15,439.64) circle (  1.49);

\path[fill=fillColor] ( 46.22,437.31) circle (  1.49);

\path[fill=fillColor] ( 72.95,462.97) circle (  1.49);

\path[fill=fillColor] ( 38.58,430.43) circle (  1.49);

\path[fill=fillColor] ( 41.71,433.50) circle (  1.49);

\path[fill=fillColor] ( 47.05,438.32) circle (  1.49);

\path[fill=fillColor] ( 45.34,436.31) circle (  1.49);

\path[fill=fillColor] ( 41.60,432.28) circle (  1.49);

\path[fill=fillColor] ( 39.57,432.04) circle (  1.49);

\path[fill=fillColor] ( 46.70,436.87) circle (  1.49);

\path[fill=fillColor] ( 42.57,434.70) circle (  1.49);

\path[fill=fillColor] ( 35.05,427.71) circle (  1.49);

\path[fill=fillColor] ( 35.95,428.85) circle (  1.49);

\path[fill=fillColor] ( 40.92,432.80) circle (  1.49);

\path[fill=fillColor] ( 81.56,463.54) circle (  1.49);

\path[fill=fillColor] ( 46.39,435.92) circle (  1.49);

\path[fill=fillColor] ( 37.22,429.98) circle (  1.49);

\path[fill=fillColor] ( 62.59,451.93) circle (  1.49);

\path[fill=fillColor] ( 37.58,429.44) circle (  1.49);

\path[fill=fillColor] ( 56.72,445.53) circle (  1.49);

\path[fill=fillColor] ( 46.45,437.79) circle (  1.49);

\path[fill=fillColor] ( 43.06,433.73) circle (  1.49);

\path[fill=fillColor] ( 90.79,474.71) circle (  1.49);

\path[fill=fillColor] ( 41.18,431.49) circle (  1.49);

\path[fill=fillColor] ( 39.41,431.74) circle (  1.49);

\path[fill=fillColor] ( 56.55,445.44) circle (  1.49);

\path[fill=fillColor] ( 50.06,440.70) circle (  1.49);

\path[fill=fillColor] ( 42.99,434.51) circle (  1.49);

\path[fill=fillColor] ( 41.01,433.37) circle (  1.49);

\path[fill=fillColor] ( 57.68,446.01) circle (  1.49);

\path[fill=fillColor] ( 43.14,433.73) circle (  1.49);

\path[fill=fillColor] ( 43.04,434.51) circle (  1.49);

\path[fill=fillColor] ( 70.80,455.87) circle (  1.49);

\path[fill=fillColor] ( 35.42,428.77) circle (  1.49);

\path[fill=fillColor] ( 45.54,438.29) circle (  1.49);

\path[fill=fillColor] ( 40.45,432.71) circle (  1.49);

\path[fill=fillColor] ( 35.80,428.70) circle (  1.49);

\path[fill=fillColor] ( 43.63,436.43) circle (  1.49);

\path[fill=fillColor] ( 44.52,435.69) circle (  1.49);

\path[fill=fillColor] ( 41.22,433.33) circle (  1.49);

\path[fill=fillColor] ( 52.67,443.00) circle (  1.49);

\path[fill=fillColor] ( 40.92,432.84) circle (  1.49);

\path[fill=fillColor] ( 38.30,430.90) circle (  1.49);

\path[fill=fillColor] ( 67.25,456.24) circle (  1.49);

\path[fill=fillColor] ( 44.11,436.12) circle (  1.49);

\path[fill=fillColor] ( 36.46,429.46) circle (  1.49);

\path[fill=fillColor] ( 52.38,442.98) circle (  1.49);

\path[fill=fillColor] ( 36.40,430.65) circle (  1.49);

\path[fill=fillColor] ( 42.74,434.01) circle (  1.49);

\path[fill=fillColor] ( 45.58,437.01) circle (  1.49);

\path[fill=fillColor] ( 38.79,431.30) circle (  1.49);

\path[fill=fillColor] ( 34.77,427.86) circle (  1.49);

\path[fill=fillColor] ( 41.38,433.02) circle (  1.49);

\path[fill=fillColor] ( 65.11,453.84) circle (  1.49);

\path[fill=fillColor] ( 68.73,456.83) circle (  1.49);

\path[fill=fillColor] ( 35.68,428.23) circle (  1.49);

\path[fill=fillColor] ( 37.11,429.72) circle (  1.49);

\path[fill=fillColor] ( 61.57,476.51) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,547.27) {Coverage: 93\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- (166.85,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 32.47,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.35,425.83) -- ( 59.35,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.22,425.83) -- ( 86.22,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (113.10,425.83) -- (113.10,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.97,425.83) -- (139.97,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (166.85,425.83) -- (166.85,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47,411.58) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 59.35,411.58) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 86.22,411.58) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (113.10,411.58) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (139.97,411.58) {2.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (166.85,411.58) {2.5};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 32.47,537.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 28.51,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,448.26) -- ( 28.51,448.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,470.69) -- ( 28.51,470.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,493.12) -- ( 28.51,493.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,515.55) -- ( 28.51,515.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,537.98) -- ( 28.51,537.98);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,423.56) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,445.99) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,468.42) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,490.85) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,513.28) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,535.71) {2.5};
\end{scope}
\begin{scope}
\path[clip] (192.72,385.44) rectangle (385.44,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (297.00,559.19) {\bfseries Yule birth rate $\lambda$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (297.00,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (199.85,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (225.19,425.83) rectangle (368.81,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (225.19,425.83) --
	(368.81,545.69);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.87,428.62) --
	(232.87,431.65);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.82,431.94) --
	(239.82,440.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.76,427.94) --
	(228.76,430.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.34,428.91) --
	(231.34,432.38);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.80,427.16) --
	(228.80,428.58);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.62,430.12) --
	(232.62,435.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.92,431.89) --
	(237.92,438.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.62,434.43) --
	(240.62,444.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (287.09,459.31) --
	(287.09,500.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.27,431.85) --
	(237.27,439.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.93,434.73) --
	(240.93,444.96);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.88,431.60) --
	(239.88,437.89);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.07,440.70) --
	(245.07,458.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.73,426.73) --
	(226.73,427.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.30,429.78) --
	(231.30,434.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.42,434.30) --
	(241.42,443.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.10,428.60) --
	(230.10,432.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.21,429.18) --
	(232.21,433.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.83,441.37) --
	(263.83,459.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.71,435.55) --
	(246.71,446.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.78,430.11) --
	(234.78,434.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.23,430.40) --
	(234.23,435.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.69,428.44) --
	(229.69,431.33);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.18,428.08) --
	(231.18,430.71);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.85,429.87) --
	(233.85,434.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.30,430.44) --
	(234.30,435.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.60,434.57) --
	(239.60,446.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.07,429.62) --
	(231.07,433.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.45,427.30) --
	(228.45,428.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.87,435.67) --
	(241.87,446.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.60,429.05) --
	(230.60,432.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.79,428.33) --
	(229.79,431.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (307.77,460.47) --
	(307.77,501.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.08,429.19) --
	(232.08,432.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.73,433.14) --
	(238.73,441.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.02,432.94) --
	(242.02,441.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.81,429.93) --
	(231.81,437.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.32,435.42) --
	(246.32,446.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.35,428.42) --
	(230.35,431.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.52,429.32) --
	(231.52,433.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.63,427.82) --
	(229.63,430.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (288.07,475.68) --
	(288.07,536.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.24,430.72) --
	(233.24,436.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.34,431.00) --
	(235.34,436.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.92,427.23) --
	(227.92,428.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.83,439.91) --
	(252.83,456.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.07,433.12) --
	(240.07,441.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.32,430.28) --
	(235.32,435.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.68,431.44) --
	(235.68,437.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.93,434.93) --
	(239.93,446.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (244.38,434.52) --
	(244.38,444.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.45,429.98) --
	(233.45,434.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.88,431.31) --
	(236.88,438.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.96,448.04) --
	(257.96,475.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.76,445.08) --
	(261.76,467.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.35,432.66) --
	(237.35,440.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.88,427.66) --
	(228.88,429.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.13,430.54) --
	(236.13,436.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.40,436.86) --
	(243.40,449.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.79,427.87) --
	(228.79,430.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.49,439.90) --
	(247.49,455.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.70,428.14) --
	(230.70,430.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.41,430.02) --
	(232.41,434.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.57,432.15) --
	(234.57,439.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.70,426.93) --
	(227.70,428.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.82,434.39) --
	(245.82,451.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.76,433.11) --
	(237.76,441.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.82,428.28) --
	(229.82,430.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.66,428.83) --
	(231.66,432.23);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.27,431.30) --
	(239.27,437.43);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.25,431.23) --
	(234.25,437.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.61,428.49) --
	(231.61,431.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.04,430.65) --
	(241.04,448.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.22,431.78) --
	(240.22,438.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.72,427.33) --
	(227.72,429.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.20,440.97) --
	(255.20,462.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.41,429.32) --
	(238.41,436.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.10,433.43) --
	(239.10,442.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.34,446.41) --
	(257.34,472.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.46,431.29) --
	(235.46,439.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.19,431.03) --
	(237.19,436.86);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.23,432.08) --
	(241.23,439.21);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.14,428.82) --
	(232.14,432.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.44,433.34) --
	(243.44,443.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.47,434.91) --
	(241.47,445.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.82,427.41) --
	(227.82,429.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.43,430.47) --
	(235.43,436.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.02,440.04) --
	(253.02,463.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.12,428.06) --
	(229.12,430.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.63,433.06) --
	(254.63,462.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.15,430.76) --
	(234.15,436.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.95,430.26) --
	(234.95,435.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.71,436.87) --
	(248.71,449.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.33,452.04) --
	(261.33,483.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.97,432.35) --
	(239.97,439.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.93,428.61) --
	(228.93,431.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.43,427.44) --
	(228.43,429.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.69,449.17) --
	(275.69,476.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.90,438.65) --
	(242.90,453.61);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.33,426.77) --
	(229.33,429.03);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (232.87,430.07) circle (  1.49);

\path[fill=fillColor] (239.82,436.13) circle (  1.49);

\path[fill=fillColor] (228.76,429.06) circle (  1.49);

\path[fill=fillColor] (231.34,430.58) circle (  1.49);

\path[fill=fillColor] (228.80,427.83) circle (  1.49);

\path[fill=fillColor] (232.62,432.59) circle (  1.49);

\path[fill=fillColor] (237.92,435.23) circle (  1.49);

\path[fill=fillColor] (240.62,439.39) circle (  1.49);

\path[fill=fillColor] (287.09,479.35) circle (  1.49);

\path[fill=fillColor] (237.27,435.38) circle (  1.49);

\path[fill=fillColor] (240.93,439.71) circle (  1.49);

\path[fill=fillColor] (239.88,434.53) circle (  1.49);

\path[fill=fillColor] (245.07,449.29) circle (  1.49);

\path[fill=fillColor] (226.73,427.23) circle (  1.49);

\path[fill=fillColor] (231.30,432.02) circle (  1.49);

\path[fill=fillColor] (241.42,438.88) circle (  1.49);

\path[fill=fillColor] (230.10,430.45) circle (  1.49);

\path[fill=fillColor] (232.21,431.01) circle (  1.49);

\path[fill=fillColor] (263.83,450.27) circle (  1.49);

\path[fill=fillColor] (246.71,440.60) circle (  1.49);

\path[fill=fillColor] (234.78,432.46) circle (  1.49);

\path[fill=fillColor] (234.23,432.99) circle (  1.49);

\path[fill=fillColor] (229.69,429.86) circle (  1.49);

\path[fill=fillColor] (231.18,429.36) circle (  1.49);

\path[fill=fillColor] (233.85,431.99) circle (  1.49);

\path[fill=fillColor] (234.30,432.98) circle (  1.49);

\path[fill=fillColor] (239.60,440.35) circle (  1.49);

\path[fill=fillColor] (231.07,431.74) circle (  1.49);

\path[fill=fillColor] (228.45,428.08) circle (  1.49);

\path[fill=fillColor] (241.87,441.11) circle (  1.49);

\path[fill=fillColor] (230.60,430.86) circle (  1.49);

\path[fill=fillColor] (229.79,429.72) circle (  1.49);

\path[fill=fillColor] (307.77,480.52) circle (  1.49);

\path[fill=fillColor] (232.08,430.99) circle (  1.49);

\path[fill=fillColor] (238.73,437.10) circle (  1.49);

\path[fill=fillColor] (242.02,436.88) circle (  1.49);

\path[fill=fillColor] (231.81,433.43) circle (  1.49);

\path[fill=fillColor] (246.32,440.76) circle (  1.49);

\path[fill=fillColor] (230.35,429.91) circle (  1.49);

\path[fill=fillColor] (231.52,431.16) circle (  1.49);

\path[fill=fillColor] (229.63,428.88) circle (  1.49);

\path[fill=fillColor] (288.07,505.74) circle (  1.49);

\path[fill=fillColor] (233.24,433.43) circle (  1.49);

\path[fill=fillColor] (235.34,433.73) circle (  1.49);

\path[fill=fillColor] (227.92,427.98) circle (  1.49);

\path[fill=fillColor] (252.83,447.80) circle (  1.49);

\path[fill=fillColor] (240.07,437.32) circle (  1.49);

\path[fill=fillColor] (235.32,432.75) circle (  1.49);

\path[fill=fillColor] (235.68,434.26) circle (  1.49);

\path[fill=fillColor] (239.93,440.20) circle (  1.49);

\path[fill=fillColor] (244.38,439.39) circle (  1.49);

\path[fill=fillColor] (233.45,432.19) circle (  1.49);

\path[fill=fillColor] (236.88,434.69) circle (  1.49);

\path[fill=fillColor] (257.96,461.02) circle (  1.49);

\path[fill=fillColor] (261.76,456.23) circle (  1.49);

\path[fill=fillColor] (237.35,436.38) circle (  1.49);

\path[fill=fillColor] (228.88,428.63) circle (  1.49);

\path[fill=fillColor] (236.13,433.32) circle (  1.49);

\path[fill=fillColor] (243.40,443.04) circle (  1.49);

\path[fill=fillColor] (228.79,428.96) circle (  1.49);

\path[fill=fillColor] (247.49,447.44) circle (  1.49);

\path[fill=fillColor] (230.70,429.41) circle (  1.49);

\path[fill=fillColor] (232.41,432.17) circle (  1.49);

\path[fill=fillColor] (234.57,435.54) circle (  1.49);

\path[fill=fillColor] (227.70,427.53) circle (  1.49);

\path[fill=fillColor] (245.82,442.60) circle (  1.49);

\path[fill=fillColor] (237.76,437.14) circle (  1.49);

\path[fill=fillColor] (229.82,429.60) circle (  1.49);

\path[fill=fillColor] (231.66,430.52) circle (  1.49);

\path[fill=fillColor] (239.27,434.29) circle (  1.49);

\path[fill=fillColor] (234.25,434.22) circle (  1.49);

\path[fill=fillColor] (231.61,429.93) circle (  1.49);

\path[fill=fillColor] (241.04,438.73) circle (  1.49);

\path[fill=fillColor] (240.22,435.02) circle (  1.49);

\path[fill=fillColor] (227.72,428.17) circle (  1.49);

\path[fill=fillColor] (255.20,451.35) circle (  1.49);

\path[fill=fillColor] (238.41,432.95) circle (  1.49);

\path[fill=fillColor] (239.10,437.53) circle (  1.49);

\path[fill=fillColor] (257.34,458.82) circle (  1.49);

\path[fill=fillColor] (235.46,435.25) circle (  1.49);

\path[fill=fillColor] (237.19,433.82) circle (  1.49);

\path[fill=fillColor] (241.23,435.57) circle (  1.49);

\path[fill=fillColor] (232.14,430.47) circle (  1.49);

\path[fill=fillColor] (243.44,437.99) circle (  1.49);

\path[fill=fillColor] (241.47,440.05) circle (  1.49);

\path[fill=fillColor] (227.82,428.27) circle (  1.49);

\path[fill=fillColor] (235.43,433.35) circle (  1.49);

\path[fill=fillColor] (253.02,451.41) circle (  1.49);

\path[fill=fillColor] (229.12,429.32) circle (  1.49);

\path[fill=fillColor] (254.63,447.41) circle (  1.49);

\path[fill=fillColor] (234.15,433.78) circle (  1.49);

\path[fill=fillColor] (234.95,432.57) circle (  1.49);

\path[fill=fillColor] (248.71,443.06) circle (  1.49);

\path[fill=fillColor] (261.33,467.22) circle (  1.49);

\path[fill=fillColor] (239.97,435.80) circle (  1.49);

\path[fill=fillColor] (228.93,430.14) circle (  1.49);

\path[fill=fillColor] (228.43,428.31) circle (  1.49);

\path[fill=fillColor] (275.69,462.26) circle (  1.49);

\path[fill=fillColor] (242.90,445.93) circle (  1.49);

\path[fill=fillColor] (229.33,427.82) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (297.00,547.27) {Coverage: 93\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (341.37,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (225.19,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.92,425.83) -- (263.92,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.64,425.83) -- (302.64,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (341.37,425.83) -- (341.37,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (225.19,411.58) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (263.92,411.58) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (302.64,411.58) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (341.37,411.58) {150};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (225.19,522.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (221.23,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,458.15) -- (221.23,458.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,490.47) -- (221.23,490.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,522.79) -- (221.23,522.79);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,423.56) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,455.88) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,488.19) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,520.51) {150};
\end{scope}
\begin{scope}
\path[clip] (385.44,385.44) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (489.72,559.19) {\bfseries HKY model $\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (489.72,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (392.57,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (417.91,425.83) rectangle (561.53,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (417.91,425.83) --
	(561.53,545.69);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.07,426.78) --
	(419.07,426.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.37,464.31) --
	(467.37,476.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.56,432.83) --
	(426.56,433.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.82,426.56) --
	(418.82,426.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.76,437.65) --
	(433.76,439.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.75,428.08) --
	(420.75,428.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.86,429.96) --
	(422.86,430.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.13,430.84) --
	(424.13,431.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (441.24,440.99) --
	(441.24,449.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.44,427.03) --
	(419.44,427.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.83,429.03) --
	(421.83,429.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.41,427.81) --
	(420.41,428.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.50,430.89) --
	(424.50,432.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.62,428.81) --
	(421.62,429.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.39,426.96) --
	(419.39,427.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.45,427.81) --
	(420.45,428.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.38,434.86) --
	(429.38,435.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.50,427.92) --
	(420.50,428.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (512.46,486.12) --
	(512.46,527.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.82,427.34) --
	(419.82,427.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.88,427.39) --
	(419.88,427.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.92,428.88) --
	(421.92,429.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.55,429.39) --
	(422.55,429.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.30,436.61) --
	(431.30,437.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.20,430.01) --
	(423.20,430.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.24,426.07) --
	(418.24,426.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.69,428.83) --
	(421.69,429.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.19,432.17) --
	(426.19,432.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.93,426.67) --
	(418.93,426.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.14,428.40) --
	(421.14,428.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.79,428.15) --
	(420.79,428.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.17,426.02) --
	(418.17,426.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.05,430.89) --
	(424.05,432.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.81,429.60) --
	(422.81,430.04);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.44,429.71) --
	(422.44,430.32);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.62,428.74) --
	(421.62,429.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.13,426.81) --
	(419.13,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.77,432.90) --
	(426.77,434.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.99,431.23) --
	(424.99,431.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.62,427.89) --
	(420.62,428.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.49,431.81) --
	(425.49,432.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.24,432.22) --
	(427.24,435.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.25,433.12) --
	(427.25,434.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.70,426.48) --
	(418.70,426.60);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (451.72,454.07) --
	(451.72,457.60);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.29,433.72) --
	(428.29,435.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.52,434.77) --
	(429.52,436.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.04,427.53) --
	(420.04,427.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.57,427.80) --
	(420.57,428.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.60,454.32) --
	(452.60,464.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.65,427.16) --
	(419.65,427.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.65,434.67) --
	(428.65,435.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.19,426.80) --
	(419.19,426.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.75,431.71) --
	(425.75,433.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.31,427.60) --
	(420.31,428.08);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.79,431.28) --
	(425.79,432.18);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.54,429.59) --
	(422.54,429.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.09,427.57) --
	(420.09,427.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.72,427.20) --
	(419.72,427.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.22,426.06) --
	(418.22,426.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.71,426.36) --
	(418.71,426.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.03,430.73) --
	(424.03,431.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.51,453.11) --
	(454.51,458.43);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.71,429.93) --
	(423.71,430.56);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (448.74,449.49) --
	(448.74,451.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.82,426.47) --
	(418.82,426.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.00,427.40) --
	(420.00,427.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.69,426.45) --
	(418.69,426.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.27,428.48) --
	(421.27,428.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.87,428.10) --
	(420.87,428.42);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.80,430.05) --
	(422.80,430.64);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.64,428.70) --
	(421.64,429.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.36,431.38) --
	(425.36,432.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.15,432.36) --
	(426.15,433.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.85,426.56) --
	(418.85,426.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.02,438.39) --
	(435.02,442.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.34,428.47) --
	(421.34,428.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (441.84,443.15) --
	(441.84,447.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (449.27,446.51) --
	(449.27,456.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.41,426.19) --
	(418.41,426.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.68,431.71) --
	(425.68,432.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.11,454.72) --
	(461.11,462.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.60,428.80) --
	(421.60,429.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.73,432.46) --
	(426.73,433.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.20,434.07) --
	(428.20,435.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.81,427.37) --
	(419.81,427.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.95,432.89) --
	(426.95,433.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.86,428.06) --
	(420.86,428.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.18,441.04) --
	(437.18,442.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.67,428.80) --
	(421.67,429.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.24,439.78) --
	(436.24,442.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.21,427.66) --
	(420.21,427.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.75,433.83) --
	(427.75,435.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.43,430.83) --
	(424.43,432.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.74,428.75) --
	(421.74,429.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.32,430.16) --
	(423.32,430.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.05,429.10) --
	(422.05,429.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.18,431.06) --
	(425.18,432.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.35,426.17) --
	(418.35,426.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.24,429.32) --
	(422.24,429.70);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (419.07,426.85) circle (  1.49);

\path[fill=fillColor] (467.37,470.48) circle (  1.49);

\path[fill=fillColor] (426.56,433.17) circle (  1.49);

\path[fill=fillColor] (418.82,426.62) circle (  1.49);

\path[fill=fillColor] (433.76,438.35) circle (  1.49);

\path[fill=fillColor] (420.75,428.22) circle (  1.49);

\path[fill=fillColor] (422.86,430.27) circle (  1.49);

\path[fill=fillColor] (424.13,431.29) circle (  1.49);

\path[fill=fillColor] (441.24,444.75) circle (  1.49);

\path[fill=fillColor] (419.44,427.13) circle (  1.49);

\path[fill=fillColor] (421.83,429.32) circle (  1.49);

\path[fill=fillColor] (420.41,427.95) circle (  1.49);

\path[fill=fillColor] (424.50,431.50) circle (  1.49);

\path[fill=fillColor] (421.62,428.95) circle (  1.49);

\path[fill=fillColor] (419.39,427.04) circle (  1.49);

\path[fill=fillColor] (420.45,427.98) circle (  1.49);

\path[fill=fillColor] (429.38,435.38) circle (  1.49);

\path[fill=fillColor] (420.50,428.05) circle (  1.49);

\path[fill=fillColor] (512.46,505.74) circle (  1.49);

\path[fill=fillColor] (419.82,427.48) circle (  1.49);

\path[fill=fillColor] (419.88,427.50) circle (  1.49);

\path[fill=fillColor] (421.92,429.08) circle (  1.49);

\path[fill=fillColor] (422.55,429.58) circle (  1.49);

\path[fill=fillColor] (431.30,437.24) circle (  1.49);

\path[fill=fillColor] (423.20,430.27) circle (  1.49);

\path[fill=fillColor] (418.24,426.11) circle (  1.49);

\path[fill=fillColor] (421.69,429.12) circle (  1.49);

\path[fill=fillColor] (426.19,432.58) circle (  1.49);

\path[fill=fillColor] (418.93,426.72) circle (  1.49);

\path[fill=fillColor] (421.14,428.65) circle (  1.49);

\path[fill=fillColor] (420.79,428.28) circle (  1.49);

\path[fill=fillColor] (418.17,426.05) circle (  1.49);

\path[fill=fillColor] (424.05,431.91) circle (  1.49);

\path[fill=fillColor] (422.81,429.82) circle (  1.49);

\path[fill=fillColor] (422.44,430.01) circle (  1.49);

\path[fill=fillColor] (421.62,428.96) circle (  1.49);

\path[fill=fillColor] (419.13,426.90) circle (  1.49);

\path[fill=fillColor] (426.77,433.60) circle (  1.49);

\path[fill=fillColor] (424.99,431.51) circle (  1.49);

\path[fill=fillColor] (420.62,428.01) circle (  1.49);

\path[fill=fillColor] (425.49,432.11) circle (  1.49);

\path[fill=fillColor] (427.24,433.82) circle (  1.49);

\path[fill=fillColor] (427.25,433.63) circle (  1.49);

\path[fill=fillColor] (418.70,426.54) circle (  1.49);

\path[fill=fillColor] (451.72,455.83) circle (  1.49);

\path[fill=fillColor] (428.29,434.72) circle (  1.49);

\path[fill=fillColor] (429.52,435.59) circle (  1.49);

\path[fill=fillColor] (420.04,427.64) circle (  1.49);

\path[fill=fillColor] (420.57,427.94) circle (  1.49);

\path[fill=fillColor] (452.60,459.60) circle (  1.49);

\path[fill=fillColor] (419.65,427.29) circle (  1.49);

\path[fill=fillColor] (428.65,435.25) circle (  1.49);

\path[fill=fillColor] (419.19,426.88) circle (  1.49);

\path[fill=fillColor] (425.75,432.60) circle (  1.49);

\path[fill=fillColor] (420.31,427.83) circle (  1.49);

\path[fill=fillColor] (425.79,431.73) circle (  1.49);

\path[fill=fillColor] (422.54,429.78) circle (  1.49);

\path[fill=fillColor] (420.09,427.69) circle (  1.49);

\path[fill=fillColor] (419.72,427.35) circle (  1.49);

\path[fill=fillColor] (418.22,426.09) circle (  1.49);

\path[fill=fillColor] (418.71,426.45) circle (  1.49);

\path[fill=fillColor] (424.03,430.97) circle (  1.49);

\path[fill=fillColor] (454.51,455.68) circle (  1.49);

\path[fill=fillColor] (423.71,430.25) circle (  1.49);

\path[fill=fillColor] (448.74,450.73) circle (  1.49);

\path[fill=fillColor] (418.82,426.55) circle (  1.49);

\path[fill=fillColor] (420.00,427.53) circle (  1.49);

\path[fill=fillColor] (418.69,426.50) circle (  1.49);

\path[fill=fillColor] (421.27,428.62) circle (  1.49);

\path[fill=fillColor] (420.87,428.26) circle (  1.49);

\path[fill=fillColor] (422.80,430.34) circle (  1.49);

\path[fill=fillColor] (421.64,428.85) circle (  1.49);

\path[fill=fillColor] (425.36,431.94) circle (  1.49);

\path[fill=fillColor] (426.15,432.85) circle (  1.49);

\path[fill=fillColor] (418.85,426.61) circle (  1.49);

\path[fill=fillColor] (435.02,440.52) circle (  1.49);

\path[fill=fillColor] (421.34,428.66) circle (  1.49);

\path[fill=fillColor] (441.84,445.08) circle (  1.49);

\path[fill=fillColor] (449.27,451.13) circle (  1.49);

\path[fill=fillColor] (418.41,426.24) circle (  1.49);

\path[fill=fillColor] (425.68,432.14) circle (  1.49);

\path[fill=fillColor] (461.11,458.59) circle (  1.49);

\path[fill=fillColor] (421.60,428.96) circle (  1.49);

\path[fill=fillColor] (426.73,433.06) circle (  1.49);

\path[fill=fillColor] (428.20,434.85) circle (  1.49);

\path[fill=fillColor] (419.81,427.44) circle (  1.49);

\path[fill=fillColor] (426.95,433.41) circle (  1.49);

\path[fill=fillColor] (420.86,428.34) circle (  1.49);

\path[fill=fillColor] (437.18,441.87) circle (  1.49);

\path[fill=fillColor] (421.67,429.19) circle (  1.49);

\path[fill=fillColor] (436.24,440.96) circle (  1.49);

\path[fill=fillColor] (420.21,427.78) circle (  1.49);

\path[fill=fillColor] (427.75,434.68) circle (  1.49);

\path[fill=fillColor] (424.43,431.66) circle (  1.49);

\path[fill=fillColor] (421.74,428.97) circle (  1.49);

\path[fill=fillColor] (423.32,430.39) circle (  1.49);

\path[fill=fillColor] (422.05,429.25) circle (  1.49);

\path[fill=fillColor] (425.18,431.86) circle (  1.49);

\path[fill=fillColor] (418.35,426.23) circle (  1.49);

\path[fill=fillColor] (422.24,429.51) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (489.72,547.27) {Coverage: 95\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (552.76,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (417.91,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.18,425.83) -- (437.18,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.44,425.83) -- (456.44,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.70,425.83) -- (475.70,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.97,425.83) -- (494.97,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (514.23,425.83) -- (514.23,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (533.49,425.83) -- (533.49,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.76,425.83) -- (552.76,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (417.91,411.58) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (437.18,411.58) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (456.44,411.58) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (475.70,411.58) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (494.97,411.58) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (514.23,411.58) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (533.49,411.58) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (552.76,411.58) {70};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (417.91,538.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (413.95,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,441.91) -- (413.95,441.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,457.98) -- (413.95,457.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,474.06) -- (413.95,474.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,490.14) -- (413.95,490.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,506.21) -- (413.95,506.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,522.29) -- (413.95,522.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,538.37) -- (413.95,538.37);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,423.56) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,439.64) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,455.71) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,471.79) {30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,487.87) {40};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,503.94) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,520.02) {60};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,536.09) {70};
\end{scope}
\begin{scope}
\path[clip] (  0.00,192.72) rectangle (192.72,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,366.47) {\bfseries Nucleotide frequency $f$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47,233.11) rectangle (176.09,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47,233.11) --
	(176.09,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.52,253.76) --
	( 59.52,256.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 80.98,271.74) --
	( 80.98,275.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.46,265.67) --
	( 73.46,268.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (106.47,294.82) --
	(106.47,298.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.59,282.91) --
	( 93.59,285.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.32,275.30) --
	( 85.32,279.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.14,298.38) --
	(114.14,303.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.03,296.78) --
	(114.03,302.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.32,286.59) --
	( 99.32,293.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.70,275.05) --
	( 85.70,279.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (119.28,301.72) --
	(119.28,307.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.17,265.49) --
	( 74.17,269.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.64,276.28) --
	( 86.64,281.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.12,268.61) --
	( 76.12,270.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.26,272.24) --
	( 81.26,275.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.22,305.31) --
	(121.22,310.96);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.79,256.93) --
	( 60.79,259.35);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.73,277.88) --
	( 88.73,281.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (106.97,290.23) --
	(106.97,296.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.24,290.35) --
	(105.24,295.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.44,263.99) --
	( 72.44,267.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.86,281.52) --
	( 92.86,285.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.03,297.50) --
	(112.03,301.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.55,273.90) --
	( 84.55,276.87);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.64,279.31) --
	( 92.64,283.22);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.89,284.75) --
	( 96.89,288.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (108.82,292.09) --
	(108.82,297.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.57,261.09) --
	( 68.57,264.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.03,286.22) --
	( 98.03,289.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.02,290.76) --
	(104.02,296.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.85,281.54) --
	( 90.85,285.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.22,310.98) --
	(128.22,315.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.10,282.10) --
	( 95.10,288.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.15,256.58) --
	( 61.15,259.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.84,287.78) --
	(100.84,292.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.64,288.40) --
	(101.64,293.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.37,273.00) --
	( 81.37,277.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (108.47,292.87) --
	(108.47,298.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.02,267.71) --
	( 76.02,270.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.04,283.85) --
	( 95.04,287.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 82.97,275.05) --
	( 82.97,277.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.04,278.23) --
	( 89.04,284.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.13,288.33) --
	(100.13,292.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.15,277.50) --
	( 86.15,281.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.23,260.98) --
	( 66.23,263.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.83,275.12) --
	( 84.83,280.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.01,296.71) --
	(110.01,302.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.64,265.65) --
	( 72.64,269.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.63,290.31) --
	(103.63,294.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.17,264.75) --
	( 75.17,269.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.42,304.96) --
	(121.42,310.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.69,284.41) --
	( 95.69,288.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.56,266.10) --
	( 73.56,269.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.69,296.27) --
	(114.69,302.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.50,280.46) --
	( 92.50,286.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.37,288.76) --
	(101.37,293.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.95,276.20) --
	( 86.95,279.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.83,268.60) --
	( 77.83,272.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.72,275.11) --
	( 86.72,280.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (113.57,299.01) --
	(113.57,302.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.07,258.35) --
	( 66.07,262.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.56,284.63) --
	( 95.56,287.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.40,291.16) --
	(103.40,295.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.68,269.78) --
	( 78.68,274.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.18,284.30) --
	( 95.18,287.08);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.11,277.47) --
	( 85.11,282.66);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (107.03,293.66) --
	(107.03,298.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.85,263.16) --
	( 70.85,265.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.58,267.87) --
	( 74.58,270.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.56,282.90) --
	( 96.56,287.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.98,278.44) --
	( 89.98,282.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.98,291.20) --
	(104.98,294.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.19,267.27) --
	( 75.19,271.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.27,293.47) --
	(110.27,298.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.48,260.04) --
	( 65.48,262.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.91,264.08) --
	( 72.91,269.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.11,278.89) --
	( 90.11,283.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 94.00,280.99) --
	( 94.00,285.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.13,289.20) --
	(105.13,295.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.06,272.87) --
	( 84.06,276.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.09,283.64) --
	( 96.09,288.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.77,278.96) --
	( 89.77,284.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.41,300.31) --
	(114.41,304.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.74,292.13) --
	(103.74,297.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.57,295.70) --
	(112.57,301.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.47,271.45) --
	( 79.47,273.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.02,278.38) --
	( 88.02,282.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.94,264.32) --
	( 74.94,269.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.72,268.10) --
	( 76.72,270.87);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.57,274.20) --
	( 81.57,279.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 71.87,266.76) --
	( 71.87,270.48);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.18,296.67) --
	(110.18,301.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.41,264.83) --
	( 73.41,269.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.28,273.66) --
	( 85.28,279.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.24,274.87) --
	( 86.24,279.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.02,273.03) --
	( 83.02,276.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.89,264.97) --
	( 72.89,267.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.64,280.22) --
	( 93.64,286.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.83,289.66) --
	(103.83,295.48);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 64.27,260.24) --
	( 64.27,262.42);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 59.52,254.96) circle (  1.49);

\path[fill=fillColor] ( 80.98,273.78) circle (  1.49);

\path[fill=fillColor] ( 73.46,266.95) circle (  1.49);

\path[fill=fillColor] (106.47,296.73) circle (  1.49);

\path[fill=fillColor] ( 93.59,284.10) circle (  1.49);

\path[fill=fillColor] ( 85.32,277.27) circle (  1.49);

\path[fill=fillColor] (114.14,300.94) circle (  1.49);

\path[fill=fillColor] (114.03,299.46) circle (  1.49);

\path[fill=fillColor] ( 99.32,289.90) circle (  1.49);

\path[fill=fillColor] ( 85.70,277.26) circle (  1.49);

\path[fill=fillColor] (119.28,304.42) circle (  1.49);

\path[fill=fillColor] ( 74.17,267.36) circle (  1.49);

\path[fill=fillColor] ( 86.64,278.83) circle (  1.49);

\path[fill=fillColor] ( 76.12,269.68) circle (  1.49);

\path[fill=fillColor] ( 81.26,273.96) circle (  1.49);

\path[fill=fillColor] (121.22,308.01) circle (  1.49);

\path[fill=fillColor] ( 60.79,258.18) circle (  1.49);

\path[fill=fillColor] ( 88.73,279.66) circle (  1.49);

\path[fill=fillColor] (106.97,293.56) circle (  1.49);

\path[fill=fillColor] (105.24,292.91) circle (  1.49);

\path[fill=fillColor] ( 72.44,265.72) circle (  1.49);

\path[fill=fillColor] ( 92.86,283.61) circle (  1.49);

\path[fill=fillColor] (112.03,299.44) circle (  1.49);

\path[fill=fillColor] ( 84.55,275.40) circle (  1.49);

\path[fill=fillColor] ( 92.64,281.33) circle (  1.49);

\path[fill=fillColor] ( 96.89,286.84) circle (  1.49);

\path[fill=fillColor] (108.82,295.08) circle (  1.49);

\path[fill=fillColor] ( 68.57,262.60) circle (  1.49);

\path[fill=fillColor] ( 98.03,287.67) circle (  1.49);

\path[fill=fillColor] (104.02,293.62) circle (  1.49);

\path[fill=fillColor] ( 90.85,283.31) circle (  1.49);

\path[fill=fillColor] (128.22,312.91) circle (  1.49);

\path[fill=fillColor] ( 95.10,285.47) circle (  1.49);

\path[fill=fillColor] ( 61.15,257.89) circle (  1.49);

\path[fill=fillColor] (100.84,290.29) circle (  1.49);

\path[fill=fillColor] (101.64,290.96) circle (  1.49);

\path[fill=fillColor] ( 81.37,275.07) circle (  1.49);

\path[fill=fillColor] (108.47,295.65) circle (  1.49);

\path[fill=fillColor] ( 76.02,269.27) circle (  1.49);

\path[fill=fillColor] ( 95.04,285.69) circle (  1.49);

\path[fill=fillColor] ( 82.97,276.49) circle (  1.49);

\path[fill=fillColor] ( 89.04,281.42) circle (  1.49);

\path[fill=fillColor] (100.13,290.62) circle (  1.49);

\path[fill=fillColor] ( 86.15,279.52) circle (  1.49);

\path[fill=fillColor] ( 66.23,262.06) circle (  1.49);

\path[fill=fillColor] ( 84.83,277.89) circle (  1.49);

\path[fill=fillColor] (110.01,299.35) circle (  1.49);

\path[fill=fillColor] ( 72.64,267.35) circle (  1.49);

\path[fill=fillColor] (103.63,292.60) circle (  1.49);

\path[fill=fillColor] ( 75.17,267.32) circle (  1.49);

\path[fill=fillColor] (121.42,307.80) circle (  1.49);

\path[fill=fillColor] ( 95.69,286.43) circle (  1.49);

\path[fill=fillColor] ( 73.56,268.02) circle (  1.49);

\path[fill=fillColor] (114.69,299.49) circle (  1.49);

\path[fill=fillColor] ( 92.50,283.21) circle (  1.49);

\path[fill=fillColor] (101.37,291.28) circle (  1.49);

\path[fill=fillColor] ( 86.95,277.61) circle (  1.49);

\path[fill=fillColor] ( 77.83,270.42) circle (  1.49);

\path[fill=fillColor] ( 86.72,277.66) circle (  1.49);

\path[fill=fillColor] (113.57,300.85) circle (  1.49);

\path[fill=fillColor] ( 66.07,260.59) circle (  1.49);

\path[fill=fillColor] ( 95.56,286.34) circle (  1.49);

\path[fill=fillColor] (103.40,293.50) circle (  1.49);

\path[fill=fillColor] ( 78.68,271.94) circle (  1.49);

\path[fill=fillColor] ( 95.18,285.68) circle (  1.49);

\path[fill=fillColor] ( 85.11,280.00) circle (  1.49);

\path[fill=fillColor] (107.03,296.17) circle (  1.49);

\path[fill=fillColor] ( 70.85,264.47) circle (  1.49);

\path[fill=fillColor] ( 74.58,269.31) circle (  1.49);

\path[fill=fillColor] ( 96.56,285.14) circle (  1.49);

\path[fill=fillColor] ( 89.98,280.50) circle (  1.49);

\path[fill=fillColor] (104.98,293.09) circle (  1.49);

\path[fill=fillColor] ( 75.19,269.40) circle (  1.49);

\path[fill=fillColor] (110.27,295.97) circle (  1.49);

\path[fill=fillColor] ( 65.48,261.04) circle (  1.49);

\path[fill=fillColor] ( 72.91,266.61) circle (  1.49);

\path[fill=fillColor] ( 90.11,280.93) circle (  1.49);

\path[fill=fillColor] ( 94.00,283.29) circle (  1.49);

\path[fill=fillColor] (105.13,292.43) circle (  1.49);

\path[fill=fillColor] ( 84.06,274.97) circle (  1.49);

\path[fill=fillColor] ( 96.09,285.76) circle (  1.49);

\path[fill=fillColor] ( 89.77,281.62) circle (  1.49);

\path[fill=fillColor] (114.41,302.30) circle (  1.49);

\path[fill=fillColor] (103.74,294.89) circle (  1.49);

\path[fill=fillColor] (112.57,298.41) circle (  1.49);

\path[fill=fillColor] ( 79.47,272.68) circle (  1.49);

\path[fill=fillColor] ( 88.02,280.35) circle (  1.49);

\path[fill=fillColor] ( 74.94,266.71) circle (  1.49);

\path[fill=fillColor] ( 76.72,269.54) circle (  1.49);

\path[fill=fillColor] ( 81.57,276.93) circle (  1.49);

\path[fill=fillColor] ( 71.87,268.65) circle (  1.49);

\path[fill=fillColor] (110.18,298.83) circle (  1.49);

\path[fill=fillColor] ( 73.41,267.06) circle (  1.49);

\path[fill=fillColor] ( 85.28,276.57) circle (  1.49);

\path[fill=fillColor] ( 86.24,276.98) circle (  1.49);

\path[fill=fillColor] ( 83.02,274.73) circle (  1.49);

\path[fill=fillColor] ( 72.89,266.19) circle (  1.49);

\path[fill=fillColor] ( 93.64,283.39) circle (  1.49);

\path[fill=fillColor] (103.83,292.68) circle (  1.49);

\path[fill=fillColor] ( 64.27,261.28) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,354.55) {Coverage: 94\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- (174.42,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 32.47,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.13,233.11) -- ( 56.13,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.79,233.11) -- ( 79.79,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.45,233.11) -- (103.45,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.10,233.11) -- (127.10,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (150.76,233.11) -- (150.76,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (174.42,233.11) -- (174.42,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 56.13,218.86) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 79.79,218.86) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (103.45,218.86) {0.3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (127.10,218.86) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (150.76,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (174.42,218.86) {0.6};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 32.47,351.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 28.51,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,252.86) -- ( 28.51,252.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,272.60) -- ( 28.51,272.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,292.34) -- ( 28.51,292.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,312.09) -- ( 28.51,312.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,331.83) -- ( 28.51,331.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,351.58) -- ( 28.51,351.58);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,250.58) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,270.33) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,290.07) {0.3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,309.82) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,329.56) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,349.30) {0.6};
\end{scope}
\begin{scope}
\path[clip] (192.72,192.72) rectangle (385.44,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (297.00,366.47) {\bfseries Clock standard deviation $\sigma$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (297.00,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (199.85,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (225.19,233.11) rectangle (368.81,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (225.19,233.11) --
	(368.81,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.20,233.11) --
	(225.20,235.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.97,249.69) --
	(255.97,264.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.49,233.11) --
	(225.49,235.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.92,235.92) --
	(230.92,241.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.36,233.11) --
	(227.36,238.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.05,233.11) --
	(227.05,235.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.68,240.30) --
	(241.68,247.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.79,237.31) --
	(232.79,243.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.76,233.11) --
	(233.76,243.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.43,239.68) --
	(238.43,246.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.51,233.11) --
	(225.51,237.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.28,233.11) --
	(226.28,235.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.23,236.27) --
	(230.23,244.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.36,242.13) --
	(240.36,250.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.44,233.11) --
	(225.44,234.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.07,233.11) --
	(228.07,237.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.35,248.54) --
	(261.35,263.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.24,240.64) --
	(238.24,247.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.48,233.11) --
	(228.48,239.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.46,237.06) --
	(232.46,243.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.38,237.51) --
	(232.38,242.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.77,240.05) --
	(240.77,247.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.24,233.11) --
	(228.24,236.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.92,240.29) --
	(238.92,247.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.37,233.11) --
	(227.37,237.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.43,235.84) --
	(232.43,240.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.87,244.94) --
	(241.87,256.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.04,233.11) --
	(226.04,237.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.22,233.11) --
	(225.22,234.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.42,233.11) --
	(230.42,241.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.26,240.39) --
	(239.26,246.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.68,233.11) --
	(229.68,237.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.93,233.11) --
	(230.93,242.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.41,233.11) --
	(227.41,237.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.67,233.11) --
	(225.67,236.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.72,233.11) --
	(226.72,238.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (269.99,264.24) --
	(269.99,281.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.21,233.11) --
	(228.21,240.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.33,233.11) --
	(226.33,235.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.49,233.11) --
	(226.49,236.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.33,233.11) --
	(225.33,237.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.90,233.11) --
	(225.90,244.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.62,233.11) --
	(226.62,237.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) --
	(225.19,235.06);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.10,234.04) --
	(226.10,239.52);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.74,233.11) --
	(234.74,242.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.62,233.11) --
	(226.62,235.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.14,233.11) --
	(226.14,235.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.54,233.11) --
	(225.54,239.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.54,244.57) --
	(246.54,254.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.84,233.11) --
	(225.84,237.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.12,233.11) --
	(227.12,235.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.15,243.78) --
	(242.15,251.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.68,240.21) --
	(241.68,249.78);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.26,233.11) --
	(231.26,237.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.82,233.11) --
	(230.82,237.77);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.58,233.11) --
	(229.58,238.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.35,243.90) --
	(246.35,252.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.43,233.11) --
	(225.43,237.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.47,233.11) --
	(226.47,238.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.55,233.11) --
	(229.55,239.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.27,237.36) --
	(232.27,241.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.16,233.11) --
	(228.16,236.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.25,235.52) --
	(233.25,241.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.14,233.11) --
	(226.14,235.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.31,262.88) --
	(279.31,283.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.32,233.11) --
	(226.32,235.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.25,233.11) --
	(227.25,235.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.67,233.11) --
	(225.67,235.49);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.55,233.11) --
	(229.55,236.48);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.73,236.00) --
	(230.73,240.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.64,234.18) --
	(230.64,238.84);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (320.94,273.05) --
	(320.94,305.28);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.95,233.11) --
	(226.95,236.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.77,236.79) --
	(232.77,240.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.27,247.98) --
	(250.27,263.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.09,268.05) --
	(280.09,286.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.09,236.27) --
	(233.09,242.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.20,240.59) --
	(240.20,251.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.20,250.94) --
	(248.20,265.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.16,233.11) --
	(227.16,235.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.87,233.11) --
	(226.87,235.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) --
	(225.19,235.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.60,246.49) --
	(250.60,257.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.30,233.11) --
	(226.30,239.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.41,239.03) --
	(234.41,243.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.08,241.50) --
	(236.08,248.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.66,254.20) --
	(255.66,271.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.27,233.11) --
	(225.27,235.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.90,274.73) --
	(286.90,304.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.36,243.88) --
	(243.36,252.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.80,239.51) --
	(235.80,245.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.41,233.11) --
	(226.41,237.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.44,233.11) --
	(226.44,240.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.65,233.11) --
	(227.65,235.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.20,233.11) --
	(225.20,236.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.96,235.43) --
	(229.96,239.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.58,233.11) --
	(226.58,238.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.84,233.11) --
	(226.84,236.46);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.92,269.00) --
	(263.92,292.59);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (225.20,233.90) circle (  1.49);

\path[fill=fillColor] (255.97,256.77) circle (  1.49);

\path[fill=fillColor] (225.49,233.84) circle (  1.49);

\path[fill=fillColor] (230.92,238.53) circle (  1.49);

\path[fill=fillColor] (227.36,236.11) circle (  1.49);

\path[fill=fillColor] (227.05,234.01) circle (  1.49);

\path[fill=fillColor] (241.68,243.48) circle (  1.49);

\path[fill=fillColor] (232.79,240.07) circle (  1.49);

\path[fill=fillColor] (233.76,237.44) circle (  1.49);

\path[fill=fillColor] (238.43,242.92) circle (  1.49);

\path[fill=fillColor] (225.51,234.64) circle (  1.49);

\path[fill=fillColor] (226.28,233.75) circle (  1.49);

\path[fill=fillColor] (230.23,240.36) circle (  1.49);

\path[fill=fillColor] (240.36,246.15) circle (  1.49);

\path[fill=fillColor] (225.44,233.67) circle (  1.49);

\path[fill=fillColor] (228.07,234.66) circle (  1.49);

\path[fill=fillColor] (261.35,255.47) circle (  1.49);

\path[fill=fillColor] (238.24,243.83) circle (  1.49);

\path[fill=fillColor] (228.48,235.21) circle (  1.49);

\path[fill=fillColor] (232.46,240.22) circle (  1.49);

\path[fill=fillColor] (232.38,240.05) circle (  1.49);

\path[fill=fillColor] (240.77,243.35) circle (  1.49);

\path[fill=fillColor] (228.24,234.52) circle (  1.49);

\path[fill=fillColor] (238.92,243.53) circle (  1.49);

\path[fill=fillColor] (227.37,235.26) circle (  1.49);

\path[fill=fillColor] (232.43,238.09) circle (  1.49);

\path[fill=fillColor] (241.87,250.24) circle (  1.49);

\path[fill=fillColor] (226.04,234.96) circle (  1.49);

\path[fill=fillColor] (225.22,233.60) circle (  1.49);

\path[fill=fillColor] (230.42,237.99) circle (  1.49);

\path[fill=fillColor] (239.26,243.34) circle (  1.49);

\path[fill=fillColor] (229.68,235.91) circle (  1.49);

\path[fill=fillColor] (230.93,236.66) circle (  1.49);

\path[fill=fillColor] (227.41,235.08) circle (  1.49);

\path[fill=fillColor] (225.67,234.24) circle (  1.49);

\path[fill=fillColor] (226.72,235.40) circle (  1.49);

\path[fill=fillColor] (269.99,272.75) circle (  1.49);

\path[fill=fillColor] (228.21,237.55) circle (  1.49);

\path[fill=fillColor] (226.33,233.80) circle (  1.49);

\path[fill=fillColor] (226.49,234.55) circle (  1.49);

\path[fill=fillColor] (225.33,235.18) circle (  1.49);

\path[fill=fillColor] (225.90,237.00) circle (  1.49);

\path[fill=fillColor] (226.62,234.43) circle (  1.49);

\path[fill=fillColor] (225.19,233.72) circle (  1.49);

\path[fill=fillColor] (226.10,236.79) circle (  1.49);

\path[fill=fillColor] (234.74,238.21) circle (  1.49);

\path[fill=fillColor] (226.62,234.01) circle (  1.49);

\path[fill=fillColor] (226.14,233.74) circle (  1.49);

\path[fill=fillColor] (225.54,236.55) circle (  1.49);

\path[fill=fillColor] (246.54,249.40) circle (  1.49);

\path[fill=fillColor] (225.84,234.89) circle (  1.49);

\path[fill=fillColor] (227.12,233.78) circle (  1.49);

\path[fill=fillColor] (242.15,247.57) circle (  1.49);

\path[fill=fillColor] (241.68,244.71) circle (  1.49);

\path[fill=fillColor] (231.26,234.36) circle (  1.49);

\path[fill=fillColor] (230.82,235.22) circle (  1.49);

\path[fill=fillColor] (229.58,235.75) circle (  1.49);

\path[fill=fillColor] (246.35,247.96) circle (  1.49);

\path[fill=fillColor] (225.43,234.59) circle (  1.49);

\path[fill=fillColor] (226.47,236.00) circle (  1.49);

\path[fill=fillColor] (229.55,235.34) circle (  1.49);

\path[fill=fillColor] (232.27,239.45) circle (  1.49);

\path[fill=fillColor] (228.16,234.23) circle (  1.49);

\path[fill=fillColor] (233.25,238.42) circle (  1.49);

\path[fill=fillColor] (226.14,234.08) circle (  1.49);

\path[fill=fillColor] (279.31,272.70) circle (  1.49);

\path[fill=fillColor] (226.32,233.87) circle (  1.49);

\path[fill=fillColor] (227.25,234.28) circle (  1.49);

\path[fill=fillColor] (225.67,233.89) circle (  1.49);

\path[fill=fillColor] (229.55,234.26) circle (  1.49);

\path[fill=fillColor] (230.73,238.29) circle (  1.49);

\path[fill=fillColor] (230.64,236.42) circle (  1.49);

\path[fill=fillColor] (320.94,288.25) circle (  1.49);

\path[fill=fillColor] (226.95,234.18) circle (  1.49);

\path[fill=fillColor] (232.77,238.61) circle (  1.49);

\path[fill=fillColor] (250.27,255.51) circle (  1.49);

\path[fill=fillColor] (280.09,277.06) circle (  1.49);

\path[fill=fillColor] (233.09,239.11) circle (  1.49);

\path[fill=fillColor] (240.20,245.98) circle (  1.49);

\path[fill=fillColor] (248.20,257.65) circle (  1.49);

\path[fill=fillColor] (227.16,233.77) circle (  1.49);

\path[fill=fillColor] (226.87,233.80) circle (  1.49);

\path[fill=fillColor] (225.19,233.80) circle (  1.49);

\path[fill=fillColor] (250.60,251.75) circle (  1.49);

\path[fill=fillColor] (226.30,236.94) circle (  1.49);

\path[fill=fillColor] (234.41,241.03) circle (  1.49);

\path[fill=fillColor] (236.08,245.01) circle (  1.49);

\path[fill=fillColor] (255.66,262.18) circle (  1.49);

\path[fill=fillColor] (225.27,233.86) circle (  1.49);

\path[fill=fillColor] (286.90,289.39) circle (  1.49);

\path[fill=fillColor] (243.36,247.76) circle (  1.49);

\path[fill=fillColor] (235.80,242.53) circle (  1.49);

\path[fill=fillColor] (226.41,234.78) circle (  1.49);

\path[fill=fillColor] (226.44,235.87) circle (  1.49);

\path[fill=fillColor] (227.65,234.05) circle (  1.49);

\path[fill=fillColor] (225.20,234.26) circle (  1.49);

\path[fill=fillColor] (229.96,237.38) circle (  1.49);

\path[fill=fillColor] (226.58,234.93) circle (  1.49);

\path[fill=fillColor] (226.84,234.15) circle (  1.49);

\path[fill=fillColor] (263.92,280.15) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (297.00,354.55) {Coverage: 94\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (356.16,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (225.19,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.02,233.11) -- (247.02,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.85,233.11) -- (268.85,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (290.68,233.11) -- (290.68,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.51,233.11) -- (312.51,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.33,233.11) -- (334.33,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.16,233.11) -- (356.16,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (225.19,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (247.02,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (268.85,218.86) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (290.68,218.86) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (312.51,218.86) {2.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (334.33,218.86) {2.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (356.16,218.86) {3.0};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (225.19,342.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (221.23,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,251.33) -- (221.23,251.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,269.55) -- (221.23,269.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,287.76) -- (221.23,287.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,305.98) -- (221.23,305.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,324.20) -- (221.23,324.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,342.41) -- (221.23,342.41);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,249.06) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,267.27) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,285.49) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,303.71) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,321.92) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,340.14) {3.0};
\end{scope}
\begin{scope}
\path[clip] (385.44,192.72) rectangle (578.16,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (489.72,366.47) {\bfseries Branch rate mean};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (489.72,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (392.57,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (417.91,233.11) rectangle (561.53,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (417.91,233.11) --
	(561.53,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.90,286.57) --
	(482.90,288.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.90,270.72) --
	(479.90,293.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.01,286.75) --
	(483.01,287.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.67,283.37) --
	(482.67,289.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.26,285.72) --
	(483.26,289.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.10,286.57) --
	(483.10,288.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.42,281.19) --
	(479.42,292.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (486.59,283.60) --
	(486.59,290.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (484.74,283.70) --
	(484.74,289.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.32,283.47) --
	(482.32,293.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.78,286.07) --
	(482.78,288.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.06,286.75) --
	(483.06,287.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.90,282.80) --
	(482.90,290.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.99,279.76) --
	(481.99,293.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.85,286.83) --
	(482.85,287.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.17,285.91) --
	(483.17,288.51);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (466.60,276.38) --
	(466.60,299.58);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.23,282.36) --
	(485.23,293.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.03,285.65) --
	(483.03,288.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.66,282.97) --
	(479.66,290.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.80,283.49) --
	(482.80,290.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.18,282.88) --
	(485.18,294.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.38,285.95) --
	(482.38,288.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.26,281.18) --
	(483.26,292.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.27,285.42) --
	(482.27,288.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (484.20,285.06) --
	(484.20,290.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.01,273.96) --
	(479.01,290.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.01,286.13) --
	(483.01,288.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.90,286.90) --
	(482.90,287.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.76,284.49) --
	(482.76,289.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.11,282.31) --
	(483.11,292.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.24,285.70) --
	(481.24,289.51);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.74,283.81) --
	(485.74,289.42);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.08,285.66) --
	(482.08,288.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.02,286.43) --
	(483.02,288.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.62,285.99) --
	(482.62,288.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.60,262.81) --
	(470.60,290.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.32,284.01) --
	(483.32,289.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.98,286.89) --
	(482.98,288.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.04,286.46) --
	(483.04,288.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.97,285.86) --
	(482.97,288.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.86,284.14) --
	(482.86,290.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.36,286.00) --
	(482.36,288.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.90,286.84) --
	(482.90,287.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.30,285.55) --
	(482.30,289.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.48,284.00) --
	(479.48,290.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.16,286.58) --
	(482.16,288.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.94,286.80) --
	(482.94,287.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.78,285.24) --
	(482.78,289.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.27,278.35) --
	(477.27,293.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.18,285.84) --
	(483.18,288.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.24,286.79) --
	(482.24,287.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (484.28,280.51) --
	(484.28,294.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.90,281.01) --
	(476.90,292.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.16,286.00) --
	(482.16,288.24);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.87,285.88) --
	(480.87,288.79);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.15,285.73) --
	(481.15,289.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.93,278.54) --
	(476.93,292.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.95,286.07) --
	(482.95,288.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.93,285.79) --
	(482.93,289.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.45,285.32) --
	(482.45,288.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.59,284.31) --
	(481.59,290.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.56,286.46) --
	(483.56,288.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.55,283.92) --
	(481.55,289.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.69,286.48) --
	(482.69,287.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.35,266.86) --
	(478.35,308.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.82,286.74) --
	(482.82,288.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.95,286.32) --
	(482.95,288.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.80,286.65) --
	(482.80,287.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.62,286.40) --
	(482.62,288.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.28,284.36) --
	(483.28,289.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.36,285.37) --
	(482.36,288.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.27,253.57) --
	(464.27,308.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.17,286.41) --
	(483.17,288.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.66,284.46) --
	(482.66,289.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (486.20,273.15) --
	(486.20,296.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (489.89,268.29) --
	(489.89,311.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.23,284.01) --
	(482.23,289.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.22,279.98) --
	(480.22,294.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.07,274.18) --
	(482.07,299.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.47,286.67) --
	(482.47,287.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.26,286.75) --
	(483.26,287.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.90,286.76) --
	(482.90,287.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.38,276.94) --
	(479.38,295.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.46,285.08) --
	(483.46,289.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.63,283.61) --
	(483.63,291.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.80,280.43) --
	(482.80,291.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.75,269.66) --
	(480.75,300.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.89,286.77) --
	(482.89,287.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (486.51,250.72) --
	(486.51,303.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.43,280.71) --
	(479.43,295.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.58,283.04) --
	(483.58,292.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.99,285.82) --
	(482.99,288.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.91,285.16) --
	(482.91,289.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.33,286.64) --
	(483.33,288.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.90,286.23) --
	(482.90,288.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.03,284.68) --
	(482.03,289.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.80,285.73) --
	(482.80,288.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.23,286.49) --
	(483.23,288.26);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (513.66,258.51) --
	(513.66,299.82);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (482.90,287.34) circle (  1.49);

\path[fill=fillColor] (479.90,282.04) circle (  1.49);

\path[fill=fillColor] (483.01,287.33) circle (  1.49);

\path[fill=fillColor] (482.67,286.46) circle (  1.49);

\path[fill=fillColor] (483.26,287.52) circle (  1.49);

\path[fill=fillColor] (483.10,287.34) circle (  1.49);

\path[fill=fillColor] (479.42,286.84) circle (  1.49);

\path[fill=fillColor] (486.59,287.13) circle (  1.49);

\path[fill=fillColor] (484.74,287.01) circle (  1.49);

\path[fill=fillColor] (482.32,288.55) circle (  1.49);

\path[fill=fillColor] (482.78,287.37) circle (  1.49);

\path[fill=fillColor] (483.06,287.36) circle (  1.49);

\path[fill=fillColor] (482.90,286.66) circle (  1.49);

\path[fill=fillColor] (481.99,286.67) circle (  1.49);

\path[fill=fillColor] (482.85,287.34) circle (  1.49);

\path[fill=fillColor] (483.17,287.28) circle (  1.49);

\path[fill=fillColor] (466.60,288.00) circle (  1.49);

\path[fill=fillColor] (485.23,287.81) circle (  1.49);

\path[fill=fillColor] (483.03,287.22) circle (  1.49);

\path[fill=fillColor] (479.66,286.59) circle (  1.49);

\path[fill=fillColor] (482.80,287.20) circle (  1.49);

\path[fill=fillColor] (485.18,288.78) circle (  1.49);

\path[fill=fillColor] (482.38,287.20) circle (  1.49);

\path[fill=fillColor] (483.26,286.93) circle (  1.49);

\path[fill=fillColor] (482.27,287.11) circle (  1.49);

\path[fill=fillColor] (484.20,287.60) circle (  1.49);

\path[fill=fillColor] (479.01,282.47) circle (  1.49);

\path[fill=fillColor] (483.01,287.41) circle (  1.49);

\path[fill=fillColor] (482.90,287.37) circle (  1.49);

\path[fill=fillColor] (482.76,287.14) circle (  1.49);

\path[fill=fillColor] (483.11,287.28) circle (  1.49);

\path[fill=fillColor] (481.24,287.56) circle (  1.49);

\path[fill=fillColor] (485.74,287.00) circle (  1.49);

\path[fill=fillColor] (482.08,287.28) circle (  1.49);

\path[fill=fillColor] (483.02,287.37) circle (  1.49);

\path[fill=fillColor] (482.62,287.38) circle (  1.49);

\path[fill=fillColor] (470.60,276.47) circle (  1.49);

\path[fill=fillColor] (483.32,286.83) circle (  1.49);

\path[fill=fillColor] (482.98,287.41) circle (  1.49);

\path[fill=fillColor] (483.04,287.40) circle (  1.49);

\path[fill=fillColor] (482.97,287.24) circle (  1.49);

\path[fill=fillColor] (482.86,287.20) circle (  1.49);

\path[fill=fillColor] (482.36,287.37) circle (  1.49);

\path[fill=fillColor] (482.90,287.35) circle (  1.49);

\path[fill=fillColor] (482.30,287.48) circle (  1.49);

\path[fill=fillColor] (479.48,287.21) circle (  1.49);

\path[fill=fillColor] (482.16,287.37) circle (  1.49);

\path[fill=fillColor] (482.94,287.35) circle (  1.49);

\path[fill=fillColor] (482.78,287.43) circle (  1.49);

\path[fill=fillColor] (477.27,285.98) circle (  1.49);

\path[fill=fillColor] (483.18,287.29) circle (  1.49);

\path[fill=fillColor] (482.24,287.36) circle (  1.49);

\path[fill=fillColor] (484.28,287.39) circle (  1.49);

\path[fill=fillColor] (476.90,286.55) circle (  1.49);

\path[fill=fillColor] (482.16,287.22) circle (  1.49);

\path[fill=fillColor] (480.87,287.33) circle (  1.49);

\path[fill=fillColor] (481.15,287.36) circle (  1.49);

\path[fill=fillColor] (476.93,285.56) circle (  1.49);

\path[fill=fillColor] (482.95,287.32) circle (  1.49);

\path[fill=fillColor] (482.93,287.52) circle (  1.49);

\path[fill=fillColor] (482.45,287.22) circle (  1.49);

\path[fill=fillColor] (481.59,287.19) circle (  1.49);

\path[fill=fillColor] (483.56,287.36) circle (  1.49);

\path[fill=fillColor] (481.55,286.74) circle (  1.49);

\path[fill=fillColor] (482.69,287.30) circle (  1.49);

\path[fill=fillColor] (478.35,287.73) circle (  1.49);

\path[fill=fillColor] (482.82,287.37) circle (  1.49);

\path[fill=fillColor] (482.95,287.29) circle (  1.49);

\path[fill=fillColor] (482.80,287.32) circle (  1.49);

\path[fill=fillColor] (482.62,287.35) circle (  1.49);

\path[fill=fillColor] (483.28,287.04) circle (  1.49);

\path[fill=fillColor] (482.36,287.16) circle (  1.49);

\path[fill=fillColor] (464.27,279.64) circle (  1.49);

\path[fill=fillColor] (483.17,287.31) circle (  1.49);

\path[fill=fillColor] (482.66,286.97) circle (  1.49);

\path[fill=fillColor] (486.20,284.80) circle (  1.49);

\path[fill=fillColor] (489.89,288.30) circle (  1.49);

\path[fill=fillColor] (482.23,287.12) circle (  1.49);

\path[fill=fillColor] (480.22,287.01) circle (  1.49);

\path[fill=fillColor] (482.07,287.18) circle (  1.49);

\path[fill=fillColor] (482.47,287.31) circle (  1.49);

\path[fill=fillColor] (483.26,287.36) circle (  1.49);

\path[fill=fillColor] (482.90,287.36) circle (  1.49);

\path[fill=fillColor] (479.38,286.13) circle (  1.49);

\path[fill=fillColor] (483.46,287.32) circle (  1.49);

\path[fill=fillColor] (483.63,287.35) circle (  1.49);

\path[fill=fillColor] (482.80,286.15) circle (  1.49);

\path[fill=fillColor] (480.75,285.06) circle (  1.49);

\path[fill=fillColor] (482.89,287.34) circle (  1.49);

\path[fill=fillColor] (486.51,278.64) circle (  1.49);

\path[fill=fillColor] (479.43,287.88) circle (  1.49);

\path[fill=fillColor] (483.58,288.05) circle (  1.49);

\path[fill=fillColor] (482.99,287.30) circle (  1.49);

\path[fill=fillColor] (482.91,287.29) circle (  1.49);

\path[fill=fillColor] (483.33,287.37) circle (  1.49);

\path[fill=fillColor] (482.90,287.37) circle (  1.49);

\path[fill=fillColor] (482.03,286.93) circle (  1.49);

\path[fill=fillColor] (482.80,287.26) circle (  1.49);

\path[fill=fillColor] (483.23,287.36) circle (  1.49);

\path[fill=fillColor] (513.66,277.98) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (489.72,354.55) {Coverage: 96\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (547.89,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (417.91,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (450.41,233.11) -- (450.41,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.90,233.11) -- (482.90,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.40,233.11) -- (515.40,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.89,233.11) -- (547.89,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (417.91,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (450.41,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (482.90,218.86) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (515.40,218.86) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (547.89,218.86) {2.0};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (417.91,341.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (413.95,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,260.23) -- (413.95,260.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,287.35) -- (413.95,287.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,314.47) -- (413.95,314.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,341.59) -- (413.95,341.59);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,257.96) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,285.08) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,312.20) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,339.32) {2.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (192.72,192.72);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,173.75) {\bfseries Branch rate variance};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28, 10.30) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,100.32) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47, 40.39) rectangle (176.09,160.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47, 40.39) --
	(176.09,160.25);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.92, 45.33) --
	( 49.92, 61.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.99, 40.53) --
	( 32.99, 41.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.55, 40.39) --
	( 32.55, 40.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.53, 40.39) --
	( 32.53, 40.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.02, 41.41) --
	( 36.02, 44.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.76, 40.74) --
	( 33.76, 42.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.80, 40.39) --
	( 33.80, 42.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.06, 41.26) --
	( 35.06, 43.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.91, 40.51) --
	( 32.91, 42.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.74, 42.15) --
	( 36.74, 46.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.59, 40.39) --
	( 32.59, 40.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.13, 44.60) --
	( 51.13, 60.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.68, 41.55) --
	( 34.68, 44.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.64, 40.39) --
	( 32.64, 41.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.34, 40.64) --
	( 33.34, 42.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.54, 40.77) --
	( 33.54, 42.12);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.84, 41.37) --
	( 36.84, 43.88);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.60, 40.39) --
	( 32.60, 40.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.73, 41.49) --
	( 35.73, 44.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.54, 40.39) --
	( 32.54, 40.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.35, 40.51) --
	( 33.35, 41.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.29, 43.14) --
	( 37.29, 50.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.00, 40.39) --
	( 33.00, 41.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.94, 41.44) --
	( 35.94, 43.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.92, 40.39) --
	( 32.92, 40.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.92, 40.39) --
	( 32.92, 42.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.55, 40.39) --
	( 32.55, 40.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.51, 40.39) --
	( 32.51, 40.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.88, 48.38) --
	( 74.88, 88.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.63, 40.39) --
	( 32.63, 41.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 42.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.51, 40.39) --
	( 32.51, 40.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 41.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.49, 40.39) --
	( 33.49, 42.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 41.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.42, 42.90) --
	( 40.42, 49.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.54, 40.39) --
	( 32.54, 40.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.69, 42.67) --
	( 35.69, 47.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.07, 41.26) --
	( 36.07, 45.52);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.01, 40.39) --
	( 33.01, 40.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.07, 40.39) --
	( 33.07, 40.80);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.74, 40.39) --
	( 32.74, 40.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.81, 42.69) --
	( 37.81, 47.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.82, 40.39) --
	( 32.82, 41.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.46, 40.75) --
	( 33.46, 41.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.62, 40.39) --
	( 32.62, 40.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.79, 40.39) --
	( 33.79, 41.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.58, 51.36) --
	( 96.58,117.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.54, 40.39) --
	( 32.54, 40.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.50);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.75, 40.39) --
	( 32.75, 40.61);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.97, 40.52) --
	( 32.97, 41.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.84, 40.39) --
	( 32.84, 40.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.34, 46.98) --
	( 73.34,160.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.52, 40.39) --
	( 32.52, 40.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.47, 40.65) --
	( 33.47, 41.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.43, 44.33) --
	( 50.43, 59.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.92, 72.17) --
	( 92.92,138.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.39, 40.56) --
	( 33.39, 41.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.02, 41.27) --
	( 36.02, 46.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 47.90, 47.07) --
	( 47.90, 64.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.54, 40.39) --
	( 32.54, 40.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.52, 40.39) --
	( 32.52, 40.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.18, 43.71) --
	( 40.18, 52.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 41.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.93, 41.08) --
	( 33.93, 42.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.39, 41.86) --
	( 35.39, 44.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.37, 46.72) --
	( 46.37, 69.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.22, 44.56) --
	(128.22,133.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.12, 42.73) --
	( 37.12, 47.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.10, 41.26) --
	( 34.10, 43.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.51, 40.39) --
	( 32.51, 40.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 41.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.57, 40.39) --
	( 32.57, 40.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.79, 40.48) --
	( 32.79, 41.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.51, 40.39) --
	( 32.51, 40.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.52, 40.39) --
	( 32.52, 40.61);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.72, 47.66) --
	( 96.72, 90.93);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 32.47, 40.42) circle (  1.49);

\path[fill=fillColor] ( 49.92, 52.19) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.99, 41.00) circle (  1.49);

\path[fill=fillColor] ( 32.55, 40.61) circle (  1.49);

\path[fill=fillColor] ( 32.53, 40.42) circle (  1.49);

\path[fill=fillColor] ( 36.02, 42.64) circle (  1.49);

\path[fill=fillColor] ( 33.76, 41.39) circle (  1.49);

\path[fill=fillColor] ( 33.80, 40.95) circle (  1.49);

\path[fill=fillColor] ( 35.06, 42.33) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.48) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.91, 41.49) circle (  1.49);

\path[fill=fillColor] ( 36.74, 43.90) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.59, 40.48) circle (  1.49);

\path[fill=fillColor] ( 51.13, 51.17) circle (  1.49);

\path[fill=fillColor] ( 34.68, 42.70) circle (  1.49);

\path[fill=fillColor] ( 32.64, 40.55) circle (  1.49);

\path[fill=fillColor] ( 33.34, 41.43) circle (  1.49);

\path[fill=fillColor] ( 33.54, 41.38) circle (  1.49);

\path[fill=fillColor] ( 36.84, 42.49) circle (  1.49);

\path[fill=fillColor] ( 32.60, 40.45) circle (  1.49);

\path[fill=fillColor] ( 35.73, 42.64) circle (  1.49);

\path[fill=fillColor] ( 32.54, 40.52) circle (  1.49);

\path[fill=fillColor] ( 33.35, 40.90) circle (  1.49);

\path[fill=fillColor] ( 37.29, 46.43) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.49) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 33.00, 40.92) circle (  1.49);

\path[fill=fillColor] ( 35.94, 42.42) circle (  1.49);

\path[fill=fillColor] ( 32.92, 40.57) circle (  1.49);

\path[fill=fillColor] ( 32.92, 40.83) circle (  1.49);

\path[fill=fillColor] ( 32.55, 40.50) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.51, 40.54) circle (  1.49);

\path[fill=fillColor] ( 74.88, 64.10) circle (  1.49);

\path[fill=fillColor] ( 32.63, 40.85) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.46) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.50) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.94) circle (  1.49);

\path[fill=fillColor] ( 32.51, 40.46) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.69) circle (  1.49);

\path[fill=fillColor] ( 33.49, 41.02) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.67) circle (  1.49);

\path[fill=fillColor] ( 40.42, 45.98) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.50) circle (  1.49);

\path[fill=fillColor] ( 32.54, 40.41) circle (  1.49);

\path[fill=fillColor] ( 35.69, 44.68) circle (  1.49);

\path[fill=fillColor] ( 36.07, 43.16) circle (  1.49);

\path[fill=fillColor] ( 33.01, 40.46) circle (  1.49);

\path[fill=fillColor] ( 33.07, 40.52) circle (  1.49);

\path[fill=fillColor] ( 32.74, 40.56) circle (  1.49);

\path[fill=fillColor] ( 37.81, 44.88) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.47) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.58) circle (  1.49);

\path[fill=fillColor] ( 32.82, 40.56) circle (  1.49);

\path[fill=fillColor] ( 33.46, 41.18) circle (  1.49);

\path[fill=fillColor] ( 32.62, 40.44) circle (  1.49);

\path[fill=fillColor] ( 33.79, 40.99) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.43) circle (  1.49);

\path[fill=fillColor] ( 96.58, 81.28) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.54, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.75, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.97, 40.94) circle (  1.49);

\path[fill=fillColor] ( 32.84, 40.63) circle (  1.49);

\path[fill=fillColor] ( 73.34, 93.86) circle (  1.49);

\path[fill=fillColor] ( 32.52, 40.44) circle (  1.49);

\path[fill=fillColor] ( 33.47, 40.99) circle (  1.49);

\path[fill=fillColor] ( 50.43, 50.54) circle (  1.49);

\path[fill=fillColor] ( 92.92,102.54) circle (  1.49);

\path[fill=fillColor] ( 33.39, 41.13) circle (  1.49);

\path[fill=fillColor] ( 36.02, 43.77) circle (  1.49);

\path[fill=fillColor] ( 47.90, 54.73) circle (  1.49);

\path[fill=fillColor] ( 32.54, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.52, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.41) circle (  1.49);

\path[fill=fillColor] ( 40.18, 47.53) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.74) circle (  1.49);

\path[fill=fillColor] ( 33.93, 41.64) circle (  1.49);

\path[fill=fillColor] ( 35.39, 43.22) circle (  1.49);

\path[fill=fillColor] ( 46.37, 57.05) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.41) circle (  1.49);

\path[fill=fillColor] (128.22, 81.40) circle (  1.49);

\path[fill=fillColor] ( 37.12, 44.88) circle (  1.49);

\path[fill=fillColor] ( 34.10, 42.20) circle (  1.49);

\path[fill=fillColor] ( 32.51, 40.49) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.65) circle (  1.49);

\path[fill=fillColor] ( 32.57, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.79, 40.75) circle (  1.49);

\path[fill=fillColor] ( 32.51, 40.52) circle (  1.49);

\path[fill=fillColor] ( 32.52, 40.44) circle (  1.49);

\path[fill=fillColor] ( 96.72, 68.23) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,161.83) {Coverage: 95\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- (161.43, 40.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 32.47, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 64.71, 40.39) -- ( 64.71, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.95, 40.39) -- ( 96.95, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.19, 40.39) -- (129.19, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.43, 40.39) -- (161.43, 36.43);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47, 26.14) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 64.71, 26.14) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 96.95, 26.14) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (129.19, 26.14) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (161.43, 26.14) {4};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 32.47,148.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 28.51, 40.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 67.30) -- ( 28.51, 67.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 94.20) -- ( 28.51, 94.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,121.11) -- ( 28.51,121.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,148.01) -- ( 28.51,148.01);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 38.12) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 65.02) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 91.93) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,118.83) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,145.74) {4};
\end{scope}
\end{tikzpicture}

\end{document}