### 2020-09-01 RatesWCSS/WCSS_quant_cached_adaptive.tex

% Created by tikzDevice version 0.12.3.1 on 2020-09-01 10:41:06
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,385.44) rectangle (192.72,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,559.19) {\bfseries Tree height $h$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47,425.83) rectangle (176.09,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47,425.83) --
	(176.09,545.69);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.35,447.61) --
	( 59.35,449.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.26,431.45) --
	( 40.26,439.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.86,447.48) --
	( 59.86,449.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 55.07,442.95) --
	( 55.07,446.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.02,455.23) --
	( 70.02,458.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.00,438.40) --
	( 49.00,439.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.59,432.68) --
	( 43.59,439.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.36,429.42) --
	( 37.36,430.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.77,426.72) --
	( 33.77,427.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.97,432.02) --
	( 41.97,435.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.28,430.45) --
	( 38.28,431.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.65,435.18) --
	( 44.65,436.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.43,428.58) --
	( 36.43,429.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.38,481.82) --
	(110.38,532.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.05,434.69) --
	( 44.05,435.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.01,432.00) --
	( 41.01,433.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.87,439.72) --
	( 56.87,471.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.70,437.73) --
	( 50.70,443.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.13,428.66) --
	( 36.13,429.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.24,429.46) --
	( 37.24,430.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.82,438.18) --
	( 48.82,441.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.20,433.23) --
	( 48.20,440.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 55.23,443.59) --
	( 55.23,447.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.95,448.46) --
	( 68.95,460.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.06,439.06) --
	( 50.06,441.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.40,433.24) --
	( 42.40,434.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.55,430.30) --
	( 38.55,432.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.00,441.09) --
	( 52.00,443.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.63,454.83) --
	( 68.63,458.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.82,431.18) --
	( 39.82,433.04);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.29,437.39) --
	( 45.29,445.03);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.83,452.17) --
	( 66.83,459.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.94,426.91) --
	( 33.94,427.20);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.49,438.28) --
	( 49.49,439.95);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.58,431.61) --
	( 39.58,432.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.87,430.68) --
	( 38.87,431.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.93,432.40) --
	( 43.93,471.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.20,429.62) --
	( 37.20,430.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.17,446.43) --
	( 60.17,449.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.26,440.00) --
	( 50.26,441.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.26,450.64) --
	( 62.26,453.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.49,426.51) --
	( 33.49,426.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.38,439.07) --
	( 48.38,440.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.84,435.47) --
	( 44.84,436.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.91,454.56) --
	( 68.91,457.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.97,429.68) --
	( 37.97,431.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.79,431.70) --
	( 40.79,433.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.59,436.50) --
	( 45.59,437.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.05,435.02) --
	( 44.05,435.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.69,430.50) --
	( 40.69,434.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.86,430.86) --
	( 38.86,431.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.28,435.95) --
	( 45.28,437.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.56,431.52) --
	( 41.56,436.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.80,427.55) --
	( 34.80,428.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.60,428.22) --
	( 35.60,428.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.08,431.67) --
	( 40.08,433.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.66,458.91) --
	( 76.66,463.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.00,433.95) --
	( 45.00,440.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.74,429.00) --
	( 36.74,429.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.58,447.54) --
	( 59.58,449.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.07,429.17) --
	( 37.07,430.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.30,442.17) --
	( 54.30,445.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.05,435.79) --
	( 45.05,437.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.01,432.05) --
	( 42.01,435.15);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.96,464.12) --
	( 84.96,469.15);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.31,428.95) --
	( 40.31,439.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.72,430.74) --
	( 38.72,431.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.14,443.05) --
	( 54.14,445.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.31,438.42) --
	( 48.31,439.50);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.94,433.76) --
	( 41.94,434.70);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.16,432.06) --
	( 40.16,432.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 55.16,442.53) --
	( 55.16,445.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.07,427.02) --
	( 42.07,516.87);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.99,432.93) --
	( 41.99,433.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.97,449.77) --
	( 66.97,454.12);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.13,427.66) --
	( 35.13,429.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.24,434.88) --
	( 44.24,522.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.65,430.98) --
	( 39.65,432.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.47,428.28) --
	( 35.47,429.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.51,432.28) --
	( 42.51,440.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.32,434.47) --
	( 43.32,435.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.35,431.93) --
	( 40.35,432.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.65,441.00) --
	( 50.65,442.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.08,430.82) --
	( 40.08,433.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.72,429.92) --
	( 37.72,430.63);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.77,452.11) --
	( 63.77,460.87);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.95,433.55) --
	( 42.95,436.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.06,428.42) --
	( 36.06,432.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.39,440.44) --
	( 50.39,441.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.01,427.09) --
	( 36.01,439.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.72,431.57) --
	( 41.72,435.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.27,433.81) --
	( 44.27,441.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.16,429.94) --
	( 38.16,430.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.54,427.37) --
	( 34.54,427.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.49,432.39) --
	( 40.49,433.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.85,449.15) --
	( 61.85,452.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.11,451.25) --
	( 65.11,454.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.36,428.09) --
	( 35.36,428.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.65,429.00) --
	( 36.65,429.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.66,444.22) --
	( 58.66,490.90);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 59.35,448.70) circle (  1.49);

\path[fill=fillColor] ( 40.26,435.31) circle (  1.49);

\path[fill=fillColor] ( 59.86,448.50) circle (  1.49);

\path[fill=fillColor] ( 55.07,444.62) circle (  1.49);

\path[fill=fillColor] ( 70.02,456.76) circle (  1.49);

\path[fill=fillColor] ( 49.00,439.10) circle (  1.49);

\path[fill=fillColor] ( 43.59,436.08) circle (  1.49);

\path[fill=fillColor] ( 37.36,429.83) circle (  1.49);

\path[fill=fillColor] ( 33.77,426.97) circle (  1.49);

\path[fill=fillColor] ( 41.97,433.61) circle (  1.49);

\path[fill=fillColor] ( 38.28,430.84) circle (  1.49);

\path[fill=fillColor] ( 44.65,436.04) circle (  1.49);

\path[fill=fillColor] ( 36.43,429.11) circle (  1.49);

\path[fill=fillColor] (110.38,505.74) circle (  1.49);

\path[fill=fillColor] ( 44.05,435.18) circle (  1.49);

\path[fill=fillColor] ( 41.01,432.74) circle (  1.49);

\path[fill=fillColor] ( 56.87,454.12) circle (  1.49);

\path[fill=fillColor] ( 50.70,440.25) circle (  1.49);

\path[fill=fillColor] ( 36.13,428.94) circle (  1.49);

\path[fill=fillColor] ( 37.24,429.80) circle (  1.49);

\path[fill=fillColor] ( 48.82,439.84) circle (  1.49);

\path[fill=fillColor] ( 48.20,436.62) circle (  1.49);

\path[fill=fillColor] ( 55.23,445.34) circle (  1.49);

\path[fill=fillColor] ( 68.95,454.44) circle (  1.49);

\path[fill=fillColor] ( 50.06,440.01) circle (  1.49);

\path[fill=fillColor] ( 42.40,433.92) circle (  1.49);

\path[fill=fillColor] ( 38.55,431.59) circle (  1.49);

\path[fill=fillColor] ( 52.00,442.25) circle (  1.49);

\path[fill=fillColor] ( 68.63,456.38) circle (  1.49);

\path[fill=fillColor] ( 39.82,432.10) circle (  1.49);

\path[fill=fillColor] ( 45.29,441.14) circle (  1.49);

\path[fill=fillColor] ( 66.83,455.56) circle (  1.49);

\path[fill=fillColor] ( 33.94,427.05) circle (  1.49);

\path[fill=fillColor] ( 49.49,439.12) circle (  1.49);

\path[fill=fillColor] ( 39.58,431.96) circle (  1.49);

\path[fill=fillColor] ( 38.87,431.04) circle (  1.49);

\path[fill=fillColor] ( 43.93,447.50) circle (  1.49);

\path[fill=fillColor] ( 37.20,429.95) circle (  1.49);

\path[fill=fillColor] ( 60.17,448.20) circle (  1.49);

\path[fill=fillColor] ( 50.26,440.66) circle (  1.49);

\path[fill=fillColor] ( 62.26,451.81) circle (  1.49);

\path[fill=fillColor] ( 33.49,426.69) circle (  1.49);

\path[fill=fillColor] ( 48.38,439.92) circle (  1.49);

\path[fill=fillColor] ( 44.84,436.00) circle (  1.49);

\path[fill=fillColor] ( 68.91,456.23) circle (  1.49);

\path[fill=fillColor] ( 37.97,430.50) circle (  1.49);

\path[fill=fillColor] ( 40.79,432.45) circle (  1.49);

\path[fill=fillColor] ( 45.59,437.20) circle (  1.49);

\path[fill=fillColor] ( 44.05,435.50) circle (  1.49);

\path[fill=fillColor] ( 40.69,432.29) circle (  1.49);

\path[fill=fillColor] ( 38.86,431.29) circle (  1.49);

\path[fill=fillColor] ( 45.28,436.72) circle (  1.49);

\path[fill=fillColor] ( 41.56,434.14) circle (  1.49);

\path[fill=fillColor] ( 34.80,428.12) circle (  1.49);

\path[fill=fillColor] ( 35.60,428.45) circle (  1.49);

\path[fill=fillColor] ( 40.08,432.35) circle (  1.49);

\path[fill=fillColor] ( 76.66,461.26) circle (  1.49);

\path[fill=fillColor] ( 45.00,437.05) circle (  1.49);

\path[fill=fillColor] ( 36.74,429.28) circle (  1.49);

\path[fill=fillColor] ( 59.58,448.64) circle (  1.49);

\path[fill=fillColor] ( 37.07,429.97) circle (  1.49);

\path[fill=fillColor] ( 54.30,443.78) circle (  1.49);

\path[fill=fillColor] ( 45.05,436.52) circle (  1.49);

\path[fill=fillColor] ( 42.01,433.61) circle (  1.49);

\path[fill=fillColor] ( 84.96,466.59) circle (  1.49);

\path[fill=fillColor] ( 40.31,433.27) circle (  1.49);

\path[fill=fillColor] ( 38.72,431.05) circle (  1.49);

\path[fill=fillColor] ( 54.14,444.01) circle (  1.49);

\path[fill=fillColor] ( 48.31,438.93) circle (  1.49);

\path[fill=fillColor] ( 41.94,434.21) circle (  1.49);

\path[fill=fillColor] ( 40.16,432.47) circle (  1.49);

\path[fill=fillColor] ( 55.16,444.15) circle (  1.49);

\path[fill=fillColor] ( 42.07,449.76) circle (  1.49);

\path[fill=fillColor] ( 41.99,433.33) circle (  1.49);

\path[fill=fillColor] ( 66.97,451.83) circle (  1.49);

\path[fill=fillColor] ( 35.13,428.66) circle (  1.49);

\path[fill=fillColor] ( 44.24,462.43) circle (  1.49);

\path[fill=fillColor] ( 39.65,431.99) circle (  1.49);

\path[fill=fillColor] ( 35.47,428.93) circle (  1.49);

\path[fill=fillColor] ( 42.51,435.99) circle (  1.49);

\path[fill=fillColor] ( 43.32,435.03) circle (  1.49);

\path[fill=fillColor] ( 40.35,432.36) circle (  1.49);

\path[fill=fillColor] ( 50.65,441.76) circle (  1.49);

\path[fill=fillColor] ( 40.08,432.22) circle (  1.49);

\path[fill=fillColor] ( 37.72,430.27) circle (  1.49);

\path[fill=fillColor] ( 63.77,456.18) circle (  1.49);

\path[fill=fillColor] ( 42.95,434.87) circle (  1.49);

\path[fill=fillColor] ( 36.06,430.27) circle (  1.49);

\path[fill=fillColor] ( 50.39,441.15) circle (  1.49);

\path[fill=fillColor] ( 36.01,431.04) circle (  1.49);

\path[fill=fillColor] ( 41.72,433.19) circle (  1.49);

\path[fill=fillColor] ( 44.27,437.05) circle (  1.49);

\path[fill=fillColor] ( 38.16,430.44) circle (  1.49);

\path[fill=fillColor] ( 34.54,427.61) circle (  1.49);

\path[fill=fillColor] ( 40.49,432.80) circle (  1.49);

\path[fill=fillColor] ( 61.85,450.88) circle (  1.49);

\path[fill=fillColor] ( 65.11,452.79) circle (  1.49);

\path[fill=fillColor] ( 35.36,428.40) circle (  1.49);

\path[fill=fillColor] ( 36.65,429.29) circle (  1.49);

\path[fill=fillColor] ( 58.66,463.72) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,547.27) {Coverage: 93\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- (153.43,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 32.47,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.66,425.83) -- ( 56.66,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 80.85,425.83) -- ( 80.85,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.05,425.83) -- (105.05,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.24,425.83) -- (129.24,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (153.43,425.83) -- (153.43,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47,411.58) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 56.66,411.58) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 80.85,411.58) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (105.05,411.58) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (129.24,411.58) {2.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (153.43,411.58) {2.5};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 32.47,526.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 28.51,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,446.02) -- ( 28.51,446.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,466.21) -- ( 28.51,466.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,486.40) -- ( 28.51,486.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,506.59) -- ( 28.51,506.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,526.78) -- ( 28.51,526.78);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,423.56) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,443.75) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,463.94) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,484.13) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,504.32) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,524.50) {2.5};
\end{scope}
\begin{scope}
\path[clip] (192.72,385.44) rectangle (385.44,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (297.00,559.19) {\bfseries Yule birth rate $\lambda$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (297.00,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (199.85,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (225.19,425.83) rectangle (368.81,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (225.19,425.83) --
	(368.81,545.69);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.80,428.58) --
	(232.80,431.61);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.69,430.84) --
	(239.69,439.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.72,427.88) --
	(228.72,430.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.29,428.97) --
	(231.29,432.35);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.77,427.13) --
	(228.77,428.57);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.55,430.33) --
	(232.55,435.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.81,431.00) --
	(237.81,437.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.48,435.07) --
	(240.48,445.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.53,459.37) --
	(286.53,501.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.16,431.90) --
	(237.16,439.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.78,434.54) --
	(240.78,444.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.75,431.62) --
	(239.75,438.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (244.89,440.41) --
	(244.89,457.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.72,426.66) --
	(226.72,427.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.24,429.79) --
	(231.24,434.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.27,434.21) --
	(241.27,443.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.06,427.64) --
	(230.06,431.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.15,429.17) --
	(232.15,433.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.48,441.65) --
	(263.48,460.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.52,434.44) --
	(246.52,444.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.70,430.15) --
	(234.70,434.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.15,430.37) --
	(234.15,435.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.65,428.36) --
	(229.65,431.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.12,428.20) --
	(231.12,431.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.78,429.86) --
	(233.78,434.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.22,430.77) --
	(234.22,436.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.47,435.28) --
	(239.47,447.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.02,429.65) --
	(231.02,433.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.42,427.33) --
	(228.42,428.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.72,435.52) --
	(241.72,446.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.55,428.87) --
	(230.55,432.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.75,428.25) --
	(229.75,430.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (307.02,461.00) --
	(307.02,502.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.02,429.10) --
	(232.02,432.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.61,432.87) --
	(238.61,440.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.87,433.15) --
	(241.87,441.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.75,427.01) --
	(231.75,434.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.12,435.62) --
	(246.12,446.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.31,428.46) --
	(230.31,431.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.46,429.26) --
	(231.46,433.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.59,427.76) --
	(229.59,429.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (287.50,476.24) --
	(287.50,538.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.17,430.48) --
	(233.17,435.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.25,431.04) --
	(235.25,436.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.89,427.24) --
	(227.89,428.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.58,438.69) --
	(252.58,454.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.94,433.23) --
	(239.94,441.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.23,430.23) --
	(235.23,435.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.59,431.20) --
	(235.59,437.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.79,434.51) --
	(239.79,445.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (244.20,434.68) --
	(244.20,444.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.38,429.78) --
	(233.38,434.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.78,431.29) --
	(236.78,437.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.67,446.53) --
	(257.67,472.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.43,445.20) --
	(261.43,468.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.24,432.36) --
	(237.24,439.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.85,427.64) --
	(228.85,429.61);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.03,429.68) --
	(236.03,434.60);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.23,437.65) --
	(243.23,450.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.76,427.90) --
	(228.76,430.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.28,439.07) --
	(247.28,454.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.65,428.11) --
	(230.65,430.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.34,429.93) --
	(232.34,434.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.48,431.58) --
	(234.48,438.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.68,426.91) --
	(227.68,428.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.63,431.17) --
	(245.63,451.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.64,432.94) --
	(237.64,440.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.78,428.24) --
	(229.78,430.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.60,428.93) --
	(231.60,432.36);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.15,431.06) --
	(239.15,436.96);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.17,431.09) --
	(234.17,437.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.55,428.44) --
	(231.55,431.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.89,425.99) --
	(240.89,446.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.09,431.81) --
	(240.09,438.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.70,427.38) --
	(227.70,429.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.93,442.02) --
	(254.93,465.45);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.29,426.27) --
	(238.29,431.74);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.98,433.23) --
	(238.98,441.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.05,444.18) --
	(257.05,467.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.37,430.60) --
	(235.37,437.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.08,430.93) --
	(237.08,436.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.09,432.05) --
	(241.09,439.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.08,428.83) --
	(232.08,432.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.28,433.04) --
	(243.28,442.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.33,434.97) --
	(241.33,445.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.80,427.31) --
	(227.80,429.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.34,430.55) --
	(235.34,435.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.77,437.42) --
	(252.77,457.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.09,428.07) --
	(229.09,430.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.36,428.83) --
	(254.36,469.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.07,430.58) --
	(234.07,436.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.87,429.90) --
	(234.87,434.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.50,437.32) --
	(248.50,450.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.00,453.45) --
	(261.00,485.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.83,432.03) --
	(239.83,439.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.90,428.69) --
	(228.90,431.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.40,427.44) --
	(228.40,429.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.24,447.48) --
	(275.24,472.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.74,438.13) --
	(242.74,452.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.30,427.04) --
	(229.30,430.28);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (232.80,430.08) circle (  1.49);

\path[fill=fillColor] (239.69,434.81) circle (  1.49);

\path[fill=fillColor] (228.72,429.02) circle (  1.49);

\path[fill=fillColor] (231.29,430.56) circle (  1.49);

\path[fill=fillColor] (228.77,427.84) circle (  1.49);

\path[fill=fillColor] (232.55,432.73) circle (  1.49);

\path[fill=fillColor] (237.81,434.32) circle (  1.49);

\path[fill=fillColor] (240.48,440.35) circle (  1.49);

\path[fill=fillColor] (286.53,480.29) circle (  1.49);

\path[fill=fillColor] (237.16,435.43) circle (  1.49);

\path[fill=fillColor] (240.78,439.31) circle (  1.49);

\path[fill=fillColor] (239.75,434.77) circle (  1.49);

\path[fill=fillColor] (244.89,448.74) circle (  1.49);

\path[fill=fillColor] (226.72,427.11) circle (  1.49);

\path[fill=fillColor] (231.24,431.97) circle (  1.49);

\path[fill=fillColor] (241.27,438.94) circle (  1.49);

\path[fill=fillColor] (230.06,429.41) circle (  1.49);

\path[fill=fillColor] (232.15,431.10) circle (  1.49);

\path[fill=fillColor] (263.48,450.77) circle (  1.49);

\path[fill=fillColor] (246.52,439.49) circle (  1.49);

\path[fill=fillColor] (234.70,432.46) circle (  1.49);

\path[fill=fillColor] (234.15,432.93) circle (  1.49);

\path[fill=fillColor] (229.65,429.72) circle (  1.49);

\path[fill=fillColor] (231.12,429.62) circle (  1.49);

\path[fill=fillColor] (233.78,432.03) circle (  1.49);

\path[fill=fillColor] (234.22,433.33) circle (  1.49);

\path[fill=fillColor] (239.47,441.03) circle (  1.49);

\path[fill=fillColor] (231.02,431.72) circle (  1.49);

\path[fill=fillColor] (228.42,428.09) circle (  1.49);

\path[fill=fillColor] (241.72,440.72) circle (  1.49);

\path[fill=fillColor] (230.55,430.53) circle (  1.49);

\path[fill=fillColor] (229.75,429.61) circle (  1.49);

\path[fill=fillColor] (307.02,480.88) circle (  1.49);

\path[fill=fillColor] (232.02,430.91) circle (  1.49);

\path[fill=fillColor] (238.61,436.56) circle (  1.49);

\path[fill=fillColor] (241.87,437.10) circle (  1.49);

\path[fill=fillColor] (231.75,430.83) circle (  1.49);

\path[fill=fillColor] (246.12,440.99) circle (  1.49);

\path[fill=fillColor] (230.31,429.92) circle (  1.49);

\path[fill=fillColor] (231.46,431.16) circle (  1.49);

\path[fill=fillColor] (229.59,428.81) circle (  1.49);

\path[fill=fillColor] (287.50,505.74) circle (  1.49);

\path[fill=fillColor] (233.17,432.95) circle (  1.49);

\path[fill=fillColor] (235.25,433.80) circle (  1.49);

\path[fill=fillColor] (227.89,428.01) circle (  1.49);

\path[fill=fillColor] (252.58,446.23) circle (  1.49);

\path[fill=fillColor] (239.94,437.07) circle (  1.49);

\path[fill=fillColor] (235.23,432.67) circle (  1.49);

\path[fill=fillColor] (235.59,434.02) circle (  1.49);

\path[fill=fillColor] (239.79,439.68) circle (  1.49);

\path[fill=fillColor] (244.20,439.72) circle (  1.49);

\path[fill=fillColor] (233.38,431.90) circle (  1.49);

\path[fill=fillColor] (236.78,434.43) circle (  1.49);

\path[fill=fillColor] (257.67,458.82) circle (  1.49);

\path[fill=fillColor] (261.43,456.33) circle (  1.49);

\path[fill=fillColor] (237.24,436.07) circle (  1.49);

\path[fill=fillColor] (228.85,428.58) circle (  1.49);

\path[fill=fillColor] (236.03,432.01) circle (  1.49);

\path[fill=fillColor] (243.23,443.87) circle (  1.49);

\path[fill=fillColor] (228.76,429.00) circle (  1.49);

\path[fill=fillColor] (247.28,446.08) circle (  1.49);

\path[fill=fillColor] (230.65,429.38) circle (  1.49);

\path[fill=fillColor] (232.34,432.12) circle (  1.49);

\path[fill=fillColor] (234.48,434.80) circle (  1.49);

\path[fill=fillColor] (227.68,427.50) circle (  1.49);

\path[fill=fillColor] (245.63,441.15) circle (  1.49);

\path[fill=fillColor] (237.64,436.82) circle (  1.49);

\path[fill=fillColor] (229.78,429.51) circle (  1.49);

\path[fill=fillColor] (231.60,430.58) circle (  1.49);

\path[fill=fillColor] (239.15,433.95) circle (  1.49);

\path[fill=fillColor] (234.17,433.97) circle (  1.49);

\path[fill=fillColor] (231.55,429.88) circle (  1.49);

\path[fill=fillColor] (240.89,433.95) circle (  1.49);

\path[fill=fillColor] (240.09,435.05) circle (  1.49);

\path[fill=fillColor] (227.70,428.19) circle (  1.49);

\path[fill=fillColor] (254.93,453.15) circle (  1.49);

\path[fill=fillColor] (238.29,429.03) circle (  1.49);

\path[fill=fillColor] (238.98,437.29) circle (  1.49);

\path[fill=fillColor] (257.05,455.51) circle (  1.49);

\path[fill=fillColor] (235.37,434.05) circle (  1.49);

\path[fill=fillColor] (237.08,433.71) circle (  1.49);

\path[fill=fillColor] (241.09,435.66) circle (  1.49);

\path[fill=fillColor] (232.08,430.45) circle (  1.49);

\path[fill=fillColor] (243.28,437.44) circle (  1.49);

\path[fill=fillColor] (241.33,440.16) circle (  1.49);

\path[fill=fillColor] (227.80,428.15) circle (  1.49);

\path[fill=fillColor] (235.34,433.14) circle (  1.49);

\path[fill=fillColor] (252.77,446.89) circle (  1.49);

\path[fill=fillColor] (229.09,429.30) circle (  1.49);

\path[fill=fillColor] (254.36,449.14) circle (  1.49);

\path[fill=fillColor] (234.07,433.47) circle (  1.49);

\path[fill=fillColor] (234.87,432.23) circle (  1.49);

\path[fill=fillColor] (248.50,443.56) circle (  1.49);

\path[fill=fillColor] (261.00,468.81) circle (  1.49);

\path[fill=fillColor] (239.83,435.53) circle (  1.49);

\path[fill=fillColor] (228.90,430.20) circle (  1.49);

\path[fill=fillColor] (228.40,428.30) circle (  1.49);

\path[fill=fillColor] (275.24,459.57) circle (  1.49);

\path[fill=fillColor] (242.74,444.89) circle (  1.49);

\path[fill=fillColor] (229.30,428.64) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (297.00,547.27) {Coverage: 95\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (340.32,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (225.19,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.57,425.83) -- (263.57,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (301.94,425.83) -- (301.94,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.32,425.83) -- (340.32,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (225.19,411.58) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (263.57,411.58) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (301.94,411.58) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (340.32,411.58) {150};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (225.19,521.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (221.23,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,457.86) -- (221.23,457.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,489.88) -- (221.23,489.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,521.91) -- (221.23,521.91);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,423.56) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,455.59) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,487.61) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,519.64) {150};
\end{scope}
\begin{scope}
\path[clip] (385.44,385.44) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (489.72,559.19) {\bfseries HKY model $\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (489.72,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (392.57,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (417.91,425.83) rectangle (561.53,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (417.91,425.83) --
	(561.53,545.69);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.03,426.70) --
	(419.03,426.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.54,459.67) --
	(465.54,469.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.24,432.35) --
	(426.24,432.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.79,426.50) --
	(418.79,426.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.17,437.95) --
	(433.17,439.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.65,427.96) --
	(420.65,428.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.68,429.74) --
	(422.68,430.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.90,430.22) --
	(423.90,431.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (440.37,441.88) --
	(440.37,451.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.38,426.99) --
	(419.38,427.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.69,428.62) --
	(421.69,429.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.32,427.73) --
	(420.32,428.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.25,430.48) --
	(424.25,431.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.48,428.61) --
	(421.48,428.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.34,426.95) --
	(419.34,427.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.36,427.66) --
	(420.36,427.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.96,434.68) --
	(428.96,435.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.41,427.76) --
	(420.41,427.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (508.96,484.50) --
	(508.96,526.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.75,427.23) --
	(419.75,427.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.81,427.36) --
	(419.81,427.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.77,428.75) --
	(421.77,429.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.37,429.30) --
	(422.37,429.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (430.81,436.58) --
	(430.81,437.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.00,429.67) --
	(423.00,430.15);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.22,426.09) --
	(418.22,426.17);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.55,428.53) --
	(421.55,429.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.89,432.16) --
	(425.89,432.97);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.89,426.66) --
	(418.89,426.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.02,428.46) --
	(421.02,428.96);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.69,428.13) --
	(420.69,428.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.16,426.00) --
	(418.16,426.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.82,429.76) --
	(423.82,431.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.63,429.59) --
	(422.63,430.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.27,429.22) --
	(422.27,429.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.48,428.54) --
	(421.48,428.98);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.09,426.84) --
	(419.09,427.00);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.44,432.32) --
	(426.44,433.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.73,431.17) --
	(424.73,431.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.52,427.89) --
	(420.52,428.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.21,431.69) --
	(425.21,432.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.89,432.18) --
	(426.89,435.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.90,432.85) --
	(426.90,433.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.67,426.40) --
	(418.67,426.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (450.47,451.60) --
	(450.47,454.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.91,433.09) --
	(427.91,434.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.09,434.49) --
	(429.09,436.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.96,427.43) --
	(419.96,427.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.47,427.84) --
	(420.47,428.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (451.31,450.90) --
	(451.31,459.45);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.58,427.25) --
	(419.58,427.52);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.25,434.08) --
	(428.25,435.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.15,426.83) --
	(419.15,426.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.46,431.20) --
	(425.46,432.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.22,427.55) --
	(420.22,428.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.49,431.41) --
	(425.49,432.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.37,429.31) --
	(422.37,429.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.01,427.51) --
	(420.01,427.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.65,427.10) --
	(419.65,427.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.21,426.04) --
	(418.21,426.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.68,426.40) --
	(418.68,426.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.80,430.55) --
	(423.80,431.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.16,451.64) --
	(453.16,456.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.49,430.09) --
	(423.49,430.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (447.60,449.25) --
	(447.60,451.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.79,426.47) --
	(418.79,426.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.93,427.43) --
	(419.93,427.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.66,426.39) --
	(418.66,426.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.14,428.36) --
	(421.14,428.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.76,427.97) --
	(420.76,428.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.62,429.40) --
	(422.62,429.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.50,428.74) --
	(421.50,429.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.09,431.16) --
	(425.09,432.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.84,431.86) --
	(425.84,432.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.81,426.57) --
	(418.81,426.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (434.39,436.79) --
	(434.39,440.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.22,428.36) --
	(421.22,428.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (440.95,444.24) --
	(440.95,448.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (448.11,447.41) --
	(448.11,457.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.39,426.19) --
	(418.39,426.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.39,431.62) --
	(425.39,432.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.52,460.41) --
	(459.52,470.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.47,428.69) --
	(421.47,429.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.41,432.37) --
	(426.41,433.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.82,433.29) --
	(427.82,434.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.74,427.29) --
	(419.74,427.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.62,432.63) --
	(426.62,433.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.75,428.00) --
	(420.75,428.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.47,440.22) --
	(436.47,441.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.53,428.73) --
	(421.53,429.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.57,439.19) --
	(435.57,441.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.13,427.58) --
	(420.13,427.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.39,433.41) --
	(427.39,435.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.19,430.40) --
	(424.19,431.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.60,428.78) --
	(421.60,429.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.12,429.90) --
	(423.12,430.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.89,428.99) --
	(421.89,429.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.91,430.49) --
	(424.91,431.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.33,426.12) --
	(418.33,426.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.08,429.11) --
	(422.08,429.48);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (419.03,426.76) circle (  1.49);

\path[fill=fillColor] (465.54,464.87) circle (  1.49);

\path[fill=fillColor] (426.24,432.65) circle (  1.49);

\path[fill=fillColor] (418.79,426.56) circle (  1.49);

\path[fill=fillColor] (433.17,438.64) circle (  1.49);

\path[fill=fillColor] (420.65,428.10) circle (  1.49);

\path[fill=fillColor] (422.68,430.03) circle (  1.49);

\path[fill=fillColor] (423.90,430.66) circle (  1.49);

\path[fill=fillColor] (440.37,446.76) circle (  1.49);

\path[fill=fillColor] (419.38,427.09) circle (  1.49);

\path[fill=fillColor] (421.69,428.87) circle (  1.49);

\path[fill=fillColor] (420.32,427.88) circle (  1.49);

\path[fill=fillColor] (424.25,431.02) circle (  1.49);

\path[fill=fillColor] (421.48,428.75) circle (  1.49);

\path[fill=fillColor] (419.34,427.03) circle (  1.49);

\path[fill=fillColor] (420.36,427.82) circle (  1.49);

\path[fill=fillColor] (428.96,435.22) circle (  1.49);

\path[fill=fillColor] (420.41,427.88) circle (  1.49);

\path[fill=fillColor] (508.96,505.74) circle (  1.49);

\path[fill=fillColor] (419.75,427.36) circle (  1.49);

\path[fill=fillColor] (419.81,427.46) circle (  1.49);

\path[fill=fillColor] (421.77,428.94) circle (  1.49);

\path[fill=fillColor] (422.37,429.48) circle (  1.49);

\path[fill=fillColor] (430.81,437.25) circle (  1.49);

\path[fill=fillColor] (423.00,429.91) circle (  1.49);

\path[fill=fillColor] (418.22,426.13) circle (  1.49);

\path[fill=fillColor] (421.55,428.79) circle (  1.49);

\path[fill=fillColor] (425.89,432.56) circle (  1.49);

\path[fill=fillColor] (418.89,426.71) circle (  1.49);

\path[fill=fillColor] (421.02,428.71) circle (  1.49);

\path[fill=fillColor] (420.69,428.25) circle (  1.49);

\path[fill=fillColor] (418.16,426.03) circle (  1.49);

\path[fill=fillColor] (423.82,430.51) circle (  1.49);

\path[fill=fillColor] (422.63,429.80) circle (  1.49);

\path[fill=fillColor] (422.27,429.48) circle (  1.49);

\path[fill=fillColor] (421.48,428.76) circle (  1.49);

\path[fill=fillColor] (419.09,426.92) circle (  1.49);

\path[fill=fillColor] (426.44,433.02) circle (  1.49);

\path[fill=fillColor] (424.73,431.46) circle (  1.49);

\path[fill=fillColor] (420.52,428.01) circle (  1.49);

\path[fill=fillColor] (425.21,431.97) circle (  1.49);

\path[fill=fillColor] (426.89,433.85) circle (  1.49);

\path[fill=fillColor] (426.90,433.36) circle (  1.49);

\path[fill=fillColor] (418.67,426.45) circle (  1.49);

\path[fill=fillColor] (450.47,453.15) circle (  1.49);

\path[fill=fillColor] (427.91,433.96) circle (  1.49);

\path[fill=fillColor] (429.09,435.23) circle (  1.49);

\path[fill=fillColor] (419.96,427.54) circle (  1.49);

\path[fill=fillColor] (420.47,427.97) circle (  1.49);

\path[fill=fillColor] (451.31,455.23) circle (  1.49);

\path[fill=fillColor] (419.58,427.39) circle (  1.49);

\path[fill=fillColor] (428.25,434.62) circle (  1.49);

\path[fill=fillColor] (419.15,426.91) circle (  1.49);

\path[fill=fillColor] (425.46,432.01) circle (  1.49);

\path[fill=fillColor] (420.22,427.78) circle (  1.49);

\path[fill=fillColor] (425.49,431.88) circle (  1.49);

\path[fill=fillColor] (422.37,429.48) circle (  1.49);

\path[fill=fillColor] (420.01,427.62) circle (  1.49);

\path[fill=fillColor] (419.65,427.24) circle (  1.49);

\path[fill=fillColor] (418.21,426.07) circle (  1.49);

\path[fill=fillColor] (418.68,426.49) circle (  1.49);

\path[fill=fillColor] (423.80,430.79) circle (  1.49);

\path[fill=fillColor] (453.16,454.16) circle (  1.49);

\path[fill=fillColor] (423.49,430.41) circle (  1.49);

\path[fill=fillColor] (447.60,450.46) circle (  1.49);

\path[fill=fillColor] (418.79,426.57) circle (  1.49);

\path[fill=fillColor] (419.93,427.56) circle (  1.49);

\path[fill=fillColor] (418.66,426.44) circle (  1.49);

\path[fill=fillColor] (421.14,428.49) circle (  1.49);

\path[fill=fillColor] (420.76,428.14) circle (  1.49);

\path[fill=fillColor] (422.62,429.66) circle (  1.49);

\path[fill=fillColor] (421.50,428.89) circle (  1.49);

\path[fill=fillColor] (425.09,431.79) circle (  1.49);

\path[fill=fillColor] (425.84,432.33) circle (  1.49);

\path[fill=fillColor] (418.81,426.61) circle (  1.49);

\path[fill=fillColor] (434.39,438.84) circle (  1.49);

\path[fill=fillColor] (421.22,428.51) circle (  1.49);

\path[fill=fillColor] (440.95,446.45) circle (  1.49);

\path[fill=fillColor] (448.11,452.30) circle (  1.49);

\path[fill=fillColor] (418.39,426.24) circle (  1.49);

\path[fill=fillColor] (425.39,432.03) circle (  1.49);

\path[fill=fillColor] (459.52,465.57) circle (  1.49);

\path[fill=fillColor] (421.47,428.84) circle (  1.49);

\path[fill=fillColor] (426.41,432.94) circle (  1.49);

\path[fill=fillColor] (427.82,434.05) circle (  1.49);

\path[fill=fillColor] (419.74,427.37) circle (  1.49);

\path[fill=fillColor] (426.62,433.12) circle (  1.49);

\path[fill=fillColor] (420.75,428.26) circle (  1.49);

\path[fill=fillColor] (436.47,441.04) circle (  1.49);

\path[fill=fillColor] (421.53,429.20) circle (  1.49);

\path[fill=fillColor] (435.57,440.31) circle (  1.49);

\path[fill=fillColor] (420.13,427.69) circle (  1.49);

\path[fill=fillColor] (427.39,434.26) circle (  1.49);

\path[fill=fillColor] (424.19,431.18) circle (  1.49);

\path[fill=fillColor] (421.60,429.00) circle (  1.49);

\path[fill=fillColor] (423.12,430.13) circle (  1.49);

\path[fill=fillColor] (421.89,429.13) circle (  1.49);

\path[fill=fillColor] (424.91,431.18) circle (  1.49);

\path[fill=fillColor] (418.33,426.18) circle (  1.49);

\path[fill=fillColor] (422.08,429.29) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (489.72,547.27) {Coverage: 95\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (547.77,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (417.91,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.46,425.83) -- (436.46,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (455.01,425.83) -- (455.01,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (473.56,425.83) -- (473.56,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (492.11,425.83) -- (492.11,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.67,425.83) -- (510.67,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (529.22,425.83) -- (529.22,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.77,425.83) -- (547.77,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (417.91,411.58) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (436.46,411.58) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (455.01,411.58) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (473.56,411.58) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (492.11,411.58) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (510.67,411.58) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (529.22,411.58) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (547.77,411.58) {70};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (417.91,534.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (413.95,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,441.31) -- (413.95,441.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,456.80) -- (413.95,456.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,472.28) -- (413.95,472.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,487.76) -- (413.95,487.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,503.24) -- (413.95,503.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,518.72) -- (413.95,518.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,534.20) -- (413.95,534.20);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,423.56) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,439.04) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,454.52) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,470.00) {30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,485.49) {40};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,500.97) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,516.45) {60};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,531.93) {70};
\end{scope}
\begin{scope}
\path[clip] (  0.00,192.72) rectangle (192.72,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,366.47) {\bfseries Nucleotide frequency $f$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47,233.11) rectangle (176.09,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47,233.11) --
	(176.09,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.33,254.08) --
	( 59.33,256.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 80.63,271.74) --
	( 80.63,275.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.17,266.71) --
	( 73.17,269.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.94,291.97) --
	(105.94,295.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.16,281.84) --
	( 93.16,284.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.94,275.30) --
	( 84.94,279.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (113.56,296.56) --
	(113.56,301.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (113.44,297.60) --
	(113.44,303.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.84,285.56) --
	( 98.84,292.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.32,275.04) --
	( 85.32,279.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (118.66,301.28) --
	(118.66,306.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.87,264.55) --
	( 73.87,268.22);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.25,271.76) --
	( 86.25,276.98);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.81,268.83) --
	( 75.81,270.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 80.91,272.07) --
	( 80.91,275.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.58,303.49) --
	(120.58,309.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.58,256.44) --
	( 60.58,258.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.33,277.33) --
	( 88.33,280.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (106.44,293.72) --
	(106.44,300.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.72,290.82) --
	(104.72,296.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.16,264.69) --
	( 72.16,267.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.42,282.20) --
	( 92.42,286.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (111.46,295.79) --
	(111.46,299.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.18,273.96) --
	( 84.18,277.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.21,280.57) --
	( 92.21,284.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.43,285.22) --
	( 96.43,289.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (108.28,291.45) --
	(108.28,297.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.31,261.26) --
	( 68.31,264.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 97.56,285.25) --
	( 97.56,288.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.51,290.19) --
	(103.51,295.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.43,280.32) --
	( 90.43,283.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.53,310.97) --
	(127.53,314.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 94.65,282.42) --
	( 94.65,288.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.94,255.34) --
	( 60.94,257.90);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.35,289.98) --
	(100.35,294.88);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.15,288.35) --
	(101.15,293.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.01,270.90) --
	( 81.01,274.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (107.92,292.85) --
	(107.92,298.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.71,267.54) --
	( 75.71,270.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 94.59,282.13) --
	( 94.59,285.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 82.61,272.60) --
	( 82.61,275.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.63,276.72) --
	( 88.63,283.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.65,286.81) --
	( 99.65,291.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.76,275.12) --
	( 85.76,279.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.99,259.09) --
	( 65.99,261.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.46,273.24) --
	( 84.46,278.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.45,293.44) --
	(109.45,298.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.35,263.72) --
	( 72.35,267.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.12,289.90) --
	(103.12,294.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.86,266.38) --
	( 74.86,271.24);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.79,300.74) --
	(120.79,306.24);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.24,283.37) --
	( 95.24,287.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.27,264.69) --
	( 73.27,268.22);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.10,294.78) --
	(114.10,301.21);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.07,279.75) --
	( 92.07,285.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.88,289.47) --
	(100.88,294.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.56,276.10) --
	( 86.56,278.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.50,269.33) --
	( 77.50,272.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.33,275.25) --
	( 86.33,280.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.99,298.27) --
	(112.99,301.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.83,258.04) --
	( 65.83,262.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.11,283.39) --
	( 95.11,286.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (102.89,290.21) --
	(102.89,294.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.35,268.59) --
	( 78.35,272.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 94.74,284.09) --
	( 94.74,286.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.73,273.90) --
	( 84.73,279.15);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (106.49,289.89) --
	(106.49,294.77);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.57,264.41) --
	( 70.57,267.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.28,266.55) --
	( 74.28,269.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.10,284.68) --
	( 96.10,289.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.57,278.04) --
	( 89.57,282.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.46,291.30) --
	(104.46,295.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.89,266.30) --
	( 74.89,270.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.71,294.80) --
	(109.71,299.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.25,260.21) --
	( 65.25,262.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.62,263.65) --
	( 72.62,269.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.70,279.70) --
	( 89.70,283.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.56,281.90) --
	( 93.56,286.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.61,290.60) --
	(104.61,296.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.69,274.56) --
	( 83.69,278.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.63,283.67) --
	( 95.63,288.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.36,277.96) --
	( 89.36,283.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (113.82,299.81) --
	(113.82,303.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.23,288.47) --
	(103.23,293.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (111.99,298.59) --
	(111.99,304.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.13,269.61) --
	( 79.13,272.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.62,277.54) --
	( 87.62,281.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.64,265.27) --
	( 74.64,270.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.40,268.70) --
	( 76.40,271.45);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.22,273.88) --
	( 81.22,279.93);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 71.59,264.18) --
	( 71.59,267.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.62,295.48) --
	(109.62,299.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.11,264.80) --
	( 73.11,269.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.90,274.47) --
	( 84.90,280.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.86,275.99) --
	( 85.86,280.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 82.66,273.44) --
	( 82.66,276.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.60,266.22) --
	( 72.60,268.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.20,281.16) --
	( 93.20,287.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.32,290.70) --
	(103.32,296.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 64.04,258.94) --
	( 64.04,261.45);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 59.33,255.22) circle (  1.49);

\path[fill=fillColor] ( 80.63,273.70) circle (  1.49);

\path[fill=fillColor] ( 73.17,268.05) circle (  1.49);

\path[fill=fillColor] (105.94,293.98) circle (  1.49);

\path[fill=fillColor] ( 93.16,282.98) circle (  1.49);

\path[fill=fillColor] ( 84.94,277.19) circle (  1.49);

\path[fill=fillColor] (113.56,299.21) circle (  1.49);

\path[fill=fillColor] (113.44,300.47) circle (  1.49);

\path[fill=fillColor] ( 98.84,288.83) circle (  1.49);

\path[fill=fillColor] ( 85.32,277.06) circle (  1.49);

\path[fill=fillColor] (118.66,304.11) circle (  1.49);

\path[fill=fillColor] ( 73.87,266.38) circle (  1.49);

\path[fill=fillColor] ( 86.25,274.44) circle (  1.49);

\path[fill=fillColor] ( 75.81,269.77) circle (  1.49);

\path[fill=fillColor] ( 80.91,273.81) circle (  1.49);

\path[fill=fillColor] (120.58,306.20) circle (  1.49);

\path[fill=fillColor] ( 60.58,257.59) circle (  1.49);

\path[fill=fillColor] ( 88.33,279.15) circle (  1.49);

\path[fill=fillColor] (106.44,296.80) circle (  1.49);

\path[fill=fillColor] (104.72,293.47) circle (  1.49);

\path[fill=fillColor] ( 72.16,266.32) circle (  1.49);

\path[fill=fillColor] ( 92.42,284.17) circle (  1.49);

\path[fill=fillColor] (111.46,297.60) circle (  1.49);

\path[fill=fillColor] ( 84.18,275.50) circle (  1.49);

\path[fill=fillColor] ( 92.21,282.50) circle (  1.49);

\path[fill=fillColor] ( 96.43,287.31) circle (  1.49);

\path[fill=fillColor] (108.28,294.22) circle (  1.49);

\path[fill=fillColor] ( 68.31,262.72) circle (  1.49);

\path[fill=fillColor] ( 97.56,286.66) circle (  1.49);

\path[fill=fillColor] (103.51,292.77) circle (  1.49);

\path[fill=fillColor] ( 90.43,282.06) circle (  1.49);

\path[fill=fillColor] (127.53,313.02) circle (  1.49);

\path[fill=fillColor] ( 94.65,285.65) circle (  1.49);

\path[fill=fillColor] ( 60.94,256.66) circle (  1.49);

\path[fill=fillColor] (100.35,292.39) circle (  1.49);

\path[fill=fillColor] (101.15,290.82) circle (  1.49);

\path[fill=fillColor] ( 81.01,272.68) circle (  1.49);

\path[fill=fillColor] (107.92,295.57) circle (  1.49);

\path[fill=fillColor] ( 75.71,269.04) circle (  1.49);

\path[fill=fillColor] ( 94.59,283.98) circle (  1.49);

\path[fill=fillColor] ( 82.61,274.05) circle (  1.49);

\path[fill=fillColor] ( 88.63,279.95) circle (  1.49);

\path[fill=fillColor] ( 99.65,288.96) circle (  1.49);

\path[fill=fillColor] ( 85.76,277.07) circle (  1.49);

\path[fill=fillColor] ( 65.99,260.15) circle (  1.49);

\path[fill=fillColor] ( 84.46,275.80) circle (  1.49);

\path[fill=fillColor] (109.45,296.06) circle (  1.49);

\path[fill=fillColor] ( 72.35,265.31) circle (  1.49);

\path[fill=fillColor] (103.12,292.24) circle (  1.49);

\path[fill=fillColor] ( 74.86,268.86) circle (  1.49);

\path[fill=fillColor] (120.79,303.54) circle (  1.49);

\path[fill=fillColor] ( 95.24,285.31) circle (  1.49);

\path[fill=fillColor] ( 73.27,266.47) circle (  1.49);

\path[fill=fillColor] (114.10,298.07) circle (  1.49);

\path[fill=fillColor] ( 92.07,282.70) circle (  1.49);

\path[fill=fillColor] (100.88,291.71) circle (  1.49);

\path[fill=fillColor] ( 86.56,277.46) circle (  1.49);

\path[fill=fillColor] ( 77.50,271.05) circle (  1.49);

\path[fill=fillColor] ( 86.33,277.79) circle (  1.49);

\path[fill=fillColor] (112.99,300.12) circle (  1.49);

\path[fill=fillColor] ( 65.83,260.12) circle (  1.49);

\path[fill=fillColor] ( 95.11,285.13) circle (  1.49);

\path[fill=fillColor] (102.89,292.35) circle (  1.49);

\path[fill=fillColor] ( 78.35,270.69) circle (  1.49);

\path[fill=fillColor] ( 94.74,285.45) circle (  1.49);

\path[fill=fillColor] ( 84.73,276.37) circle (  1.49);

\path[fill=fillColor] (106.49,292.26) circle (  1.49);

\path[fill=fillColor] ( 70.57,265.77) circle (  1.49);

\path[fill=fillColor] ( 74.28,268.06) circle (  1.49);

\path[fill=fillColor] ( 96.10,286.96) circle (  1.49);

\path[fill=fillColor] ( 89.57,279.95) circle (  1.49);

\path[fill=fillColor] (104.46,293.11) circle (  1.49);

\path[fill=fillColor] ( 74.89,268.58) circle (  1.49);

\path[fill=fillColor] (109.71,297.28) circle (  1.49);

\path[fill=fillColor] ( 65.25,261.25) circle (  1.49);

\path[fill=fillColor] ( 72.62,266.42) circle (  1.49);

\path[fill=fillColor] ( 89.70,281.42) circle (  1.49);

\path[fill=fillColor] ( 93.56,284.18) circle (  1.49);

\path[fill=fillColor] (104.61,293.87) circle (  1.49);

\path[fill=fillColor] ( 83.69,276.64) circle (  1.49);

\path[fill=fillColor] ( 95.63,285.85) circle (  1.49);

\path[fill=fillColor] ( 89.36,280.58) circle (  1.49);

\path[fill=fillColor] (113.82,301.76) circle (  1.49);

\path[fill=fillColor] (103.23,291.03) circle (  1.49);

\path[fill=fillColor] (111.99,301.31) circle (  1.49);

\path[fill=fillColor] ( 79.13,270.89) circle (  1.49);

\path[fill=fillColor] ( 87.62,279.58) circle (  1.49);

\path[fill=fillColor] ( 74.64,267.74) circle (  1.49);

\path[fill=fillColor] ( 76.40,270.07) circle (  1.49);

\path[fill=fillColor] ( 81.22,276.89) circle (  1.49);

\path[fill=fillColor] ( 71.59,265.91) circle (  1.49);

\path[fill=fillColor] (109.62,297.52) circle (  1.49);

\path[fill=fillColor] ( 73.11,266.98) circle (  1.49);

\path[fill=fillColor] ( 84.90,277.48) circle (  1.49);

\path[fill=fillColor] ( 85.86,278.17) circle (  1.49);

\path[fill=fillColor] ( 82.66,275.14) circle (  1.49);

\path[fill=fillColor] ( 72.60,267.38) circle (  1.49);

\path[fill=fillColor] ( 93.20,284.39) circle (  1.49);

\path[fill=fillColor] (103.32,293.36) circle (  1.49);

\path[fill=fillColor] ( 64.04,260.17) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,354.55) {Coverage: 94\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- (173.40,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 32.47,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 55.96,233.11) -- ( 55.96,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.45,233.11) -- ( 79.45,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (102.94,233.11) -- (102.94,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.43,233.11) -- (126.43,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (149.92,233.11) -- (149.92,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (173.40,233.11) -- (173.40,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 55.96,218.86) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 79.45,218.86) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (102.94,218.86) {0.3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (126.43,218.86) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (149.92,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (173.40,218.86) {0.6};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 32.47,350.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 28.51,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,252.71) -- ( 28.51,252.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,272.32) -- ( 28.51,272.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,291.92) -- ( 28.51,291.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,311.52) -- ( 28.51,311.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,331.13) -- ( 28.51,331.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,350.73) -- ( 28.51,350.73);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,250.44) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,270.04) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,289.65) {0.3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,309.25) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,328.85) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,348.46) {0.6};
\end{scope}
\begin{scope}
\path[clip] (192.72,192.72) rectangle (385.44,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (297.00,366.47) {\bfseries Clock standard deviation $\sigma$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (297.00,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (199.85,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (225.19,233.11) rectangle (368.81,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (225.19,233.11) --
	(368.81,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.20,233.11) --
	(225.20,235.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.64,252.30) --
	(255.64,269.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.49,233.11) --
	(225.49,235.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.86,234.52) --
	(230.86,239.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.34,233.11) --
	(227.34,235.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.03,233.11) --
	(227.03,235.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.50,244.96) --
	(241.50,254.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.70,236.42) --
	(232.70,242.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.67,237.81) --
	(233.67,250.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.29,240.77) --
	(238.29,247.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.50,233.11) --
	(225.50,236.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.26,233.11) --
	(226.26,236.95);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.18,238.07) --
	(230.18,246.32);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.20,241.20) --
	(240.20,249.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.44,233.11) --
	(225.44,235.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.04,233.11) --
	(228.04,240.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (260.96,253.37) --
	(260.96,276.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.10,241.22) --
	(238.10,248.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.45,233.11) --
	(228.45,236.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.38,235.64) --
	(232.38,241.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.30,237.70) --
	(232.30,243.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.60,241.23) --
	(240.60,248.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.21,234.46) --
	(228.21,239.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.77,240.75) --
	(238.77,248.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.35,233.11) --
	(227.35,237.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.35,234.87) --
	(232.35,239.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.69,242.17) --
	(241.69,251.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.03,233.11) --
	(226.03,236.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.22,233.11) --
	(225.22,235.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.36,235.18) --
	(230.36,243.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.11,239.96) --
	(239.11,245.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.64,235.16) --
	(229.64,239.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.87,233.11) --
	(230.87,238.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.38,233.11) --
	(227.38,235.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.67,233.11) --
	(225.67,235.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.70,233.11) --
	(226.70,235.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (269.51,261.13) --
	(269.51,299.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.18,233.11) --
	(228.18,238.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.32,233.11) --
	(226.32,236.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.48,233.11) --
	(226.48,235.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.33,233.11) --
	(225.33,235.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.89,233.11) --
	(225.89,247.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.60,233.11) --
	(226.60,235.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) --
	(225.19,235.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.09,233.11) --
	(226.09,235.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.64,238.99) --
	(234.64,247.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.60,233.11) --
	(226.60,240.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.13,233.11) --
	(226.13,235.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.53,233.11) --
	(225.53,236.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.32,243.30) --
	(246.32,253.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.83,233.11) --
	(225.83,239.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.10,233.11) --
	(227.10,235.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.97,242.33) --
	(241.97,250.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.51,240.77) --
	(241.51,251.29);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.20,233.11) --
	(231.20,237.01);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.75,236.78) --
	(230.75,242.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.53,233.11) --
	(229.53,237.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.13,246.26) --
	(246.13,256.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.42,233.11) --
	(225.42,235.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.46,233.11) --
	(226.46,236.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.51,233.11) --
	(229.51,241.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.19,238.02) --
	(232.19,242.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.13,233.11) --
	(228.13,237.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.16,237.73) --
	(233.16,244.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.13,233.11) --
	(226.13,235.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (278.73,261.48) --
	(278.73,291.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.31,233.11) --
	(226.31,236.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.23,233.11) --
	(227.23,236.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.66,233.11) --
	(225.66,234.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.51,233.11) --
	(229.51,237.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.67,233.11) --
	(230.67,237.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.58,235.33) --
	(230.58,239.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (319.91,287.40) --
	(319.91,343.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.93,233.11) --
	(226.93,236.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.69,236.92) --
	(232.69,240.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.00,247.58) --
	(250.00,263.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.51,268.48) --
	(279.51,308.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.00,238.12) --
	(233.00,243.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.04,238.68) --
	(240.04,252.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.95,249.75) --
	(247.95,266.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.14,233.11) --
	(227.14,236.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.86,233.11) --
	(226.86,238.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) --
	(225.19,235.65);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.33,242.02) --
	(250.33,250.90);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.29,233.11) --
	(226.29,239.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.31,239.93) --
	(234.31,246.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.96,240.75) --
	(235.96,247.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.33,253.45) --
	(255.33,271.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.27,233.11) --
	(225.27,236.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.24,266.42) --
	(286.24,311.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.17,242.90) --
	(243.17,251.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.68,241.16) --
	(235.68,248.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.40,233.11) --
	(226.40,239.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.42,233.11) --
	(226.42,241.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.62,233.11) --
	(227.62,237.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.20,233.11) --
	(225.20,235.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.91,235.03) --
	(229.91,239.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.57,233.11) --
	(226.57,239.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.83,233.11) --
	(226.83,238.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.51,260.71) --
	(263.51,283.87);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (225.20,233.82) circle (  1.49);

\path[fill=fillColor] (255.64,260.34) circle (  1.49);

\path[fill=fillColor] (225.49,233.97) circle (  1.49);

\path[fill=fillColor] (230.86,237.04) circle (  1.49);

\path[fill=fillColor] (227.34,234.01) circle (  1.49);

\path[fill=fillColor] (227.03,234.05) circle (  1.49);

\path[fill=fillColor] (241.50,249.64) circle (  1.49);

\path[fill=fillColor] (232.70,239.36) circle (  1.49);

\path[fill=fillColor] (233.67,243.87) circle (  1.49);

\path[fill=fillColor] (238.29,244.18) circle (  1.49);

\path[fill=fillColor] (225.50,234.28) circle (  1.49);

\path[fill=fillColor] (226.26,234.59) circle (  1.49);

\path[fill=fillColor] (230.18,242.08) circle (  1.49);

\path[fill=fillColor] (240.20,245.03) circle (  1.49);

\path[fill=fillColor] (225.44,234.08) circle (  1.49);

\path[fill=fillColor] (228.04,237.26) circle (  1.49);

\path[fill=fillColor] (260.96,264.06) circle (  1.49);

\path[fill=fillColor] (238.10,244.53) circle (  1.49);

\path[fill=fillColor] (228.45,234.28) circle (  1.49);

\path[fill=fillColor] (232.38,238.46) circle (  1.49);

\path[fill=fillColor] (232.30,240.49) circle (  1.49);

\path[fill=fillColor] (240.60,244.93) circle (  1.49);

\path[fill=fillColor] (228.21,236.75) circle (  1.49);

\path[fill=fillColor] (238.77,244.21) circle (  1.49);

\path[fill=fillColor] (227.35,235.33) circle (  1.49);

\path[fill=fillColor] (232.35,237.11) circle (  1.49);

\path[fill=fillColor] (241.69,246.45) circle (  1.49);

\path[fill=fillColor] (226.03,234.24) circle (  1.49);

\path[fill=fillColor] (225.22,234.39) circle (  1.49);

\path[fill=fillColor] (230.36,239.13) circle (  1.49);

\path[fill=fillColor] (239.11,242.67) circle (  1.49);

\path[fill=fillColor] (229.64,237.11) circle (  1.49);

\path[fill=fillColor] (230.87,234.90) circle (  1.49);

\path[fill=fillColor] (227.38,233.84) circle (  1.49);

\path[fill=fillColor] (225.67,233.84) circle (  1.49);

\path[fill=fillColor] (226.70,233.90) circle (  1.49);

\path[fill=fillColor] (269.51,277.42) circle (  1.49);

\path[fill=fillColor] (228.18,235.68) circle (  1.49);

\path[fill=fillColor] (226.32,234.96) circle (  1.49);

\path[fill=fillColor] (226.48,233.79) circle (  1.49);

\path[fill=fillColor] (225.33,233.86) circle (  1.49);

\path[fill=fillColor] (225.89,239.09) circle (  1.49);

\path[fill=fillColor] (226.60,233.96) circle (  1.49);

\path[fill=fillColor] (225.19,233.91) circle (  1.49);

\path[fill=fillColor] (226.09,233.82) circle (  1.49);

\path[fill=fillColor] (234.64,243.16) circle (  1.49);

\path[fill=fillColor] (226.60,237.34) circle (  1.49);

\path[fill=fillColor] (226.13,233.77) circle (  1.49);

\path[fill=fillColor] (225.53,234.38) circle (  1.49);

\path[fill=fillColor] (246.32,248.30) circle (  1.49);

\path[fill=fillColor] (225.83,235.98) circle (  1.49);

\path[fill=fillColor] (227.10,233.90) circle (  1.49);

\path[fill=fillColor] (241.97,245.90) circle (  1.49);

\path[fill=fillColor] (241.51,245.67) circle (  1.49);

\path[fill=fillColor] (231.20,234.33) circle (  1.49);

\path[fill=fillColor] (230.75,239.36) circle (  1.49);

\path[fill=fillColor] (229.53,235.04) circle (  1.49);

\path[fill=fillColor] (246.13,251.25) circle (  1.49);

\path[fill=fillColor] (225.42,233.84) circle (  1.49);

\path[fill=fillColor] (226.46,234.66) circle (  1.49);

\path[fill=fillColor] (229.51,237.59) circle (  1.49);

\path[fill=fillColor] (232.19,240.45) circle (  1.49);

\path[fill=fillColor] (228.13,235.36) circle (  1.49);

\path[fill=fillColor] (233.16,240.73) circle (  1.49);

\path[fill=fillColor] (226.13,234.11) circle (  1.49);

\path[fill=fillColor] (278.73,275.22) circle (  1.49);

\path[fill=fillColor] (226.31,234.20) circle (  1.49);

\path[fill=fillColor] (227.23,234.79) circle (  1.49);

\path[fill=fillColor] (225.66,233.70) circle (  1.49);

\path[fill=fillColor] (229.51,234.67) circle (  1.49);

\path[fill=fillColor] (230.67,235.69) circle (  1.49);

\path[fill=fillColor] (230.58,237.26) circle (  1.49);

\path[fill=fillColor] (319.91,313.02) circle (  1.49);

\path[fill=fillColor] (226.93,234.40) circle (  1.49);

\path[fill=fillColor] (232.69,238.84) circle (  1.49);

\path[fill=fillColor] (250.00,255.21) circle (  1.49);

\path[fill=fillColor] (279.51,286.04) circle (  1.49);

\path[fill=fillColor] (233.00,240.94) circle (  1.49);

\path[fill=fillColor] (240.04,245.33) circle (  1.49);

\path[fill=fillColor] (247.95,257.65) circle (  1.49);

\path[fill=fillColor] (227.14,234.62) circle (  1.49);

\path[fill=fillColor] (226.86,235.97) circle (  1.49);

\path[fill=fillColor] (225.19,234.01) circle (  1.49);

\path[fill=fillColor] (250.33,246.18) circle (  1.49);

\path[fill=fillColor] (226.29,236.53) circle (  1.49);

\path[fill=fillColor] (234.31,242.83) circle (  1.49);

\path[fill=fillColor] (235.96,244.19) circle (  1.49);

\path[fill=fillColor] (255.33,262.16) circle (  1.49);

\path[fill=fillColor] (225.27,234.89) circle (  1.49);

\path[fill=fillColor] (286.24,286.56) circle (  1.49);

\path[fill=fillColor] (243.17,247.23) circle (  1.49);

\path[fill=fillColor] (235.68,244.51) circle (  1.49);

\path[fill=fillColor] (226.40,235.72) circle (  1.49);

\path[fill=fillColor] (226.42,235.76) circle (  1.49);

\path[fill=fillColor] (227.62,234.85) circle (  1.49);

\path[fill=fillColor] (225.20,233.78) circle (  1.49);

\path[fill=fillColor] (229.91,236.98) circle (  1.49);

\path[fill=fillColor] (226.57,235.29) circle (  1.49);

\path[fill=fillColor] (226.83,234.91) circle (  1.49);

\path[fill=fillColor] (263.51,271.73) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (297.00,354.55) {Coverage: 97\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (354.76,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (225.19,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.79,233.11) -- (246.79,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.38,233.11) -- (268.38,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (289.98,233.11) -- (289.98,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (311.57,233.11) -- (311.57,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.17,233.11) -- (333.17,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.76,233.11) -- (354.76,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (225.19,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (246.79,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (268.38,218.86) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (289.98,218.86) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (311.57,218.86) {2.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (333.17,218.86) {2.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (354.76,218.86) {3.0};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (225.19,341.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (221.23,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,251.13) -- (221.23,251.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,269.16) -- (221.23,269.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,287.18) -- (221.23,287.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,305.20) -- (221.23,305.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,323.22) -- (221.23,323.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,341.25) -- (221.23,341.25);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,248.86) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,266.88) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,284.91) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,302.93) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,320.95) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,338.97) {3.0};
\end{scope}
\begin{scope}
\path[clip] (385.44,192.72) rectangle (578.16,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (489.72,366.47) {\bfseries Branch rate mean};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (489.72,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (392.57,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (417.91,233.11) rectangle (561.53,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (417.91,233.11) --
	(561.53,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.02,281.83) --
	(477.02,283.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.35,263.58) --
	(481.35,289.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.91,281.78) --
	(476.91,282.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.58,280.30) --
	(475.58,284.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.92,281.74) --
	(476.92,283.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.48,281.76) --
	(476.48,283.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.09,273.49) --
	(482.09,291.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.99,279.26) --
	(474.99,285.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (473.68,276.43) --
	(473.68,287.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.79,276.98) --
	(475.79,287.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.04,281.43) --
	(477.04,283.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.08,281.51) --
	(477.08,283.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.69,278.06) --
	(475.69,286.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.12,278.30) --
	(482.12,291.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.95,281.57) --
	(476.95,283.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.76,280.37) --
	(476.76,284.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (473.21,262.78) --
	(473.21,298.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.51,277.91) --
	(478.51,289.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.66,281.64) --
	(476.66,283.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.87,280.06) --
	(475.87,285.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.25,278.66) --
	(475.25,285.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.64,277.39) --
	(477.64,289.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.51,280.00) --
	(477.51,283.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.99,276.47) --
	(470.99,288.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.38,281.25) --
	(476.38,284.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.15,280.41) --
	(476.15,284.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.54,276.25) --
	(475.54,289.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.90,281.45) --
	(476.90,283.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.02,281.62) --
	(477.02,283.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.19,279.54) --
	(477.19,285.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.19,277.38) --
	(481.19,286.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.18,280.19) --
	(476.18,284.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.41,280.74) --
	(478.41,283.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.60,281.89) --
	(476.60,283.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.04,281.89) --
	(477.04,283.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.69,281.89) --
	(476.69,283.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.40,245.90) --
	(500.40,302.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.81,280.52) --
	(477.81,283.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.31,281.13) --
	(477.31,283.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.83,281.94) --
	(476.83,282.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.05,281.98) --
	(477.05,283.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.19,277.35) --
	(477.19,285.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.28,281.64) --
	(477.28,283.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.02,281.87) --
	(477.02,283.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.11,281.96) --
	(477.11,283.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.49,277.28) --
	(479.49,287.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.51,280.35) --
	(477.51,285.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.94,281.90) --
	(476.94,282.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.98,281.49) --
	(476.98,283.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.13,273.12) --
	(474.13,287.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.11,281.05) --
	(477.11,284.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.14,281.90) --
	(477.14,283.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.82,275.90) --
	(479.82,287.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.50,276.39) --
	(478.50,289.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.49,281.43) --
	(477.49,283.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.42,279.76) --
	(478.42,286.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.82,281.16) --
	(475.82,283.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.82,270.99) --
	(477.82,287.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.95,281.88) --
	(476.95,283.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.05,281.58) --
	(477.05,283.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.09,280.05) --
	(478.09,285.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.63,278.26) --
	(475.63,284.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.25,281.33) --
	(477.25,284.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.86,278.43) --
	(480.86,285.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.96,281.75) --
	(476.96,283.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.08,247.70) --
	(459.08,288.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.71,281.51) --
	(476.71,283.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.55,281.17) --
	(476.55,283.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.88,281.96) --
	(476.88,282.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.45,281.33) --
	(477.45,283.66);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.98,281.30) --
	(478.98,284.04);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.97,280.59) --
	(476.97,284.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (482.57,233.94) --
	(482.57,333.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.52,281.60) --
	(477.52,283.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.74,280.25) --
	(476.74,285.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (472.31,266.92) --
	(472.31,287.33);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (513.66,238.30) --
	(513.66,290.31);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.37,278.62) --
	(476.37,285.68);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (484.58,274.95) --
	(484.58,287.92);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.53,268.47) --
	(479.53,292.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.67,281.17) --
	(476.67,283.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.73,281.09) --
	(476.73,284.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.02,281.66) --
	(477.02,283.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.44,276.20) --
	(478.44,288.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.92,280.41) --
	(476.92,284.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.29,278.03) --
	(481.29,286.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.47,275.41) --
	(476.47,284.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.63,260.12) --
	(476.63,285.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.02,281.40) --
	(477.02,283.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (460.99,237.47) --
	(460.99,284.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (473.50,273.33) --
	(473.50,286.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.09,275.62) --
	(476.09,288.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.57,280.65) --
	(476.57,284.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.78,279.91) --
	(476.78,283.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.92,281.36) --
	(476.92,283.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.02,281.80) --
	(477.02,283.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.44,280.54) --
	(477.44,284.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.43,280.80) --
	(477.43,284.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.82,281.16) --
	(476.82,283.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (468.31,250.76) --
	(468.31,281.94);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (477.02,282.44) circle (  1.49);

\path[fill=fillColor] (481.35,276.42) circle (  1.49);

\path[fill=fillColor] (476.91,282.41) circle (  1.49);

\path[fill=fillColor] (475.58,282.42) circle (  1.49);

\path[fill=fillColor] (476.92,282.45) circle (  1.49);

\path[fill=fillColor] (476.48,282.45) circle (  1.49);

\path[fill=fillColor] (482.09,282.31) circle (  1.49);

\path[fill=fillColor] (474.99,282.22) circle (  1.49);

\path[fill=fillColor] (473.68,281.96) circle (  1.49);

\path[fill=fillColor] (475.79,282.02) circle (  1.49);

\path[fill=fillColor] (477.04,282.40) circle (  1.49);

\path[fill=fillColor] (477.08,282.61) circle (  1.49);

\path[fill=fillColor] (475.69,282.30) circle (  1.49);

\path[fill=fillColor] (482.12,284.96) circle (  1.49);

\path[fill=fillColor] (476.95,282.38) circle (  1.49);

\path[fill=fillColor] (476.76,282.56) circle (  1.49);

\path[fill=fillColor] (473.21,280.45) circle (  1.49);

\path[fill=fillColor] (478.51,283.57) circle (  1.49);

\path[fill=fillColor] (476.66,282.44) circle (  1.49);

\path[fill=fillColor] (475.87,282.63) circle (  1.49);

\path[fill=fillColor] (475.25,282.17) circle (  1.49);

\path[fill=fillColor] (477.64,283.59) circle (  1.49);

\path[fill=fillColor] (477.51,281.98) circle (  1.49);

\path[fill=fillColor] (470.99,282.25) circle (  1.49);

\path[fill=fillColor] (476.38,282.57) circle (  1.49);

\path[fill=fillColor] (476.15,282.35) circle (  1.49);

\path[fill=fillColor] (475.54,282.91) circle (  1.49);

\path[fill=fillColor] (476.90,282.42) circle (  1.49);

\path[fill=fillColor] (477.02,282.50) circle (  1.49);

\path[fill=fillColor] (477.19,282.61) circle (  1.49);

\path[fill=fillColor] (481.19,281.96) circle (  1.49);

\path[fill=fillColor] (476.18,282.44) circle (  1.49);

\path[fill=fillColor] (478.41,282.37) circle (  1.49);

\path[fill=fillColor] (476.60,282.45) circle (  1.49);

\path[fill=fillColor] (477.04,282.45) circle (  1.49);

\path[fill=fillColor] (476.69,282.44) circle (  1.49);

\path[fill=fillColor] (500.40,276.39) circle (  1.49);

\path[fill=fillColor] (477.81,282.30) circle (  1.49);

\path[fill=fillColor] (477.31,282.35) circle (  1.49);

\path[fill=fillColor] (476.83,282.43) circle (  1.49);

\path[fill=fillColor] (477.05,282.52) circle (  1.49);

\path[fill=fillColor] (477.19,281.83) circle (  1.49);

\path[fill=fillColor] (477.28,282.43) circle (  1.49);

\path[fill=fillColor] (477.02,282.45) circle (  1.49);

\path[fill=fillColor] (477.11,282.46) circle (  1.49);

\path[fill=fillColor] (479.49,282.12) circle (  1.49);

\path[fill=fillColor] (477.51,282.73) circle (  1.49);

\path[fill=fillColor] (476.94,282.45) circle (  1.49);

\path[fill=fillColor] (476.98,282.43) circle (  1.49);

\path[fill=fillColor] (474.13,280.30) circle (  1.49);

\path[fill=fillColor] (477.11,282.73) circle (  1.49);

\path[fill=fillColor] (477.14,282.47) circle (  1.49);

\path[fill=fillColor] (479.82,282.04) circle (  1.49);

\path[fill=fillColor] (478.50,283.08) circle (  1.49);

\path[fill=fillColor] (477.49,282.45) circle (  1.49);

\path[fill=fillColor] (478.42,283.02) circle (  1.49);

\path[fill=fillColor] (475.82,282.43) circle (  1.49);

\path[fill=fillColor] (477.82,279.25) circle (  1.49);

\path[fill=fillColor] (476.95,282.45) circle (  1.49);

\path[fill=fillColor] (477.05,282.54) circle (  1.49);

\path[fill=fillColor] (478.09,282.85) circle (  1.49);

\path[fill=fillColor] (475.63,281.60) circle (  1.49);

\path[fill=fillColor] (477.25,282.67) circle (  1.49);

\path[fill=fillColor] (480.86,282.11) circle (  1.49);

\path[fill=fillColor] (476.96,282.46) circle (  1.49);

\path[fill=fillColor] (459.08,268.25) circle (  1.49);

\path[fill=fillColor] (476.71,282.38) circle (  1.49);

\path[fill=fillColor] (476.55,282.31) circle (  1.49);

\path[fill=fillColor] (476.88,282.43) circle (  1.49);

\path[fill=fillColor] (477.45,282.46) circle (  1.49);

\path[fill=fillColor] (478.98,282.59) circle (  1.49);

\path[fill=fillColor] (476.97,282.57) circle (  1.49);

\path[fill=fillColor] (482.57,273.83) circle (  1.49);

\path[fill=fillColor] (477.52,282.47) circle (  1.49);

\path[fill=fillColor] (476.74,282.74) circle (  1.49);

\path[fill=fillColor] (472.31,277.07) circle (  1.49);

\path[fill=fillColor] (513.66,265.80) circle (  1.49);

\path[fill=fillColor] (476.37,282.13) circle (  1.49);

\path[fill=fillColor] (484.58,281.67) circle (  1.49);

\path[fill=fillColor] (479.53,280.12) circle (  1.49);

\path[fill=fillColor] (476.67,282.32) circle (  1.49);

\path[fill=fillColor] (476.73,282.63) circle (  1.49);

\path[fill=fillColor] (477.02,282.42) circle (  1.49);

\path[fill=fillColor] (478.44,282.57) circle (  1.49);

\path[fill=fillColor] (476.92,282.34) circle (  1.49);

\path[fill=fillColor] (481.29,282.46) circle (  1.49);

\path[fill=fillColor] (476.47,280.13) circle (  1.49);

\path[fill=fillColor] (476.63,272.47) circle (  1.49);

\path[fill=fillColor] (477.02,282.41) circle (  1.49);

\path[fill=fillColor] (460.99,262.09) circle (  1.49);

\path[fill=fillColor] (473.50,279.78) circle (  1.49);

\path[fill=fillColor] (476.09,281.84) circle (  1.49);

\path[fill=fillColor] (476.57,282.42) circle (  1.49);

\path[fill=fillColor] (476.78,282.21) circle (  1.49);

\path[fill=fillColor] (476.92,282.43) circle (  1.49);

\path[fill=fillColor] (477.02,282.44) circle (  1.49);

\path[fill=fillColor] (477.44,282.36) circle (  1.49);

\path[fill=fillColor] (477.43,282.44) circle (  1.49);

\path[fill=fillColor] (476.82,282.49) circle (  1.49);

\path[fill=fillColor] (468.31,266.38) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (489.72,354.55) {Coverage: 97\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (536.13,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (417.91,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (447.47,233.11) -- (447.47,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.02,233.11) -- (477.02,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (506.58,233.11) -- (506.58,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.13,233.11) -- (536.13,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (417.91,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (447.47,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (477.02,218.86) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (506.58,218.86) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (536.13,218.86) {2.0};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (417.91,331.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (413.95,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,257.78) -- (413.95,257.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,282.44) -- (413.95,282.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,307.11) -- (413.95,307.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,331.77) -- (413.95,331.77);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,255.50) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,280.17) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,304.83) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,329.50) {2.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (192.72,192.72);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,173.75) {\bfseries Branch rate variance};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28, 10.30) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,100.32) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47, 40.39) rectangle (176.09,160.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47, 40.39) --
	(176.09,160.25);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.12, 41.21) --
	( 36.12, 44.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.55, 40.39) --
	( 32.55, 40.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.39, 40.88) --
	( 33.39, 42.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.60, 40.42) --
	( 32.60, 40.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.62, 40.43) --
	( 32.62, 41.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.93, 40.58) --
	( 32.93, 41.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.57, 40.47) --
	( 32.57, 40.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.09, 40.60) --
	( 33.09, 41.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.84, 41.17) --
	( 34.84, 46.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.88, 40.61) --
	( 32.88, 41.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.51, 40.39) --
	( 32.51, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.60, 40.41) --
	( 32.60, 40.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.59, 40.45) --
	( 32.59, 40.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.09, 40.62) --
	( 33.09, 41.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.91, 40.58) --
	( 32.91, 41.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.58, 40.40) --
	( 32.58, 40.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.45, 40.67) --
	( 33.45, 41.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.54, 40.39) --
	( 32.54, 40.65);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.09, 40.55) --
	( 33.09, 40.88);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.52, 40.40) --
	( 32.52, 40.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.56, 40.39) --
	( 32.56, 40.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.44, 40.48) --
	( 45.44, 58.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 41.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.82, 40.50) --
	( 32.82, 41.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.40, 40.72) --
	( 33.40, 41.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.24, 40.67) --
	( 33.24, 41.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.10, 40.57) --
	( 33.10, 41.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.53, 40.39) --
	( 32.53, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.53, 40.43) --
	( 32.53, 40.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.53, 40.39) --
	( 32.53, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.83, 40.97) --
	( 33.83, 42.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.52, 40.39) --
	( 32.52, 40.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.60, 40.48) --
	( 32.60, 40.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.71, 40.45) --
	( 32.71, 40.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.92, 40.62) --
	( 35.92, 55.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.53, 40.39) --
	( 32.53, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.55, 40.39) --
	( 32.55, 40.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.56, 40.41) --
	( 32.56, 40.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.14, 40.40) --
	(121.14,316.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.59, 40.44) --
	( 32.59, 40.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.55, 40.91) --
	( 33.55, 43.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.81, 40.45) --
	( 48.81, 58.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.62, 40.48) --
	( 32.62, 40.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.96, 40.45) --
	( 32.96, 41.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.09, 41.13) --
	( 34.09, 44.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.91, 40.65) --
	( 33.91, 41.38);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.75, 40.55) --
	( 32.75, 40.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.89, 40.58) --
	( 32.89, 41.06);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 53.60, 41.22) --
	( 53.60, 45.23);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.65, 40.40) --
	( 35.65, 65.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.07, 40.71) --
	( 33.07, 41.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.80, 40.63) --
	( 32.80, 41.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.54, 40.40) --
	( 32.54, 40.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.22, 41.13) --
	( 41.22, 55.76);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 36.12, 42.75) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.55, 40.45) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 33.39, 41.39) circle (  1.49);

\path[fill=fillColor] ( 32.60, 40.52) circle (  1.49);

\path[fill=fillColor] ( 32.62, 40.80) circle (  1.49);

\path[fill=fillColor] ( 32.93, 40.79) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.57, 40.66) circle (  1.49);

\path[fill=fillColor] ( 33.09, 40.88) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.46) circle (  1.49);

\path[fill=fillColor] ( 34.84, 43.33) circle (  1.49);

\path[fill=fillColor] ( 32.88, 40.81) circle (  1.49);

\path[fill=fillColor] ( 32.51, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.60, 40.49) circle (  1.49);

\path[fill=fillColor] ( 32.59, 40.58) circle (  1.49);

\path[fill=fillColor] ( 33.09, 40.87) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.91, 40.79) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.58, 40.45) circle (  1.49);

\path[fill=fillColor] ( 33.45, 40.99) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.54, 40.52) circle (  1.49);

\path[fill=fillColor] ( 33.09, 40.70) circle (  1.49);

\path[fill=fillColor] ( 32.52, 40.45) circle (  1.49);

\path[fill=fillColor] ( 32.56, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 45.44, 48.28) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.58) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.82, 40.73) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.46) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 33.40, 41.14) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 33.24, 40.95) circle (  1.49);

\path[fill=fillColor] ( 33.10, 40.94) circle (  1.49);

\path[fill=fillColor] ( 32.53, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.53, 40.52) circle (  1.49);

\path[fill=fillColor] ( 32.53, 40.41) circle (  1.49);

\path[fill=fillColor] ( 33.83, 41.57) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.52, 40.47) circle (  1.49);

\path[fill=fillColor] ( 32.60, 40.57) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.71, 40.59) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 35.92, 46.66) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.53, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.55, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.56, 40.45) circle (  1.49);

\path[fill=fillColor] (121.14,120.30) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.59, 40.50) circle (  1.49);

\path[fill=fillColor] ( 33.55, 41.85) circle (  1.49);

\path[fill=fillColor] ( 48.81, 47.34) circle (  1.49);

\path[fill=fillColor] ( 32.62, 40.61) circle (  1.49);

\path[fill=fillColor] ( 32.96, 40.91) circle (  1.49);

\path[fill=fillColor] ( 34.09, 42.40) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 33.91, 40.97) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.75, 40.69) circle (  1.49);

\path[fill=fillColor] ( 32.89, 40.79) circle (  1.49);

\path[fill=fillColor] ( 53.60, 42.87) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.41) circle (  1.49);

\path[fill=fillColor] ( 35.65, 49.15) circle (  1.49);

\path[fill=fillColor] ( 33.07, 41.07) circle (  1.49);

\path[fill=fillColor] ( 32.80, 40.86) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.39) circle (  1.49);

\path[fill=fillColor] ( 32.54, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.41) circle (  1.49);

\path[fill=fillColor] ( 41.22, 46.77) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,161.83) {Coverage: 97\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- (159.00, 40.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 32.47, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.78, 40.39) -- ( 57.78, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.08, 40.39) -- ( 83.08, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (108.39, 40.39) -- (108.39, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.70, 40.39) -- (133.70, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.00, 40.39) -- (159.00, 36.43);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47, 26.14) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 57.78, 26.14) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 83.08, 26.14) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (108.39, 26.14) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (133.70, 26.14) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (159.00, 26.14) {25};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 32.47,145.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 28.51, 40.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 61.51) -- ( 28.51, 61.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 82.63) -- ( 28.51, 82.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,103.75) -- ( 28.51,103.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,124.87) -- ( 28.51,124.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,145.99) -- ( 28.51,145.99);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 38.12) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 59.24) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 80.36) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,101.48) {15};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,122.60) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,143.72) {25};
\end{scope}
\end{tikzpicture}

\end{document}