### 2020-09-01 RatesWCSS/WCSS_real_adaptive.tex

% Created by tikzDevice version 0.12.3.1 on 2020-09-01 10:39:43
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,385.44) rectangle (192.72,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,559.19) {\bfseries Tree height $h$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47,425.83) rectangle (176.09,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47,425.83) --
	(176.09,545.69);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.62,447.90) --
	( 60.62,450.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.63,431.21) --
	( 40.63,436.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.15,448.77) --
	( 61.15,451.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.13,443.42) --
	( 56.13,446.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 71.79,457.11) --
	( 71.79,460.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.78,439.55) --
	( 49.78,440.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.12,433.88) --
	( 44.12,442.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.60,429.89) --
	( 37.60,430.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.83,426.82) --
	( 33.83,427.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.42,433.74) --
	( 42.42,436.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.56,430.24) --
	( 38.56,430.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.22,435.14) --
	( 45.22,436.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.61,429.07) --
	( 36.61,429.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.06,479.32) --
	(114.06,534.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.60,435.36) --
	( 44.60,436.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.42,432.57) --
	( 41.42,433.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.02,438.12) --
	( 58.02,455.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.55,439.71) --
	( 51.55,444.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.30,428.99) --
	( 36.30,429.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.47,429.36) --
	( 37.47,430.05);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.59,436.83) --
	( 49.59,438.93);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.94,433.47) --
	( 48.94,445.44);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.30,443.62) --
	( 56.30,445.44);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.67,451.52) --
	( 70.67,465.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.89,439.99) --
	( 50.89,441.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.87,433.77) --
	( 42.87,436.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.84,430.08) --
	( 38.84,432.07);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.92,442.96) --
	( 52.92,446.30);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.33,456.68) --
	( 70.33,459.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.17,431.83) --
	( 40.17,434.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.89,436.29) --
	( 45.89,445.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.45,452.95) --
	( 68.45,460.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.00,426.96) --
	( 34.00,427.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.29,439.94) --
	( 50.29,441.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.92,431.88) --
	( 39.92,432.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.17,430.85) --
	( 39.17,431.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.47,429.60) --
	( 44.47,437.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.42,429.86) --
	( 37.42,430.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.48,448.52) --
	( 61.48,452.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.09,440.96) --
	( 51.09,442.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.67,450.80) --
	( 63.67,452.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.54,426.68) --
	( 33.54,427.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.13,438.63) --
	( 49.13,440.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.43,436.37) --
	( 45.43,437.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.62,455.88) --
	( 70.62,459.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.23,429.96) --
	( 38.23,432.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.18,432.06) --
	( 41.18,433.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.21,436.78) --
	( 46.21,438.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.60,435.18) --
	( 44.60,436.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.08,430.11) --
	( 41.08,433.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.16,431.01) --
	( 39.16,431.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.88,436.62) --
	( 45.88,438.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.99,431.56) --
	( 41.99,436.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.91,427.24) --
	( 34.91,428.07);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.75,428.05) --
	( 35.75,428.51);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.44,432.45) --
	( 40.44,433.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.74,460.74) --
	( 78.74,466.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.59,435.56) --
	( 45.59,445.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.94,429.22) --
	( 36.94,429.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.86,449.07) --
	( 60.86,451.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.28,429.33) --
	( 37.28,430.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 55.33,443.78) --
	( 55.33,445.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.65,435.83) --
	( 45.65,437.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.45,432.03) --
	( 42.45,435.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.44,467.93) --
	( 87.44,473.74);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.68,427.84) --
	( 40.68,431.66);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.01,430.98) --
	( 39.01,431.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 55.16,444.15) --
	( 55.16,446.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.05,439.46) --
	( 49.05,440.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.38,433.66) --
	( 42.38,434.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.52,432.24) --
	( 40.52,433.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.23,444.08) --
	( 56.23,447.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.52,426.73) --
	( 42.52,438.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.44,433.94) --
	( 42.44,434.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.59,454.67) --
	( 68.59,460.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.25,427.65) --
	( 35.25,429.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.79,430.34) --
	( 44.79,443.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.99,431.45) --
	( 39.99,433.12);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.61,427.52) --
	( 35.61,428.31);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.98,432.73) --
	( 42.98,439.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.83,434.98) --
	( 43.83,436.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.72,432.35) --
	( 40.72,433.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.51,441.13) --
	( 51.51,442.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.43,431.36) --
	( 40.43,436.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.97,430.04) --
	( 37.97,430.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.25,450.84) --
	( 65.25,456.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.44,433.94) --
	( 43.44,437.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.23,428.14) --
	( 36.23,431.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.23,440.87) --
	( 51.23,442.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.18,426.82) --
	( 36.18,430.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.15,432.34) --
	( 42.15,438.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.83,434.97) --
	( 44.83,439.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.43,430.38) --
	( 38.43,431.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.64,427.43) --
	( 34.64,427.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87,432.33) --
	( 40.87,433.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.23,449.54) --
	( 63.23,452.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.64,453.36) --
	( 66.64,456.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.50,428.09) --
	( 35.50,428.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 36.85,429.30) --
	( 36.85,429.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.90,440.19) --
	( 59.90,455.62);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 60.62,448.95) circle (  1.49);

\path[fill=fillColor] ( 40.63,433.67) circle (  1.49);

\path[fill=fillColor] ( 61.15,449.89) circle (  1.49);

\path[fill=fillColor] ( 56.13,445.15) circle (  1.49);

\path[fill=fillColor] ( 71.79,458.70) circle (  1.49);

\path[fill=fillColor] ( 49.78,440.24) circle (  1.49);

\path[fill=fillColor] ( 44.12,438.26) circle (  1.49);

\path[fill=fillColor] ( 37.60,430.19) circle (  1.49);

\path[fill=fillColor] ( 33.83,427.01) circle (  1.49);

\path[fill=fillColor] ( 42.42,435.05) circle (  1.49);

\path[fill=fillColor] ( 38.56,430.58) circle (  1.49);

\path[fill=fillColor] ( 45.22,436.07) circle (  1.49);

\path[fill=fillColor] ( 36.61,429.35) circle (  1.49);

\path[fill=fillColor] (114.06,505.74) circle (  1.49);

\path[fill=fillColor] ( 44.60,435.86) circle (  1.49);

\path[fill=fillColor] ( 41.42,433.20) circle (  1.49);

\path[fill=fillColor] ( 58.02,445.65) circle (  1.49);

\path[fill=fillColor] ( 51.55,442.14) circle (  1.49);

\path[fill=fillColor] ( 36.30,429.31) circle (  1.49);

\path[fill=fillColor] ( 37.47,429.70) circle (  1.49);

\path[fill=fillColor] ( 49.59,437.86) circle (  1.49);

\path[fill=fillColor] ( 48.94,439.04) circle (  1.49);

\path[fill=fillColor] ( 56.30,444.51) circle (  1.49);

\path[fill=fillColor] ( 70.67,458.33) circle (  1.49);

\path[fill=fillColor] ( 50.89,440.77) circle (  1.49);

\path[fill=fillColor] ( 42.87,434.90) circle (  1.49);

\path[fill=fillColor] ( 38.84,431.03) circle (  1.49);

\path[fill=fillColor] ( 52.92,444.67) circle (  1.49);

\path[fill=fillColor] ( 70.33,458.30) circle (  1.49);

\path[fill=fillColor] ( 40.17,433.09) circle (  1.49);

\path[fill=fillColor] ( 45.89,440.68) circle (  1.49);

\path[fill=fillColor] ( 68.45,456.85) circle (  1.49);

\path[fill=fillColor] ( 34.00,427.15) circle (  1.49);

\path[fill=fillColor] ( 50.29,440.87) circle (  1.49);

\path[fill=fillColor] ( 39.92,432.28) circle (  1.49);

\path[fill=fillColor] ( 39.17,431.15) circle (  1.49);

\path[fill=fillColor] ( 44.47,433.00) circle (  1.49);

\path[fill=fillColor] ( 37.42,430.24) circle (  1.49);

\path[fill=fillColor] ( 61.48,450.45) circle (  1.49);

\path[fill=fillColor] ( 51.09,441.67) circle (  1.49);

\path[fill=fillColor] ( 63.67,451.93) circle (  1.49);

\path[fill=fillColor] ( 33.54,426.85) circle (  1.49);

\path[fill=fillColor] ( 49.13,439.44) circle (  1.49);

\path[fill=fillColor] ( 45.43,436.96) circle (  1.49);

\path[fill=fillColor] ( 70.62,457.57) circle (  1.49);

\path[fill=fillColor] ( 38.23,431.00) circle (  1.49);

\path[fill=fillColor] ( 41.18,432.60) circle (  1.49);

\path[fill=fillColor] ( 46.21,437.69) circle (  1.49);

\path[fill=fillColor] ( 44.60,435.64) circle (  1.49);

\path[fill=fillColor] ( 41.08,431.57) circle (  1.49);

\path[fill=fillColor] ( 39.16,431.37) circle (  1.49);

\path[fill=fillColor] ( 45.88,437.44) circle (  1.49);

\path[fill=fillColor] ( 41.99,433.90) circle (  1.49);

\path[fill=fillColor] ( 34.91,427.62) circle (  1.49);

\path[fill=fillColor] ( 35.75,428.28) circle (  1.49);

\path[fill=fillColor] ( 40.44,432.83) circle (  1.49);

\path[fill=fillColor] ( 78.74,463.74) circle (  1.49);

\path[fill=fillColor] ( 45.59,440.02) circle (  1.49);

\path[fill=fillColor] ( 36.94,429.56) circle (  1.49);

\path[fill=fillColor] ( 60.86,450.27) circle (  1.49);

\path[fill=fillColor] ( 37.28,429.85) circle (  1.49);

\path[fill=fillColor] ( 55.33,444.88) circle (  1.49);

\path[fill=fillColor] ( 45.65,436.65) circle (  1.49);

\path[fill=fillColor] ( 42.45,433.68) circle (  1.49);

\path[fill=fillColor] ( 87.44,470.83) circle (  1.49);

\path[fill=fillColor] ( 40.68,429.56) circle (  1.49);

\path[fill=fillColor] ( 39.01,431.36) circle (  1.49);

\path[fill=fillColor] ( 55.16,445.15) circle (  1.49);

\path[fill=fillColor] ( 49.05,440.06) circle (  1.49);

\path[fill=fillColor] ( 42.38,434.25) circle (  1.49);

\path[fill=fillColor] ( 40.52,432.70) circle (  1.49);

\path[fill=fillColor] ( 56.23,445.77) circle (  1.49);

\path[fill=fillColor] ( 42.52,430.91) circle (  1.49);

\path[fill=fillColor] ( 42.44,434.40) circle (  1.49);

\path[fill=fillColor] ( 68.59,457.59) circle (  1.49);

\path[fill=fillColor] ( 35.25,428.35) circle (  1.49);

\path[fill=fillColor] ( 44.79,435.83) circle (  1.49);

\path[fill=fillColor] ( 39.99,432.24) circle (  1.49);

\path[fill=fillColor] ( 35.61,427.87) circle (  1.49);

\path[fill=fillColor] ( 42.98,435.83) circle (  1.49);

\path[fill=fillColor] ( 43.83,435.51) circle (  1.49);

\path[fill=fillColor] ( 40.72,432.76) circle (  1.49);

\path[fill=fillColor] ( 51.51,441.88) circle (  1.49);

\path[fill=fillColor] ( 40.43,433.64) circle (  1.49);

\path[fill=fillColor] ( 37.97,430.32) circle (  1.49);

\path[fill=fillColor] ( 65.25,453.48) circle (  1.49);

\path[fill=fillColor] ( 43.44,435.84) circle (  1.49);

\path[fill=fillColor] ( 36.23,429.64) circle (  1.49);

\path[fill=fillColor] ( 51.23,441.61) circle (  1.49);

\path[fill=fillColor] ( 36.18,428.50) circle (  1.49);

\path[fill=fillColor] ( 42.15,434.94) circle (  1.49);

\path[fill=fillColor] ( 44.83,437.14) circle (  1.49);

\path[fill=fillColor] ( 38.43,430.79) circle (  1.49);

\path[fill=fillColor] ( 34.64,427.67) circle (  1.49);

\path[fill=fillColor] ( 40.87,432.71) circle (  1.49);

\path[fill=fillColor] ( 63.23,451.26) circle (  1.49);

\path[fill=fillColor] ( 66.64,454.92) circle (  1.49);

\path[fill=fillColor] ( 35.50,428.38) circle (  1.49);

\path[fill=fillColor] ( 36.85,429.58) circle (  1.49);

\path[fill=fillColor] ( 59.90,447.00) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,547.27) {Coverage: 94\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- (159.13,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 32.47,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.80,425.83) -- ( 57.80,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.14,425.83) -- ( 83.14,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (108.47,425.83) -- (108.47,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.80,425.83) -- (133.80,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.13,425.83) -- (159.13,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47,411.58) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 57.80,411.58) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 83.14,411.58) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (108.47,411.58) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (133.80,411.58) {2.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (159.13,411.58) {2.5};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 32.47,531.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,425.83) -- ( 28.51,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,446.97) -- ( 28.51,446.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,468.11) -- ( 28.51,468.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,489.25) -- ( 28.51,489.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,510.39) -- ( 28.51,510.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,531.54) -- ( 28.51,531.54);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,423.56) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,444.70) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,465.84) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,486.98) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,508.12) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,529.26) {2.5};
\end{scope}
\begin{scope}
\path[clip] (192.72,385.44) rectangle (385.44,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (297.00,559.19) {\bfseries Yule birth rate $\lambda$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (297.00,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (199.85,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (225.19,425.83) rectangle (368.81,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (225.19,425.83) --
	(368.81,545.69);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.08,428.69) --
	(233.08,431.85);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.22,431.73) --
	(240.22,440.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.85,428.04) --
	(228.85,430.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.51,428.90) --
	(231.51,432.42);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.90,427.17) --
	(228.90,428.63);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.82,430.33) --
	(232.82,435.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.27,431.70) --
	(238.27,438.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.04,435.27) --
	(241.04,445.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (288.80,459.50) --
	(288.80,500.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.60,431.55) --
	(237.60,438.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.36,434.91) --
	(241.36,445.49);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.29,431.65) --
	(240.29,438.16);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.62,440.55) --
	(245.62,457.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.78,426.67) --
	(226.78,427.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.47,429.93) --
	(231.47,434.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.87,434.57) --
	(241.87,444.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.24,428.37) --
	(230.24,432.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.40,429.16) --
	(232.40,432.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.90,442.27) --
	(264.90,460.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.31,435.78) --
	(247.31,447.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.05,430.52) --
	(235.05,435.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.48,430.16) --
	(234.48,435.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.81,428.56) --
	(229.81,431.54);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.34,428.14) --
	(231.34,430.91);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.09,429.97) --
	(234.09,434.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.56,430.71) --
	(234.56,436.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.00,435.25) --
	(240.00,446.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.24,429.71) --
	(231.24,434.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.54,427.36) --
	(228.54,429.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.33,435.30) --
	(242.33,446.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.75,429.12) --
	(230.75,432.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.92,428.38) --
	(229.92,431.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.05,465.13) --
	(310.05,511.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.27,429.16) --
	(232.27,432.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.10,433.14) --
	(239.10,441.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.49,433.11) --
	(242.49,441.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.99,431.17) --
	(231.99,443.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.90,435.81) --
	(246.90,447.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.49,428.58) --
	(230.49,431.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.69,429.44) --
	(231.69,433.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.75,427.90) --
	(229.75,430.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (289.81,476.34) --
	(289.81,536.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.46,430.85) --
	(233.46,436.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.63,431.03) --
	(235.63,436.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.99,427.33) --
	(227.99,428.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.60,439.32) --
	(253.60,455.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.48,433.67) --
	(240.48,442.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.60,430.46) --
	(235.60,435.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.97,431.47) --
	(235.97,437.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.33,434.49) --
	(240.33,445.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (244.91,435.12) --
	(244.91,445.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.68,429.86) --
	(233.68,434.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.21,431.92) --
	(237.21,438.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (258.87,448.93) --
	(258.87,477.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.77,447.38) --
	(262.77,472.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.68,432.45) --
	(237.68,439.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.98,427.63) --
	(228.98,429.64);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.43,429.72) --
	(236.43,434.71);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.90,437.61) --
	(243.90,451.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.89,427.92) --
	(228.89,430.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.10,440.05) --
	(248.10,456.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.86,428.17) --
	(230.86,430.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.61,430.07) --
	(232.61,434.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.83,432.02) --
	(234.83,439.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.77,426.97) --
	(227.77,428.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.39,436.88) --
	(246.39,461.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.10,433.10) --
	(238.10,441.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.95,428.28) --
	(229.95,431.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.83,429.01) --
	(231.83,432.49);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.66,431.39) --
	(239.66,437.70);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.50,431.56) --
	(234.50,438.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.79,428.52) --
	(231.79,431.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.48,427.67) --
	(241.48,477.39);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.64,431.90) --
	(240.64,438.64);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.79,427.38) --
	(227.79,429.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (256.03,441.94) --
	(256.03,463.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.78,429.47) --
	(238.78,439.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.49,432.86) --
	(239.49,441.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (258.23,451.93) --
	(258.23,484.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.75,431.46) --
	(235.75,439.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.52,431.17) --
	(237.52,437.07);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.68,432.30) --
	(241.68,439.57);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.33,428.99) --
	(232.33,432.50);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.95,431.76) --
	(243.95,440.12);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.92,435.64) --
	(241.92,446.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.90,427.50) --
	(227.90,429.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.72,430.69) --
	(235.72,436.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.79,439.70) --
	(253.79,461.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.23,428.16) --
	(229.23,430.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.44,437.43) --
	(255.44,488.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.39,430.78) --
	(234.39,436.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.22,430.13) --
	(235.22,435.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.36,437.90) --
	(249.36,451.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.32,453.16) --
	(262.32,485.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.37,432.62) --
	(240.37,440.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.03,428.82) --
	(229.03,432.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.52,427.48) --
	(228.52,429.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (277.09,448.78) --
	(277.09,475.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.39,438.59) --
	(243.39,453.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.45,428.42) --
	(229.45,431.95);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (233.08,430.24) circle (  1.49);

\path[fill=fillColor] (240.22,435.94) circle (  1.49);

\path[fill=fillColor] (228.85,429.19) circle (  1.49);

\path[fill=fillColor] (231.51,430.64) circle (  1.49);

\path[fill=fillColor] (228.90,427.89) circle (  1.49);

\path[fill=fillColor] (232.82,432.85) circle (  1.49);

\path[fill=fillColor] (238.27,435.17) circle (  1.49);

\path[fill=fillColor] (241.04,440.47) circle (  1.49);

\path[fill=fillColor] (288.80,479.45) circle (  1.49);

\path[fill=fillColor] (237.60,434.83) circle (  1.49);

\path[fill=fillColor] (241.36,440.20) circle (  1.49);

\path[fill=fillColor] (240.29,434.81) circle (  1.49);

\path[fill=fillColor] (245.62,448.40) circle (  1.49);

\path[fill=fillColor] (226.78,427.15) circle (  1.49);

\path[fill=fillColor] (231.47,432.14) circle (  1.49);

\path[fill=fillColor] (241.87,439.48) circle (  1.49);

\path[fill=fillColor] (230.24,430.33) circle (  1.49);

\path[fill=fillColor] (232.40,431.01) circle (  1.49);

\path[fill=fillColor] (264.90,450.98) circle (  1.49);

\path[fill=fillColor] (247.31,441.51) circle (  1.49);

\path[fill=fillColor] (235.05,433.13) circle (  1.49);

\path[fill=fillColor] (234.48,432.83) circle (  1.49);

\path[fill=fillColor] (229.81,430.00) circle (  1.49);

\path[fill=fillColor] (231.34,429.48) circle (  1.49);

\path[fill=fillColor] (234.09,432.21) circle (  1.49);

\path[fill=fillColor] (234.56,433.36) circle (  1.49);

\path[fill=fillColor] (240.00,440.67) circle (  1.49);

\path[fill=fillColor] (231.24,431.81) circle (  1.49);

\path[fill=fillColor] (228.54,428.14) circle (  1.49);

\path[fill=fillColor] (242.33,440.62) circle (  1.49);

\path[fill=fillColor] (230.75,430.95) circle (  1.49);

\path[fill=fillColor] (229.92,429.74) circle (  1.49);

\path[fill=fillColor] (310.05,486.51) circle (  1.49);

\path[fill=fillColor] (232.27,431.01) circle (  1.49);

\path[fill=fillColor] (239.10,437.30) circle (  1.49);

\path[fill=fillColor] (242.49,437.21) circle (  1.49);

\path[fill=fillColor] (231.99,436.81) circle (  1.49);

\path[fill=fillColor] (246.90,441.32) circle (  1.49);

\path[fill=fillColor] (230.49,430.03) circle (  1.49);

\path[fill=fillColor] (231.69,431.32) circle (  1.49);

\path[fill=fillColor] (229.75,428.97) circle (  1.49);

\path[fill=fillColor] (289.81,505.74) circle (  1.49);

\path[fill=fillColor] (233.46,433.62) circle (  1.49);

\path[fill=fillColor] (235.63,433.80) circle (  1.49);

\path[fill=fillColor] (227.99,428.12) circle (  1.49);

\path[fill=fillColor] (253.60,447.48) circle (  1.49);

\path[fill=fillColor] (240.48,437.95) circle (  1.49);

\path[fill=fillColor] (235.60,433.02) circle (  1.49);

\path[fill=fillColor] (235.97,434.58) circle (  1.49);

\path[fill=fillColor] (240.33,439.95) circle (  1.49);

\path[fill=fillColor] (244.91,440.52) circle (  1.49);

\path[fill=fillColor] (233.68,432.06) circle (  1.49);

\path[fill=fillColor] (237.21,435.22) circle (  1.49);

\path[fill=fillColor] (258.87,462.43) circle (  1.49);

\path[fill=fillColor] (262.77,459.44) circle (  1.49);

\path[fill=fillColor] (237.68,436.12) circle (  1.49);

\path[fill=fillColor] (228.98,428.62) circle (  1.49);

\path[fill=fillColor] (236.43,432.11) circle (  1.49);

\path[fill=fillColor] (243.90,444.12) circle (  1.49);

\path[fill=fillColor] (228.89,429.10) circle (  1.49);

\path[fill=fillColor] (248.10,447.68) circle (  1.49);

\path[fill=fillColor] (230.86,429.43) circle (  1.49);

\path[fill=fillColor] (232.61,432.43) circle (  1.49);

\path[fill=fillColor] (234.83,435.45) circle (  1.49);

\path[fill=fillColor] (227.77,427.56) circle (  1.49);

\path[fill=fillColor] (246.39,448.70) circle (  1.49);

\path[fill=fillColor] (238.10,437.33) circle (  1.49);

\path[fill=fillColor] (229.95,429.63) circle (  1.49);

\path[fill=fillColor] (231.83,430.65) circle (  1.49);

\path[fill=fillColor] (239.66,434.50) circle (  1.49);

\path[fill=fillColor] (234.50,434.71) circle (  1.49);

\path[fill=fillColor] (231.79,430.01) circle (  1.49);

\path[fill=fillColor] (241.48,450.07) circle (  1.49);

\path[fill=fillColor] (240.64,435.13) circle (  1.49);

\path[fill=fillColor] (227.79,428.18) circle (  1.49);

\path[fill=fillColor] (256.03,452.14) circle (  1.49);

\path[fill=fillColor] (238.78,434.37) circle (  1.49);

\path[fill=fillColor] (239.49,437.04) circle (  1.49);

\path[fill=fillColor] (258.23,467.75) circle (  1.49);

\path[fill=fillColor] (235.75,435.17) circle (  1.49);

\path[fill=fillColor] (237.52,433.95) circle (  1.49);

\path[fill=fillColor] (241.68,435.80) circle (  1.49);

\path[fill=fillColor] (232.33,430.71) circle (  1.49);

\path[fill=fillColor] (243.95,435.87) circle (  1.49);

\path[fill=fillColor] (241.92,441.03) circle (  1.49);

\path[fill=fillColor] (227.90,428.40) circle (  1.49);

\path[fill=fillColor] (235.72,433.54) circle (  1.49);

\path[fill=fillColor] (253.79,450.13) circle (  1.49);

\path[fill=fillColor] (229.23,429.46) circle (  1.49);

\path[fill=fillColor] (255.44,461.58) circle (  1.49);

\path[fill=fillColor] (234.39,433.74) circle (  1.49);

\path[fill=fillColor] (235.22,432.58) circle (  1.49);

\path[fill=fillColor] (249.36,444.28) circle (  1.49);

\path[fill=fillColor] (262.32,468.98) circle (  1.49);

\path[fill=fillColor] (240.37,436.16) circle (  1.49);

\path[fill=fillColor] (229.03,430.41) circle (  1.49);

\path[fill=fillColor] (228.52,428.36) circle (  1.49);

\path[fill=fillColor] (277.09,462.38) circle (  1.49);

\path[fill=fillColor] (243.39,445.61) circle (  1.49);

\path[fill=fillColor] (229.45,430.05) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (297.00,547.27) {Coverage: 91\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (344.58,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (225.19,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.99,425.83) -- (264.99,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.78,425.83) -- (304.78,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (344.58,425.83) -- (344.58,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (225.19,411.58) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (264.99,411.58) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (304.78,411.58) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (344.58,411.58) {150};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (225.19,525.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,425.83) -- (221.23,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,459.04) -- (221.23,459.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,492.25) -- (221.23,492.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,525.47) -- (221.23,525.47);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,423.56) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,456.77) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,489.98) {100};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,523.19) {150};
\end{scope}
\begin{scope}
\path[clip] (385.44,385.44) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (489.72,559.19) {\bfseries HKY model $\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (489.72,395.74) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (392.57,485.76) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (417.91,425.83) rectangle (561.53,545.69);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (417.91,425.83) --
	(561.53,545.69);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.09,426.81) --
	(419.09,426.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (468.00,461.66) --
	(468.00,472.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.67,432.69) --
	(426.67,433.36);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.83,426.48) --
	(418.83,426.59);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.96,438.30) --
	(433.96,439.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.79,428.09) --
	(420.79,428.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.93,429.73) --
	(422.93,430.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.21,430.73) --
	(424.21,431.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (441.53,443.53) --
	(441.53,453.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.46,427.01) --
	(419.46,427.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.88,428.85) --
	(421.88,429.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.44,427.90) --
	(420.44,428.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.58,430.69) --
	(424.58,431.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.67,428.88) --
	(421.67,429.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.41,427.07) --
	(419.41,427.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.49,427.92) --
	(420.49,428.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.53,435.11) --
	(429.53,436.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.54,427.93) --
	(420.54,428.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (513.66,482.60) --
	(513.66,519.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.84,427.18) --
	(419.84,427.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.91,427.49) --
	(419.91,427.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.97,429.05) --
	(421.97,429.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.60,429.67) --
	(422.60,430.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.47,436.80) --
	(431.47,438.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.26,430.18) --
	(423.26,430.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.24,426.09) --
	(418.24,426.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.74,428.82) --
	(421.74,429.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.30,432.49) --
	(426.30,433.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.94,426.61) --
	(418.94,426.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.18,428.42) --
	(421.18,428.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.83,428.09) --
	(420.83,428.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.18,426.02) --
	(418.18,426.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.12,430.07) --
	(424.12,431.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.88,429.73) --
	(422.88,430.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.49,429.34) --
	(422.49,429.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.67,428.70) --
	(421.67,429.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.15,426.70) --
	(419.15,426.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.88,432.79) --
	(426.88,434.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.08,431.67) --
	(425.08,432.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.65,428.04) --
	(420.65,428.31);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.59,431.67) --
	(425.59,432.23);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.36,432.05) --
	(427.36,435.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.37,432.86) --
	(427.37,433.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.71,426.44) --
	(418.71,426.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.15,452.12) --
	(452.15,455.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.42,433.32) --
	(428.42,435.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.67,434.33) --
	(429.67,435.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.07,427.53) --
	(420.07,427.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.60,427.87) --
	(420.60,428.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.04,453.13) --
	(453.04,462.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.67,427.22) --
	(419.67,427.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.78,434.36) --
	(428.78,435.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.21,426.83) --
	(419.21,427.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.85,431.89) --
	(425.85,433.84);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.34,427.96) --
	(420.34,428.56);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.89,432.10) --
	(425.89,433.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.60,429.61) --
	(422.60,429.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.12,427.59) --
	(420.12,427.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.74,427.14) --
	(419.74,427.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.23,426.06) --
	(418.23,426.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.72,426.40) --
	(418.72,426.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.10,430.95) --
	(424.10,431.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.98,455.10) --
	(454.98,461.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.78,430.36) --
	(423.78,431.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (449.13,450.43) --
	(449.13,453.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.83,426.42) --
	(418.83,426.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.03,427.41) --
	(420.03,427.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.70,426.42) --
	(418.70,426.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.31,428.58) --
	(421.31,428.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.90,428.33) --
	(420.90,428.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.87,429.69) --
	(422.87,430.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.69,428.78) --
	(421.69,429.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.46,431.23) --
	(425.46,432.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.25,432.68) --
	(426.25,433.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.86,426.53) --
	(418.86,426.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.24,439.25) --
	(435.24,444.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.39,428.51) --
	(421.39,428.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (442.14,444.64) --
	(442.14,449.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (449.66,445.65) --
	(449.66,456.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.42,426.20) --
	(418.42,426.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.78,432.27) --
	(425.78,433.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.66,456.10) --
	(461.66,464.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.65,428.94) --
	(421.65,429.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.85,433.20) --
	(426.85,434.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.33,434.04) --
	(428.33,435.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.83,427.38) --
	(419.83,427.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.07,432.86) --
	(427.07,433.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.90,428.13) --
	(420.90,428.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.42,441.23) --
	(437.42,443.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.71,428.62) --
	(421.71,429.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.48,439.16) --
	(436.48,441.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.24,427.73) --
	(420.24,427.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.88,433.32) --
	(427.88,434.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.52,430.31) --
	(424.52,431.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.79,429.00) --
	(421.79,429.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.39,430.26) --
	(423.39,430.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.10,429.17) --
	(422.10,429.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.27,431.27) --
	(425.27,432.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.36,426.19) --
	(418.36,426.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.29,429.28) --
	(422.29,429.68);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (419.09,426.88) circle (  1.49);

\path[fill=fillColor] (468.00,466.98) circle (  1.49);

\path[fill=fillColor] (426.67,433.02) circle (  1.49);

\path[fill=fillColor] (418.83,426.54) circle (  1.49);

\path[fill=fillColor] (433.96,439.01) circle (  1.49);

\path[fill=fillColor] (420.79,428.24) circle (  1.49);

\path[fill=fillColor] (422.93,430.02) circle (  1.49);

\path[fill=fillColor] (424.21,431.20) circle (  1.49);

\path[fill=fillColor] (441.53,448.54) circle (  1.49);

\path[fill=fillColor] (419.46,427.11) circle (  1.49);

\path[fill=fillColor] (421.88,429.14) circle (  1.49);

\path[fill=fillColor] (420.44,428.06) circle (  1.49);

\path[fill=fillColor] (424.58,431.24) circle (  1.49);

\path[fill=fillColor] (421.67,429.01) circle (  1.49);

\path[fill=fillColor] (419.41,427.16) circle (  1.49);

\path[fill=fillColor] (420.49,428.12) circle (  1.49);

\path[fill=fillColor] (429.53,435.65) circle (  1.49);

\path[fill=fillColor] (420.54,428.05) circle (  1.49);

\path[fill=fillColor] (513.66,499.91) circle (  1.49);

\path[fill=fillColor] (419.84,427.32) circle (  1.49);

\path[fill=fillColor] (419.91,427.61) circle (  1.49);

\path[fill=fillColor] (421.97,429.25) circle (  1.49);

\path[fill=fillColor] (422.60,429.87) circle (  1.49);

\path[fill=fillColor] (431.47,437.46) circle (  1.49);

\path[fill=fillColor] (423.26,430.46) circle (  1.49);

\path[fill=fillColor] (418.24,426.13) circle (  1.49);

\path[fill=fillColor] (421.74,429.08) circle (  1.49);

\path[fill=fillColor] (426.30,432.93) circle (  1.49);

\path[fill=fillColor] (418.94,426.67) circle (  1.49);

\path[fill=fillColor] (421.18,428.66) circle (  1.49);

\path[fill=fillColor] (420.83,428.21) circle (  1.49);

\path[fill=fillColor] (418.18,426.05) circle (  1.49);

\path[fill=fillColor] (424.12,430.90) circle (  1.49);

\path[fill=fillColor] (422.88,429.95) circle (  1.49);

\path[fill=fillColor] (422.49,429.61) circle (  1.49);

\path[fill=fillColor] (421.67,428.93) circle (  1.49);

\path[fill=fillColor] (419.15,426.79) circle (  1.49);

\path[fill=fillColor] (426.88,433.50) circle (  1.49);

\path[fill=fillColor] (425.08,431.98) circle (  1.49);

\path[fill=fillColor] (420.65,428.17) circle (  1.49);

\path[fill=fillColor] (425.59,431.96) circle (  1.49);

\path[fill=fillColor] (427.36,433.57) circle (  1.49);

\path[fill=fillColor] (427.37,433.37) circle (  1.49);

\path[fill=fillColor] (418.71,426.49) circle (  1.49);

\path[fill=fillColor] (452.15,453.71) circle (  1.49);

\path[fill=fillColor] (428.42,434.19) circle (  1.49);

\path[fill=fillColor] (429.67,435.08) circle (  1.49);

\path[fill=fillColor] (420.07,427.65) circle (  1.49);

\path[fill=fillColor] (420.60,428.01) circle (  1.49);

\path[fill=fillColor] (453.04,457.49) circle (  1.49);

\path[fill=fillColor] (419.67,427.35) circle (  1.49);

\path[fill=fillColor] (428.78,434.93) circle (  1.49);

\path[fill=fillColor] (419.21,426.91) circle (  1.49);

\path[fill=fillColor] (425.85,432.83) circle (  1.49);

\path[fill=fillColor] (420.34,428.27) circle (  1.49);

\path[fill=fillColor] (425.89,432.59) circle (  1.49);

\path[fill=fillColor] (422.60,429.80) circle (  1.49);

\path[fill=fillColor] (420.12,427.71) circle (  1.49);

\path[fill=fillColor] (419.74,427.28) circle (  1.49);

\path[fill=fillColor] (418.23,426.09) circle (  1.49);

\path[fill=fillColor] (418.72,426.49) circle (  1.49);

\path[fill=fillColor] (424.10,431.19) circle (  1.49);

\path[fill=fillColor] (454.98,457.98) circle (  1.49);

\path[fill=fillColor] (423.78,430.68) circle (  1.49);

\path[fill=fillColor] (449.13,451.68) circle (  1.49);

\path[fill=fillColor] (418.83,426.51) circle (  1.49);

\path[fill=fillColor] (420.03,427.55) circle (  1.49);

\path[fill=fillColor] (418.70,426.47) circle (  1.49);

\path[fill=fillColor] (421.31,428.73) circle (  1.49);

\path[fill=fillColor] (420.90,428.51) circle (  1.49);

\path[fill=fillColor] (422.87,429.96) circle (  1.49);

\path[fill=fillColor] (421.69,428.94) circle (  1.49);

\path[fill=fillColor] (425.46,431.98) circle (  1.49);

\path[fill=fillColor] (426.25,433.23) circle (  1.49);

\path[fill=fillColor] (418.86,426.58) circle (  1.49);

\path[fill=fillColor] (435.24,441.66) circle (  1.49);

\path[fill=fillColor] (421.39,428.73) circle (  1.49);

\path[fill=fillColor] (442.14,446.77) circle (  1.49);

\path[fill=fillColor] (449.66,450.64) circle (  1.49);

\path[fill=fillColor] (418.42,426.25) circle (  1.49);

\path[fill=fillColor] (425.78,432.76) circle (  1.49);

\path[fill=fillColor] (461.66,460.20) circle (  1.49);

\path[fill=fillColor] (421.65,429.12) circle (  1.49);

\path[fill=fillColor] (426.85,433.79) circle (  1.49);

\path[fill=fillColor] (428.33,434.83) circle (  1.49);

\path[fill=fillColor] (419.83,427.45) circle (  1.49);

\path[fill=fillColor] (427.07,433.35) circle (  1.49);

\path[fill=fillColor] (420.90,428.40) circle (  1.49);

\path[fill=fillColor] (437.42,442.12) circle (  1.49);

\path[fill=fillColor] (421.71,429.08) circle (  1.49);

\path[fill=fillColor] (436.48,440.27) circle (  1.49);

\path[fill=fillColor] (420.24,427.86) circle (  1.49);

\path[fill=fillColor] (427.88,434.12) circle (  1.49);

\path[fill=fillColor] (424.52,430.96) circle (  1.49);

\path[fill=fillColor] (421.79,429.23) circle (  1.49);

\path[fill=fillColor] (423.39,430.51) circle (  1.49);

\path[fill=fillColor] (422.10,429.32) circle (  1.49);

\path[fill=fillColor] (425.27,432.09) circle (  1.49);

\path[fill=fillColor] (418.36,426.25) circle (  1.49);

\path[fill=fillColor] (422.29,429.48) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (489.72,547.27) {Coverage: 97\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (554.47,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (417.91,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.42,425.83) -- (437.42,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.93,425.83) -- (456.93,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.44,425.83) -- (476.44,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.94,425.83) -- (495.94,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.45,425.83) -- (515.45,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (534.96,425.83) -- (534.96,421.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (554.47,425.83) -- (554.47,421.87);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (417.91,411.58) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (437.42,411.58) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (456.93,411.58) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (476.44,411.58) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (495.94,411.58) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (515.45,411.58) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (534.96,411.58) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (554.47,411.58) {70};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (417.91,539.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,425.83) -- (413.95,425.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,442.11) -- (413.95,442.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,458.39) -- (413.95,458.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,474.67) -- (413.95,474.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,490.95) -- (413.95,490.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,507.23) -- (413.95,507.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,523.51) -- (413.95,523.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,539.80) -- (413.95,539.80);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,423.56) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,439.84) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,456.12) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,472.40) {30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,488.68) {40};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,504.96) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,521.24) {60};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,537.52) {70};
\end{scope}
\begin{scope}
\path[clip] (  0.00,192.72) rectangle (192.72,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,366.47) {\bfseries Nucleotide frequency $f$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47,233.11) rectangle (176.09,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47,233.11) --
	(176.09,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.52,255.16) --
	( 59.52,257.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 80.98,270.03) --
	( 80.98,273.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.46,266.60) --
	( 73.46,269.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (106.47,294.56) --
	(106.47,298.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.59,283.32) --
	( 93.59,285.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.32,276.11) --
	( 85.32,280.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.14,298.49) --
	(114.14,303.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.03,298.18) --
	(114.03,303.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.32,285.43) --
	( 99.32,292.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.70,274.91) --
	( 85.70,279.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (119.28,303.70) --
	(119.28,309.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.17,267.05) --
	( 74.17,270.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.64,275.03) --
	( 86.64,280.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.12,268.00) --
	( 76.12,270.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.26,272.84) --
	( 81.26,276.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.22,303.64) --
	(121.22,309.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.79,255.94) --
	( 60.79,258.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.73,277.61) --
	( 88.73,281.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (106.97,292.34) --
	(106.97,298.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.24,291.74) --
	(105.24,297.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.44,266.29) --
	( 72.44,269.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.86,281.03) --
	( 92.86,285.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.03,298.03) --
	(112.03,301.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.55,275.25) --
	( 84.55,278.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.64,282.00) --
	( 92.64,286.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.89,284.83) --
	( 96.89,289.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (108.82,294.55) --
	(108.82,300.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.57,260.48) --
	( 68.57,263.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.03,286.40) --
	( 98.03,289.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.02,288.23) --
	(104.02,293.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.85,281.07) --
	( 90.85,284.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.22,310.00) --
	(128.22,314.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.10,283.26) --
	( 95.10,290.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.15,256.51) --
	( 61.15,259.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.84,288.09) --
	(100.84,293.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.64,288.34) --
	(101.64,293.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.37,271.57) --
	( 81.37,275.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (108.47,295.28) --
	(108.47,300.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.02,267.15) --
	( 76.02,270.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.04,282.68) --
	( 95.04,286.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 82.97,274.28) --
	( 82.97,277.23);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.04,272.19) --
	( 89.04,278.52);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.13,287.56) --
	(100.13,292.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.15,276.82) --
	( 86.15,280.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.23,260.73) --
	( 66.23,263.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.83,275.11) --
	( 84.83,280.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.01,293.67) --
	(110.01,298.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.64,265.07) --
	( 72.64,268.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.63,291.49) --
	(103.63,296.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.17,266.01) --
	( 75.17,270.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.42,302.97) --
	(121.42,308.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.69,283.77) --
	( 95.69,287.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.56,265.83) --
	( 73.56,269.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.69,299.29) --
	(114.69,306.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.50,281.02) --
	( 92.50,287.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.37,288.52) --
	(101.37,293.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.95,277.38) --
	( 86.95,280.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.83,269.50) --
	( 77.83,272.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.72,276.86) --
	( 86.72,282.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (113.57,299.06) --
	(113.57,302.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.07,259.65) --
	( 66.07,263.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.56,283.79) --
	( 95.56,287.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.40,291.16) --
	(103.40,295.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.68,270.79) --
	( 78.68,275.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.18,284.39) --
	( 95.18,287.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.11,276.10) --
	( 85.11,281.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (107.03,291.97) --
	(107.03,297.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.85,263.59) --
	( 70.85,266.34);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.58,265.02) --
	( 74.58,267.95);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.56,286.31) --
	( 96.56,290.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.98,280.60) --
	( 89.98,284.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.98,291.56) --
	(104.98,295.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.19,268.02) --
	( 75.19,273.30);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.27,292.49) --
	(110.27,297.27);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.48,260.42) --
	( 65.48,262.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.91,265.05) --
	( 72.91,270.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 90.11,278.33) --
	( 90.11,282.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 94.00,282.19) --
	( 94.00,286.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (105.13,292.19) --
	(105.13,298.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.06,273.37) --
	( 84.06,277.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 96.09,284.22) --
	( 96.09,288.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.77,278.41) --
	( 89.77,283.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.41,300.94) --
	(114.41,304.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.74,290.88) --
	(103.74,295.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.57,295.98) --
	(112.57,301.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.47,271.65) --
	( 79.47,274.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.02,278.41) --
	( 88.02,282.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.94,268.31) --
	( 74.94,273.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.72,268.06) --
	( 76.72,270.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.57,271.33) --
	( 81.57,277.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 71.87,264.63) --
	( 71.87,268.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.18,296.46) --
	(110.18,300.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 73.41,265.08) --
	( 73.41,269.64);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 85.28,270.76) --
	( 85.28,276.46);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.24,275.69) --
	( 86.24,280.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.02,272.68) --
	( 83.02,276.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 72.89,265.90) --
	( 72.89,268.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.64,281.36) --
	( 93.64,287.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.83,289.39) --
	(103.83,295.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 64.27,258.16) --
	( 64.27,260.60);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 59.52,256.31) circle (  1.49);

\path[fill=fillColor] ( 80.98,271.97) circle (  1.49);

\path[fill=fillColor] ( 73.46,267.99) circle (  1.49);

\path[fill=fillColor] (106.47,296.47) circle (  1.49);

\path[fill=fillColor] ( 93.59,284.61) circle (  1.49);

\path[fill=fillColor] ( 85.32,278.07) circle (  1.49);

\path[fill=fillColor] (114.14,300.88) circle (  1.49);

\path[fill=fillColor] (114.03,300.80) circle (  1.49);

\path[fill=fillColor] ( 99.32,288.89) circle (  1.49);

\path[fill=fillColor] ( 85.70,276.90) circle (  1.49);

\path[fill=fillColor] (119.28,306.56) circle (  1.49);

\path[fill=fillColor] ( 74.17,268.90) circle (  1.49);

\path[fill=fillColor] ( 86.64,277.85) circle (  1.49);

\path[fill=fillColor] ( 76.12,269.00) circle (  1.49);

\path[fill=fillColor] ( 81.26,274.66) circle (  1.49);

\path[fill=fillColor] (121.22,306.42) circle (  1.49);

\path[fill=fillColor] ( 60.79,257.10) circle (  1.49);

\path[fill=fillColor] ( 88.73,279.39) circle (  1.49);

\path[fill=fillColor] (106.97,295.64) circle (  1.49);

\path[fill=fillColor] (105.24,294.46) circle (  1.49);

\path[fill=fillColor] ( 72.44,267.96) circle (  1.49);

\path[fill=fillColor] ( 92.86,283.14) circle (  1.49);

\path[fill=fillColor] (112.03,299.77) circle (  1.49);

\path[fill=fillColor] ( 84.55,276.80) circle (  1.49);

\path[fill=fillColor] ( 92.64,284.02) circle (  1.49);

\path[fill=fillColor] ( 96.89,286.92) circle (  1.49);

\path[fill=fillColor] (108.82,297.30) circle (  1.49);

\path[fill=fillColor] ( 68.57,262.05) circle (  1.49);

\path[fill=fillColor] ( 98.03,287.92) circle (  1.49);

\path[fill=fillColor] (104.02,290.80) circle (  1.49);

\path[fill=fillColor] ( 90.85,282.88) circle (  1.49);

\path[fill=fillColor] (128.22,312.01) circle (  1.49);

\path[fill=fillColor] ( 95.10,286.47) circle (  1.49);

\path[fill=fillColor] ( 61.15,257.76) circle (  1.49);

\path[fill=fillColor] (100.84,290.49) circle (  1.49);

\path[fill=fillColor] (101.64,290.96) circle (  1.49);

\path[fill=fillColor] ( 81.37,273.72) circle (  1.49);

\path[fill=fillColor] (108.47,298.20) circle (  1.49);

\path[fill=fillColor] ( 76.02,268.67) circle (  1.49);

\path[fill=fillColor] ( 95.04,284.57) circle (  1.49);

\path[fill=fillColor] ( 82.97,275.76) circle (  1.49);

\path[fill=fillColor] ( 89.04,275.52) circle (  1.49);

\path[fill=fillColor] (100.13,289.80) circle (  1.49);

\path[fill=fillColor] ( 86.15,278.76) circle (  1.49);

\path[fill=fillColor] ( 66.23,261.83) circle (  1.49);

\path[fill=fillColor] ( 84.83,277.83) circle (  1.49);

\path[fill=fillColor] (110.01,296.25) circle (  1.49);

\path[fill=fillColor] ( 72.64,266.61) circle (  1.49);

\path[fill=fillColor] (103.63,293.95) circle (  1.49);

\path[fill=fillColor] ( 75.17,268.24) circle (  1.49);

\path[fill=fillColor] (121.42,305.81) circle (  1.49);

\path[fill=fillColor] ( 95.69,285.73) circle (  1.49);

\path[fill=fillColor] ( 73.56,267.64) circle (  1.49);

\path[fill=fillColor] (114.69,302.69) circle (  1.49);

\path[fill=fillColor] ( 92.50,283.92) circle (  1.49);

\path[fill=fillColor] (101.37,290.87) circle (  1.49);

\path[fill=fillColor] ( 86.95,278.72) circle (  1.49);

\path[fill=fillColor] ( 77.83,271.17) circle (  1.49);

\path[fill=fillColor] ( 86.72,279.55) circle (  1.49);

\path[fill=fillColor] (113.57,300.94) circle (  1.49);

\path[fill=fillColor] ( 66.07,261.77) circle (  1.49);

\path[fill=fillColor] ( 95.56,285.41) circle (  1.49);

\path[fill=fillColor] (103.40,293.30) circle (  1.49);

\path[fill=fillColor] ( 78.68,272.94) circle (  1.49);

\path[fill=fillColor] ( 95.18,285.77) circle (  1.49);

\path[fill=fillColor] ( 85.11,278.89) circle (  1.49);

\path[fill=fillColor] (107.03,294.56) circle (  1.49);

\path[fill=fillColor] ( 70.85,264.94) circle (  1.49);

\path[fill=fillColor] ( 74.58,266.46) circle (  1.49);

\path[fill=fillColor] ( 96.56,288.61) circle (  1.49);

\path[fill=fillColor] ( 89.98,282.70) circle (  1.49);

\path[fill=fillColor] (104.98,293.46) circle (  1.49);

\path[fill=fillColor] ( 75.19,270.71) circle (  1.49);

\path[fill=fillColor] (110.27,294.94) circle (  1.49);

\path[fill=fillColor] ( 65.48,261.45) circle (  1.49);

\path[fill=fillColor] ( 72.91,267.74) circle (  1.49);

\path[fill=fillColor] ( 90.11,280.35) circle (  1.49);

\path[fill=fillColor] ( 94.00,284.47) circle (  1.49);

\path[fill=fillColor] (105.13,295.40) circle (  1.49);

\path[fill=fillColor] ( 84.06,275.35) circle (  1.49);

\path[fill=fillColor] ( 96.09,286.43) circle (  1.49);

\path[fill=fillColor] ( 89.77,280.90) circle (  1.49);

\path[fill=fillColor] (114.41,302.93) circle (  1.49);

\path[fill=fillColor] (103.74,293.51) circle (  1.49);

\path[fill=fillColor] (112.57,298.70) circle (  1.49);

\path[fill=fillColor] ( 79.47,272.94) circle (  1.49);

\path[fill=fillColor] ( 88.02,280.29) circle (  1.49);

\path[fill=fillColor] ( 74.94,270.80) circle (  1.49);

\path[fill=fillColor] ( 76.72,269.50) circle (  1.49);

\path[fill=fillColor] ( 81.57,274.16) circle (  1.49);

\path[fill=fillColor] ( 71.87,266.32) circle (  1.49);

\path[fill=fillColor] (110.18,298.57) circle (  1.49);

\path[fill=fillColor] ( 73.41,267.37) circle (  1.49);

\path[fill=fillColor] ( 85.28,273.71) circle (  1.49);

\path[fill=fillColor] ( 86.24,277.82) circle (  1.49);

\path[fill=fillColor] ( 83.02,274.45) circle (  1.49);

\path[fill=fillColor] ( 72.89,267.16) circle (  1.49);

\path[fill=fillColor] ( 93.64,284.46) circle (  1.49);

\path[fill=fillColor] (103.83,292.30) circle (  1.49);

\path[fill=fillColor] ( 64.27,259.36) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,354.55) {Coverage: 96\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- (174.42,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 32.47,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.13,233.11) -- ( 56.13,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.79,233.11) -- ( 79.79,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.45,233.11) -- (103.45,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.10,233.11) -- (127.10,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (150.76,233.11) -- (150.76,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (174.42,233.11) -- (174.42,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 56.13,218.86) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 79.79,218.86) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (103.45,218.86) {0.3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (127.10,218.86) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (150.76,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (174.42,218.86) {0.6};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 32.47,351.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,233.11) -- ( 28.51,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,252.86) -- ( 28.51,252.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,272.60) -- ( 28.51,272.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,292.34) -- ( 28.51,292.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,312.09) -- ( 28.51,312.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,331.83) -- ( 28.51,331.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,351.58) -- ( 28.51,351.58);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,250.58) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,270.33) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,290.07) {0.3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,309.82) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,329.56) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,349.30) {0.6};
\end{scope}
\begin{scope}
\path[clip] (192.72,192.72) rectangle (385.44,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (297.00,366.47) {\bfseries Clock standard deviation $\sigma$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (297.00,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (199.85,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (225.19,233.11) rectangle (368.81,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (225.19,233.11) --
	(368.81,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.20,233.11) --
	(225.20,235.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.97,251.22) --
	(255.97,265.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.49,233.11) --
	(225.49,235.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.92,235.60) --
	(230.92,240.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.36,233.11) --
	(227.36,235.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.05,233.11) --
	(227.05,235.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.68,241.59) --
	(241.68,250.18);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.79,233.11) --
	(232.79,237.60);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.76,233.11) --
	(233.76,246.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.43,239.46) --
	(238.43,245.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.51,233.11) --
	(225.51,235.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.28,233.11) --
	(226.28,237.19);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.23,233.11) --
	(230.23,236.42);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.36,243.70) --
	(240.36,253.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.44,233.11) --
	(225.44,235.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.07,233.11) --
	(228.07,239.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.35,255.95) --
	(261.35,273.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.24,239.57) --
	(238.24,245.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.48,233.11) --
	(228.48,239.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.46,235.04) --
	(232.46,241.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.38,235.15) --
	(232.38,240.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.77,244.58) --
	(240.77,254.70);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.24,233.11) --
	(228.24,235.36);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.92,241.83) --
	(238.92,249.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.37,233.11) --
	(227.37,235.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.43,238.49) --
	(232.43,243.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.87,241.81) --
	(241.87,251.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.04,233.11) --
	(226.04,237.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.22,233.11) --
	(225.22,235.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.42,236.44) --
	(230.42,244.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.26,242.21) --
	(239.26,249.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.68,235.68) --
	(229.68,239.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.93,233.11) --
	(230.93,243.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.41,233.11) --
	(227.41,235.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.67,233.11) --
	(225.67,235.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.72,233.11) --
	(226.72,236.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (269.99,261.23) --
	(269.99,284.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.21,233.11) --
	(228.21,240.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.33,233.11) --
	(226.33,236.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.49,233.11) --
	(226.49,235.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.33,233.11) --
	(225.33,234.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.90,233.11) --
	(225.90,241.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.62,233.11) --
	(226.62,235.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) --
	(225.19,235.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.10,233.11) --
	(226.10,235.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.74,239.40) --
	(234.74,248.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.62,233.11) --
	(226.62,235.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.14,233.11) --
	(226.14,236.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.54,233.11) --
	(225.54,235.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.54,245.01) --
	(246.54,256.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.84,233.11) --
	(225.84,236.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.12,233.11) --
	(227.12,235.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.15,241.10) --
	(242.15,248.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (241.68,243.14) --
	(241.68,254.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.26,233.11) --
	(231.26,238.37);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.82,233.11) --
	(230.82,236.95);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.58,234.44) --
	(229.58,238.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.35,247.08) --
	(246.35,257.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.43,233.11) --
	(225.43,238.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.47,233.11) --
	(226.47,236.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.55,233.11) --
	(229.55,239.24);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.27,235.54) --
	(232.27,238.86);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.16,233.11) --
	(228.16,238.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.25,238.99) --
	(233.25,245.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.14,233.11) --
	(226.14,235.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.31,262.62) --
	(279.31,283.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.32,233.11) --
	(226.32,237.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.25,233.11) --
	(227.25,236.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.67,233.11) --
	(225.67,235.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.55,233.11) --
	(229.55,238.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.73,233.11) --
	(230.73,238.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.64,234.09) --
	(230.64,239.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (320.94,283.48) --
	(320.94,330.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.95,233.11) --
	(226.95,236.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.77,237.27) --
	(232.77,241.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.27,245.97) --
	(250.27,259.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.09,267.31) --
	(280.09,291.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.09,237.24) --
	(233.09,243.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.20,236.30) --
	(240.20,249.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.20,248.41) --
	(248.20,263.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.16,233.11) --
	(227.16,236.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.87,233.11) --
	(226.87,237.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) --
	(225.19,235.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.60,249.81) --
	(250.60,263.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.30,233.11) --
	(226.30,235.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.41,238.10) --
	(234.41,242.72);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.08,243.02) --
	(236.08,251.25);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.66,253.41) --
	(255.66,268.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.27,233.11) --
	(225.27,237.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (286.90,270.35) --
	(286.90,301.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.36,243.98) --
	(243.36,253.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.80,240.69) --
	(235.80,247.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.41,233.11) --
	(226.41,236.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.44,233.11) --
	(226.44,240.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.65,233.11) --
	(227.65,235.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.20,233.11) --
	(225.20,234.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.96,234.59) --
	(229.96,238.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.58,233.11) --
	(226.58,236.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.84,233.11) --
	(226.84,236.61);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.92,250.57) --
	(263.92,263.20);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (225.20,233.78) circle (  1.49);

\path[fill=fillColor] (255.97,257.95) circle (  1.49);

\path[fill=fillColor] (225.49,233.86) circle (  1.49);

\path[fill=fillColor] (230.92,237.80) circle (  1.49);

\path[fill=fillColor] (227.36,233.95) circle (  1.49);

\path[fill=fillColor] (227.05,233.81) circle (  1.49);

\path[fill=fillColor] (241.68,245.65) circle (  1.49);

\path[fill=fillColor] (232.79,234.68) circle (  1.49);

\path[fill=fillColor] (233.76,240.98) circle (  1.49);

\path[fill=fillColor] (238.43,242.42) circle (  1.49);

\path[fill=fillColor] (225.51,233.87) circle (  1.49);

\path[fill=fillColor] (226.28,234.74) circle (  1.49);

\path[fill=fillColor] (230.23,234.16) circle (  1.49);

\path[fill=fillColor] (240.36,248.49) circle (  1.49);

\path[fill=fillColor] (225.44,233.86) circle (  1.49);

\path[fill=fillColor] (228.07,235.59) circle (  1.49);

\path[fill=fillColor] (261.35,264.37) circle (  1.49);

\path[fill=fillColor] (238.24,242.45) circle (  1.49);

\path[fill=fillColor] (228.48,235.22) circle (  1.49);

\path[fill=fillColor] (232.46,238.23) circle (  1.49);

\path[fill=fillColor] (232.38,237.99) circle (  1.49);

\path[fill=fillColor] (240.77,249.39) circle (  1.49);

\path[fill=fillColor] (228.24,233.85) circle (  1.49);

\path[fill=fillColor] (238.92,245.67) circle (  1.49);

\path[fill=fillColor] (227.37,234.01) circle (  1.49);

\path[fill=fillColor] (232.43,241.08) circle (  1.49);

\path[fill=fillColor] (241.87,246.46) circle (  1.49);

\path[fill=fillColor] (226.04,234.96) circle (  1.49);

\path[fill=fillColor] (225.22,233.75) circle (  1.49);

\path[fill=fillColor] (230.42,240.31) circle (  1.49);

\path[fill=fillColor] (239.26,245.58) circle (  1.49);

\path[fill=fillColor] (229.68,237.54) circle (  1.49);

\path[fill=fillColor] (230.93,237.30) circle (  1.49);

\path[fill=fillColor] (227.41,233.90) circle (  1.49);

\path[fill=fillColor] (225.67,233.92) circle (  1.49);

\path[fill=fillColor] (226.72,234.30) circle (  1.49);

\path[fill=fillColor] (269.99,272.39) circle (  1.49);

\path[fill=fillColor] (228.21,237.54) circle (  1.49);

\path[fill=fillColor] (226.33,234.98) circle (  1.49);

\path[fill=fillColor] (226.49,233.84) circle (  1.49);

\path[fill=fillColor] (225.33,233.64) circle (  1.49);

\path[fill=fillColor] (225.90,235.89) circle (  1.49);

\path[fill=fillColor] (226.62,233.78) circle (  1.49);

\path[fill=fillColor] (225.19,233.74) circle (  1.49);

\path[fill=fillColor] (226.10,234.10) circle (  1.49);

\path[fill=fillColor] (234.74,244.00) circle (  1.49);

\path[fill=fillColor] (226.62,233.89) circle (  1.49);

\path[fill=fillColor] (226.14,234.52) circle (  1.49);

\path[fill=fillColor] (225.54,233.93) circle (  1.49);

\path[fill=fillColor] (246.54,250.33) circle (  1.49);

\path[fill=fillColor] (225.84,234.35) circle (  1.49);

\path[fill=fillColor] (227.12,233.73) circle (  1.49);

\path[fill=fillColor] (242.15,244.37) circle (  1.49);

\path[fill=fillColor] (241.68,248.52) circle (  1.49);

\path[fill=fillColor] (231.26,234.80) circle (  1.49);

\path[fill=fillColor] (230.82,234.65) circle (  1.49);

\path[fill=fillColor] (229.58,236.46) circle (  1.49);

\path[fill=fillColor] (246.35,251.99) circle (  1.49);

\path[fill=fillColor] (225.43,235.18) circle (  1.49);

\path[fill=fillColor] (226.47,235.07) circle (  1.49);

\path[fill=fillColor] (229.55,235.50) circle (  1.49);

\path[fill=fillColor] (232.27,237.13) circle (  1.49);

\path[fill=fillColor] (228.16,236.11) circle (  1.49);

\path[fill=fillColor] (233.25,242.19) circle (  1.49);

\path[fill=fillColor] (226.14,234.13) circle (  1.49);

\path[fill=fillColor] (279.31,272.54) circle (  1.49);

\path[fill=fillColor] (226.32,235.04) circle (  1.49);

\path[fill=fillColor] (227.25,234.70) circle (  1.49);

\path[fill=fillColor] (225.67,233.77) circle (  1.49);

\path[fill=fillColor] (229.55,235.46) circle (  1.49);

\path[fill=fillColor] (230.73,236.23) circle (  1.49);

\path[fill=fillColor] (230.64,236.67) circle (  1.49);

\path[fill=fillColor] (320.94,306.02) circle (  1.49);

\path[fill=fillColor] (226.95,234.20) circle (  1.49);

\path[fill=fillColor] (232.77,239.15) circle (  1.49);

\path[fill=fillColor] (250.27,252.29) circle (  1.49);

\path[fill=fillColor] (280.09,278.51) circle (  1.49);

\path[fill=fillColor] (233.09,240.12) circle (  1.49);

\path[fill=fillColor] (240.20,242.37) circle (  1.49);

\path[fill=fillColor] (248.20,255.66) circle (  1.49);

\path[fill=fillColor] (227.16,234.13) circle (  1.49);

\path[fill=fillColor] (226.87,235.30) circle (  1.49);

\path[fill=fillColor] (225.19,233.73) circle (  1.49);

\path[fill=fillColor] (250.60,256.38) circle (  1.49);

\path[fill=fillColor] (226.30,233.89) circle (  1.49);

\path[fill=fillColor] (234.41,240.35) circle (  1.49);

\path[fill=fillColor] (236.08,246.93) circle (  1.49);

\path[fill=fillColor] (255.66,260.83) circle (  1.49);

\path[fill=fillColor] (225.27,235.26) circle (  1.49);

\path[fill=fillColor] (286.90,285.54) circle (  1.49);

\path[fill=fillColor] (243.36,248.40) circle (  1.49);

\path[fill=fillColor] (235.80,243.89) circle (  1.49);

\path[fill=fillColor] (226.41,234.10) circle (  1.49);

\path[fill=fillColor] (226.44,235.64) circle (  1.49);

\path[fill=fillColor] (227.65,234.04) circle (  1.49);

\path[fill=fillColor] (225.20,233.65) circle (  1.49);

\path[fill=fillColor] (229.96,236.31) circle (  1.49);

\path[fill=fillColor] (226.58,234.27) circle (  1.49);

\path[fill=fillColor] (226.84,234.21) circle (  1.49);

\path[fill=fillColor] (263.92,256.34) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (297.00,354.55) {Coverage: 93\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (356.16,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (225.19,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.02,233.11) -- (247.02,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.85,233.11) -- (268.85,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (290.68,233.11) -- (290.68,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.51,233.11) -- (312.51,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.33,233.11) -- (334.33,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.16,233.11) -- (356.16,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (225.19,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (247.02,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (268.85,218.86) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (290.68,218.86) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (312.51,218.86) {2.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (334.33,218.86) {2.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (356.16,218.86) {3.0};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (225.19,342.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,233.11) -- (221.23,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,251.33) -- (221.23,251.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,269.55) -- (221.23,269.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,287.76) -- (221.23,287.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,305.98) -- (221.23,305.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,324.20) -- (221.23,324.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.19,342.41) -- (221.23,342.41);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,249.06) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,267.27) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,285.49) {1.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,303.71) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,321.92) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (217.27,340.14) {3.0};
\end{scope}
\begin{scope}
\path[clip] (385.44,192.72) rectangle (578.16,385.44);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (489.72,366.47) {\bfseries Branch rate mean};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (489.72,203.02) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (392.57,293.04) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] (417.91,233.11) rectangle (561.53,352.97);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] (417.91,233.11) --
	(561.53,352.97);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.34,300.35) --
	(499.34,301.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.07,279.07) --
	(498.07,308.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.37,300.20) --
	(499.37,301.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.75,297.47) --
	(499.75,303.91);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (497.93,300.22) --
	(497.93,302.00);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.66,300.38) --
	(499.66,301.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.94,297.00) --
	(511.94,318.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.42,299.34) --
	(499.42,302.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.78,293.72) --
	(498.78,305.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.51,295.19) --
	(502.51,307.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.30,300.24) --
	(499.30,301.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.47,299.33) --
	(499.47,302.80);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.50,299.93) --
	(500.50,301.99);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.61,291.01) --
	(502.61,311.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.14,300.37) --
	(499.14,302.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.54,298.84) --
	(499.54,303.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (488.59,285.43) --
	(488.59,330.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (503.34,293.94) --
	(503.34,305.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.21,298.89) --
	(500.21,302.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.62,297.86) --
	(495.62,304.37);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.16,297.32) --
	(494.16,304.09);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (503.26,290.87) --
	(503.26,314.41);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.40,300.36) --
	(498.40,302.00);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (497.73,292.77) --
	(497.73,309.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.97,300.00) --
	(499.97,301.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.06,296.50) --
	(499.06,306.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (504.00,293.09) --
	(504.00,310.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.17,299.22) --
	(499.17,302.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.32,300.30) --
	(499.32,301.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.08,296.43) --
	(501.08,306.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.59,293.01) --
	(502.59,308.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.41,297.94) --
	(500.41,304.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.83,296.91) --
	(498.83,304.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.76,300.18) --
	(498.76,301.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.23,300.18) --
	(499.23,301.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.15,299.88) --
	(499.15,302.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.58,267.14) --
	(470.58,316.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.07,297.39) --
	(500.07,303.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.76,299.66) --
	(498.76,302.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.94,300.34) --
	(499.94,301.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.40,300.51) --
	(499.40,301.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.56,298.32) --
	(499.56,304.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.01,300.21) --
	(499.01,302.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.33,300.39) --
	(499.33,301.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.97,300.05) --
	(499.97,302.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.11,293.03) --
	(499.11,307.20);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.27,300.19) --
	(498.27,301.90);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.49,299.91) --
	(499.49,302.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.34,300.13) --
	(499.34,301.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.75,290.14) --
	(495.75,310.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.33,299.37) --
	(499.33,302.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.00,300.35) --
	(500.00,301.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.64,293.15) --
	(495.64,306.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (493.70,290.58) --
	(493.70,309.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.04,299.13) --
	(500.04,302.88);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.67,299.98) --
	(501.67,303.00);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.94,298.45) --
	(498.94,303.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (513.66,290.03) --
	(513.66,313.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.36,298.70) --
	(499.36,302.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.13,299.55) --
	(500.13,302.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.94,298.82) --
	(499.94,303.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.06,298.80) --
	(501.06,303.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.48,298.43) --
	(500.48,303.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.03,293.80) --
	(499.03,305.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.09,300.03) --
	(499.09,302.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.77,273.15) --
	(471.77,327.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.13,299.42) --
	(499.13,302.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.78,299.61) --
	(499.78,302.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.48,300.30) --
	(499.48,301.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.07,298.92) --
	(500.07,303.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (497.09,298.27) --
	(497.09,302.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.20,298.51) --
	(499.20,303.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (445.97,237.33) --
	(445.97,347.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.16,299.78) --
	(499.16,301.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.31,296.44) --
	(498.31,303.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (506.39,287.49) --
	(506.39,312.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (503.68,269.92) --
	(503.68,347.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.97,295.67) --
	(499.97,304.30);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.26,292.64) --
	(485.26,306.08);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (504.97,288.15) --
	(504.97,319.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.56,299.88) --
	(499.56,302.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.39,299.14) --
	(500.39,302.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.33,300.29) --
	(499.33,301.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (512.41,284.47) --
	(512.41,313.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.72,300.16) --
	(499.72,301.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (496.68,296.20) --
	(496.68,304.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.71,292.12) --
	(501.71,308.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.78,280.06) --
	(501.78,314.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.34,299.09) --
	(499.34,302.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.92,255.86) --
	(474.92,330.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.91,292.59) --
	(500.91,313.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.40,295.70) --
	(500.40,309.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.35,299.95) --
	(499.35,302.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.96,298.24) --
	(498.96,303.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.65,300.16) --
	(498.65,302.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.33,300.37) --
	(499.33,301.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.11,299.10) --
	(500.11,303.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.80,299.79) --
	(499.80,302.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (498.47,299.89) --
	(498.47,302.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.56,285.14) --
	(487.56,315.11);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (499.34,301.11) circle (  1.49);

\path[fill=fillColor] (498.07,294.03) circle (  1.49);

\path[fill=fillColor] (499.37,301.02) circle (  1.49);

\path[fill=fillColor] (499.75,300.60) circle (  1.49);

\path[fill=fillColor] (497.93,301.08) circle (  1.49);

\path[fill=fillColor] (499.66,301.09) circle (  1.49);

\path[fill=fillColor] (511.94,307.52) circle (  1.49);

\path[fill=fillColor] (499.42,300.98) circle (  1.49);

\path[fill=fillColor] (498.78,299.70) circle (  1.49);

\path[fill=fillColor] (502.51,301.26) circle (  1.49);

\path[fill=fillColor] (499.30,301.06) circle (  1.49);

\path[fill=fillColor] (499.47,301.05) circle (  1.49);

\path[fill=fillColor] (500.50,301.03) circle (  1.49);

\path[fill=fillColor] (502.61,301.05) circle (  1.49);

\path[fill=fillColor] (499.14,301.12) circle (  1.49);

\path[fill=fillColor] (499.54,301.13) circle (  1.49);

\path[fill=fillColor] (488.59,307.43) circle (  1.49);

\path[fill=fillColor] (503.34,299.86) circle (  1.49);

\path[fill=fillColor] (500.21,300.95) circle (  1.49);

\path[fill=fillColor] (495.62,301.14) circle (  1.49);

\path[fill=fillColor] (494.16,300.82) circle (  1.49);

\path[fill=fillColor] (503.26,302.70) circle (  1.49);

\path[fill=fillColor] (498.40,301.12) circle (  1.49);

\path[fill=fillColor] (497.73,301.34) circle (  1.49);

\path[fill=fillColor] (499.97,301.01) circle (  1.49);

\path[fill=fillColor] (499.06,301.66) circle (  1.49);

\path[fill=fillColor] (504.00,301.70) circle (  1.49);

\path[fill=fillColor] (499.17,301.10) circle (  1.49);

\path[fill=fillColor] (499.32,301.07) circle (  1.49);

\path[fill=fillColor] (501.08,301.42) circle (  1.49);

\path[fill=fillColor] (502.59,300.86) circle (  1.49);

\path[fill=fillColor] (500.41,301.22) circle (  1.49);

\path[fill=fillColor] (498.83,300.84) circle (  1.49);

\path[fill=fillColor] (498.76,301.06) circle (  1.49);

\path[fill=fillColor] (499.23,301.05) circle (  1.49);

\path[fill=fillColor] (499.15,301.03) circle (  1.49);

\path[fill=fillColor] (470.58,292.44) circle (  1.49);

\path[fill=fillColor] (500.07,300.64) circle (  1.49);

\path[fill=fillColor] (498.76,301.18) circle (  1.49);

\path[fill=fillColor] (499.94,301.06) circle (  1.49);

\path[fill=fillColor] (499.40,301.08) circle (  1.49);

\path[fill=fillColor] (499.56,301.12) circle (  1.49);

\path[fill=fillColor] (499.01,301.09) circle (  1.49);

\path[fill=fillColor] (499.33,301.05) circle (  1.49);

\path[fill=fillColor] (499.97,301.07) circle (  1.49);

\path[fill=fillColor] (499.11,300.47) circle (  1.49);

\path[fill=fillColor] (498.27,301.07) circle (  1.49);

\path[fill=fillColor] (499.49,301.19) circle (  1.49);

\path[fill=fillColor] (499.34,301.05) circle (  1.49);

\path[fill=fillColor] (495.75,300.48) circle (  1.49);

\path[fill=fillColor] (499.33,300.90) circle (  1.49);

\path[fill=fillColor] (500.00,301.03) circle (  1.49);

\path[fill=fillColor] (495.64,300.14) circle (  1.49);

\path[fill=fillColor] (493.70,300.41) circle (  1.49);

\path[fill=fillColor] (500.04,301.04) circle (  1.49);

\path[fill=fillColor] (501.67,301.24) circle (  1.49);

\path[fill=fillColor] (498.94,300.79) circle (  1.49);

\path[fill=fillColor] (513.66,301.64) circle (  1.49);

\path[fill=fillColor] (499.36,300.92) circle (  1.49);

\path[fill=fillColor] (500.13,301.14) circle (  1.49);

\path[fill=fillColor] (499.94,301.19) circle (  1.49);

\path[fill=fillColor] (501.06,301.27) circle (  1.49);

\path[fill=fillColor] (500.48,300.82) circle (  1.49);

\path[fill=fillColor] (499.03,299.57) circle (  1.49);

\path[fill=fillColor] (499.09,301.05) circle (  1.49);

\path[fill=fillColor] (471.77,299.97) circle (  1.49);

\path[fill=fillColor] (499.13,301.11) circle (  1.49);

\path[fill=fillColor] (499.78,301.03) circle (  1.49);

\path[fill=fillColor] (499.48,301.04) circle (  1.49);

\path[fill=fillColor] (500.07,301.08) circle (  1.49);

\path[fill=fillColor] (497.09,300.67) circle (  1.49);

\path[fill=fillColor] (499.20,300.89) circle (  1.49);

\path[fill=fillColor] (445.97,288.93) circle (  1.49);

\path[fill=fillColor] (499.16,300.98) circle (  1.49);

\path[fill=fillColor] (498.31,300.00) circle (  1.49);

\path[fill=fillColor] (506.39,299.45) circle (  1.49);

\path[fill=fillColor] (503.68,310.30) circle (  1.49);

\path[fill=fillColor] (499.97,300.03) circle (  1.49);

\path[fill=fillColor] (485.26,299.86) circle (  1.49);

\path[fill=fillColor] (504.97,303.91) circle (  1.49);

\path[fill=fillColor] (499.56,301.02) circle (  1.49);

\path[fill=fillColor] (500.39,301.02) circle (  1.49);

\path[fill=fillColor] (499.33,301.04) circle (  1.49);

\path[fill=fillColor] (512.41,298.79) circle (  1.49);

\path[fill=fillColor] (499.72,301.01) circle (  1.49);

\path[fill=fillColor] (496.68,300.47) circle (  1.49);

\path[fill=fillColor] (501.71,300.36) circle (  1.49);

\path[fill=fillColor] (501.78,297.25) circle (  1.49);

\path[fill=fillColor] (499.34,300.76) circle (  1.49);

\path[fill=fillColor] (474.92,291.97) circle (  1.49);

\path[fill=fillColor] (500.91,303.14) circle (  1.49);

\path[fill=fillColor] (500.40,302.50) circle (  1.49);

\path[fill=fillColor] (499.35,301.06) circle (  1.49);

\path[fill=fillColor] (498.96,300.92) circle (  1.49);

\path[fill=fillColor] (498.65,301.08) circle (  1.49);

\path[fill=fillColor] (499.33,301.08) circle (  1.49);

\path[fill=fillColor] (500.11,301.24) circle (  1.49);

\path[fill=fillColor] (499.80,301.03) circle (  1.49);

\path[fill=fillColor] (498.47,301.07) circle (  1.49);

\path[fill=fillColor] (487.56,300.48) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (489.72,354.55) {Coverage: 93\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (540.05,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (417.91,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.62,233.11) -- (458.62,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (499.33,233.11) -- (499.33,229.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (540.05,233.11) -- (540.05,229.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (417.91,218.86) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (458.62,218.86) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (499.33,218.86) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (540.05,218.86) {1.5};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (417.91,335.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,233.11) -- (413.95,233.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,267.09) -- (413.95,267.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,301.06) -- (413.95,301.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.91,335.04) -- (413.95,335.04);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,230.84) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,264.81) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,298.79) {1.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at (409.99,332.77) {1.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (192.72,192.72);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.79] at (104.28,173.75) {\bfseries Branch rate variance};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (104.28, 10.30) {True value};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (  7.13,100.32) {Estimated value};
\end{scope}
\begin{scope}
\path[clip] ( 32.47, 40.39) rectangle (176.09,160.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 32.47, 40.39) --
	(176.09,160.25);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.21, 42.42) --
	( 43.21, 50.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.67, 40.43) --
	( 32.67, 40.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.14, 40.91) --
	( 34.14, 42.59);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.81, 40.39) --
	( 32.81, 40.54);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.99, 40.39) --
	( 32.99, 41.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.37, 40.69) --
	( 33.37, 41.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.51);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.66, 40.39) --
	( 32.66, 40.47);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.20, 41.18) --
	( 34.20, 43.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.52, 40.39) --
	( 32.52, 40.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.76, 43.24) --
	( 48.76, 63.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.41, 40.71) --
	( 33.41, 41.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.56, 40.39) --
	( 32.56, 40.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.69, 40.39) --
	( 32.69, 40.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.77, 40.39) --
	( 32.77, 40.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.26, 41.41) --
	( 34.26, 44.88);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.54, 40.39) --
	( 32.54, 40.43);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.69, 40.99) --
	( 33.69, 42.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.51, 40.39) --
	( 32.51, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.90, 40.62) --
	( 32.90, 41.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.95, 40.92) --
	( 33.95, 42.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.70, 40.44) --
	( 32.70, 41.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.69, 41.09) --
	( 33.69, 42.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.62, 40.44) --
	( 32.62, 40.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.72, 40.39) --
	( 32.72, 41.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.51, 40.39) --
	( 32.51, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.92, 43.50) --
	( 38.92, 88.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.53, 40.39) --
	( 32.53, 40.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.94, 40.66) --
	( 32.94, 42.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.77, 41.38) --
	( 34.77, 44.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.13, 40.85) --
	( 34.13, 41.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.11, 41.10) --
	( 34.11, 43.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.61, 40.39) --
	( 32.61, 40.59);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.62, 40.39) --
	( 32.62, 40.50);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.60, 40.39) --
	( 32.60, 40.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.00, 41.70) --
	( 37.00, 45.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.55, 40.39) --
	( 32.55, 40.66);
\definecolor{drawColor}{RGB}{186,46,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.75, 40.43) --
	( 32.75, 40.61);
\definecolor{drawColor}{RGB}{0,140,186}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.56, 40.39) --
	( 32.56, 40.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.91, 40.66) --
	( 32.91, 41.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.10, 43.14) --
	( 46.10, 76.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.59, 40.39) --
	( 32.59, 40.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.65, 40.39) --
	( 32.65, 40.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.62, 40.39) --
	( 32.62, 40.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 46.23, 40.44) --
	( 46.23,292.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.81, 40.52) --
	( 32.81, 40.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.92, 41.43) --
	( 35.92, 45.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.73, 43.34) --
	( 49.73,120.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.88, 40.50) --
	( 32.88, 41.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.31, 40.39) --
	( 33.31, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.73, 42.18) --
	( 35.73, 49.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.18, 42.66) --
	( 41.18, 52.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.01, 40.59) --
	( 33.01, 41.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.93, 41.21) --
	( 33.93, 43.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.21, 43.31) --
	( 38.21, 56.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.94, 41.55) --
	( 43.94,142.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 34.22, 41.27) --
	( 34.22, 43.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.42, 40.80) --
	( 33.42, 41.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.50, 40.39) --
	( 32.50, 40.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) --
	( 32.47, 40.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.59, 40.41) --
	( 32.59, 40.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.48, 40.39) --
	( 32.48, 40.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.49, 40.39) --
	( 32.49, 40.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.60, 42.52) --
	( 38.60, 49.84);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 43.21, 45.80) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.67, 40.56) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.40) circle (  1.49);

\path[fill=fillColor] ( 34.14, 41.67) circle (  1.49);

\path[fill=fillColor] ( 32.81, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.99, 40.92) circle (  1.49);

\path[fill=fillColor] ( 33.37, 41.05) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.66, 40.41) circle (  1.49);

\path[fill=fillColor] ( 34.20, 42.23) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.52, 40.46) circle (  1.49);

\path[fill=fillColor] ( 48.76, 51.18) circle (  1.49);

\path[fill=fillColor] ( 33.41, 41.04) circle (  1.49);

\path[fill=fillColor] ( 32.56, 40.45) circle (  1.49);

\path[fill=fillColor] ( 32.69, 40.60) circle (  1.49);

\path[fill=fillColor] ( 32.77, 40.58) circle (  1.49);

\path[fill=fillColor] ( 34.26, 42.88) circle (  1.49);

\path[fill=fillColor] ( 32.54, 40.40) circle (  1.49);

\path[fill=fillColor] ( 33.69, 41.64) circle (  1.49);

\path[fill=fillColor] ( 32.51, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.90, 40.88) circle (  1.49);

\path[fill=fillColor] ( 33.95, 41.69) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.70, 40.80) circle (  1.49);

\path[fill=fillColor] ( 33.69, 41.64) circle (  1.49);

\path[fill=fillColor] ( 32.62, 40.54) circle (  1.49);

\path[fill=fillColor] ( 32.72, 40.60) circle (  1.49);

\path[fill=fillColor] ( 32.51, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.41) circle (  1.49);

\path[fill=fillColor] ( 38.92, 60.72) circle (  1.49);

\path[fill=fillColor] ( 32.53, 40.56) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.51) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.94, 41.31) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 34.77, 42.65) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.40) circle (  1.49);

\path[fill=fillColor] ( 34.13, 41.35) circle (  1.49);

\path[fill=fillColor] ( 34.11, 42.28) circle (  1.49);

\path[fill=fillColor] ( 32.61, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.62, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.60, 40.48) circle (  1.49);

\path[fill=fillColor] ( 37.00, 43.14) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.45) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.55, 40.46) circle (  1.49);

\path[fill=fillColor] ( 32.75, 40.51) circle (  1.49);

\path[fill=fillColor] ( 32.56, 40.47) circle (  1.49);

\path[fill=fillColor] ( 32.91, 41.03) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.41) circle (  1.49);

\path[fill=fillColor] ( 46.10, 56.76) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.43) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.42) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.59, 40.45) circle (  1.49);

\path[fill=fillColor] ( 32.65, 40.48) circle (  1.49);

\path[fill=fillColor] ( 32.62, 40.49) circle (  1.49);

\path[fill=fillColor] ( 46.23,120.30) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.81, 40.67) circle (  1.49);

\path[fill=fillColor] ( 35.92, 43.18) circle (  1.49);

\path[fill=fillColor] ( 49.73, 72.06) circle (  1.49);

\path[fill=fillColor] ( 32.88, 40.77) circle (  1.49);

\path[fill=fillColor] ( 33.31, 41.12) circle (  1.49);

\path[fill=fillColor] ( 35.73, 45.14) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.44) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 41.18, 46.58) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.40) circle (  1.49);

\path[fill=fillColor] ( 33.01, 40.79) circle (  1.49);

\path[fill=fillColor] ( 33.93, 42.01) circle (  1.49);

\path[fill=fillColor] ( 38.21, 48.65) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.44) circle (  1.49);

\path[fill=fillColor] ( 43.94, 76.74) circle (  1.49);

\path[fill=fillColor] ( 34.22, 42.20) circle (  1.49);

\path[fill=fillColor] ( 33.42, 41.24) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.48) circle (  1.49);

\path[fill=fillColor] ( 32.50, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.47, 40.40) circle (  1.49);

\path[fill=fillColor] ( 32.59, 40.47) circle (  1.49);

\path[fill=fillColor] ( 32.48, 40.41) circle (  1.49);

\path[fill=fillColor] ( 32.49, 40.41) circle (  1.49);

\path[fill=fillColor] ( 38.60, 45.75) circle (  1.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (104.28,161.83) {Coverage: 95\%};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- (174.17, 40.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 32.47, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 56.09, 40.39) -- ( 56.09, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.70, 40.39) -- ( 79.70, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.32, 40.39) -- (103.32, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.93, 40.39) -- (126.93, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (150.55, 40.39) -- (150.55, 36.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (174.17, 40.39) -- (174.17, 36.43);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 32.47, 26.14) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 56.09, 26.14) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 79.70, 26.14) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (103.32, 26.14) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (126.93, 26.14) {8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (150.55, 26.14) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.66] at (174.17, 26.14) {12};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 32.47,158.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 40.39) -- ( 28.51, 40.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 60.10) -- ( 28.51, 60.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 79.81) -- ( 28.51, 79.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47, 99.52) -- ( 28.51, 99.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,119.23) -- ( 28.51,119.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,138.93) -- ( 28.51,138.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.47,158.64) -- ( 28.51,158.64);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 38.12) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 57.83) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 77.54) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55, 97.24) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,116.95) {8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,136.66) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 24.55,156.37) {12};
\end{scope}
\end{tikzpicture}

\end{document}