### 2020-09-01 benchmarking/benchmarkingVM/ESS_round3.tex

% Created by tikzDevice version 0.12.3.1 on 2020-09-01 15:30:15
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,357.33);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (301.08,335.19) {\bfseries Comparison of Bactrian+adapt(\textit{real}) and adapt(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (  9.60,190.67) {Relative ESS/hr};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 48.00,173.19) --
	(476.29,173.19);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 53.56,210.61) --
	( 53.56,170.08);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] ( 53.56,190.34) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.12,246.09) --
	( 59.12,233.67);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] ( 59.12,239.88) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 64.69,249.96) --
	( 64.69,218.26);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] ( 64.69,234.11) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.25,226.82) --
	( 70.25,207.10);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] ( 70.25,216.96) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.81,209.65) --
	( 75.81,192.67);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] ( 75.81,201.16) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.37,205.70) --
	( 81.37,138.28);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] ( 81.37,171.99) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.94,250.06) --
	( 86.94,239.46);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] ( 86.94,244.76) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.50,223.96) --
	( 92.50,209.76);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] ( 92.50,216.86) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.06,229.79) --
	( 98.06,207.30);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] ( 98.06,218.55) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] ( 53.56,215.85) --
	(103.62,215.85);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 78.59,284.14) {17\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (109.18,  0.00) --
	(109.18,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.75,234.94) --
	(114.75,216.63);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (114.75,225.78) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.31,231.18) --
	(120.31,216.59);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (120.31,223.89) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (125.87,210.48) --
	(125.87,196.00);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (125.87,203.24) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (131.43,217.85) --
	(131.43,208.48);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (131.43,213.17) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.00,207.88) --
	(137.00,192.85);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (137.00,200.36) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (142.56,184.21) --
	(142.56,163.53);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (142.56,173.87) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.12,274.87) --
	(148.12,264.59);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (148.12,269.73) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (153.68,345.05) --
	(153.68,220.65);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (153.68,282.85) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.24,222.89) --
	(159.24,199.51);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (159.24,211.20) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (114.75,224.57) --
	(164.81,224.57);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (139.78,284.14) {20\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (170.37,  0.00) --
	(170.37,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (175.93,201.00) --
	(175.93,192.97);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (175.93,196.98) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.49,237.79) --
	(181.49,235.79);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (181.49,236.79) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.05,213.87) --
	(187.05,210.98);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (187.05,212.42) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (192.62,213.40) --
	(192.62,210.38);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (192.62,211.89) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.18,192.27) --
	(198.18,187.06);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (198.18,189.67) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.74,187.16) --
	(203.74,185.57);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (203.74,186.36) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.30,240.68) --
	(209.30,238.19);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (209.30,239.44) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (214.87,223.05) --
	(214.87,219.07);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (214.87,221.06) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.43,215.97) --
	(220.43,213.85);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (220.43,214.91) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (175.93,212.74) --
	(225.99,212.74);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (200.96,284.14) {15\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (231.55,  0.00) --
	(231.55,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.11,198.51) --
	(237.11,180.05);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (237.11,189.28) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.68,250.44) --
	(242.68,241.85);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (242.68,246.14) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.24,239.13) --
	(248.24,208.60);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (248.24,223.86) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.80,213.14) --
	(253.80,198.87);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (253.80,206.01) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (259.36,215.64) --
	(259.36,185.82);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (259.36,200.73) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.93,175.34) --
	(264.93,160.05);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (264.93,167.69) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (270.49,231.48) --
	(270.49,205.22);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (270.49,218.35) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.05,253.96) --
	(276.05,239.75);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (276.05,246.86) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (281.61,236.72) --
	(281.61,225.81);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (281.61,231.26) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (237.11,215.55) --
	(287.17,215.55);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (262.14,284.14) {16\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (292.74,  0.00) --
	(292.74,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.30,199.13) --
	(298.30,180.61);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (298.30,189.87) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (303.86,240.03) --
	(303.86,230.01);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (303.86,235.02) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (309.42,238.59) --
	(309.42,208.20);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (309.42,223.39) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (314.99,210.81) --
	(314.99,195.70);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (314.99,203.25) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (320.55,225.30) --
	(320.55,196.46);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (320.55,210.88) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (326.11,175.84) --
	(326.11,163.67);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (326.11,169.76) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (331.67,230.74) --
	(331.67,204.43);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (331.67,217.58) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (337.23,232.28) --
	(337.23,208.40);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (337.23,220.34) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.80,233.20) --
	(342.80,221.37);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (342.80,227.29) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (298.30,211.47) --
	(348.36,211.47);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (323.33,284.14) {15\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (353.92,  0.00) --
	(353.92,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.48,201.98) --
	(359.48,175.41);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (359.48,188.70) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.05,237.99) --
	(365.05,223.51);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (365.05,230.75) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (370.61,244.31) --
	(370.61,211.56);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (370.61,227.93) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.17,216.84) --
	(376.17,196.62);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (376.17,206.73) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.73,223.80) --
	(381.73,193.65);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (381.73,208.73) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.29,197.62) --
	(387.29,172.56);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (387.29,185.09) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (392.86,237.58) --
	(392.86,211.29);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (392.86,224.44) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.42,213.92) --
	(398.42,200.71);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (398.42,207.32) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (403.98,230.89) --
	(403.98,207.97);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (403.98,219.43) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (359.48,211.43) --
	(409.54,211.43);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (384.51,284.14) {15\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (415.11,  0.00) --
	(415.11,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.67,179.14) --
	(420.67,173.06);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (420.67,176.10) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.23,182.77) --
	(426.23,177.72);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (426.23,180.25) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.79,177.03) --
	(431.79,173.95);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (431.79,175.49) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.35,175.37) --
	(437.35,171.54);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (437.35,173.45) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (442.92,175.20) --
	(442.92,172.49);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (442.92,173.84) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (448.48,182.05) --
	(448.48,177.75);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (448.48,179.90) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.04,176.59) --
	(454.04,169.19);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (454.04,172.89) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.60,177.97) --
	(459.60,173.15);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (459.60,175.56) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.16,183.83) --
	(465.16,176.41);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (465.16,180.12) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (420.67,176.41) --
	(470.73,176.41);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (445.70,284.14) {1.2\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (476.29,  0.00) --
	(476.29,357.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- (554.16, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.59, 60.00) -- ( 78.59, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.78, 60.00) -- (139.78, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.96, 60.00) -- (200.96, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.14, 60.00) -- (262.14, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.33, 60.00) -- (323.33, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.51, 60.00) -- (384.51, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (445.70, 60.00) -- (445.70, 54.00);

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at ( 80.66, 48.00) {Prior};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (141.84, 48.00) {Likelihood};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (203.03, 48.00) {Leaf rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (264.21, 48.00) {Mean branch rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (325.39, 48.00) {Tree length};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (386.58, 48.00) {Clock SD $\sigma$};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (447.76, 48.00) {HKY model $\kappa$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 48.00,286.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 42.00, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 92.88) -- ( 42.00, 92.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,135.92) -- ( 42.00,135.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,173.19) -- ( 42.00,173.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,210.47) -- ( 42.00,210.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,253.51) -- ( 42.00,253.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,286.39) -- ( 42.00,286.39);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 56.56) {$\frac{2}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 89.44) {$\frac{3}{4}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,132.47) {$\frac{7}{8}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,169.75) {$1\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,207.03) {$\frac{8}{7}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,250.06) {$\frac{4}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,282.94) {$\frac{3}{2}\times$};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,241.67) rectangle (494.39,235.67);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,229.67) rectangle (494.39,223.67);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,217.67) rectangle (494.39,211.67);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,205.67) rectangle (494.39,199.67);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,193.67) rectangle (494.39,187.67);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,181.67) rectangle (494.39,175.67);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,169.67) rectangle (494.39,163.67);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,157.67) rectangle (494.39,151.67);
\definecolor{fillColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,145.67) rectangle (494.39,139.67);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,235.22) {Ran};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,223.22) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,211.22) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,199.22) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,187.22) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,175.22) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,163.22) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,151.22) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,139.22) {Moyle};
\end{scope}
\end{tikzpicture}

\end{document}