### 2020-09-01 benchmarking/benchmarkingVM/ESS_round4b.tex

% Created by tikzDevice version 0.12.3.1 on 2020-09-01 15:30:41
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,357.33);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (301.08,335.19) {\bfseries Comparison of Bactrian+adapt(\textit{real}) and NER+Bactrian+adapt(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (  9.60,190.67) {Relative ESS/hr};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 48.00,173.19) --
	(476.29,173.19);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 53.56,175.23) --
	( 53.56,146.36);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] ( 53.56,160.80) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.12,186.05) --
	( 59.12,174.62);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] ( 59.12,180.33) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 64.69,164.25) --
	( 64.69,141.62);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] ( 64.69,152.94) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.25,167.28) --
	( 70.25,150.64);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] ( 70.25,158.96) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.81,179.53) --
	( 75.81,164.10);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] ( 75.81,171.82) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.37,162.02) --
	( 81.37, 98.30);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] ( 81.37,130.16) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.94,202.23) --
	( 86.94,189.46);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] ( 86.94,195.84) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.50,203.28) --
	( 92.50,181.04);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] ( 92.50,192.16) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.06,240.71) --
	( 98.06,222.93);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] ( 98.06,231.82) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] ( 53.56,176.39) --
	(103.62,176.39);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 78.59,284.14) {1.1\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (109.18,  0.00) --
	(109.18,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.75,200.32) --
	(114.75,179.86);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (114.75,190.09) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.31,192.84) --
	(120.31,179.45);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (120.31,186.14) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (125.87,182.97) --
	(125.87,171.24);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (125.87,177.11) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (131.43,184.85) --
	(131.43,171.12);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (131.43,177.99) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.00,184.14) --
	(137.00,167.55);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (137.00,175.84) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (142.56,169.54) --
	(142.56,147.07);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (142.56,158.31) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.12,195.26) --
	(148.12,186.84);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (148.12,191.05) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (153.68,257.34) --
	(153.68,213.88);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (153.68,235.61) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.24,242.16) --
	(159.24,219.28);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (159.24,230.72) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (114.75,192.50) --
	(164.81,192.50);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (139.78,284.14) {7.2\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (170.37,  0.00) --
	(170.37,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (175.93,178.18) --
	(175.93,171.14);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (175.93,174.66) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.49,183.40) --
	(181.49,181.91);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (181.49,182.66) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.05,178.32) --
	(187.05,176.29);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (187.05,177.30) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (192.62,173.36) --
	(192.62,171.68);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (192.62,172.52) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.18,173.13) --
	(198.18,167.55);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (198.18,170.34) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.74,175.58) --
	(203.74,173.46);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (203.74,174.52) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.30,188.62) --
	(209.30,187.01);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (209.30,187.82) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (214.87,191.38) --
	(214.87,187.66);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (214.87,189.52) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.43,221.94) --
	(220.43,219.83);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (220.43,220.89) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (175.93,183.76) --
	(225.99,183.76);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (200.96,284.14) {3.9\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (231.55,  0.00) --
	(231.55,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.11,175.10) --
	(237.11,147.62);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (237.11,161.36) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.68,181.65) --
	(242.68,172.54);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (242.68,177.10) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.24,159.01) --
	(248.24,132.62);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (248.24,145.81) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.80,177.17) --
	(253.80,157.93);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (253.80,167.55) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (259.36,184.40) --
	(259.36,154.83);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (259.36,169.62) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.93,177.18) --
	(264.93,164.80);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (264.93,170.99) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (270.49,208.69) --
	(270.49,185.32);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (270.49,197.01) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.05,218.20) --
	(276.05,198.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (276.05,208.22) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (281.61,223.20) --
	(281.61,214.78);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (281.61,218.99) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (237.11,180.53) --
	(287.17,180.53);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (262.14,284.14) {2.7\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (292.74,  0.00) --
	(292.74,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.30,171.47) --
	(298.30,146.80);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (298.30,159.14) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (303.86,178.61) --
	(303.86,167.68);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (303.86,173.15) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (309.42,162.68) --
	(309.42,139.42);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (309.42,151.05) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (314.99,176.48) --
	(314.99,158.35);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (314.99,167.41) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (320.55,197.57) --
	(320.55,164.96);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (320.55,181.26) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (326.11,176.14) --
	(326.11,165.55);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (326.11,170.84) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (331.67,208.50) --
	(331.67,185.79);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (331.67,197.14) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (337.23,224.16) --
	(337.23,192.88);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (337.23,208.52) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.80,226.54) --
	(342.80,218.12);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (342.80,222.33) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (298.30,182.10) --
	(348.36,182.10);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (323.33,284.14) {3.2\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (353.92,  0.00) --
	(353.92,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.48,183.45) --
	(359.48,166.61);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (359.48,175.03) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.05,182.24) --
	(365.05,167.41);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (365.05,174.82) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (370.61,157.24) --
	(370.61,131.60);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (370.61,144.42) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.17,169.99) --
	(376.17,153.88);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (376.17,161.94) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.73,185.79) --
	(381.73,156.55);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (381.73,171.17) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.29,174.11) --
	(387.29,150.84);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (387.29,162.48) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (392.86,193.90) --
	(392.86,166.47);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (392.86,180.18) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.42,203.16) --
	(398.42,181.99);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (398.42,192.57) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (403.98,239.78) --
	(403.98,221.53);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (403.98,230.66) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (359.48,177.99) --
	(409.54,177.99);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (384.51,284.14) {1.7\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (415.11,  0.00) --
	(415.11,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.67,175.06) --
	(420.67,169.51);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (420.67,172.29) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.23,189.40) --
	(426.23,183.99);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (426.23,186.70) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.79,179.73) --
	(431.79,176.02);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (431.79,177.87) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.35,178.20) --
	(437.35,174.37);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (437.35,176.28) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (442.92,174.49) --
	(442.92,171.77);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (442.92,173.13) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (448.48,177.55) --
	(448.48,172.82);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (448.48,175.18) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.04,198.72) --
	(454.04,192.07);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (454.04,195.40) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.60,191.62) --
	(459.60,186.36);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (459.60,188.99) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.16,229.29) --
	(465.16,223.00);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (465.16,226.15) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (420.67,186.26) --
	(470.73,186.26);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (445.70,284.14) {4.8\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (476.29,  0.00) --
	(476.29,357.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- (554.16, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.59, 60.00) -- ( 78.59, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.78, 60.00) -- (139.78, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.96, 60.00) -- (200.96, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.14, 60.00) -- (262.14, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.33, 60.00) -- (323.33, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.51, 60.00) -- (384.51, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (445.70, 60.00) -- (445.70, 54.00);

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at ( 80.66, 48.00) {Prior};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (141.84, 48.00) {Likelihood};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (203.03, 48.00) {Leaf rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (264.21, 48.00) {Mean branch rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (325.39, 48.00) {Tree length};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (386.58, 48.00) {Clock SD $\sigma$};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (447.76, 48.00) {HKY model $\kappa$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 48.00,286.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 42.00, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 92.88) -- ( 42.00, 92.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,135.92) -- ( 42.00,135.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,173.19) -- ( 42.00,173.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,210.47) -- ( 42.00,210.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,253.51) -- ( 42.00,253.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,286.39) -- ( 42.00,286.39);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 56.56) {$\frac{2}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 89.44) {$\frac{3}{4}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,132.47) {$\frac{7}{8}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,169.75) {$1\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,207.03) {$\frac{8}{7}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,250.06) {$\frac{4}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,282.94) {$\frac{3}{2}\times$};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,241.67) rectangle (494.39,235.67);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,229.67) rectangle (494.39,223.67);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,217.67) rectangle (494.39,211.67);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,205.67) rectangle (494.39,199.67);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,193.67) rectangle (494.39,187.67);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,181.67) rectangle (494.39,175.67);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,169.67) rectangle (494.39,163.67);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,157.67) rectangle (494.39,151.67);
\definecolor{fillColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,145.67) rectangle (494.39,139.67);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,235.22) {Ran};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,223.22) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,211.22) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,199.22) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,187.22) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,175.22) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,163.22) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,151.22) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,139.22) {Moyle};
\end{scope}
\end{tikzpicture}

\end{document}