### 2020-09-01 benchmarking/benchmarkingVM/ESS_round5.tex

% Created by tikzDevice version 0.12.3.1 on 2020-09-01 15:30:48
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,357.33);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (301.08,335.19) {\bfseries Comparison of AVMVN+NER+Bactrian+adapt(\textit{real}) and NER+Bactrian+adapt(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (  9.60,190.67) {Relative ESS/hr};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 48.00,173.19) --
	(476.29,173.19);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 53.56,197.87) --
	( 53.56,173.22);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] ( 53.56,185.54) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.12,170.57) --
	( 59.12,160.89);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] ( 59.12,165.73) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 64.69,193.03) --
	( 64.69,169.27);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] ( 64.69,181.15) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.25,191.40) --
	( 70.25,174.61);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] ( 70.25,183.01) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.81,194.06) --
	( 75.81,177.12);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] ( 75.81,185.59) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.37,141.00) --
	( 81.37, 76.27);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] ( 81.37,108.63) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.94,183.38) --
	( 86.94,168.50);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] ( 86.94,175.94) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.50,170.18) --
	( 92.50,143.82);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] ( 92.50,157.00) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.06,196.84) --
	( 98.06,181.84);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] ( 98.06,189.34) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] ( 53.56,171.19) --
	(103.62,171.19);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 78.59,284.14) {0.72\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (109.18,  0.00) --
	(109.18,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.75,229.74) --
	(114.75,203.57);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (114.75,216.66) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.31,194.11) --
	(120.31,179.13);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (120.31,186.62) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (125.87,185.55) --
	(125.87,175.29);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (125.87,180.42) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (131.43,214.73) --
	(131.43,201.99);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (131.43,208.36) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.00,239.63) --
	(137.00,227.48);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (137.00,233.56) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (142.56,159.92) --
	(142.56,140.75);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (142.56,150.33) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.12,214.01) --
	(148.12,205.08);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (148.12,209.54) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (153.68,200.10) --
	(153.68,107.60);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (153.68,153.85) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.24,178.44) --
	(159.24,154.84);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (159.24,166.64) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (114.75,190.92) --
	(164.81,190.92);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (139.78,284.14) {6.6\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (170.37,  0.00) --
	(170.37,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (175.93,168.43) --
	(175.93,160.62);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (175.93,164.53) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.49,172.66) --
	(181.49,170.84);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (181.49,171.75) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.05,169.79) --
	(187.05,167.75);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (187.05,168.77) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (192.62,189.75) --
	(192.62,187.44);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (192.62,188.59) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.18,201.95) --
	(198.18,198.81);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (198.18,200.38) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.74,171.28) --
	(203.74,169.79);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (203.74,170.53) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.30,184.57) --
	(209.30,182.27);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (209.30,183.42) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (214.87,163.32) --
	(214.87,160.43);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (214.87,161.87) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.43,155.78) --
	(220.43,154.17);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (220.43,154.97) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (175.93,174.20) --
	(225.99,174.20);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (200.96,284.14) {0.36\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (231.55,  0.00) --
	(231.55,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.11,171.43) --
	(237.11,152.78);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (237.11,162.10) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.68,218.83) --
	(242.68,206.46);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (242.68,212.65) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.24,242.97) --
	(248.24,220.02);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (248.24,231.49) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.80,208.35) --
	(253.80,192.32);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (253.80,200.33) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (259.36,179.26) --
	(259.36,152.83);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (259.36,166.04) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.93,187.62) --
	(264.93,174.61);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (264.93,181.12) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (270.49,206.77) --
	(270.49,181.77);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (270.49,194.27) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.05,229.57) --
	(276.05,209.79);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (276.05,219.68) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (281.61,216.31) --
	(281.61,206.19);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (281.61,211.25) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (237.11,198.57) --
	(287.17,198.57);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (262.14,284.14) {9.5\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (292.74,  0.00) --
	(292.74,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.30,166.69) --
	(298.30,150.30);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (298.30,158.50) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (303.86,209.96) --
	(303.86,198.22);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (303.86,204.09) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (309.42,238.47) --
	(309.42,217.31);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (309.42,227.89) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (314.99,204.06) --
	(314.99,189.08);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (314.99,196.57) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (320.55,183.38) --
	(320.55,154.32);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (320.55,168.85) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (326.11,183.17) --
	(326.11,172.59);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (326.11,177.88) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (331.67,206.07) --
	(331.67,181.21);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (331.67,193.64) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (337.23,210.76) --
	(337.23,182.47);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (337.23,196.61) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.80,208.61) --
	(342.80,197.60);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (342.80,203.11) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (298.30,192.59) --
	(348.36,192.59);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (323.33,284.14) {7.2\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (353.92,  0.00) --
	(353.92,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.48,175.89) --
	(359.48,161.24);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (359.48,168.57) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.05,174.36) --
	(365.05,162.95);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (365.05,168.66) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (370.61,189.08) --
	(370.61,165.88);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (370.61,177.48) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.17,190.22) --
	(376.17,174.54);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (376.17,182.38) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.73,191.32) --
	(381.73,161.99);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (381.73,176.66) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.29,146.90) --
	(387.29,126.33);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (387.29,136.61) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (392.86,194.96) --
	(392.86,169.70);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (392.86,182.33) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.42,175.08) --
	(398.42,145.39);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (398.42,160.24) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (403.98,198.72) --
	(403.98,181.55);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (403.98,190.14) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (359.48,171.84) --
	(409.54,171.84);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (384.51,284.14) {0.48\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (415.11,  0.00) --
	(415.11,357.33);
\definecolor{drawColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.67,166.71) --
	(420.67,162.11);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (420.67,164.41) circle (  2.25);
\definecolor{drawColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.23,167.61) --
	(426.23,161.93);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (426.23,164.77) circle (  2.25);
\definecolor{drawColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.79,164.25) --
	(431.79,160.41);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (431.79,162.33) circle (  2.25);
\definecolor{drawColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.35,168.36) --
	(437.35,165.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (437.35,166.80) circle (  2.25);
\definecolor{drawColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (442.92,169.61) --
	(442.92,166.59);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (442.92,168.10) circle (  2.25);
\definecolor{drawColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (448.48,161.65) --
	(448.48,156.22);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (448.48,158.94) circle (  2.25);
\definecolor{drawColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.04,162.17) --
	(454.04,154.52);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (454.04,158.35) circle (  2.25);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.60,155.49) --
	(459.60,149.68);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (459.60,152.58) circle (  2.25);
\definecolor{drawColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (465.16,166.82) --
	(465.16,162.55);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (465.16,164.69) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (420.67,162.37) --
	(470.73,162.37);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (445.70,284.14) {3.8\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (476.29,  0.00) --
	(476.29,357.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- (554.16, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.59, 60.00) -- ( 78.59, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (139.78, 60.00) -- (139.78, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.96, 60.00) -- (200.96, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.14, 60.00) -- (262.14, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.33, 60.00) -- (323.33, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.51, 60.00) -- (384.51, 54.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (445.70, 60.00) -- (445.70, 54.00);

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at ( 80.66, 48.00) {Prior};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (141.84, 48.00) {Likelihood};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (203.03, 48.00) {Leaf rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (264.21, 48.00) {Mean branch rate};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (325.39, 48.00) {Tree length};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (386.58, 48.00) {Clock SD $\sigma$};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (447.76, 48.00) {HKY model $\kappa$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 48.00,286.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 60.00) -- ( 42.00, 60.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 92.88) -- ( 42.00, 92.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,135.92) -- ( 42.00,135.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,173.19) -- ( 42.00,173.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,210.47) -- ( 42.00,210.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,253.51) -- ( 42.00,253.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,286.39) -- ( 42.00,286.39);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 56.56) {$\frac{2}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00, 89.44) {$\frac{3}{4}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,132.47) {$\frac{7}{8}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,169.75) {$1\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,207.03) {$\frac{8}{7}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,250.06) {$\frac{4}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 36.00,282.94) {$\frac{3}{2}\times$};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 60.00) rectangle (554.16,321.33);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,241.67) rectangle (494.39,235.67);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,229.67) rectangle (494.39,223.67);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,217.67) rectangle (494.39,211.67);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,205.67) rectangle (494.39,199.67);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,193.67) rectangle (494.39,187.67);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,181.67) rectangle (494.39,175.67);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,169.67) rectangle (494.39,163.67);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,157.67) rectangle (494.39,151.67);
\definecolor{fillColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.19,145.67) rectangle (494.39,139.67);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,235.22) {Ran};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,223.22) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,211.22) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,199.22) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,187.22) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,175.22) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,163.22) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,151.22) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (503.39,139.22) {Moyle};
\end{scope}
\end{tikzpicture}

\end{document}