### 2020-09-01 benchmarking/benchmarkingVM/ESS_timesettings.tex

% Created by tikzDevice version 0.12.3.1 on 2020-09-01 15:31:18
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,357.33);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (301.08,328.59) {\bfseries Relative speed compared with \textit{cat}};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,184.67) {Relative samples/hr};
\end{scope}
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (552.96,308.13);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 2pt off 5pt on 2pt off 5pt ,line join=round,line cap=round] ( 49.20,184.67) --
	(459.67,184.67);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] ( 52.93,182.48) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] ( 56.66,177.12) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] ( 60.39,177.00) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] ( 64.13,177.26) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] ( 67.86,179.92) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] ( 71.59,167.44) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] ( 75.32,177.00) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] ( 79.05,175.65) circle (  2.25);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] ( 82.78,174.58) circle (  2.25);

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] ( 52.93,176.54) --
	( 86.52,176.54);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at ( 69.72,297.24) {4.5\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] ( 90.25,  0.00) --
	( 90.25,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] ( 93.98,212.46) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] ( 97.71,238.18) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (101.44,226.82) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (105.17,232.21) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (108.90,235.82) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (112.64,254.03) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (116.37,265.04) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (120.10,247.48) circle (  2.25);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (123.83,282.32) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] ( 93.98,244.94) --
	(127.56,244.94);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (110.77,297.24) {40\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (131.29,  0.00) --
	(131.29,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (135.03,179.49) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (138.76,159.08) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (142.49,169.55) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (146.22,168.89) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (149.95,173.84) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (153.68,156.37) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (157.42,135.12) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (161.15,158.88) circle (  2.25);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (164.88,140.87) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (135.03,160.77) --
	(168.61,160.77);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (151.82,297.24) {13\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (172.34,  0.00) --
	(172.34,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (176.07,216.47) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (179.80,231.86) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (183.54,221.99) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (187.27,229.14) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (191.00,229.26) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (194.73,228.99) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (198.46,265.68) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (202.19,246.83) circle (  2.25);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (205.93,256.19) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (176.07,236.95) --
	(209.66,236.95);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (192.86,297.24) {34\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (213.39,  0.00) --
	(213.39,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (217.12,166.67) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (220.85,147.39) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (224.58,147.20) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (228.31,150.68) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (232.05,161.99) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (235.78,151.74) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (239.51,170.65) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (243.24,155.66) circle (  2.25);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (246.97,179.55) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (217.12,159.38) --
	(250.70,159.38);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (233.91,297.24) {13\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (254.44,  0.00) --
	(254.44,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (258.17,167.17) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (261.90,139.53) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (265.63,141.49) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (269.36,140.86) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (273.09,152.88) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (276.82,131.96) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (280.56,147.94) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (284.29,149.69) circle (  2.25);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (288.02,156.57) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (258.17,147.85) --
	(291.75,147.85);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (274.96,297.24) {19\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (295.48,  0.00) --
	(295.48,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (299.21,164.07) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (302.95,125.97) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (306.68,129.45) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (310.41,129.08) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (314.14,140.95) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (317.87,118.44) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (321.60,121.35) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (325.34,136.49) circle (  2.25);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (329.07,144.64) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (299.21,135.00) --
	(332.80,135.00);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (316.01,297.24) {24\% slower};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (336.53,  0.00) --
	(336.53,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (340.26,216.42) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (343.99,235.48) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (347.72,222.58) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (351.46,228.09) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (355.19,229.31) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (358.92,233.69) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (362.65,265.46) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (366.38,249.84) circle (  2.25);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (370.11,260.21) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (340.26,238.62) --
	(373.85,238.62);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (357.05,297.24) {35\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (377.58,  0.00) --
	(377.58,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (381.31,218.13) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (385.04,228.31) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (388.77,220.20) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (392.50,226.36) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (396.23,229.72) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (399.97,231.76) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (403.70,254.94) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (407.43,240.20) circle (  2.25);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (411.16,228.85) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (381.31,231.25) --
	(414.89,231.25);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (398.10,297.24) {30\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (418.62,  0.00) --
	(418.62,357.33);
\definecolor{fillColor}{RGB}{255,0,109}

\path[fill=fillColor] (422.36,214.96) circle (  2.25);
\definecolor{fillColor}{RGB}{253,104,0}

\path[fill=fillColor] (426.09,222.09) circle (  2.25);
\definecolor{fillColor}{RGB}{143,161,0}

\path[fill=fillColor] (429.82,212.46) circle (  2.25);
\definecolor{fillColor}{RGB}{0,195,0}

\path[fill=fillColor] (433.55,222.04) circle (  2.25);
\definecolor{fillColor}{RGB}{0,219,171}

\path[fill=fillColor] (437.28,225.80) circle (  2.25);
\definecolor{fillColor}{RGB}{0,216,255}

\path[fill=fillColor] (441.01,224.41) circle (  2.25);
\definecolor{fillColor}{RGB}{151,0,255}

\path[fill=fillColor] (444.74,244.63) circle (  2.25);
\definecolor{fillColor}{RGB}{255,0,255}

\path[fill=fillColor] (448.48,226.72) circle (  2.25);
\definecolor{fillColor}{RGB}{68,68,68}

\path[fill=fillColor] (452.21,222.15) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.2pt,line join=round,line cap=round] (422.36,224.13) --
	(455.94,224.13);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.70] at (439.15,297.24) {25\% faster};
\definecolor{drawColor}{RGB}{105,105,105}

\path[draw=drawColor,draw opacity=0.60,line width= 0.2pt,line join=round,line cap=round] (459.67,  0.00) --
	(459.67,357.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,357.33);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 61.20) -- (552.96, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 69.72, 61.20) -- ( 69.72, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.77, 61.20) -- (110.77, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.82, 61.20) -- (151.82, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (192.86, 61.20) -- (192.86, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.91, 61.20) -- (233.91, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.96, 61.20) -- (274.96, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.01, 61.20) -- (316.01, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.05, 61.20) -- (357.05, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.10, 61.20) -- (398.10, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (439.15, 61.20) -- (439.15, 55.20);

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at ( 71.79, 49.20) {adapt(\textit{cat})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (112.84, 49.20) {nocons(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (153.88, 49.20) {cons(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (194.93, 49.20) {adapt(\textit{real})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (235.98, 49.20) {nocons(\textit{quant})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (277.03, 49.20) {cons(\textit{quant})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (318.07, 49.20) {adapt(\textit{quant})};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (359.12, 49.20) {Bactrian};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (400.17, 49.20) {NER};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.60] at (441.21, 49.20) {AVMVN};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 61.20) -- ( 49.20,308.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 61.20) -- ( 43.20, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,133.42) -- ( 43.20,133.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,184.67) -- ( 43.20,184.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,235.91) -- ( 43.20,235.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,308.13) -- ( 43.20,308.13);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20, 57.76) {$\frac{1}{2}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,129.98) {$\frac{3}{4}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,181.22) {$1\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,232.46) {$\frac{4}{3}\times$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 37.20,304.69) {$\frac{2}{1}\times$};
\end{scope}
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (552.96,308.13);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{255,0,109}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,235.67) rectangle (493.19,229.67);
\definecolor{fillColor}{RGB}{253,104,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,223.67) rectangle (493.19,217.67);
\definecolor{fillColor}{RGB}{143,161,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,211.67) rectangle (493.19,205.67);
\definecolor{fillColor}{RGB}{0,195,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,199.67) rectangle (493.19,193.67);
\definecolor{fillColor}{RGB}{0,219,171}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,187.67) rectangle (493.19,181.67);
\definecolor{fillColor}{RGB}{0,216,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,175.67) rectangle (493.19,169.67);
\definecolor{fillColor}{RGB}{151,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,163.67) rectangle (493.19,157.67);
\definecolor{fillColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,151.67) rectangle (493.19,145.67);
\definecolor{fillColor}{RGB}{68,68,68}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.99,139.67) rectangle (493.19,133.67);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,229.22) {Ran};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,217.22) {Cognato};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,205.22) {Dornburg};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,193.22) {Sauquet};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,181.22) {Broughton};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,169.22) {Kawahara};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,157.22) {simulated};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,145.22) {Rightmyer};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (502.19,133.22) {Moyle};
\end{scope}
\end{tikzpicture}

\end{document}